### 2023-06-01 manuscript.Rmd
group all scripts and text into one Rmd file

---
title: "Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass"
author: "Marguerite Cogné, Régis Peltier, Vincent Freycon, Alexis Toumazeau, Camille Piponiot"
date: "2023-05-10"
output: 
  bookdown::word_document2:
  fig_caption: yes
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide", tinytex.clean = FALSE)
packages_needed = c("ggplot2", "dataverse", "readxl", "data.table")

packages_to_install = packages_needed[!( packages_needed %in% rownames(installed.packages()))]

if (length(packages_to_install) > 0)
  install.packages(packages_to_install)

lapply(packages_needed, require, character.only = TRUE)
```

# Introduction

Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO_2_e/yr over 2010-2014; @Pendrill2019]. The demand for tropical agricultural land is expected to continue to increase in the tropics as the population grows [@UN2022]. In addition, tropical areas are expected to be disproportionately affected by deforestation and climate change, putting their agriculture at risk [@Lawrence2015]. There is therefore an urgent need to promote tropical agricultural landscapes that are both more resilient to climate change, while maintaining high productivity and ecological value [@Harvey2014]. In particular, increasing the carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of the food production system [@Lal2004].

Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Verchot2007]. Agroforestry is a collective name for land use systems and technologies where woody perennials (trees, shrubs, palms, bamboo, etc.) are deliberately used on the same land management units as agricultural crops and/or animals, in some form of spatial arrangement or temporal sequence [@Lundgren1982]. Agroforestry allows for diversification of agricultural production, making agricultural systems more resilient [@Duffy2021]; it also increases the carbon sequestration in the vegetation but also in the soil [@DeStefano2018].

Carbon stocks in AFS [VF]

Importance of the ecological/social context in the success of agroforestry techniques - existence of traditional agroforestry techniques that could be more adapted to local conditions, [MC]
Agroforestry needs to be adapted to the context [(Coe et al., 2014, Cur Opinion in Env Sust) @Coe2014]

Existence of traditional agroforestry systems [examples from Marguerite’s intro]

Context in Timor Leste: current knowledge on their agroforestry practices + knowledge gap [MC]

Local typology of production systems and why it is not adapted to traditional agroforestry

No prior description of traditional agroforestry systems in TL

In Timor Leste, a typology of production systems taking into account different agroecological zoning was carried out in 2017 based on the results of the 2010 national census (R. Williams et al, 2018). However, this typology does not highlight the agroforestry component of Timorese agriculture by focusing on other differentiating criteria such as livestock, subsistence crops (rice) or the island's historical cash crops (coconut, coffee).

Need to better understand how these systems work and quantify their carbon storage to include them in climate change mitigation policies 
Research questions  [VF]

What are the characteristics of traditional agroforestry systems in Timor Leste?

How much carbon do they store in both their biomass and in the soil

Are there measurable differences between AFS?

Hypotheses [VF]

Proposition: lien entre typologie et stocks de carbone

# Materials and methods

```{r download-data}
## download data from dataverse - once published
dataverse_url <- "https://dataverse.cirad.fr/privateurl.xhtml?token=f8eabb2a-3635-499b-aee5-80fd2b7495a0"
dataverse_doi <- "10.18167/DVN1/QCXWIY"

# for now : download manually from temporary link
# utils::unzip("dataverse_files.zip")
```

```{r agb-calculation}
# get agb data
source("script-agb.R")
```

```{r open-soil-data}
# open organic carbon data
# bulk density in Gariuai according to SoilGrids = 125 cg/cm3
data_soil <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")

# add bulk density and coarse fraction
data_soil2 <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")

# bulk density from direct measurements
data_soil2$BD <- data_soil2$Sample_DRY/data_soil2$V_cylinder

# coarse fraction
data_soil2$CF <- data_soil2$V_rocks_tot/data_soil2$V_cylinder

# change format of depth to correspond to other soil table
data_soil2$SAMPLE_DEPTH <- factor(data_soil2$DEPTH_SAMPLE_C, levels = c("2_7", "12_17", "22_27"))
levels(data_soil2$SAMPLE_DEPTH) <- c("0-10", "10-20", "20-30")

data_soil <- merge(data_soil, data_soil2, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
data.table::setDT(data_soil)
```

```{r test-bd-pedotransfer}
# bulk density from pedotransfer function: Minasny and Hartemink, 2011

data_soil$depth_mean <- unlist(lapply(strsplit(data_soil$SAMPLE_DEPTH, "-"), function(x) mean(as.numeric(x))))
data_soil$depth_max <- unlist(lapply(strsplit(data_soil$SAMPLE_DEPTH, "-"), function(x) as.numeric(x[2])))

data_soil$sand <- as.numeric(data_soil$`FINE SAND`) + as.numeric(data_soil$`COARSE SAND`)/100

data_soil$BDmin <- 
  0.935 + 
  0.049 * log(data_soil$depth_mean) + 
  0.0055 * data_soil$sand + 
  0.000065 * (data_soil$sand - 38.96)^2

data_soil$BDptf <- 
  100/(
    data_soil$OM/0.224 + 
      (100 - data_soil$OM)/data_soil$BDmin
  )

# g1 = ggplot(data_soil, aes(x = BD, y = BDptf, color = ID_TYPE)) + 
#   geom_abline(slope = 1, intercept = 0, lty=2) + 
#   geom_point() + 
#   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
#   lims(x = c(0.8, 1.6), y = c(0.8, 1.6)) +
#   facet_wrap(~ID_VILLAGE)
# #+
#   # theme(legend.position = "none")
# g2 = ggplot(data_soil, aes(x = BD, y = BDmin, color = SAMPLE_DEPTH)) +
#   geom_abline(slope = 1, intercept = 0, lty=2) + 
#   geom_point() + 
#   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
#   lims(x = c(0.8, 1.6), y = c(0.8, 1.6)) +
#   theme(legend.position = c(0.2, 0.8))
# 
# ggpubr::ggarrange(g1, g2) 
# ggsave("figures/pedotransfer-test.png", height = 5, width = 10)
```

```{r add-missing-values}
## add intermediate soil carbon in 10-20 cm depth (for now: no value)
data_soil[, Corg := mean(`Corg\n`, na.rm = TRUE), .(N_PLOT)]
data_soil[!is.na(`Corg\n`), Corg := `Corg\n`]

# replace one missing value of coarse fraction with plot level mean
data_soil[, CF_site := mean(CF, na.rm = TRUE), .(N_PLOT)]
data_soil[is.na(CF), CF := CF_site]
```

```{r total-soc}
# calculate total soil carbon stock per depth
data_soil[, c("h0", "h1") := lapply(data.table::tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
data_soil[, Cstock :=          ## in Mg/ha
            (h1 - h0) * 1e8 *  ## equivalent volume of a 1 ha soil sample, in cm3/ha
            BD *               ## bulk density 
            Corg/100 *         ## carbon organic content in soil
            (1-CF) *           ## remove coarse fraction
            1e-6               ## convert from g/ha to Mg/ha
]

data_soil_C <- data_soil[, .(meanSC = sum(Cstock), clay = mean(as.numeric(CLAY), na.rm = TRUE)), .(N_PLOT)]
```

```{r soc-error-propagation}
## error propagation using a Monte Carlo method -------------
data_soil_unc = data_soil[, .(BD = rnorm(1000, BD, BD*0.1), 
                              Corg = rnorm(1000, Corg, Corg*0.12), 
                              CF = rnorm(1000, CF, CF*0.15), 
                              iter = 1:1000), 
                          .(N_PLOT, SAMPLE_DEPTH, h1, h0)]

data_soil_unc = data_soil_unc[, .(Cstock = sum(
  (h1 - h0) * 1e8 *  
    BD * Corg/100 * (1-CF) * 1e-6  
)),
.(N_PLOT, iter)
]

data_soil_unc = data_soil_unc[, .(
  lwrSC = quantile(Cstock, 0.025), 
  uprSC = quantile(Cstock, 0.975)), 
  .(N_PLOT)]
```

```{r combine-agb-soc-data}
# combine AGB and soil C
dataC <- merge(data_agb, data_soil_C)

# add soil carbon uncertainty
dataC <- merge(dataC, data_soil_unc)

dataC[, village := substr(ID_VILLAGE, 1, 1)]
dataC$ID_AF <- factor(dataC$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
```

# Results

```{r tab-mean, results = "show"}
# average values per village and SAF type
dataC_mean <- dataC[, .(AGC = paste0(round(mean(meanAGC), 2), " +/- ", round(sd(meanAGC)/sqrt(length(meanAGC)), 2)), 
                        SC = paste0(round(mean(meanSC), 2), " +/- ", round(sd(meanSC)/sqrt(length(meanSC)), 2))),
                        .(ID_AF)]
dataC_mean$ID_AF <- factor(dataC_mean$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
dataC_mean = dataC_mean[order(dataC_mean$ID_AF),]
knitr::kable(dataC_mean, col.names = c("Agroforestry system", "Aboveground carbon (Mg/ha)", "Soil organic carbon (Mg/ha)"))
write.csv(dataC_mean, file = "tables/C_mean_SAF.csv", row.names = FALSE)
```

```{r fig-1-map, fig.height =6, fig.width = 7}
# required packages
# req_packages <- c("BIOMASS", "readxl", "data.table", "ggmap", "ggplot2")

## 2. Load data from excel file ####
data_coord <- read_xlsx("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = "GPS")
data.table::setDT(data_coord)

data_coord[, N_PLOT := as.numeric(data.table::tstrsplit(N_Q_ID, "_")[[1]])]

data_coord <- data_coord[, .(long = mean(as.numeric(LONG)), lat = mean(as.numeric(LAT))), .(N_PLOT)]

data_plot <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")

data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])

data_plot[, village := substr(ID_VILLAGE, 1, 1)]
data_plot[village == "S", village := "Samalari"]
data_plot[village == "G", village := "Gariuai"]

data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(village)]
data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
data_box[village == "Samalari", w := 0.05]
data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]

## create Timor map ####

bbox <- c(124.5, -9.2, 127.5,  -8.3)
# basemap <- ggmap::get_stamenmap(bbox, maptype = "terrain", zoom = 10)

# geodata::elevation_3s(lon= c(data_box[village == "Samalari"]$x1), 
#                       lat= c(data_box[village == "Samalari"]$y1), 
#                       path = "data")
timor_map <- map_data("world", region = "Timor-Leste")

# get elevation information
elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
elevS <- subset(elev, x > data_box[village == "Samalari"]$x1 & 
                  x < data_box[village == "Samalari"]$x2 & 
                  y > data_box[village == "Samalari"]$y1 & 
                  y < data_box[village == "Samalari"]$y2)
elevG <- subset(elev, x > data_box[village == "Gariuai"]$x1 & 
                  x < data_box[village == "Gariuai"]$x2 & 
                  y > data_box[village == "Gariuai"]$y1 & 
                  y < data_box[village == "Gariuai"]$y2)

# Timor map (fig 1a)
g1 = ggplot(timor_map) + 
  # lims(x = c(124.5, 127.5), y = c(-9.2, -8.3)) +
  #ggmap::ggmap(basemap) +
  geom_polygon(aes(x = long, y = lat, group = group), fill="lightgray", colour = "grey") +
  geom_tile(data = data_box, aes(x = xc, y = yc, width = w, height = w), fill = NA, color = 1, linewidth = 0.5) +
  ggrepel::geom_text_repel(data = data_box, aes(x = xc, y = yc, label = village)) + 
  labs(x = "", y = "") +
  theme(legend.position = "top", axis.title = element_blank(), 
        axis.ticks = element_blank(), axis.text = element_blank(), 
        panel.background = element_blank())+
  coord_sf(crs = 'WGS84')

# Gariuai map (fig 1b)
g2 = ggplot(data = elevG) +
  geom_raster(aes(x = x, y = y, fill=srtm_62_14)) +
  geom_point(data = subset(data_plot, village == "Gariuai"), 
             aes(x = long, y = lat, color = ID_AF), size = 2) + 
  geom_point(data = subset(data_plot, village == "Gariuai"), 
             aes(x = long, y = lat), shape = 1, size = 2) + 
  # coord_equal() +
  viridis::scale_fill_viridis(option = "D",
                              limits = range(c(elevS$srtm_62_14, elevG$srtm_62_14))
  ) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  labs(x = "", y = "", title = "Gariuai", color = "", fill="Elevation (m)") +
  guides(color = FALSE, fill=guide_colorbar(title.position = "top")) + 
  theme(legend.position = "top", axis.title = element_blank(), 
        axis.ticks = element_blank(), axis.text = element_blank(), 
        plot.title = element_text(hjust = 0.5), 
        panel.background = element_blank()) +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
  ggspatial::annotation_scale() +
  coord_sf(crs = 'WGS84')

l1 = ggpubr::get_legend(g2)

g2 = g2 + theme(legend.position = "none")
# 
# g2 + ggspatial::annotation_scale() +
#   coord_sf(crs = 'WGS84')

# Samalari map (fig 1c)
g3 = ggplot(data = elevS) +
  geom_raster(aes(x = x, y = y, fill=srtm_62_14)) +

  geom_point(data = subset(data_plot, village == "Samalari"), 
             aes(x = long, y = lat, color = ID_AF), size = 2) + 
  geom_point(data = subset(data_plot, village == "Samalari"), 
             aes(x = long, y = lat), shape = 1, size = 2) + 
  coord_equal() +
  labs(x = "", y = "", title = "Samalari", color = "AF\nsystem", fill="Elevation (m)") +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  viridis::scale_fill_viridis(option = "D",
                              limits = range(c(elevS$srtm_62_14, elevG$srtm_62_14)))+
  guides(color=guide_legend(nrow=2,byrow=TRUE), fill=FALSE) +
  theme(legend.position = "bottom", axis.title = element_blank(), 
        axis.ticks = element_blank(), axis.text = element_blank(), 
        plot.title = element_text(hjust = 0.5), 
        panel.background = element_blank()) + 
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
  ggspatial::annotation_scale() +
  coord_sf(crs = 'WGS84')

l2 = ggpubr::get_legend(g3)
g3 = g3 + theme(legend.position = "none")

ggpubr::ggarrange(g1, 
                  ggpubr::ggarrange(g2, 
                                    ggpubr::ggarrange(g3, l1, l2, ncol=1, heights = c(4,1,1)), 
                                    widths = c(4,3), labels = c("b", "c")), 
                  heights = c(1, 3), nrow = 2, labels = c("a", ""))

ggsave("figures/mapPlotsSRTM.png", height =6, width = 7)
```

```{r anova}
## identify otuliers
# library(magrittr)
library(rstatix)
dataC %>% 
  group_by(ID_AF) %>%
  identify_outliers(meanAGC)

dataC %>% 
  group_by(ID_AF) %>%
  identify_outliers(meanSC)

# remove extreme outlier and test normality
dataC[N_PLOT != 11] %>%
  group_by(ID_AF) %>%
  shapiro_test(meanAGC)

dataC %>%
  group_by(ID_AF) %>%
  shapiro_test(meanSC)

# non parametric kruskal wallis test 
kruskalAGC = kruskal.test(meanAGC ~ ID_AF,  data = dataC)
kruskal.test(meanSC ~ ID_AF,  data = dataC)
kruskal.test(meanAGC ~ village,  data = dataC)
kruskal.test(meanSC ~ village,  data = dataC)

# pairwise group comparison with Tukey HSD method
# https://arc.lib.montana.edu/book/statistics-with-r-textbook/item/59
library(multcomp)
aovAGC = aov(meanAGC ~ ID_AF, data = dataC[N_PLOT != 11])
tuk <- glht(aovAGC, linfct = mcp(ID_AF = "Tukey"))
summary(tuk)          # standard display
tuk.cld <- cld(tuk)   # letter-based display

plot(tuk.cld)

# soc
aovSOC = aov(meanSC ~ ID_AF, data = dataC)
tuk <- glht(aovSOC, linfct = mcp(ID_AF = "Tukey"))
summary(tuk)          # standard display
tuk.cld <- cld(tuk)   # letter-based display

opar <- par(mai=c(1,1,1.5,1))
plot(tuk.cld)
```

```{r fig-2-barplot, fig.height = 5, fig.width = 6}
## melt data.table
dataC_all <- melt(dataC, id.vars = c("N_PLOT", "ID_VILLAGE", "ID_AF", "village"), 
              measure.vars = c("meanAGC", "meanBGC", "meanSC"))

dataC_all <- dataC_all[, .(value = mean(value)), .(village, ID_AF, variable)]

## reorder factors

dataC_all$ID_AF <- factor(dataC_all$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))

# uncertainty fig 2: bootstrap by MC iteration?

dataC_all[, variable := factor(variable, levels = c("meanAGC", "meanBGC", "meanSC"))]
levels(dataC_all$variable) <- c("AGC", "BGC", "SOC")

ggplot(dataC_all, aes(x = ID_AF, y = value, fill = variable)) +
  geom_bar(position="stack", stat="identity") + 
  facet_wrap(~village, nrow = 1, strip.position = "bottom") +
  scale_fill_manual(values = c("forestgreen", "darkgoldenrod1", "coral4")) +
  theme_classic() +
  theme(strip.placement = "outside") +
  labs(x = "", y = "Carbon stocks (Mg C/ha)", fill = "Carbon\npool") +
  scale_y_continuous(expand = c(0, 0))
ggsave("figures/fig-carbon-stocks-stack.png", height = 5, width = 6)
```

```{r fig-3-reg, fig.height = 5, fig.width = 7}
## mixed effect linear regression
lmeReg <- lmerTest::lmer(meanSC ~ meanAGC + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R2 = performance::r2_nakagawa(lmeReg)

## figure
ggplot(dataC, aes(color = ID_AF)) +
  geom_point(aes(x = meanAGC, y = meanSC, shape = village), size = 2) + 
  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 1, parse = TRUE,
           label = paste0("R^{2}*' (marg.) = '*", round(R2$R2_marginal, 3))) +
  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 2, parse = TRUE,
           label = paste0("R^{2}*' (cond.) = '*", round(R2$R2_conditional, 3))) +
  geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
  geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
  geom_abline(intercept = summary(lmeReg)$coefficients[1,1], slope = summary(lmeReg)$coefficients[2,1], lty = 2) +
  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
  expand_limits(x = 0, y = 0) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  # scale_linetype_manual(values=c("twodash", "dotted"))+
  theme_classic() 
ggsave("figures/fig-scatterplot-carbon.png", height = 5, width = 7)
```

# Supplementary material

```{r fig-distr-sol, fig.height = 5, fig.width = 15}
## diagnostic données sol pour marguerite
g1 <- ggplot(data_soil, aes(x = CF*100, col = SAMPLE_DEPTH)) + 
  geom_histogram(bins = 30) +
  geom_text(data = subset(data_soil, CF > 0.3), angle = -60, 
            aes(x = CF*100-5, y = 15, 
                label = paste("Village:", ID_VILLAGE, "/ SAF:", ID_TYPE, "/ Plot:", N_PLOT))) +
  labs(x = "Elements grossiers (% du volume)", 
       y = "Nombre d'échantillons (site x profondeur)", 
       col = "Profondeur (cm)") + 
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic()  +
  theme(legend.position = c(0.8, 0.9))

g2 <- ggplot(data_soil, aes(x = Corg, col = SAMPLE_DEPTH)) + 
  geom_histogram(bins = 20) +
  labs(x = "Teneur en carbone (%)", 
       y = "Nombre d'échantillons (site x profondeur)", 
       col = "Profondeur (cm)") + 
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.9))

g3 <- ggplot(data_soil, aes(x = BD, col = SAMPLE_DEPTH)) + 
  geom_histogram(bins = 20) +
  labs(x = "Densité apparente", 
       y = "Nombre d'échantillons (site x profondeur)", 
       col = "Profondeur (cm)") + 
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic()+
  theme(legend.position = c(0.8, 0.9))

ggpubr::ggarrange(g1, g2, g3, labels = "auto", nrow = 1)
ggsave("figures/fig-distr-sol.pdf", height = 5, width = 15)
```

```{r fig-clay-corg, fig.height = 4, fig.width = 6}
ggplot(data_soil, aes(x = as.numeric(CLAY), y = Corg, col = SAMPLE_DEPTH)) +
  geom_point() +
  labs(col = "Profondeur (cm)", x = "Argile (%)", y = "Teneur en carbone (%)") + 
  geom_smooth(method = "lm") +
  theme_classic()
ggsave("figures/fig-clay-corg.pdf", height = 4, width = 6)
```

# References

### 2023-06-01 making-figures-from-raw-data.R

@@ -1,260 +1,0 @@
-
-# set working directory to match the source file
-setwd(dirname(rstudioapi::getSourceEditorContext()$path))
-
-# install and load packages
-
-packages_needed = c("ggplot2", "dataverse", "readxl", "data.table")
-
-packages_to_install = packages_needed[!( packages_needed %in% rownames(installed.packages()))]
-
-if (length(packages_to_install) > 0)
-  install.packages(packages_to_install)
-
-lapply(packages_needed, require, character.only = TRUE)
-
-## download data from dataverse - once published
-dataverse_url <- "https://dataverse.cirad.fr/privateurl.xhtml?token=f8eabb2a-3635-499b-aee5-80fd2b7495a0"
-dataverse_doi <- "10.18167/DVN1/QCXWIY"
-
-
-# for now : download manually from temporary link
-# utils::unzip("dataverse_files.zip")
-
-# get agb data
-source("script-agb.R")
-
-
-# Prepare the soil carbon data ####
-
-# open organic carbon data
-# bulk density in Gariuai according to SoilGrids = 125 cg/cm3
-data_soil <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")
-
-# add bulk density and coars fraction
-data_soil2 <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")
-
-# bulk density from direct measurements
-data_soil2$BD <- data_soil2$Sample_DRY/data_soil2$V_cylinder
-
-# coarse fraction
-data_soil2$CF <- data_soil2$V_rocks_tot/data_soil2$V_cylinder
-
-# change format of depth to correspond to other soil table
-data_soil2$SAMPLE_DEPTH <- factor(data_soil2$DEPTH_SAMPLE_C, levels = c("2_7", "12_17", "22_27"))
-levels(data_soil2$SAMPLE_DEPTH) <- c("0-10", "10-20", "20-30")
-
-data_soil <- merge(data_soil, data_soil2, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
-
-# bulk density from pedotransfer function: Minasny and Hartemink, 2011
-
[…]

### 2023-06-01 plot-map.R

@@ -1,138 +1,0 @@
-### make map of plots for publication ------------
-
-## 1. Prepare R environment ####
-
-# R packages - installation
-
-# required packages
-req_packages <- c("BIOMASS", "readxl", "data.table", "ggmap", "ggplot2")
-
-# install missing packages
-inst_packages <- req_packages[!( req_packages %in% rownames(installed.packages()))]
-
-if (length(inst_packages) > 0) install.packages(inst_packages)
-
-library(ggmap)
-library(ggplot2)
-
-# Set working directory to source file location
-setwd(dirname(rstudioapi::getSourceEditorContext()$path))
-
-
-## 2. Load data from excel file ####
-
-data_coord <- read.csv("data/coords-timor.csv")
-data.table::setDT(data_coord)
-
-data_coord[, N_PLOT := as.numeric(data.table::tstrsplit(N_Q_ID, "_")[[1]])]
-
-data_coord <- data_coord[, .(long = mean(LONG), lat = mean(LAT)), .(N_PLOT)]
-
-data_plot <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")
-
-data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])
-
-data_plot[, village := substr(ID_VILLAGE, 1, 1)]
-data_plot[village == "S", village := "Samalari"]
-data_plot[village == "G", village := "Gariuai"]
-
-data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(village)]
-data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
-data_box[village == "Samalari", w := 0.05]
-data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
-
-## 3. create Timor map ####
-
-bbox <- c(124.5, -9.2, 127.5,  -8.3)
-# basemap <- ggmap::get_stamenmap(bbox, maptype = "terrain", zoom = 10)
-
-# geodata::elevation_3s(lon= c(data_box[village == "Samalari"]$x1), 
-#                       lat= c(data_box[village == "Samalari"]$y1), 
-#                       path = "data")
-timor_map <- map_data("world", region = "Timor-Leste")
-
-elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
[…]

### 2024-02-09 manuscript.Rmd
new version of the figures after discussions in feb 2024

@@
 
-data_soil_C <- data_soil[, .(meanSC = sum(Cstock), clay = mean(as.numeric(CLAY), na.rm = TRUE)), .(N_PLOT)]
+data_soil_C <-
+  data_soil[, .(
+    meanSC = sum(Cstock),
+    SC10 = Cstock[SAMPLE_DEPTH == "0-10"],
+    clay = mean(as.numeric(CLAY), na.rm = TRUE)
+  ), .(N_PLOT)]
+
+setkey(data_soil_C, "N_PLOT")
 ```
@@
 
-data_soil_unc = data_soil_unc[, .(Cstock = sum(
+data_soil_unc = data_soil_unc[, .(
+  Cstock = sum(
     (h1 - h0) * 1e8 *  
       BD * Corg/100 * (1-CF) * 1e-6  
-)),
+  ), 
+  Cstock10 = (
+    (h1 - h0) * 1e8 *  
+      BD * Corg/100 * (1-CF) * 1e-6  
+  )[SAMPLE_DEPTH == "0-10"]),
   .(N_PLOT, iter)
@@
   lwrSC = quantile(Cstock, 0.025), 
-  uprSC = quantile(Cstock, 0.975)), 
+  uprSC = quantile(Cstock, 0.975),
+  lwrSC10 = quantile(Cstock10, 0.025), 
+  uprSC10 = quantile(Cstock10, 0.975)), 
   .(N_PLOT)]
@@
 # combine AGB and soil C
-dataC <- merge(data_agb, data_soil_C)
+dataC <- merge(data_agb, data_soil_C, by = "N_PLOT", all=TRUE)
 
 # add soil carbon uncertainty
-dataC <- merge(dataC, data_soil_unc)
+dataC <- merge(dataC, data_soil_unc, all=TRUE)
+
+dataC[N_PLOT==6, `:=`(ID_VILLAGE = "GD", ID_AF = "CF")]
 
@@
 # average values per village and SAF type
-dataC_mean <- dataC[, .(AGC = paste0(round(mean(meanAGC), 2), " +/- ", round(sd(meanAGC)/sqrt(length(meanAGC)), 2)), 
-                        SC = paste0(round(mean(meanSC), 2), " +/- ", round(sd(meanSC)/sqrt(length(meanSC)), 2))),
+dataC_mean <- dataC[, .(AGC = paste0(round(mean(meanAGC[!is.na(meanAGC)]), 2), " +/- ", round(sd(meanAGC[!is.na(meanAGC)])/sqrt(length(meanAGC[!is.na(meanAGC)])), 2)), 
+                        SC = paste0(round(mean(meanSC[!is.na(meanSC)]), 2), " +/- ", round(sd(meanSC[!is.na(meanSC)])/sqrt(length(meanSC[!is.na(meanSC)])), 2))),
                     .(ID_AF)]
@@
 
+```{r tab-anova, results = "show"}
+# 2-way anova 
+dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
+anov1 = anova(lm(meanAGC~ID_SUCO+ID_AF, data = dataC))
+anov2 = anova(lm(meanSC~ID_SUCO+ID_AF, data = dataC))
+tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
+                      ind_var = rep(c("SUCO", "AF"), 2), 
+                      p = round(c(anov1$`Pr(>F)`[1:2], anov2$`Pr(>F)`[1:2]), 3)
+)
+```
+
 ```{r fig-1-map, fig.height =6, fig.width = 7}
-# required packages
-# req_packages <- c("BIOMASS", "readxl", "data.table", "ggmap", "ggplot2")
 
@@
 data_plot[, village := substr(ID_VILLAGE, 1, 1)]
-data_plot[village == "S", village := "Samalari"]
+data_plot[village == "S", village := "Osso Luga"]
 data_plot[village == "G", village := "Gariuai"]
@@
 data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
-data_box[village == "Samalari", w := 0.05]
+data_box[village == "Osso Luga", w := 0.05]
 data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
@@
 
-# geodata::elevation_3s(lon= c(data_box[village == "Samalari"]$x1), 
-#                       lat= c(data_box[village == "Samalari"]$y1), 
+# geodata::elevation_3s(lon= c(data_box[village == "Osso Luga"]$x1), 
+#                       lat= c(data_box[village == "Osso Luga"]$y1), 
 #                       path = "data")
@@
 
-# get elevation information
-elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
-elevS <- subset(elev, x > data_box[village == "Samalari"]$x1 & 
-                  x < data_box[village == "Samalari"]$x2 & 
-                  y > data_box[village == "Samalari"]$y1 & 
-                  y < data_box[village == "Samalari"]$y2)
-elevG <- subset(elev, x > data_box[village == "Gariuai"]$x1 & 
-                  x < data_box[village == "Gariuai"]$x2 & 
-                  y > data_box[village == "Gariuai"]$y1 & 
-                  y < data_box[village == "Gariuai"]$y2)
-
-# Timor map (fig 1a)
 g1 = ggplot(timor_map) + 
-  # lims(x = c(124.5, 127.5), y = c(-9.2, -8.3)) +
-  #ggmap::ggmap(basemap) +
+  annotate(geom = "text", label = "a", x = -Inf, y = Inf, 
+           hjust = 0, vjust = 1, fontface = "bold", size = 6) + 
   geom_polygon(aes(x = long, y = lat, group = group), fill="lightgray", colour = "grey") +
@@
 
-# Gariuai map (fig 1b)
-g2 = ggplot(data = elevG) +
+# get elevation information
+elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
+elevS <- subset(elev, x > data_box[village == "Osso Luga"]$x1 & 
+                  x < data_box[village == "Osso Luga"]$x2 & 
+                  y > data_box[village == "Osso Luga"]$y1 & 
+                  y < data_box[village == "Osso Luga"]$y2)
+elevS$village = "Osso Luga"
+elevG <- subset(elev, x > data_box[village == "Gariuai"]$x1 & 
+                  x < data_box[village == "Gariuai"]$x2 & 
+                  y > data_box[village == "Gariuai"]$y1 & 
+                  y < data_box[village == "Gariuai"]$y2)
+elevG$village = "Gariuai"
+
+elevAll = rbind(elevS, elevG)
+
+## Map of villages and plots with elevation
+g2 = 
+  ggplot(data = elevAll) +
   geom_raster(aes(x = x, y = y, fill=srtm_62_14)) +
-  geom_point(data = subset(data_plot, village == "Gariuai"), 
+  geom_point(data = data_plot, 
              aes(x = long, y = lat, color = ID_AF), size = 2) + 
-  geom_point(data = subset(data_plot, village == "Gariuai"), 
+  geom_point(data = data_plot, 
              aes(x = long, y = lat), shape = 1, size = 2) + 
@@
   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
-  labs(x = "", y = "", title = "Gariuai", color = "", fill="Elevation (m)") +
+  labs(x = NULL, y = NULL, color = NULL, fill="Elevation (m)") +
+  annotate(geom = "text", label = "b", x = -Inf, y = Inf, 
+           hjust = -0.5, vjust = 1, fontface = "bold", size = 6) + 
+  annotate(geom = "text", label = "c", x = Inf, y = -Inf, 
+           hjust = 12, vjust = -9.8, fontface = "bold", size = 6) + 
   guides(color = FALSE, fill=guide_colorbar(title.position = "top")) +
-  theme(legend.position = "top", axis.title = element_blank(), 
+  theme(legend.position = c(0.8, 0.85), legend.direction="horizontal", 
+        axis.title = element_blank(), 
         axis.ticks = element_blank(), axis.text = element_blank(), 
@@
 
-l1 = ggpubr::get_legend(g2)
+# get AFS legend
+g3 = ggplot(data_plot) +
+  geom_point(aes(x = long, y = lat, color = ID_AF)) + 
+  labs(color = "AFS") +
+  guides(color=guide_legend(ncol=2)) +
+  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
+  theme_classic()
+l1 = ggpubr::get_legend(g3)
 
-g2 = g2 + theme(legend.position = "none")
-# 
-# g2 + ggspatial::annotation_scale() +
-#   coord_sf(crs = 'WGS84')
+ggpubr::ggarrange(ggpubr::ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))
 
-# Samalari map (fig 1c)
-g3 = ggplot(data = elevS) +
-  geom_raster(aes(x = x, y = y, fill=srtm_62_14)) +
+ggsave("figures/mapPlotsSRTM.png", height =4.5, width = 6)
 
-  geom_point(data = subset(data_plot, village == "Samalari"), 
-             aes(x = long, y = lat, color = ID_AF), size = 2) + 
-  geom_point(data = subset(data_plot, village == "Samalari"), 
-             aes(x = long, y = lat), shape = 1, size = 2) + 
-  coord_equal() +
-  labs(x = "", y = "", title = "Samalari", color = "AF\nsystem", fill="Elevation (m)") +
-  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
-  viridis::scale_fill_viridis(option = "D",
-                              limits = range(c(elevS$srtm_62_14, elevG$srtm_62_14)))+
-  guides(color=guide_legend(nrow=2,byrow=TRUE), fill=FALSE) +
-  theme(legend.position = "bottom", axis.title = element_blank(), 
-        axis.ticks = element_blank(), axis.text = element_blank(), 
-        plot.title = element_text(hjust = 0.5), 
-        panel.background = element_blank()) + 
-  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
-  ggspatial::annotation_scale() +
-  coord_sf(crs = 'WGS84')
-
-l2 = ggpubr::get_legend(g3)
-g3 = g3 + theme(legend.position = "none")
-
-ggpubr::ggarrange(g1, 
-                  ggpubr::ggarrange(g2, 
-                                    ggpubr::ggarrange(g3, l1, l2, ncol=1, heights = c(4,1,1)), 
-                                    widths = c(4,3), labels = c("b", "c")), 
-                  heights = c(1, 3), nrow = 2, labels = c("a", ""))
-
-ggsave("figures/mapPlotsSRTM.png", height =6, width = 7)
+# extract elevation and rainfall for the 2 villages
+data_plot$prec <- terra::extract(
+  terra::rast("D:/maps/CHELSA/CHELSA_bio12_1981-2010_V.2.1.tif"), 
+  cbind(data_plot$long, data_plot$lat)
+)
+data_plot$elev <- terra::extract(
+  terra::rast("data/srtm_62_14.tif"), 
+  cbind(data_plot$long, data_plot$lat)
+)
+data_plot[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
+data_plot[, .(prec = mean(prec), range_prec = paste(range(prec), collapse = " - ")), .(ID_SUCO)]
+data_plot[, .(elev = mean(elev), range_elev = paste(range(elev), collapse = " - ")), .(ID_SUCO)]
 ```
@@
 ```{r anova}
-## identify otuliers
+## identify outliers
 # library(magrittr)
@@
 plot(tuk.cld)
+
+## non parametric test
+#updated translation of help page implementation of Nemenyi-Damico-Wolfe-Dunn test
+library(coin)
+is.factor(dataC$ID_AF)
+NDWD <-
+  independence_test(meanAGC ~ ID_AF, data = dataC)
+# , distribution = approximate(B = 10000),
+# ytrafo = function(data) trafo(data, numeric_trafo = rank_trafo),
+# xtrafo = mcp_trafo(iv = "Tukey"))
+
+### global p-value
+print(pvalue(NDWD))
+
+### sites (I = II) != (III = IV) at alpha = 0.01 (page 244)
+print(pvalue(NDWD, method = "single-step"))
+
+#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
+
+library(tidyverse)
+library(ggpubr)
+library(rstatix)
+
+# Data preparation
+# 
+# Here, we’ll use the built-in R data set named PlantGrowth. It contains the weight of plants obtained under a control and two different treatment conditions.
+
+# Re-order the group levels:
+
+# dataC <- dataC %>%
+#   reorder_levels(group, order = c("ctrl", "trt1", "trt2"))
+
+dataC %>% 
+  group_by(ID_AF) %>%
+  get_summary_stats(meanAGC, type = "common")
+
+# Create a box plot of weight by group:
+
+# Question: We want to know if there is any significant difference between the
+# average weights of plants in the 3 experimental conditions.
+
+# We’ll use the pipe-friendly kruskal_test() function [rstatix package], a
+# wrapper around the R base function kruskal.test().
+
+res.kruskal <- dataC %>% kruskal_test(meanAGC ~ ID_AF)
+res.kruskal
+
+# Effect size
+#
+# The eta squared, based on the H-statistic, can be used as the measure of the
+# Kruskal-Wallis test effect size. It is calculated as follow : eta2[H] = (H - k
+# + 1)/(n - k); where H is the value obtained in the Kruskal-Wallis test; k is
+# the number of groups; n is the total number of observations (M. T. Tomczak and
+# Tomczak 2014).
+#
+# The eta-squared estimate assumes values from 0 to 1 and multiplied by 100
+# indicates the percentage of variance in the dependent variable explained by
+# the independent variable.
+#
+# The interpretation values commonly in published literature are: 0.01- < 0.06
+# (small effect), 0.06 - < 0.14 (moderate effect) and >= 0.14 (large effect).
+
+dataC %>% kruskal_effsize(meanAGC ~ ID_AF)
+
+# Multiple pairwise-comparisons
+# 
+# From the output of the Kruskal-Wallis test, we know that there is a significant difference between groups, but we don’t know which pairs of groups are different.
+# 
+# A significant Kruskal-Wallis test is generally followed up by Dunn’s test to identify which groups are different.
+It’s also possible to use the Wilcoxon’s test to calculate pairwise comparisons between group levels with corrections for multiple testing.
+
+# Pairwise comparisons
+pwc <- dataC %>% 
+  dunn_test(meanAGC ~ ID_AF, p.adjust.method = "bonferroni") 
+pwc[pwc$p.adj.signif=="*",]
 ```
@@
 
-dataC_all <- dataC_all[, .(value = mean(value)), .(village, ID_AF, variable)]
+dataC_all <- dataC_all[, .(value = mean(value, na.rm = TRUE)), .(village, ID_AF, variable)]
 
@@
 levels(dataC_all$variable) <- c("AGC", "BGC", "SOC")
+dataC_all[, village := factor(village)]
+levels(dataC_all$village) <- c("Gariuai", "Osso Luga")
 
@@
 
-```{r fig-3-reg, fig.height = 5, fig.width = 7}
+```{r fig-3-reg, fig.height = 4, fig.width = 8}
 ## mixed effect linear regression
 lmeReg <- lmerTest::lmer(meanSC ~ meanAGC + (1|village), data = dataC)
+# Nakagawa’s R2 for mixed models
+R2 = performance::r2_nakagawa(lmeReg)
+summary(lmeReg)
+
+# same for topsoil layer (0-10 cm)
+lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
+# Nakagawa’s R2 for mixed models
+R210 = performance::r2_nakagawa(lmeReg10)
+summary(lmeReg10)
+
+dfreg = 
+  data.frame(
+    depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
+    intercept = c(summary(lmeReg)$coefficients[1,1], summary(lmeReg10)$coefficients[1,1]), 
+    slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
+    R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
+    R2_marg = c(R2$R2_marginal, R210$R2_marginal)
+  )
+
+# change village names
+dataC[, village := factor(village)]
+levels(dataC$village) <- c("Gariuai", "Osso Luga")
+
+# stack data for figure
+
+dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSC", "lwrSC", "uprSC")]
+dataC_30$depth = "Soil depth: 0-30 cm"
+dataC_10 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSC10", "uprSC10")]
+dataC_10$depth = "Soil depth: 0-10 cm"
+colnames(dataC_10) = colnames(dataC_30)
+dataC_fig = rbind(dataC_10, dataC_30)
+
+## figure
+ggplot(dataC_fig, aes(color = ID_AF)) +
+  geom_point(aes(x = meanAGC, y = meanSC, shape = village), size = 2) + 
+  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
+            x = Inf, y = Inf, hjust = 1, vjust = 1.2, parse = TRUE, color = 1) +
+  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
+            x = Inf, y = Inf, hjust = 1, vjust = 2.2, parse = TRUE, color = 1) +
+  geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
+  geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
+  geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
+  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
+  expand_limits(x = 0, y = 0) +
+  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
+  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
+  facet_wrap(~depth) + 
+  theme_classic()
+
+ggsave("figures/fig-scatterplot-carbon.png", height = 4, width = 8)
+```
+
+```{r fig-3-reg-10, fig.height = 5, fig.width = 7}
+## mixed effect linear regression
+lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
 # Nakagawa’s R2 for mixed models
@@
 ggplot(dataC, aes(color = ID_AF)) +
-  geom_point(aes(x = meanAGC, y = meanSC, shape = village), size = 2) + 
+  geom_point(aes(x = meanAGC, y = SC10, shape = village), size = 2) + 
   annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 1, parse = TRUE,
@@
            label = paste0("R^{2}*' (cond.) = '*", round(R2$R2_conditional, 3))) +
-  geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
-  geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
+  # geom_linerange(aes(x = meanAGC, y = SC10, ymin = lwrSC, ymax = uprSC)) +
+  geom_errorbarh(aes(y = SC10, xmin = lwrAGC, xmax = uprAGC)) +
   geom_abline(intercept = summary(lmeReg)$coefficients[1,1], slope = summary(lmeReg)$coefficients[2,1], lty = 2) +
@@
 
+```{r test-ac-soc-by-depth}
+data_soil_agc <- merge(data_soil[, c("N_PLOT", "SAMPLE_DEPTH", "Cstock")], 
+                       dataC[, c("N_PLOT","ID_AF", "ID_VILLAGE",  "meanAGC")], by = "N_PLOT")
+
+ggplot(data_soil_agc, aes(x = meanAGC, y = Cstock, col = SAMPLE_DEPTH)) +
+  geom_point() +
+  geom_smooth(method = "lm")
+
+lm(data = data_soil_agc, Cstock ~ meanAGC * SAMPLE_DEPTH) |> summary()
+lm(data = data_soil_agc[SAMPLE_DEPTH == "0-10"], Cstock ~ meanAGC) |> summary()
+lm(data = data_soil_agc[SAMPLE_DEPTH == "10-20"], Cstock ~ meanAGC ) |> summary()
+lm(data = data_soil_agc[SAMPLE_DEPTH == "20-30"], Cstock ~ meanAGC ) |> summary()
+```
+
+```{r compare-lit-values}
+litval <- read_excel("data/literature-values.xlsx")
+colnames(litval) <- c("system", "pool", "value", "location", "ref")
+
+litval$main_crop <- NA
+litval$main_crop[grepl("palm", litval$system)] <- "oil palm"
+litval$main_crop[grepl("offee", litval$system)] <- "coffee"
+unique(litval$system[is.na(litval$main_crop)])
+
+litval$main_per <- NA
+litval$main_per[grepl("nset", litval$system)] <- "enset"
+litval$main_per[grepl("Inga", litval$system)] <- "inga"
+litval$main_per[grepl("Eucalyptus", litval$system)] <- "eucalyptus"
+litval$main_per[grepl("Pinus", litval$system)] <- "pinus"
+unique(litval$system[is.na(litval$main_per)])
+
+litval$ID_AF <- NA
+litval$ID_AF[grepl("orests|orest |orest$", litval$system)] <- "forest"
+litval$ID_AF[grepl("onoculture", litval$system)] <- "monoculture"
+litval$ID_AF[!is.na(litval$main_crop) & !is.na(litval$main_per)] <- "two-crops"
+litval$ID_AF[grepl("ome garden", litval$system)] <- "HG"
+unique(litval$system[is.na(litval$ID_AF)])
+
+litval$var <- as.numeric(gsub(" ", "", tstrsplit(litval$value, "±")[[2]]))
+litval$value <- gsub(" ", "", tstrsplit(litval$value, "±")[[1]])
+
+mult_values <- which(sapply(strsplit(litval$value, ",|and"), length)>1)
+
+new_rows = lapply(mult_values, function(i){
+  vals = as.numeric(strsplit(litval$value[i], ",|and")[[1]])
+  vals = vals[!is.na(vals)]
+
+  x = matrix(litval[i,], ncol = ncol(litval), nrow = length(vals), byrow = TRUE)
+  x[, colnames(litval) == "value"] = vals
+
+  x = as.data.frame(x)
+  colnames(x) = colnames(litval)
+  return(x)
+}
+)
+
+new_rows = do.call(rbind, new_rows)
+
+litval2 = rbind(new_rows, litval[-mult_values,])
+litval = data.frame(system = unlist(litval2$system), 
+                    pool = unlist(litval2$pool), 
+                    value = unlist(litval2$value), 
+                    location = unlist(litval2$location), 
+                    ref = unlist(litval2$ref), 
+                    main_crop = unlist(litval2$main_crop),
+                    main_per = unlist(litval2$main_per), 
+                    ID_AF = unlist(litval2$ID_AF), 
+                    var = unlist(litval2$var))
+
+```
+
 # References

### 2024-02-22 manuscript.Rmd
add script for table of species correspondance to tetum names

@@
 
+## Supplemental information
+
+### Botanical identification from Tetum names
+
+```{r tetum-names}
+sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
+sheets2 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)
+
+vern1 <- lapply(sheets1, \(x){
+  df = readxl::read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x) 
+  df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |> 
+    unique() |>  subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
+})
+vern2 <- lapply(sheets2, \(x){
+    df = readxl::read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x) 
+  df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |> 
+    unique() |>  subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
+})
+
+vern = rbind(rbindlist(vern1), rbindlist(vern2)) |>  unique()
+vern <- subset(vern, !is.na(TETUM_NAME) & !is.na(SCIENTIFIC_NAME))
+
+## correct scientific names 
+vern[, c("genus", "species") := tstrsplit(SCIENTIFIC_NAME, " ")[1:2]]
+vern[species %in% c("sp", "sp."), species := NA]
+vern[species == "purpuroides", species := NA]
+corr = BIOMASS::correctTaxo(genus = vern$genus, species = vern$species)
+vern = cbind(vern, corr)
+
+vern[, latin := paste(genusCorrected, speciesCorrected)]
+vern[, latin := gsub(" NA", " sp.", latin)]
+vern[,tetum := gsub("  ", " ", tolower(TETUM_NAME))]
+vern <- vern[, c("tetum", "latin", "genusCorrected", "speciesCorrected")]
+vern = unique(vern)
+
+# check that all names are unique
+# if one tetum name has two levels of id (same genus), keep better identification
+dupl_tetum <- vern[duplicated(tetum), unique(tetum)]
+vern = vern[!(tetum %in% dupl_tetum & is.na(speciesCorrected))]
+setorder(vern, tetum)
+
+vern = vern[, .(latin = paste(latin, collapse = ", ")), .(tetum)]
+write.csv(vern, "data/species-correspondance.csv", row.names = FALSE) 
+```
 # References

### 2024-02-22 script-agb.R

@@ -206,7 +206,7 @@ data_tot$BGC  = data_tot$RS * data_tot$AGC
 ## 8. estimate plot AGB and BGB ####
 
 # get plot area
-data_tot[, AREA := ifelse(ARTIFICIAL_PLOT, 1, AREA_TOTAL_PARCEL)]
+data_tot[, AREA := AREA_TOTAL_PARCEL]
 
 # type of plot (total or quadrat)
 data_tot[, TYPE := ifelse(grepl("Q", data_tot$N_Q_ID), "QUADRAT", "TOTAL")]

### 2024-02-22 manuscript.Rmd
reshape anova test

@@
 
+```{r group-comparison}
+## test normality of the data
+
+# log-transform data to get normal distributions (because strictly positive data)
+hist(log(dataC$meanAGC), breaks = 20)
+hist(log(dataC$meanSC))
+
+shapiro.test(log(dataC$meanAGC))
+shapiro.test(log(dataC$meanSC))
+# remove extreme outlier and test normality
+dataC[, .(
+  shap.A = shapiro.test(log(meanAGC))$p.value, 
+  shap.S = shapiro.test(log(meanSC))$p.value
+  ), .(ID_AF, ID_SUCO)]
+# levene test of homoscedasticity
+dataC[, leveneTest(log(meanAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
+dataC[, leveneTest(log(meanSC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
+
+# # pairwise group comparison with Tukey HSD method
+# # https://arc.lib.montana.edu/book/statistics-with-r-textbook/item/59
+# library(multcomp)
+# aovAGC = aov(meanAGC ~ ID_AF, data = dataC)
+# tuk <- glht(aovAGC, linfct = mcp(ID_AF = "Tukey"))
+# summary(tuk)          # standard display
+# tuk.cld <- cld(tuk)   # letter-based display
+# 
+# plot(tuk.cld)
+
+# # soc
+# aovSOC = aov(meanSC ~ ID_AF, data = dataC)
+# tuk <- glht(aovSOC, linfct = mcp(ID_AF = "Tukey"))
+# summary(tuk)          # standard display
+# tuk.cld <- cld(tuk)   # letter-based display
+# 
+# opar <- par(mai=c(1,1,1.5,1))
+# plot(tuk.cld)
+
+## non parametric test
+#updated translation of help page implementation of Nemenyi-Damico-Wolfe-Dunn test
+# library(coin)
+# is.factor(dataC$ID_AF)
+# NDWD <-
+#   independence_test(meanAGC ~ ID_AF, data = dataC)
+# , distribution = approximate(B = 10000),
+# ytrafo = function(data) trafo(data, numeric_trafo = rank_trafo),
+# xtrafo = mcp_trafo(iv = "Tukey"))
+
+### global p-value
+# print(pvalue(NDWD))
+
+### sites (I = II) != (III = IV) at alpha = 0.01 (page 244)
+# print(pvalue(NDWD, method = "single-step"))
+
+#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
+
+res.kruskal <- kruskal_test(log(meanAGC) ~ ID_AF, data = dataC)
+res.kruskal
+
+# Multiple pairwise-comparisons
+
+# Pairwise comparisons
+pwc <- dataC %>% 
+  dunn_test(meanAGC ~ ID_AF, p.adjust.method = "bonferroni") 
+pwc[pwc$p.adj.signif=="*",]
+
+```
+
 ```{r tab-anova, results = "show"}
@@
 dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
-anov1 = anova(lm(meanAGC~ID_SUCO+ID_AF, data = dataC))
-anov2 = anova(lm(meanSC~ID_SUCO+ID_AF, data = dataC))
+anov1 = anova(lm(log(meanAGC)~ID_SUCO+ID_AF, data = dataC))
+anov2 = anova(lm(log(meanSC)~ID_SUCO+ID_AF, data = dataC))
 tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
@@
 )
+
+knitr::kable(tab_anov)
+## non parametric 2 way anova: anova on ranks
+# anov1 = anova(lm(rank(meanAGC)~ID_SUCO+ID_AF, data = dataC))
+# anov2 = anova(lm(rank(meanSC)~ID_SUCO+ID_AF, data = dataC))
+# tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
+#                       ind_var = rep(c("SUCO", "AF"), 2), 
+#                       p = round(c(anov1$`Pr(>F)`[1:2], anov2$`Pr(>F)`[1:2]), 3)
+# )
 ```
@@
 data_plot[, .(elev = mean(elev), range_elev = paste(range(elev), collapse = " - ")), .(ID_SUCO)]
-```
-
-```{r anova}
-## identify outliers
-# library(magrittr)
-library(rstatix)
-dataC %>% 
-  group_by(ID_AF) %>%
-  identify_outliers(meanAGC)
-
-dataC %>% 
-  group_by(ID_AF) %>%
-  identify_outliers(meanSC)
-
-# remove extreme outlier and test normality
-dataC[N_PLOT != 11] %>%
-  group_by(ID_AF) %>%
-  shapiro_test(meanAGC)
-
-dataC %>%
-  group_by(ID_AF) %>%
-  shapiro_test(meanSC)
-
-# non parametric kruskal wallis test 
-kruskalAGC = kruskal.test(meanAGC ~ ID_AF,  data = dataC)
-kruskal.test(meanSC ~ ID_AF,  data = dataC)
-kruskal.test(meanAGC ~ village,  data = dataC)
-kruskal.test(meanSC ~ village,  data = dataC)
-
-# pairwise group comparison with Tukey HSD method
-# https://arc.lib.montana.edu/book/statistics-with-r-textbook/item/59
-library(multcomp)
-aovAGC = aov(meanAGC ~ ID_AF, data = dataC[N_PLOT != 11])
-tuk <- glht(aovAGC, linfct = mcp(ID_AF = "Tukey"))
-summary(tuk)          # standard display
-tuk.cld <- cld(tuk)   # letter-based display
-
-plot(tuk.cld)
-
-# soc
-aovSOC = aov(meanSC ~ ID_AF, data = dataC)
-tuk <- glht(aovSOC, linfct = mcp(ID_AF = "Tukey"))
-summary(tuk)          # standard display
-tuk.cld <- cld(tuk)   # letter-based display
-
-opar <- par(mai=c(1,1,1.5,1))
-plot(tuk.cld)
-
-## non parametric test
-#updated translation of help page implementation of Nemenyi-Damico-Wolfe-Dunn test
-library(coin)
-is.factor(dataC$ID_AF)
-NDWD <-
-  independence_test(meanAGC ~ ID_AF, data = dataC)
-# , distribution = approximate(B = 10000),
-# ytrafo = function(data) trafo(data, numeric_trafo = rank_trafo),
-# xtrafo = mcp_trafo(iv = "Tukey"))
-
-### global p-value
-print(pvalue(NDWD))
-
-### sites (I = II) != (III = IV) at alpha = 0.01 (page 244)
-print(pvalue(NDWD, method = "single-step"))
-
-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
-
-library(tidyverse)
-library(ggpubr)
-library(rstatix)
-
-# Data preparation
-# 
-# Here, we’ll use the built-in R data set named PlantGrowth. It contains the weight of plants obtained under a control and two different treatment conditions.
-
-# Re-order the group levels:
-
-# dataC <- dataC %>%
-#   reorder_levels(group, order = c("ctrl", "trt1", "trt2"))
-
-dataC %>% 
-  group_by(ID_AF) %>%
-  get_summary_stats(meanAGC, type = "common")
-
-# Create a box plot of weight by group:
-
-# Question: We want to know if there is any significant difference between the
-# average weights of plants in the 3 experimental conditions.
-
-# We’ll use the pipe-friendly kruskal_test() function [rstatix package], a
-# wrapper around the R base function kruskal.test().
-
-res.kruskal <- dataC %>% kruskal_test(meanAGC ~ ID_AF)
-res.kruskal
-
-# Effect size
-#
-# The eta squared, based on the H-statistic, can be used as the measure of the
-# Kruskal-Wallis test effect size. It is calculated as follow : eta2[H] = (H - k
-# + 1)/(n - k); where H is the value obtained in the Kruskal-Wallis test; k is
-# the number of groups; n is the total number of observations (M. T. Tomczak and
-# Tomczak 2014).
-#
-# The eta-squared estimate assumes values from 0 to 1 and multiplied by 100
-# indicates the percentage of variance in the dependent variable explained by
-# the independent variable.
-#
-# The interpretation values commonly in published literature are: 0.01- < 0.06
-# (small effect), 0.06 - < 0.14 (moderate effect) and >= 0.14 (large effect).
-
-dataC %>% kruskal_effsize(meanAGC ~ ID_AF)
-
-# Multiple pairwise-comparisons
-# 
-# From the output of the Kruskal-Wallis test, we know that there is a significant difference between groups, but we don’t know which pairs of groups are different.
-# 
-# A significant Kruskal-Wallis test is generally followed up by Dunn’s test to identify which groups are different.
-It’s also possible to use the Wilcoxon’s test to calculate pairwise comparisons between group levels with corrections for multiple testing.
-
-# Pairwise comparisons
-pwc <- dataC %>% 
-  dunn_test(meanAGC ~ ID_AF, p.adjust.method = "bonferroni") 
-pwc[pwc$p.adj.signif=="*",]
 ```
@@
 
-## Supplemental information
-
 ### Botanical identification from Tetum names

### 2024-02-23 manuscript.Rmd
add results

@@
 ```{r group-comparison}
+
+dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
 ## test normality of the data
@@
 # log-transform data to get normal distributions (because strictly positive data)
-hist(log(dataC$meanAGC), breaks = 20)
-hist(log(dataC$meanSC))
+hist((dataC$meanAGC), breaks = 20)
+hist((dataC$meanSC))
 
-shapiro.test(log(dataC$meanAGC))
-shapiro.test(log(dataC$meanSC))
-# remove extreme outlier and test normality
+# test normality
 dataC[, .(
@@
 # levene test of homoscedasticity
+library(car)
 dataC[, leveneTest(log(meanAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
@@
 # print(pvalue(NDWD))
-
-### sites (I = II) != (III = IV) at alpha = 0.01 (page 244)
-# print(pvalue(NDWD, method = "single-step"))
-
-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
-
-res.kruskal <- kruskal_test(log(meanAGC) ~ ID_AF, data = dataC)
-res.kruskal
-
-# Multiple pairwise-comparisons
-
-# Pairwise comparisons
-pwc <- dataC %>% 
-  dunn_test(meanAGC ~ ID_AF, p.adjust.method = "bonferroni") 
-pwc[pwc$p.adj.signif=="*",]
-
 ```
@@
 ```{r fig-1-map, fig.height =6, fig.width = 7}
-
 ## 2. Load data from excel file ####
@@
 
-```{r fig-3-reg, fig.height = 4, fig.width = 8}
+```{r tab-3-reg, results = "show"}
 ## mixed effect linear regression
-lmeReg <- lmerTest::lmer(meanSC ~ meanAGC + (1|village), data = dataC)
+lmeReg <- lmerTest::lmer(log(meanSC) ~ log(meanAGC) + (1|village), data = dataC)
 # Nakagawa’s R2 for mixed models
@@
 # same for topsoil layer (0-10 cm)
-lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
+lmeReg10 <- lmerTest::lmer(log(SC10) ~ log(meanAGC) + (1|village), data = dataC)
 # Nakagawa’s R2 for mixed models
@@
 dfreg = 
-  data.frame(
+  data.table(
     depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
@@
     slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
+    pvalue = c(summary(lmeReg)$coefficients[2,5], summary(lmeReg10)$coefficients[2,5]),
     R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
@@
   )
+
+```
+
+```{r fig-3-reg, fig.height = 4, fig.width = 8}
+dfpred = expand.grid(meanAGC = exp(seq(min(log(dataC$meanAGC), na.rm = TRUE),
+                                       max(log(dataC$meanAGC), na.rm = TRUE),
+                                       length.out = 20)),
+                     depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm")) |>
+  merge(dfreg[, c("depth", "intercept", "slope")], by = "depth")
+setDT(dfpred)
+dfpred[, meanSC := exp(intercept + slope * log(meanAGC))]
 
@@
 # stack data for figure
-
 dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSC", "lwrSC", "uprSC")]
@@
   geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
-            x = Inf, y = Inf, hjust = 1, vjust = 1.2, parse = TRUE, color = 1) +
+            x = -Inf, y = Inf, hjust = -0.1, vjust = 1.2, parse = TRUE, color = 1) +
   geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
-            x = Inf, y = Inf, hjust = 1, vjust = 2.2, parse = TRUE, color = 1) +
+            x = -Inf, y = Inf, hjust = -0.1, vjust = 2.2, parse = TRUE, color = 1) +
   geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
   geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
-  geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
-  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
+  # geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
+  geom_line(data = dfpred, aes(x = meanAGC, y = meanSC), lty = 2, col = 1) +
+  labs(color = "AFS", x = "AGC stocks (Mg C/ha)", y = "SOC stocks (Mg C/ha)", shape = "Site") +
   expand_limits(x = 0, y = 0) +
   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
-  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
+  # scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
+  scale_x_log10() +
+  scale_y_log10() +
   facet_wrap(~depth) +
@@
 
-```{r fig-3-reg-10, fig.height = 5, fig.width = 7}
-## mixed effect linear regression
-lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
-# Nakagawa’s R2 for mixed models
-R2 = performance::r2_nakagawa(lmeReg)
+## Total carbon stocks
 
-## figure
-ggplot(dataC, aes(color = ID_AF)) +
-  geom_point(aes(x = meanAGC, y = SC10, shape = village), size = 2) + 
-  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 1, parse = TRUE,
-           label = paste0("R^{2}*' (marg.) = '*", round(R2$R2_marginal, 3))) +
-  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 2, parse = TRUE,
-           label = paste0("R^{2}*' (cond.) = '*", round(R2$R2_conditional, 3))) +
-  # geom_linerange(aes(x = meanAGC, y = SC10, ymin = lwrSC, ymax = uprSC)) +
-  geom_errorbarh(aes(y = SC10, xmin = lwrAGC, xmax = uprAGC)) +
-  geom_abline(intercept = summary(lmeReg)$coefficients[1,1], slope = summary(lmeReg)$coefficients[2,1], lty = 2) +
-  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
-  expand_limits(x = 0, y = 0) +
-  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
-  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
-  # scale_linetype_manual(values=c("twodash", "dotted"))+
-  theme_classic() 
-ggsave("figures/fig-scatterplot-carbon.png", height = 5, width = 7)
-```
+Total carbon stocks varied between `r dataC[, round(min(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` and `r dataC[, round(max(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` Mg C/ha, with an average of `r dataC[, round(mean(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` Mg C/ha and a coefficient of variation of `r dataC[, round(100*sd(meanAGC + meanSC + meanBGC, na.rm = TRUE)/mean(meanAGC + meanSC + meanBGC, na.rm = TRUE))]`%  (Table 1; Figure 2).
+The AFS with the lowest total carbon stocks was the `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.min(V1)]]` with an average of `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest total carbon stocks was the `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.max(V1)]]` with an average of `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).
+
+## Aboveground carbon
+
+AGC stocks varied between `r dataC[, round(min(meanAGC, na.rm = TRUE))]` and `r dataC[, round(max(meanAGC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanAGC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanAGC, na.rm = TRUE)/mean(meanAGC, na.rm = TRUE))]`%  (Table 1; Figure 2).
+In line with our hypothesis (H1a), AGC stocks varied significantly between AFS (ANOVA p-value: `r tab_anov$p[tab_anov$response=="AGC" & tab_anov$ind_var == "AF"]`), but not between site (ANOVA p-value = `r tab_anov$p[tab_anov$response=="AGC" & tab_anov$ind_var == "SUCO"]`, Table 2).
+The AFS with the lowest AGC stocks was the `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.min(V1)]]` with an average of `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest AGC stocks was the `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.max(V1)]]` with an average of `r dataC[, mean(meanAGC , na.rm = TRUE), .(ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).
+
+## Soil organic carbon
+
+SOC stocks varied between `r dataC[, round(min(meanSC, na.rm = TRUE))]` and `r dataC[, round(max(meanSC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanSC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanSC, na.rm = TRUE)/mean(meanSC, na.rm = TRUE))]`%  (Table 1; Figure 2).
+In disagreement with our hypothesis (H1b), no significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "AF"]`).
+However, the measured SOC stocks differed from suco to suco (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "SUCO"]`; Table 2).
+The SOC stocks in Gariuai were on average higher (`r dataC[ID_SUCO=="G", round(mean(meanSC, na.rm = TRUE))]` Mg C/ha) than those in Osso Luga (`r dataC[ID_SUCO=="S", round(mean(meanSC, na.rm = TRUE))]` Mg C/ha; Figure 2).
+
+## Relationship between AGC and SOC
+
+The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure 3), as hypothesized (H2b).
+Contrary to our hypothesis (H2a), this relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
+The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure 3).
+Overall, AGC stocks were a poor predictor of SOC stocks.
 

### 2024-03-13 manuscript.Rmd
add supplementary figures and tables

@@
 Contrary to our hypothesis (H2a), this relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
-The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure 3).
+The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure 3).
 Overall, AGC stocks were a poor predictor of SOC stocks.
@@
 
-```{r fig-distr-sol, fig.height = 5, fig.width = 15}
+```{r sfig-distr-sol, fig.height = 5, fig.width = 15}
 ## diagnostic données sol pour marguerite
@@
 ggpubr::ggarrange(g1, g2, g3, labels = "auto", nrow = 1)
-ggsave("figures/fig-distr-sol.pdf", height = 5, width = 15)
+ggsave("figures/fig-distr-sol.png", height = 5, width = 15)
 ```
 
-```{r fig-clay-corg, fig.height = 4, fig.width = 6}
-ggplot(data_soil, aes(x = as.numeric(CLAY), y = Corg, col = SAMPLE_DEPTH)) +
+```{r sfig-clay-corg, fig.height = 4, fig.width = 6}
+data_soil[, village := substr(ID_VILLAGE, 1, 1)]
+data_soil[village == "S", village := "Osso Luga"]
+data_soil[village == "G", village := "Gariuai"]
+
+ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY), y = Corg, col = SAMPLE_DEPTH)) +
   geom_point() +
-  labs(col = "Profondeur (cm)", x = "Argile (%)", y = "Teneur en carbone (%)") +
-  geom_smooth(method = "lm") +
+  labs(col = "Soil depth (cm)", x = "Clay content (%)", y = "SOC content (%)") +
+  scale_color_brewer(type = "qual") +
+  # geom_smooth(method = "lm") +
   theme_classic()
 ggsave("figures/fig-clay-corg.pdf", height = 4, width = 6)
+
+ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, col = SAMPLE_DEPTH, shape = village)) +
+  geom_point() +
+  labs(col = "Soil depth (cm)", x = "Clay and fine silt content (%)", y = "SOC content (%)", shape = "Site") +
+  scale_color_brewer(type = "qual") +
+  # geom_smooth(method = "lm") +
+  theme_classic()
+ggsave("figures/fig-clay-silt-corg.png", height = 4, width = 6)
 ```
@@
 
+### Effect of altitude on SOC stocks
+
+```{r sfig-altitude}
+dataCelev = merge(dataC[, c("N_PLOT", "meanSC", "uprSC", "lwrSC")], 
+                 data_plot[, c("N_PLOT", "elev", "village")], by = "N_PLOT")
+
+ggplot(dataCelev, aes(x = elev, y = meanSC, col = village)) +
+  geom_pointrange(aes(ymin=lwrSC, ymax = uprSC)) + 
+  theme_classic() +
+  labs(x = "Elevation (m)", y = "SOC stocks (Mg C/ha)", col = "Site")
+ggsave("figures/soc-elev.png", height = 4, width = 6)
+```
+
+### Site soil characteristics
+
+```{r stab-site-soil}
+data_soil$silt <- as.numeric(data_soil$`FINE SILT`) + as.numeric(data_soil$`COARSE SILT`)
+data_soil$sand <- as.numeric(data_soil$`FINE SAND`) + as.numeric(data_soil$`COARSE SAND`)
+
+vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "CA", "Mg", "K", "Na", "CEC")
+data_soil_plot = melt(data_soil, id.vars = c("village", "N_PLOT", "SAMPLE_DEPTH"), 
+                      measure.vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n"))
+
+data_soil_plot[, variable := gsub("\n", "", variable)]
+
+# when value below detection limit: use half of detection value
+data_soil_plot[grepl("LD", value), value := as.numeric(strsplit(value, "=")[[1]][2])/2]
+data_soil_plot[, value := as.numeric(value)]
+
+# remove NA values
+data_soil_plot = subset(data_soil_plot, !is.na(value))
+
+# simulation 100 values per plot, soil depth and variable with a 10% coefficient
+# of variation (CV value from lab)
+data_soil_plot_simu = 
+  data_soil_plot[, .(
+    value = rnorm(100, value, sd = 0.1*value), n = 1:100), 
+    .(village, N_PLOT, SAMPLE_DEPTH, variable)]
+
+# bootstrap plot values to get mean value at the site level (for each simulation)
+data_soil_site_simu = 
+  data_soil_plot_simu[, .(
+    value = mean(sample(value, length(value), replace = TRUE))),
+    .(village, variable, n)]
+
+# calculate mean and 95% confidence interval at site level
+data_soil_site  = data_soil_site_simu[, .(
+  value = paste0(
+  signif(mean(value), 3), " (",
+  signif(quantile(value, 0.025), 3), "-", 
+  signif(quantile(value, 0.975), 3), ")"
+  )), 
+  .(village, variable)]
+
+# change table shape to have one column per variable
+data_soil_site = dcast(data_soil_site, village ~ variable, value.var = "value")
+data_soil_site = data_soil_site[, c("village", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "Ca", "K", "Mg", "Na", "CEC")]
+colnames(data_soil_site) = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)","N", "C/N", "pH", "Ca", "K", "Mg", "Na", "CEC")
+
+write.csv(data_soil_site, "data/soil_site.csv", row.names = FALSE)
+```
+
 ### Botanical identification from Tetum names
 
-```{r tetum-names}
+```{r tab-tetum-names}
 sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
@@
 ```
+
+```{r timer-leste-c-footprint-lulcc}
+df_eluc <- read_excel("C:/Users/piponiot-laroche/Downloads/National_LandUseChange_Carbon_Emissions_2023v1.0.xlsx", sheet = "OSCAR") 
+colnames(df_eluc) <- df_eluc[7,]
+df_eluc <- df_eluc[-(1:8),colnames(df_eluc) %in% c("unit: Tg C/year", "Timor-Leste")]
+```
+
 # References

### 2024-03-18 manuscript.Rmd
bootstrap values

@@
 
-```{r test-bd-pedotransfer}
-# bulk density from pedotransfer function: Minasny and Hartemink, 2011
-
-data_soil$depth_mean <- unlist(lapply(strsplit(data_soil$SAMPLE_DEPTH, "-"), function(x) mean(as.numeric(x))))
-data_soil$depth_max <- unlist(lapply(strsplit(data_soil$SAMPLE_DEPTH, "-"), function(x) as.numeric(x[2])))
-
-data_soil$sand <- as.numeric(data_soil$`FINE SAND`) + as.numeric(data_soil$`COARSE SAND`)/100
-
-data_soil$BDmin <- 
-  0.935 + 
-  0.049 * log(data_soil$depth_mean) + 
-  0.0055 * data_soil$sand + 
-  0.000065 * (data_soil$sand - 38.96)^2
-
-data_soil$BDptf <- 
-  100/(
-    data_soil$OM/0.224 + 
-      (100 - data_soil$OM)/data_soil$BDmin
-  )
-
-# g1 = ggplot(data_soil, aes(x = BD, y = BDptf, color = ID_TYPE)) + 
-#   geom_abline(slope = 1, intercept = 0, lty=2) + 
-#   geom_point() + 
-#   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
-#   lims(x = c(0.8, 1.6), y = c(0.8, 1.6)) +
-#   facet_wrap(~ID_VILLAGE)
-# #+
-#   # theme(legend.position = "none")
-# g2 = ggplot(data_soil, aes(x = BD, y = BDmin, color = SAMPLE_DEPTH)) +
-#   geom_abline(slope = 1, intercept = 0, lty=2) + 
-#   geom_point() + 
-#   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
-#   lims(x = c(0.8, 1.6), y = c(0.8, 1.6)) +
-#   theme(legend.position = c(0.2, 0.8))
-# 
-# ggpubr::ggarrange(g1, g2) 
-# ggsave("figures/pedotransfer-test.png", height = 5, width = 10)
-```
-
 ```{r add-missing-values}
@@
             (h1 - h0) * 1e8 *     ## equivalent volume of a 1 ha soil sample, in cm3/ha
-            BD *               ## bulk density 
-            Corg/100 *         ## carbon organic content in soil
-            (1-CF) *           ## remove coarse fraction
-            1e-6               ## convert from g/ha to Mg/ha
+            Corg/100 *            ## carbon organic content in fine soil (unitless)
+            (Sample_DRY * 1e-6) * ## mass of the fine fraction of the soil (Mg)
+            (1/V_cylinder)        ## 1/volume of the cylinder, in cm-3
 ]
@@
   data_soil[, .(
-    meanSC = sum(Cstock),
+    meanSOC = sum(Cstock),
     SC10 = Cstock[SAMPLE_DEPTH == "0-10"],
@@
 ## error propagation using a Monte Carlo method -------------
-data_soil_unc = data_soil[, .(BD = rnorm(1000, BD, BD*0.1), 
+data_soil_unc = data_soil[, .(mass = rnorm(1000, Sample_DRY, Sample_DRY*0.1), 
                               Corg = rnorm(1000, Corg, Corg*0.12), 
-                              CF = rnorm(1000, CF, CF*0.15), 
                               iter = 1:1000), 
@@
 data_soil_unc = data_soil_unc[, .(
-  Cstock = sum(
-    (h1 - h0) * 1e8 *  
-      BD * Corg/100 * (1-CF) * 1e-6  
-  ), 
-  Cstock10 = (
-    (h1 - h0) * 1e8 *  
-      BD * Corg/100 * (1-CF) * 1e-6  
-  )[SAMPLE_DEPTH == "0-10"]),
+  Cstock = sum((h1 - h0) * 1e8 / 250 * Corg/100 * (mass * 1e-6)), 
+  Cstock10 = ((h1 - h0) * 1e8 / 250 * Corg/100 * (mass * 1e-6))[SAMPLE_DEPTH == "0-10"]),
   .(N_PLOT, iter)
@@
 
-data_soil_unc = data_soil_unc[, .(
-  lwrSC = quantile(Cstock, 0.025), 
-  uprSC = quantile(Cstock, 0.975),
-  lwrSC10 = quantile(Cstock10, 0.025), 
-  uprSC10 = quantile(Cstock10, 0.975)), 
+data_soil_quantile = data_soil_unc[, .(
+  lwrSOC = quantile(Cstock, 0.025), 
+  medSOC = quantile(Cstock10, 0.5),
+  uprSOC = quantile(Cstock, 0.975),
+  lwrSOC10 = quantile(Cstock10, 0.025), 
+  medSOC10 = quantile(Cstock10, 0.5),
+  uprSOC10 = quantile(Cstock10, 0.975)), 
   .(N_PLOT)]
@@
 # add soil carbon uncertainty
-dataC <- merge(dataC, data_soil_unc, all=TRUE)
+dataC <- merge(dataC, data_soil_quantile, all=TRUE)
 
@@
 
-```{r tab-mean, results = "show"}
-# average values per village and SAF type
-dataC_mean <- dataC[, .(AGC = paste0(round(mean(meanAGC[!is.na(meanAGC)]), 2), " +/- ", round(sd(meanAGC[!is.na(meanAGC)])/sqrt(length(meanAGC[!is.na(meanAGC)])), 2)), 
-                        SC = paste0(round(mean(meanSC[!is.na(meanSC)]), 2), " +/- ", round(sd(meanSC[!is.na(meanSC)])/sqrt(length(meanSC[!is.na(meanSC)])), 2))),
-                    .(ID_AF)]
-dataC_mean$ID_AF <- factor(dataC_mean$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
-dataC_mean = dataC_mean[order(dataC_mean$ID_AF),]
-knitr::kable(dataC_mean, col.names = c("Agroforestry system", "Aboveground carbon (Mg/ha)", "Soil organic carbon (Mg/ha)"))
-write.csv(dataC_mean, file = "tables/C_mean_SAF.csv", row.names = FALSE)
+```{r tab-median, results = "show"}
+
+## bootstrapping
+
+#1. prepare data
+data_agc_boot = melt(data_agb_unc, 
+                       measure.vars = c("AGC", "BGC"), 
+                       id.vars = c("ID_VILLAGE", "N_PLOT", "ID_AF", "iter"))
+
+data_soil_boot = data_soil_unc[, c("N_PLOT", "iter", "Cstock")]
+colnames(data_soil_boot)[3] = "value"
+data_soil_boot$variable = "SOC"
+data_soil_boot$iter = paste0("V", data_soil_boot$iter)
+
+data_soil_boot = merge(data_soil_boot, data_plot[, c("N_PLOT","ID_VILLAGE", "ID_AF")])
+
+data_boot = rbind(data_agc_boot, data_soil_boot[, colnames(data_agc_boot), with = FALSE])
+
+# 2. add total stocks
+data_boot_tot = data_boot[, .(value = sum(value), variable = "TOT"), .(ID_VILLAGE, N_PLOT, ID_AF, iter)]
+data_boot = rbind(data_boot, data_boot_tot[, colnames(data_boot), with = FALSE])
+
+data_boot[, village := substr(ID_VILLAGE, 1, 1)]
+
+# 3. bootstrap
+data_boot_saf = data_boot[, .(value = mean(sample(value, length(value), replace = TRUE))),
+    .(ID_AF, variable, iter)]
+
+# calculate median and 95% confidence interval at site level
+data_boot_saf  = data_boot_saf[, .(
+  value = paste0(
+  signif(mean(value), 3), " (",
+  signif(quantile(value, 0.025), 3), "-", 
+  signif(quantile(value, 0.975), 3), ")"
+  )), 
+  .(ID_AF, variable)]
+
+data_boot_saf = dcast(data_boot_saf, ID_AF ~ variable, value = "value")
+
+knitr::kable(data_boot_saf, col.names = c("Agroforestry system", "Aboveground carbon (MgC/ha)", "Belowground carbon (MgC/ha)", "Soil organic carbon (MgC/ha)", "Total carbon (MgC/ha)"))
+write.csv(data_boot_saf, file = "tables/C_median_SAF.csv", row.names = FALSE)
 ```
@@
 ```{r group-comparison}
-
 dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
@@
 hist((dataC$meanAGC), breaks = 20)
-hist((dataC$meanSC))
+hist((dataC$meanSOC))
 
@@
 dataC[, .(
-  shap.A = shapiro.test(log(meanAGC))$p.value, 
-  shap.S = shapiro.test(log(meanSC))$p.value
+  shap.A = shapiro.test(log(medAGC))$p.value, 
+  shap.S = shapiro.test(log(medSOC))$p.value
   ), .(ID_AF, ID_SUCO)]
@@
 library(car)
-dataC[, leveneTest(log(meanAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
-dataC[, leveneTest(log(meanSC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
+dataC[, leveneTest(log(medAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
+dataC[, leveneTest(log(medSOC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
 
@@
 # # soc
-# aovSOC = aov(meanSC ~ ID_AF, data = dataC)
+# aovSOC = aov(meanSOC ~ ID_AF, data = dataC)
 # tuk <- glht(aovSOC, linfct = mcp(ID_AF = "Tukey"))
@@
 dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
-anov1 = anova(lm(log(meanAGC)~ID_SUCO+ID_AF, data = dataC))
-anov2 = anova(lm(log(meanSC)~ID_SUCO+ID_AF, data = dataC))
+anov1 = anova(lm(log(medAGC)~ID_SUCO+ID_AF, data = dataC))
+anov2 = anova(lm(log(medSOC)~ID_SUCO+ID_AF, data = dataC))
 tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
@@
 knitr::kable(tab_anov)
+write.csv(tab_anov, file = "tables/anov.csv", row.names = FALSE)
+
 ## non parametric 2 way anova: anova on ranks
 # anov1 = anova(lm(rank(meanAGC)~ID_SUCO+ID_AF, data = dataC))
-# anov2 = anova(lm(rank(meanSC)~ID_SUCO+ID_AF, data = dataC))
+# anov2 = anova(lm(rank(meanSOC)~ID_SUCO+ID_AF, data = dataC))
 # tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
@@
 # )
+
 ```
@@
   # coord_equal() +
-  viridis::scale_fill_viridis(option = "D",
-                              limits = range(c(elevS$srtm_62_14, elevG$srtm_62_14))
-  ) +
+  # viridis::scale_fill_viridis(option = "D",
+  #                             limits = range(c(elevS$srtm_62_14, elevG$srtm_62_14))
+  # ) +
+  scale_fill_gradientn(colours = terrain.colors(10)) +
   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
@@
         axis.title = element_blank(),
-        axis.ticks = element_blank(), axis.text = element_blank(), 
+        # axis.ticks = element_blank(), axis.text = element_blank(), 
         plot.title = element_text(hjust = 0.5), 
@@
 dataC_all <- melt(dataC, id.vars = c("N_PLOT", "ID_VILLAGE", "ID_AF", "village"), 
-                  measure.vars = c("meanAGC", "meanBGC", "meanSC"))
+                  measure.vars = c("meanAGC", "meanBGC", "meanSOC"))
 
@@
 
-dataC_all[, variable := factor(variable, levels = c("meanAGC", "meanBGC", "meanSC"))]
+dataC_all[, variable := factor(variable, levels = c("meanAGC", "meanBGC", "meanSOC"))]
 levels(dataC_all$variable) <- c("AGC", "BGC", "SOC")
@@
 ## mixed effect linear regression
-lmeReg <- lmerTest::lmer(log(meanSC) ~ log(meanAGC) + (1|village), data = dataC)
+lmeReg <- lmerTest::lmer(log(meanSOC) ~ log(meanAGC) + (1|village), data = dataC)
 # Nakagawa’s R2 for mixed models
@@
 setDT(dfpred)
-dfpred[, meanSC := exp(intercept + slope * log(meanAGC))]
+dfpred[, meanSOC := exp(intercept + slope * log(meanAGC))]
 
@@
 # stack data for figure
-dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSC", "lwrSC", "uprSC")]
+dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSOC", "lwrSOC", "uprSOC")]
 dataC_30$depth = "Soil depth: 0-30 cm"
-dataC_10 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSC10", "uprSC10")]
+dataC_10 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSOC10", "uprSOC10")]
 dataC_10$depth = "Soil depth: 0-10 cm"
@@
 ggplot(dataC_fig, aes(color = ID_AF)) +
-  geom_point(aes(x = meanAGC, y = meanSC, shape = village), size = 2) +
+  geom_point(aes(x = meanAGC, y = meanSOC, shape = village), size = 2) +
   geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
@@
             x = -Inf, y = Inf, hjust = -0.1, vjust = 2.2, parse = TRUE, color = 1) +
-  geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
-  geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
+  geom_linerange(aes(x = meanAGC, y = meanSOC, ymin = lwrSOC, ymax = uprSOC)) +
+  geom_errorbarh(aes(y = meanSOC, xmin = lwrAGC, xmax = uprAGC)) +
   # geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
-  geom_line(data = dfpred, aes(x = meanAGC, y = meanSC), lty = 2, col = 1) +
+  geom_line(data = dfpred, aes(x = meanAGC, y = meanSOC), lty = 2, col = 1) +
   labs(color = "AFS", x = "AGC stocks (Mg C/ha)", y = "SOC stocks (Mg C/ha)", shape = "Site") +
@@
 
-Total carbon stocks varied between `r dataC[, round(min(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` and `r dataC[, round(max(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` Mg C/ha, with an average of `r dataC[, round(mean(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` Mg C/ha and a coefficient of variation of `r dataC[, round(100*sd(meanAGC + meanSC + meanBGC, na.rm = TRUE)/mean(meanAGC + meanSC + meanBGC, na.rm = TRUE))]`%  (Table 1; Figure 2).
-The AFS with the lowest total carbon stocks was the `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.min(V1)]]` with an average of `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest total carbon stocks was the `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.max(V1)]]` with an average of `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).
+Total carbon stocks varied between `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(min(V1))]` and `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(max(V1))]` Mg C/ha, with an average of `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(mean(V1))]` Mg C/ha and a coefficient of variation of `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(sd(V1)/mean(V1)*100)]`%  (Table 1; Figure 2).
+The AFS with the lowest total carbon stocks was the `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT,ID_AF)][, ID_AF[which.min(V1)]]` with a median of `r data_boot_tot[variable=="TOT", mean(value), .(N_PLOT,ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest total carbon stocks was the `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT,ID_AF)][, ID_AF[which.max(V1)]]` with a median of `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT,ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).
 
@@
 
-SOC stocks varied between `r dataC[, round(min(meanSC, na.rm = TRUE))]` and `r dataC[, round(max(meanSC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanSC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanSC, na.rm = TRUE)/mean(meanSC, na.rm = TRUE))]`%  (Table 1; Figure 2).
+SOC stocks varied between `r dataC[, round(min(meanSOC, na.rm = TRUE))]` and `r dataC[, round(max(meanSOC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanSOC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanSOC, na.rm = TRUE)/mean(meanSOC, na.rm = TRUE))]`%  (Table 1; Figure 2).
 In disagreement with our hypothesis (H1b), no significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "AF"]`).
 However, the measured SOC stocks differed from suco to suco (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "SUCO"]`; Table 2).
-The SOC stocks in Gariuai were on average higher (`r dataC[ID_SUCO=="G", round(mean(meanSC, na.rm = TRUE))]` Mg C/ha) than those in Osso Luga (`r dataC[ID_SUCO=="S", round(mean(meanSC, na.rm = TRUE))]` Mg C/ha; Figure 2).
+The SOC stocks in Gariuai were on average higher (`r dataC[ID_SUCO=="G", round(mean(meanSOC, na.rm = TRUE))]` Mg C/ha) than those in Osso Luga (`r dataC[ID_SUCO=="S", round(mean(meanSOC, na.rm = TRUE))]` Mg C/ha; Figure 2).
 
@@
   scale_color_brewer(type = "qual") +
-  # geom_smooth(method = "lm") +
+  scale_x_log10() + 
+  scale_y_log10() +
+  geom_smooth(method = "lm") +
   theme_classic()
 ggsave("figures/fig-clay-corg.pdf", height = 4, width = 6)
+
+lmerTest::lmer(
+  log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
+     (1|village)+ 
+     (1|SAMPLE_DEPTH), 
+   data = data_soil[SAMPLE_DEPTH != "10-20"]) |> 
+  summary()
 
@@
   scale_color_brewer(type = "qual") +
+  # scale_x_log10() +
+  # scale_y_log10() +
   # geom_smooth(method = "lm") +
@@
 ```{r sfig-altitude}
-dataCelev = merge(dataC[, c("N_PLOT", "meanSC", "uprSC", "lwrSC")], 
+dataCelev = merge(dataC[, c("N_PLOT", "meanSOC", "uprSOC", "lwrSOC")], 
                  data_plot[, c("N_PLOT", "elev", "village")], by = "N_PLOT")
 
-ggplot(dataCelev, aes(x = elev, y = meanSC, col = village)) +
-  geom_pointrange(aes(ymin=lwrSC, ymax = uprSC)) + 
+lmerTest::lmer(log(meanSOC) ~ log(elev) + (1|village), data = dataCelev) |> 
+  summary()
+
+ggplot(dataCelev, aes(x = elev, y = meanSOC, col = village)) +
+  geom_pointrange(aes(ymin=lwrSOC, ymax = uprSOC)) + 
   theme_classic() +
@@
 
-# simulation 100 values per plot, soil depth and variable with a 10% coefficient
+# simulation 100 values per plot, soil depth and variable with a 12% coefficient
 # of variation (CV value from lab)
@@
   data_soil_plot[, .(
-    value = rnorm(100, value, sd = 0.1*value), n = 1:100), 
+    value = rnorm(100, value, sd = 0.12*value), n = 1:100), 
     .(village, N_PLOT, SAMPLE_DEPTH, variable)]

### 2024-03-19 manuscript.Rmd
add text and references

@@
 author: "Marguerite Cogné, Régis Peltier, Vincent Freycon, Alexis Toumazeau, Camille Piponiot"
-date: "2023-05-10"
+date: "2024-03-19"
 output: 
@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide", tinytex.clean = FALSE)
-packages_needed = c("ggplot2", "dataverse", "readxl", "data.table")
+packages_needed = c("ggplot2", "readxl", "data.table", "knitr")
 
@@
 
+# Abstract
+
+Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural fields and diversifying agricultural landscapes.
+Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their potential inclusion in climate change mitigation strategies.
+In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastoral systems, young agroforests, home gardens and forest gardens.
+Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 255 MgC/ha, similar to an old-growth tropical forest.
+Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
+We found no relationship between the amount of carbon stored in biomass and soil.
+Our results highlight the high diversity of traditional agroforestry systems in Timor-Leste and their high capacity to store carbon.
+
 # Introduction
 
-Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO_2_e/yr over 2010-2014; @Pendrill2019].
-The demand for tropical agricultural land is expected to continue to increase in the tropics as the population grows [@UN2022].
-In addition, tropical areas are expected to be disproportionately affected by deforestation and climate change, putting their agriculture at risk [@Lawrence2015].
+Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO2e/yr over 2010-2014; @Pendrill2019].
+The demand for tropical agricultural land is expected to continue to increase as the population grows [@UN2022].
+In addition, tropical areas are expected to be disproportionately affected by climate change, putting their agriculture and food systems at risk [@Lawrence2015].
 There is therefore an urgent need to promote tropical agricultural landscapes that are both more resilient to climate change, while maintaining high productivity and ecological value [@Harvey2014].
-In particular, increasing the carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of the food production system [@Lal2004].
+In particular, increasing the carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of the food production system [@Lal2004;@Minasny2017].
 
-Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Verchot2007].
-Agroforestry is a collective name for land use systems and technologies where woody perennials (trees, shrubs, palms, bamboo, etc.) are deliberately used on the same land management units as agricultural crops and/or animals, in some form of spatial arrangement or temporal sequence [@Lundgren1982].
-Agroforestry allows for diversification of agricultural production, making agricultural systems more resilient [@Duffy2021]; it also increases the carbon sequestration in the vegetation but also in the soil [@DeStefano2018].
+Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Albrecht2003;@Verchot2007;@Lasco2014;@Cardinael2021].
+Agroforestry is a collective name for land use systems and technologies where woody perennials (trees, shrubs, bamboos, creepers, etc.) are deliberately used on the same land management units as agricultural crops and/or animals, in some form of spatial arrangement or temporal sequence [@Lundgren1982].
+Agroforestry allows for diversification of agricultural production, making agricultural systems more resilient [@Duffy2021;@Terasaki2023].
 
-Carbon stocks in AFS [VF]
+Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation and soil, and this carbon storage capacity depends on the characteristics of AFS considered [@Albrecht2003;@Feliciano2018].
+As trees store large amounts of carbon, the carbon stored in the vegetation of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
+However, the effect of AFS on soil organic carbon (SOC) has been relatively little studied [@Beillouin2021], is more variable and depends on a number of factors.
+Although AFS appear to increase SOC stocks overall, some AFS such as woodlots (i.e.
+planting trees with intercropping during the establishment phase) may decrease SOC compared to non-AFS, while AFS such as silvopastoral systems have been shown to strongly increase SOC stocks [@Feliciano2018].
+Other factors such as climate and previous land use type have also been shown to strongly influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018;@Ma2020;@Martin2020].
+Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS could only be demonstrated for the topsoil layer [< 10 cm; @Ma2020].
 
-Importance of the ecological/social context in the success of agroforestry techniques - existence of traditional agroforestry techniques that could be more adapted to local conditions, [MC]
-Agroforestry needs to be adapted to the context [(Coe et al., 2014, Cur Opinion in Env Sust) @Coe2014]
+Although AFS can make agriculture more resilient while improving its carbon footprint, poor adaptation to the local context can explain many failures to introduce new agroforestry techniques [@Coe2014].
+Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
+Moreover, traditional small-scale AFS typically retain high levels of biodiversity, making them key elements for biodiversity conservation in the tropics [@Perfecto2008;@Torquebiau1992].
+A better understanding of traditional agroforestry practices and their ecological characteristics could provide better support for the local development of agroforestry.
 
-Existence of traditional agroforestry systems [examples from Marguerite’s intro]
+In Timor-Leste, a large proportion of the population depends on agriculture for their livelihoods, but food security is threatened by population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime [@WB2009;@Molyneux2012].
+In this context, productive, resilient and sustainable agriculture is critical for the country, and agroforestry has been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022;@Cogne2024].
+Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
+Better quantification of these carbon stocks would improve the understanding of the ecological importance of these AFS at both local and global scales, and help to integrate them into national and international climate change mitigation initiatives.
 
-Context in Timor Leste: current knowledge on their agroforestry practices + knowledge gap [MC]
-
-Local typology of production systems and why it is not adapted to traditional agroforestry
-
-No prior description of traditional agroforestry systems in TL
-
-In Timor Leste, a typology of production systems taking into account different agroecological zoning was carried out in 2017 based on the results of the 2010 national census (R.
-Williams et al, 2018).
-However, this typology does not highlight the agroforestry component of Timorese agriculture by focusing on other differentiating criteria such as livestock, subsistence crops (rice) or the island's historical cash crops (coconut, coffee).
-
-Need to better understand how these systems work and quantify their carbon storage to include them in climate change mitigation policies 
-Research questions  [VF]
-
-What are the characteristics of traditional agroforestry systems in Timor Leste?
-
-How much carbon do they store in both their biomass and in the soil
-
-Are there measurable differences between AFS?
-
-Hypotheses [VF]
-
-Proposition: lien entre typologie et stocks de carbone
+In this study, we aim to better understand traditional agroforestry systems in Timor-Leste, and in particular their carbon storage both in their biomass and in the soil.
+We hypothesize that different types of traditional agroforestry systems will exhibit measurable differences in aboveground carbon (AGC; H1a) and SOC (H1b) stocks.
+We also assume, following @Ma2020, that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; H2a), but no relationship with SOC below 10 cm (H2b).
 
@@
 ```
+
+## Study area
+
+```{r get-site-info}
+## Load plot location data 
+data_coord <- read_xlsx("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = "GPS")
+setDT(data_coord)
+
+# get plot number
+data_coord[, N_PLOT := as.numeric(data.table::tstrsplit(N_Q_ID, "_")[[1]])]
+
+# get coordinates
+data_coord <- data_coord[, .(long = mean(as.numeric(LONG)), lat = mean(as.numeric(LAT))), .(N_PLOT)]
+
+## Load plot general information
+data_plot <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")
+
+# merge plot coordinates with general information (village, AFS)
+data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])
+
+# simplify site names
+data_plot[, village := substr(ID_VILLAGE, 1, 1)]
+data_plot[village == "S", village := "Osso Luga"]
+data_plot[village == "G", village := "Gariuai"]
+
+## village location: create box around plots
+data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(village)]
+data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
+data_box[village == "Osso Luga", w := 0.05]
+data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
+```
+
+```{r get-env-data}
+# extract elevation and rainfall for the 2 villages
+village_env = data_plot[, as.list(expand.grid(seq(min(long), max(long), 0.001), 
+                                           seq(min(lat), max(lat), 0.001))),
+                     .(village)]
+
+colnames(village_env) = c("village", "long", "lat")
+
+# TODO add code to download prec and elev
+village_env[, prec := terra::extract(
+  terra::rast("D:/maps/CHELSA/CHELSA_bio12_1981-2010_V.2.1.tif"), 
+  cbind(long, lat)
+)]
+
+village_env[, elev := terra::extract(
+  terra::rast("data/srtm_62_14.tif"), 
+  cbind(long, lat)
+)]
+
+village_summary = village_env[, .(
+  coord = paste(round(mean(long),2), round(mean(lat),2), sep = ", "), 
+  prec = round(mean(prec)), 
+  elev = paste(round(quantile(elev, c(0.025, 0.975))), collapse = " - ")),
+  .(village)]
+```
+
+The research was carried out in the center of the Baucau district, on the north-eastern coast of Timor-Leste.
+Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r village_summary[village=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r village_summary[village=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r village_summary[village=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r village_summary[village=="Osso Luga", prec]` mm and an altitude ranging `r village_summary[village=="Osso Luga", elev]` m (Fig.
+\@ref(fig:map-sites)).
+The average soil properties differed between villages (Table \@ref(tab:stab-site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
+The two villages also have had different histories.
+Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2000.
+During the period of abandonment, the various agroforestry systems in Osso Luga lay fallow.
+
+## Typology of agroforestry systems
+
+We used the typology described in @Cogne2024, developed in the same study area.
+This typology was based on the one proposed by @Nair1993;@Nair2022, and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
+The five AFS identified were: crop system including a wooded fallow phase (CF), Sylvopastoral (SP), Young agroforest (YA), Home garden (HG) and Forest garden (FG).
+CF are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
+SP land is a vast area (generally 200 to 500 ha) where cattle are grazed under a layer of trees.
+YA are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
+HG are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
+FGs are generally former HGs (over 50 years old), which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
+
+The AFS are presented in more detail in the supplemental information (Table \@ref(tab:afs-typology)) and in @Cogne2024.
+
+## Field inventories
+
+We randomly selected 3 plots from each of the 5 traditional AFS identified in each of the two sites (30 plots in total; Fig.
+\@ref(fig:map-sites)).
+Sampling plots were defined as the entire agricultural plot (area between 0.06-1.25 ha) for all AFS except SP plots, which were generally much larger than 1 ha.
+In SP plots, a 1 ha sampling plot was delineated within the agricultural plot.
+Tree and soil measurements (described below) were carried out between June and August 2021.
+
+### Tree measurements
+
+Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
+Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot: within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
+Height measurements were done with a dendrometer.
+
+Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using inaturalist and Pl@ntnet application, Geoffrey Hull’s monograph (Hull, 2006), the data bases of “the useful tropical plants” (https://tropical.theferns.info/) as well as the native plants websites of the NT herbarium from the Australian government (https://nt.gov.au/environment/native-plants/native-plants-and-nt-herbarium).
+The correspondence between tetum names and scientific names is provided in the supplemental material (Table S2).
+
+### Soil measurements
+
+In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (see figure S1).
+Soils were sampled with an auger at 0-10 cm and 20-30 cm.
+Soils samples were then air-dried and sieved at 2 mm.
+They were the subject of standard physical and chemical analyses at Cirad laboratories in Montpellier, France.
+Among these analyses, total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
+
+In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm$^3$ at 2-7 cm, 12-17 cm and 22-27 cm.
+Each sample was dried at 105°C in a classic oven until reaching a stable dry mass, sieved to 2 mm, and then the fine fraction (< 2 mm) was weighted.
+
+## Biomass estimation
+
+We estimated tree-level AGC stocks from the measured tree CBH and height using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017].
+Each tree was assigned a mean and standard deviation of wood density at the species or genus level when at least one value was available from the Global Wood Density Database [@Zanne2009].
+Other taxa were assigned a plot-level average wood density.
+The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density, taken here as the average wood density in the plots (mean: xxx and standard deviation: xxx).
+The volume of the palm trunks was estimated as the volume of a cylinder, using the height and diameter measured in situ, with a volumetric shrinkage of 11% [@WD2023].
+The height of the palms was measured above the leaves and was therefore multiplied by a conversion factor of 0.9 to obtain the trunk height.
+
+Tree-level belowground carbon (BGC) stocks were estimated using the root-to-shoot ratio equation from @Ledo2018, which is based on tree DBH and climatic water deficit from @Chave2014.
+This ratio was then multiplied by the estimated aboveground carbon stock to obtain tree-by-tree belowground carbon stock.
+
+Plot-level AGC and BGC stocks were then obtained by summing tree-level stocks and dividing by the plot area.
+
+## Soil organic carbon estimation
+
+Soil organic carbon stocks (in Mg.ha-1) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:
+
+$$SOC_i = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10-6) \cdot (Splot \cdot \Delta d)}{Vcyl}$$
+
+Where $Corg_i$ is the organic carbon content at depth $i$ (either 0-10, 10-20 or 20-30 cm); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth $i$ (in g), converted into Mg with a factor of 10$^{-6}$; Splot is the surface of a 1-ha plot (10$^8$ cm$^2$); $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
+At depth $i$ = 10-20 cm, where the carbon content was not measured (see "Soil measurements" section), we used the average of the carbon content values at 0-10 and 20-30 cm.
+
+## Statistical analysis
+
+### Statistical tests
+
+The estimated AGC and SOC stocks were log-transformed for all statistical tests, because we expect them to be log-normally distributed as they only take positive values.
+After testing for normality (Shapiro-Wilks test) and homoscedasticity (Levene’s test), we performed a two-way ANOVA on the log-transformed estimates of AGC and SOC stocks, grouped by AFS and site.
+To test the relationship between log-transformed SOC and AGC stocks, we fitted two linear mixed models.
+We predicted SOC stocks at two different depths (either 0-10 or 0-30 cm) as a function of AGC stocks (fixed effect) and site (Samalari or Osso-Luga; random effect).
+
+### Uncertainty propagation
+
+The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo() function from the BIOMASS package, using a (rather conservative) coefficient of variation of 15% on height measurements (Hunter et al 2013); large and small errors on 5 and 95% of trees respectively (Chave et al 2004; Réjou-Méchain et al., 2017).
+To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (Cirad laboratory in Montpellier, France), 10% for the mass of the fine fraction (Page-Dumroese et al.
+1999).
+
+We then propagated these uncertainties using a Monte Carlo method (1000 iterations).
+For single plot estimates of carbon stocks, medians and 95% confidence intervals were estimated as quantiles of the results of the 1000 iterations.
+For multiplot estimates of carbon stocks, values within each grouping factor were additionally bootstrapped before calculating quantiles (median and 95% confidence intervals).
 
@@
 
-```{r fig-1-map, fig.height =6, fig.width = 7}
-## 2. Load data from excel file ####
-data_coord <- read_xlsx("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = "GPS")
-data.table::setDT(data_coord)
-
-data_coord[, N_PLOT := as.numeric(data.table::tstrsplit(N_Q_ID, "_")[[1]])]
-
-data_coord <- data_coord[, .(long = mean(as.numeric(LONG)), lat = mean(as.numeric(LAT))), .(N_PLOT)]
-
-data_plot <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")
-
-data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])
-
-data_plot[, village := substr(ID_VILLAGE, 1, 1)]
-data_plot[village == "S", village := "Osso Luga"]
-data_plot[village == "G", village := "Gariuai"]
-
-data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(village)]
-data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
-data_box[village == "Osso Luga", w := 0.05]
-data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
-
-## create Timor map ####
-
+```{r map-sites, fig.height = 6, fig.width = 7}
+## create Timor map
 bbox <- c(124.5, -9.2, 127.5,  -8.3)
-# basemap <- ggmap::get_stamenmap(bbox, maptype = "terrain", zoom = 10)
-
-# geodata::elevation_3s(lon= c(data_box[village == "Osso Luga"]$x1), 
-#                       lat= c(data_box[village == "Osso Luga"]$y1), 
-#                       path = "data")
 timor_map <- map_data("world", region = "Timor-Leste")
@@
 ggsave("figures/mapPlotsSRTM.png", height =4.5, width = 6)
-
-# extract elevation and rainfall for the 2 villages
-data_plot$prec <- terra::extract(
-  terra::rast("D:/maps/CHELSA/CHELSA_bio12_1981-2010_V.2.1.tif"), 
-  cbind(data_plot$long, data_plot$lat)
-)
-data_plot$elev <- terra::extract(
-  terra::rast("data/srtm_62_14.tif"), 
-  cbind(data_plot$long, data_plot$lat)
-)
-data_plot[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
-data_plot[, .(prec = mean(prec), range_prec = paste(range(prec), collapse = " - ")), .(ID_SUCO)]
-data_plot[, .(elev = mean(elev), range_elev = paste(range(elev), collapse = " - ")), .(ID_SUCO)]
 ```
@@
 # Supplementary material
+
+```{r afs-typology}
+typ = readxl::read_excel("tables/afs-typology.xlsx")
+kable(typ)
+```
 

### 2024-03-19 script-agb.R

@@ -215,15 +215,17 @@ data_tot[, TYPE := ifelse(grepl("Q", data_tot$N_Q_ID), "QUADRAT", "TOTAL")]
 data_tot[TYPE == "QUADRAT", AREA := 0.05]
 
 # sum AGB of trees in each plot and quadrat, and standardize by area (to Mg/ha)
-data_agb <- data_tot[, .(AGC = sum(AGC/AREA), BGC = sum(BGC/AREA)), .(N_PLOT, ID_VILLAGE, ID_AF, iter)]
+data_agb_unc <- data_tot[, .(AGC = sum(AGC/AREA), BGC = sum(BGC/AREA)), .(N_PLOT, ID_VILLAGE, ID_AF, iter)]
 
-data_agb <- data_agb[, .(meanAGC = mean(AGC), 
-                         lwrAGC = quantile(AGC, 0.025), 
-                         uprAGC = quantile(AGC, 0.975), 
-                         meanBGC = mean(BGC), 
-                         lwrBGC = quantile(BGC, 0.025), 
-                         uprBGC = quantile(BGC, 0.975)), 
-                     .(N_PLOT, ID_VILLAGE, ID_AF)]
+data_agb <- data_agb_unc[, .(meanAGC = mean(AGC), 
+                             lwrAGC = quantile(AGC, 0.025), 
+                             medAGC = quantile(AGC, 0.5), 
+                             uprAGC = quantile(AGC, 0.975), 
+                             meanBGC = mean(BGC), 
+                             lwrBGC = quantile(BGC, 0.025), 
+                             medBGC = quantile(BGC, 0.5), 
+                             uprBGC = quantile(BGC, 0.975)), 
+                         .(N_PLOT, ID_VILLAGE, ID_AF)]
 
 
 ## optional: graphs ####

### 2024-03-20 manuscript.Rmd
consolidate calculations

@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide", tinytex.clean = FALSE)
-packages_needed = c("ggplot2", "readxl", "data.table", "knitr")
+packages_needed = c("ggplot2", "readxl", "data.table", "knitr", "BIOMASS", "terra", "utils", "car")
 
@@
 ```
+
+**Journal**: Agroforestry Systems
+
+**Max words**: The total length of Original Articles, including figures, tables and references, should not exceed **10,000 words**. The number of figures and tables should not exceed 10 in total.
+
+**Word count**: xxx; **Figure and table count**: 5
 
@@
 # for now : download manually from temporary link
-# utils::unzip("dataverse_files.zip")
+if (!dir.exists("data")) create.dir("data")
+# utils::unzip("data/dataverse_files.zip")
 ```
@@
 
-# merge plot coordinates with general information (village, AFS)
-data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])
+# merge plot coordinates with general information (site, AFS)
+data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF", "AREA_TOTAL_PARCEL")])
 
 # simplify site names
-data_plot[, village := substr(ID_VILLAGE, 1, 1)]
-data_plot[village == "S", village := "Osso Luga"]
-data_plot[village == "G", village := "Gariuai"]
-
-## village location: create box around plots
-data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(village)]
-data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
-data_box[village == "Osso Luga", w := 0.05]
-data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
+data_plot[, site := substr(ID_VILLAGE, 1, 1)]
+data_plot[site == "S", site := "Osso Luga"]
+data_plot[site == "G", site := "Gariuai"]
 ```
@@
 ```{r get-env-data}
-# extract elevation and rainfall for the 2 villages
-village_env = data_plot[, as.list(expand.grid(seq(min(long), max(long), 0.001), 
+# precipitation from CHELSA
+if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
+  url <- "https://os.zhdk.cloud.switch.ch/envicloud/chelsa/chelsa_V2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio12_1981-2010_V.2.1.tif"
+  destfile <- "data/CHELSA_bio12_1981-2010_V.2.1.tif"
+  utils::download.file(url, destfile, method = "curl")
+
+  ## crop to reduce memory usage
+  box = c(124.5, 127.5, -9.2, -8.3) # Timor Leste
+  rst <- terra::rast(destfile)
+  rst <- terra::crop(rst, terra::ext(box))
+  terra::writeRaster(rst, filename = destfile, overwrite = TRUE) 
+}
+
+# 30m resolution SRTM
+if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
+  url <- "https://srtm.csi.cgiar.org/wp-content/uploads/files/srtm_5x5/TIFF/srtm_62_14.zip"
+  destfile <- "data/srtm_62_14.zip"
+  utils::download.file(url, destfile, method = "curl")
+  utils::unzip(destfile, exdir = "data", overwrite = TRUE)
+}
+
+# download cwd from Chave et al 2014
+if (!file.exists("data/CWD.tif")) {
+  url <- "http://chave.ups-tlse.fr/pantropical_allometry/CWD.tif.zip"
+  destfile <- "data/CWD.tif.zip"
+  utils::download.file(url, destfile, method = "curl")
+  utils::unzip(destfile, exdir = "data", overwrite = TRUE)
+}
+```
+
+```{r extract-env-data}
+# extract elevation and rainfall for the 2 sites
+site_env = data_plot[, as.list(expand.grid(seq(min(long), max(long), 0.001), 
                                               seq(min(lat), max(lat), 0.001))),
-                     .(village)]
+                        .(site)]
 
-colnames(village_env) = c("village", "long", "lat")
+colnames(site_env) = c("site", "long", "lat")
 
 # TODO add code to download prec and elev
-village_env[, prec := terra::extract(
-  terra::rast("D:/maps/CHELSA/CHELSA_bio12_1981-2010_V.2.1.tif"), 
+site_env[, prec := terra::extract(
+  terra::rast("data/CHELSA_bio12_1981-2010_V.2.1.tif"), 
   cbind(long, lat)
@@
 
-village_env[, elev := terra::extract(
+site_env[, elev := terra::extract(
   terra::rast("data/srtm_62_14.tif"), 
@@
 
-village_summary = village_env[, .(
+site_summary = site_env[, .(
   coord = paste(round(mean(long),2), round(mean(lat),2), sep = ", "), 
@@
   elev = paste(round(quantile(elev, c(0.025, 0.975))), collapse = " - ")),
-  .(village)]
+  .(site)]
 ```
@@
 The research was carried out in the center of the Baucau district, on the north-eastern coast of Timor-Leste.
-Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r village_summary[village=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r village_summary[village=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r village_summary[village=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r village_summary[village=="Osso Luga", prec]` mm and an altitude ranging `r village_summary[village=="Osso Luga", elev]` m (Fig.
+Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso Luga", elev]` m (Fig.
 \@ref(fig:map-sites)).
-The average soil properties differed between villages (Table \@ref(tab:stab-site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
-The two villages also have had different histories.
+The average soil properties differed between sites (Table \@ref(tab:stab-site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
+The two sites also have had different histories.
 Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2000.
@@
 Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
-Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot: within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
+Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Fig.
+\@ref(fig:design)): within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
 Height measurements were done with a dendrometer.
 
-Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using inaturalist and Pl@ntnet application, Geoffrey Hull’s monograph (Hull, 2006), the data bases of “the useful tropical plants” (https://tropical.theferns.info/) as well as the native plants websites of the NT herbarium from the Australian government (https://nt.gov.au/environment/native-plants/native-plants-and-nt-herbarium).
-The correspondence between tetum names and scientific names is provided in the supplemental material (Table S2).
+Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using iNaturalist and Pl\@ntnet application, Geoffrey Hull’s monograph [@Hull2006], the data bases of “the useful tropical plants” (https://tropical.theferns.info/) as well as the native plants websites of the NT herbarium from the Australian government (https://nt.gov.au/environment/native-plants/native-plants-and-nt-herbarium).
+The correspondence between tetum names and scientific names is provided in the supplemental material (Table \@ref(tab:tetum-names)).
 
@@
 
-In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (see figure S1).
+In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (Fig.
+\@ref(fig:design)).
 Soils were sampled with an auger at 0-10 cm and 20-30 cm.
@@
 
+```{r agc-estimation}
+### 1. Open data
+
+# Open large tree data (> 100 cm circumference)
+data_large <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV >100CM")
+
+# Open small tree data (30-100 cm circumference)
+data_small <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV 30-100CM")
+
+### 2. Combine tables
+
+# create a "N_Q_ID" column in data_large, with corresponding quadrat number =
+# "TOT" (entire plot)
+data_large$N_Q_ID <- paste(data_large$N_PLOT, "TOT", sep = "_")
+
+# common columns to be kept
+col_keep <- c("N_PLOT", "N_Q_ID", "SCIENTIFIC_NAME", "DIAM_130_CM", "HEIGHT")
+
+# stack tree tables
+data_tot <- rbind(data_large[, col_keep], data_small[, col_keep])
+
+# merge with plot information
+data_tot <- merge(data_tot, data_plot, by = "N_PLOT")
+
+# convert data_tot into data.table 
+data.table::setDT(data_tot)
+
+### 3. Clean tables
+
+# remove NA or <30 cm circumference values or < 100 cm outside subplots
+data_tot <- subset(data_tot, 
+                   !is.na(DIAM_130_CM) & 
+                     DIAM_130_CM >= 30/pi &
+                     (DIAM_130_CM >= 100/pi | grepl("Q", N_Q_ID)))
+
+# split scientific names into genus and species
+names <- data.table::tstrsplit(data_tot$SCIENTIFIC_NAME, " ")
+
+# get genera
+data_tot$GENUS <- names[[1]]
+
+# get species
+data_tot$SPECIES <- names[[2]]
+
+# Timoneus should be Timonius
+data_tot$GENUS[!is.na(data_tot$GENUS) & data_tot$GENUS == "Timoneus"] <- "Timonius"
+
+# remove one tree with DBH = 0
+data_tot <- subset(data_tot, DIAM_130_CM > 0)
+
+### 4. Get values for Chave et al., 2014 equation
+
+# get wood density
+data_wd <- BIOMASS::getWoodDensity(genus = data_tot$GENUS, species = data_tot$SPECIES)
+
+# 15% coefficient of variation on height
+cvH <- 0.15
+
+### estimate AGC
+data_sim_agc <- data.table(
+  data_tot,
+  BIOMASS::AGBmonteCarlo(
+    D = data_tot$DIAM_130_CM, 
+    WD = data_wd$meanWD, 
+    errWD = data_wd$sdWD, 
+    H = data_tot$HEIGHT, 
+    errH = data_tot$HEIGHT*cvH, 
+    Dpropag = "chave2004", 
+    n = 1000,
+    Carbon = TRUE
+  )$AGC_simu
+) |> 
+  melt(
+    measure.vars = paste0("V", 1:1000),
+    variable.name = "iter",
+    value.name =  "AGC"
+  )
+
+### 5. Apply a different allometry for palms
+palm_genera <- c("Cocos", "Areca", "Arenga", "Corypha", "Borassus")
+
+## height in cm, with a conversion factor between 0.8 and 1
+data_sim_agc[GENUS %in% palm_genera, 
+             HEIGHT := rnorm(1, HEIGHT, HEIGHT*cvH)*
+               runif(1, .8, 1)* 100]
+
+## add WD data
+data_wd_palms = unique(data_wd[data_wd$genus %in% palm_genera, c("genus", "meanWD", "sdWD")])
+data_sim_agc = merge(data_sim_agc, data_wd_palms, all.x = TRUE, by.x = "GENUS", by.y = "genus")
+data_sim_agc[GENUS %in% palm_genera, WD := rnorm(1, meanWD,sdWD)]
+
+data_sim_agc[GENUS %in% palm_genera, 
+             AGC := WD * HEIGHT * (DIAM_130_CM/2)^2 * pi * 
+               0.89 * # volumetric shrinkage
+               1e-6  # convert from g to tons
+]
+```
+
 We estimated tree-level AGC stocks from the measured tree CBH and height using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017].
@@
 Other taxa were assigned a plot-level average wood density.
-The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density, taken here as the average wood density in the plots (mean: xxx and standard deviation: xxx).
+The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density.
 The volume of the palm trunks was estimated as the volume of a cylinder, using the height and diameter measured in situ, with a volumetric shrinkage of 11% [@WD2023].
@@
 
+```{r bgc-estimation}
+# root-to-shoot ratio allometry from Ledo et al., 2018
+# use CWD values from Chave et al., 2014
+data_plot[, cwd := terra::extract(terra::rast("data/CWD.tif"), cbind(long, lat))]
+data_sim_agc <- merge(data_sim_agc, data_plot[, c("N_PLOT", "cwd")], by = "N_PLOT")
+
+# add uncertainty on R:S ratio? 
+data_sim_agc[, RS := exp(- 1.2312 - 0.0215 * DIAM_130_CM + 0.0002 * DIAM_130_CM^2 - 0.0007 * cwd)]
+data_sim_agc[, BGC := RS * AGC]
+```
+
 Plot-level AGC and BGC stocks were then obtained by summing tree-level stocks and dividing by the plot area.
+
+```{r aggregate-biomass-data}
+# replace with area = 0.05 for quadrats (5 quadrats of 0.01 ha each)
+data_sim_agc[, area := ifelse(grepl("Q", N_Q_ID), 0.01, AREA_TOTAL_PARCEL)]
+
+# melt data frame
+data_sim <- melt(data_sim_agc, id.vars = c("N_PLOT", "N_Q_ID", "ID_VILLAGE", "ID_AF", "iter", "area"), measure.vars = c("AGC", "BGC"))
+
+# estimate biomass per quadrat and plot
+data_sim <- data_sim[, .(value = sum(value/area)), .(N_PLOT, N_Q_ID, ID_VILLAGE, ID_AF, iter, variable)]
+
+## bootstrap AGC and BGC values in small plots
+data_sim = 
+  merge(
+    data_sim[!grepl("Q", N_Q_ID)], 
+    data_sim[grepl("Q", N_Q_ID), .(
+      valueSmall = ifelse(
+        length(value) == 0,   ## some quadrats have zero trees: complete their values
+        0, 
+        mean(sample(c(value, rep(0, 5 - length(value))), replace = TRUE))
+      )), .(N_PLOT, variable, iter)], 
+    by = c("N_PLOT", "variable", "iter")
+  )
+
+data_sim[, value := value + valueSmall]
+```
 
@@
 
-Soil organic carbon stocks (in Mg.ha-1) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:
+Soil organic carbon stocks (in Mg.ha$^{-1}$) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:
 
@@
 At depth $i$ = 10-20 cm, where the carbon content was not measured (see "Soil measurements" section), we used the average of the carbon content values at 0-10 and 20-30 cm.
+
+```{r open-soil-data}
+# open soil lab data
+data_soil_lab <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")
+
+# open soil field measurements
+data_soil_field <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")
+
+# change format of depth to correspond to other soil table
+data_soil_field$SAMPLE_DEPTH <- factor(data_soil_field$DEPTH_SAMPLE_C, levels = c("2_7", "12_17", "22_27"))
+levels(data_soil_field$SAMPLE_DEPTH) <- c("0-10", "10-20", "20-30")
+
+data_soil <- merge(data_soil_lab, data_soil_field, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
+data.table::setDT(data_soil)
+```
+
+```{r add-missing-values}
+## add intermediate soil carbon in 10-20 cm depth
+data_soil[, Corg := `Corg\n`]
+data_soil[, muCorg := mean(Corg, na.rm = TRUE), .(N_PLOT)]
+data_soil[is.na(Corg), Corg := muCorg]
+```
+
+```{r soc-error-propagation}
+data_soil[, c("h0", "h1") := lapply(data.table::tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
+
+## error propagation using a Monte Carlo method -------------
+data_sim_soil = data_soil[, .(mfine = rnorm(1000, Sample_DRY, Sample_DRY*0.1), # 10% error on fine soil mass
+                              Corg = rnorm(1000, Corg, Corg*0.12), # 12% error on C content measurement
+                              iter = paste0("V", 1:1000)), 
+                          .(N_PLOT, SAMPLE_DEPTH, h1, h0)]
+
+data_sim_soil[, Cstock :=             ## in Mg/ha
+                (h1 - h0) * 1e8 *     ## equivalent volume of a 1 ha soil sample, in cm3/ha
+                Corg/100 *            ## carbon organic content in fine soil (unitless)
+                (mfine * 1e-6) * ## mass of the fine fraction of the soil (Mg)
+                (1/250)        ## 1/volume of the cylinder, in cm-3
+]
+
+# sum over depths + at 0-10 cm depth
+data_sim_soil = data_sim_soil[, .(
+  SOC = sum(Cstock), 
+  SOC10 = Cstock[SAMPLE_DEPTH == "0-10"]),
+  .(N_PLOT, iter)
+]
+
+## melt data 
+data_sim_soil = melt(data_sim_soil, id.vars = c("N_PLOT", "iter"))
+```
+
+```{r merge-all}
+## merge with biomass data
+data_sim = rbind(data_sim[, colnames(data_sim_soil), with = FALSE], data_sim_soil) |> 
+  merge(data_plot[, c("N_PLOT", "site", "ID_AF")], by = "N_PLOT")
+
+# add total carbon stocks
+data_sim = rbind(data_sim, data_sim[, .(value = sum(value[variable != "SOC10"]), variable = "TOT"), .(N_PLOT, site, ID_AF, iter)])
+```
 
@@
 
-```{r agb-calculation}
-# get agb data
-source("script-agb.R")
+```{r boot-results}
+# plot level statistics
+dataC_plot = data_sim[, .(
+  mean = mean(value),
+  lwr = quantile(value, 0.025),
+  med = quantile(value, 0.5),
+  upr= quantile(value, 0.975)
+), .(plot = N_PLOT, site, afs = ID_AF, variable)]
+
+# AFS level statistics (with bootstrapping)
+dataC_afs = data_sim[, .(
+  mean = mean(value),
+  meanBs = mean(sample(value, replace = TRUE))
+), .(afs = ID_AF, variable, iter)][, .(
+  mean = mean(mean),        
+  lwr_mean = quantile(meanBs, 0.025),   
+  upr_mean = quantile(meanBs, 0.975)
+), .(afs, variable)]
+
+# site level statistics (with bootstrapping)
+dataC_site = data_sim[, .(
+  mean = mean(value),
+  meanBs = mean(sample(value, replace = TRUE))
+), .(site, variable, iter)][, .(
+  mean = mean(mean),        
+  lwr_mean = quantile(meanBs, 0.025),   
+  upr_mean = quantile(meanBs, 0.975)
+), .(site, variable)]
 ```
 
-```{r open-soil-data}
-# open organic carbon data
-# bulk density in Gariuai according to SoilGrids = 125 cg/cm3
-data_soil <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")
+```{r group-comparison}
+# shapiro wilk test - normality
+dataC_plot[variable %in% c("AGC", "SOC"), .(
+  shap = shapiro.test(log(mean))$p.value
+), .(afs, variable)]
+# levene test of homoscedasticity
+dataC_plot[variable %in% c("AGC", "SOC"), 
+           leveneTest(log(mean) ~ afs*site)$`Pr(>F)`[1], .(variable)]
 
-# add bulk density and coarse fraction
-data_soil2 <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")
-
-# bulk density from direct measurements
-data_soil2$BD <- data_soil2$Sample_DRY/data_soil2$V_cylinder
-
-# coarse fraction
-data_soil2$CF <- data_soil2$V_rocks_tot/data_soil2$V_cylinder
-
-# change format of depth to correspond to other soil table
-data_soil2$SAMPLE_DEPTH <- factor(data_soil2$DEPTH_SAMPLE_C, levels = c("2_7", "12_17", "22_27"))
-levels(data_soil2$SAMPLE_DEPTH) <- c("0-10", "10-20", "20-30")
-
-data_soil <- merge(data_soil, data_soil2, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
-data.table::setDT(data_soil)
-```
-
-```{r add-missing-values}
-## add intermediate soil carbon in 10-20 cm depth (for now: no value)
-data_soil[, Corg := mean(`Corg\n`, na.rm = TRUE), .(N_PLOT)]
-data_soil[!is.na(`Corg\n`), Corg := `Corg\n`]
-
-# replace one missing value of coarse fraction with plot level mean
-data_soil[, CF_site := mean(CF, na.rm = TRUE), .(N_PLOT)]
-data_soil[is.na(CF), CF := CF_site]
-```
-
-```{r total-soc}
-# calculate total soil carbon stock per depth
-data_soil[, c("h0", "h1") := lapply(data.table::tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
-data_soil[, Cstock :=             ## in Mg/ha
-            (h1 - h0) * 1e8 *     ## equivalent volume of a 1 ha soil sample, in cm3/ha
-            Corg/100 *            ## carbon organic content in fine soil (unitless)
-            (Sample_DRY * 1e-6) * ## mass of the fine fraction of the soil (Mg)
-            (1/V_cylinder)        ## 1/volume of the cylinder, in cm-3
-]
-
-data_soil_C <-
-  data_soil[, .(
-    meanSOC = sum(Cstock),
-    SC10 = Cstock[SAMPLE_DEPTH == "0-10"],
-    clay = mean(as.numeric(CLAY), na.rm = TRUE)
-  ), .(N_PLOT)]
-
-setkey(data_soil_C, "N_PLOT")
-```
-
-```{r soc-error-propagation}
-## error propagation using a Monte Carlo method -------------
-data_soil_unc = data_soil[, .(mass = rnorm(1000, Sample_DRY, Sample_DRY*0.1), 
-                              Corg = rnorm(1000, Corg, Corg*0.12), 
-                              iter = 1:1000), 
-                          .(N_PLOT, SAMPLE_DEPTH, h1, h0)]
-
-data_soil_unc = data_soil_unc[, .(
-  Cstock = sum((h1 - h0) * 1e8 / 250 * Corg/100 * (mass * 1e-6)), 
-  Cstock10 = ((h1 - h0) * 1e8 / 250 * Corg/100 * (mass * 1e-6))[SAMPLE_DEPTH == "0-10"]),
-  .(N_PLOT, iter)
-]
-
-data_soil_quantile = data_soil_unc[, .(
-  lwrSOC = quantile(Cstock, 0.025), 
-  medSOC = quantile(Cstock10, 0.5),
-  uprSOC = quantile(Cstock, 0.975),
-  lwrSOC10 = quantile(Cstock10, 0.025), 
-  medSOC10 = quantile(Cstock10, 0.5),
-  uprSOC10 = quantile(Cstock10, 0.975)), 
-  .(N_PLOT)]
-```
-
-```{r combine-agb-soc-data}
-# combine AGB and soil C
-dataC <- merge(data_agb, data_soil_C, by = "N_PLOT", all=TRUE)
-
-# add soil carbon uncertainty
-dataC <- merge(dataC, data_soil_quantile, all=TRUE)
-
-dataC[N_PLOT==6, `:=`(ID_VILLAGE = "GD", ID_AF = "CF")]
-
-dataC[, village := substr(ID_VILLAGE, 1, 1)]
-dataC$ID_AF <- factor(dataC$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
+# 2-way anova 
+anova_table = 
+  dataC_plot[variable %in% c("AGC", "SOC"), 
+           .(pvalue = round(anova(lm(log(mean) ~ site + afs))$`Pr(>F)`[1:2], 3)), 
+                   .(variable)]
+anova_table$factor = rep(c("site", "afs"), 2)
 ```
@@
 
-```{r tab-median, results = "show"}
-
-## bootstrapping
-
-#1. prepare data
-data_agc_boot = melt(data_agb_unc, 
-                     measure.vars = c("AGC", "BGC"), 
-                     id.vars = c("ID_VILLAGE", "N_PLOT", "ID_AF", "iter"))
-
-data_soil_boot = data_soil_unc[, c("N_PLOT", "iter", "Cstock")]
-colnames(data_soil_boot)[3] = "value"
-data_soil_boot$variable = "SOC"
-data_soil_boot$iter = paste0("V", data_soil_boot$iter)
-
-data_soil_boot = merge(data_soil_boot, data_plot[, c("N_PLOT","ID_VILLAGE", "ID_AF")])
-
-data_boot = rbind(data_agc_boot, data_soil_boot[, colnames(data_agc_boot), with = FALSE])
-
-# 2. add total stocks
-data_boot_tot = data_boot[, .(value = sum(value), variable = "TOT"), .(ID_VILLAGE, N_PLOT, ID_AF, iter)]
-data_boot = rbind(data_boot, data_boot_tot[, colnames(data_boot), with = FALSE])
-
-data_boot[, village := substr(ID_VILLAGE, 1, 1)]
-
-# 3. bootstrap
-data_boot_saf = data_boot[, .(value = mean(sample(value, length(value), replace = TRUE))),
-                          .(ID_AF, variable, iter)]
-
+```{r summary, results = "show"}
 # calculate median and 95% confidence interval at site level
-data_boot_saf  = data_boot_saf[, .(
-  value = paste0(
-    signif(mean(value), 3), " (",
-    signif(quantile(value, 0.025), 3), "-", 
-    signif(quantile(value, 0.975), 3), ")"
-  )), 
-  .(ID_AF, variable)]
-
-data_boot_saf = dcast(data_boot_saf, ID_AF ~ variable, value = "value")
-
-knitr::kable(data_boot_saf, col.names = c("Agroforestry system", "Aboveground carbon (MgC/ha)", "Belowground carbon (MgC/ha)", "Soil organic carbon (MgC/ha)", "Total carbon (MgC/ha)"))
-write.csv(data_boot_saf, file = "tables/C_median_SAF.csv", row.names = FALSE)
+dataC_afs[, value := paste0(
+  signif(mean, 3), " (",
+  signif(lwr_mean, 3), "-", 
+  signif(upr_mean, 3), ")")] |> 
+  subset(variable != "SOC10") |> 
+  dcast(afs ~ variable, value.var = "value") |> 
+  knitr::kable(col.names = c("Agroforestry system", 
+                             "Aboveground carbon (MgC/ha)", 
+                             "Belowground carbon (MgC/ha)", 
+                             "Soil organic carbon (MgC/ha)", 
+                             "Total carbon (MgC/ha)")) 
 ```
 
-```{r group-comparison}
-dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
-## test normality of the data
-
-# log-transform data to get normal distributions (because strictly positive data)
-hist((dataC$meanAGC), breaks = 20)
-hist((dataC$meanSOC))
-
-# test normality
-dataC[, .(
-  shap.A = shapiro.test(log(medAGC))$p.value, 
-  shap.S = shapiro.test(log(medSOC))$p.value
-), .(ID_AF, ID_SUCO)]
-# levene test of homoscedasticity
-library(car)
-dataC[, leveneTest(log(medAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
-dataC[, leveneTest(log(medSOC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
-
-# # pairwise group comparison with Tukey HSD method
-# # https://arc.lib.montana.edu/book/statistics-with-r-textbook/item/59
-# library(multcomp)
-# aovAGC = aov(meanAGC ~ ID_AF, data = dataC)
-# tuk <- glht(aovAGC, linfct = mcp(ID_AF = "Tukey"))
-# summary(tuk)          # standard display
-# tuk.cld <- cld(tuk)   # letter-based display
-# 
-# plot(tuk.cld)
-
-# # soc
-# aovSOC = aov(meanSOC ~ ID_AF, data = dataC)
-# tuk <- glht(aovSOC, linfct = mcp(ID_AF = "Tukey"))
-# summary(tuk)          # standard display
-# tuk.cld <- cld(tuk)   # letter-based display
-# 
-# opar <- par(mai=c(1,1,1.5,1))
-# plot(tuk.cld)
-
-## non parametric test
-#updated translation of help page implementation of Nemenyi-Damico-Wolfe-Dunn test
-# library(coin)
-# is.factor(dataC$ID_AF)
-# NDWD <-
-#   independence_test(meanAGC ~ ID_AF, data = dataC)
-# , distribution = approximate(B = 10000),
-# ytrafo = function(data) trafo(data, numeric_trafo = rank_trafo),
-# xtrafo = mcp_trafo(iv = "Tukey"))
-
-### global p-value
-# print(pvalue(NDWD))
-```
-
-```{r tab-anova, results = "show"}
-# 2-way anova 
-dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
-anov1 = anova(lm(log(medAGC)~ID_SUCO+ID_AF, data = dataC))
-anov2 = anova(lm(log(medSOC)~ID_SUCO+ID_AF, data = dataC))
-tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
-                      ind_var = rep(c("SUCO", "AF"), 2), 
-                      p = round(c(anov1$`Pr(>F)`[1:2], anov2$`Pr(>F)`[1:2]), 3)
-)
-
-knitr::kable(tab_anov)
-write.csv(tab_anov, file = "tables/anov.csv", row.names = FALSE)
-
-## non parametric 2 way anova: anova on ranks
-# anov1 = anova(lm(rank(meanAGC)~ID_SUCO+ID_AF, data = dataC))
-# anov2 = anova(lm(rank(meanSOC)~ID_SUCO+ID_AF, data = dataC))
-# tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
-#                       ind_var = rep(c("SUCO", "AF"), 2), 
-#                       p = round(c(anov1$`Pr(>F)`[1:2], anov2$`Pr(>F)`[1:2]), 3)
-# )
-
+```{r anova, results = "show"}
+knitr::kable(anova_table[, c("variable", "factor", "pvalue")])
 ```
@@
 ```{r map-sites, fig.height = 6, fig.width = 7}
+
+## site location: create box around plots
+data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(site)]
+data_box[site == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
+data_box[site == "Osso Luga", w := 0.05]
+data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
+
 ## create Timor map
@@
   geom_tile(data = data_box, aes(x = xc, y = yc, width = w, height = w), fill = NA, color = 1, linewidth = 0.5) +
-  ggrepel::geom_text_repel(data = data_box, aes(x = xc, y = yc, label = village)) + 
+  ggrepel::geom_text_repel(data = data_box, aes(x = xc, y = yc, label = site)) + 
   labs(x = "", y = "") +
@@
 elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
-elevS <- subset(elev, x > data_box[village == "Osso Luga"]$x1 & 
-                  x < data_box[village == "Osso Luga"]$x2 & 
-                  y > data_box[village == "Osso Luga"]$y1 & 
-                  y < data_box[village == "Osso Luga"]$y2)
-elevS$village = "Osso Luga"
-elevG <- subset(elev, x > data_box[village == "Gariuai"]$x1 & 
-                  x < data_box[village == "Gariuai"]$x2 & 
-                  y > data_box[village == "Gariuai"]$y1 & 
-                  y < data_box[village == "Gariuai"]$y2)
-elevG$village = "Gariuai"
+elevS <- subset(elev, x > data_box[site == "Osso Luga"]$x1 & 
+                  x < data_box[site == "Osso Luga"]$x2 & 
+                  y > data_box[site == "Osso Luga"]$y1 & 
+                  y < data_box[site == "Osso Luga"]$y2)
+elevS$site = "Osso Luga"
+elevG <- subset(elev, x > data_box[site == "Gariuai"]$x1 & 
+                  x < data_box[site == "Gariuai"]$x2 & 
+                  y > data_box[site == "Gariuai"]$y1 & 
+                  y < data_box[site == "Gariuai"]$y2)
+elevG$site = "Gariuai"
 
@@
 
-## Map of villages and plots with elevation
+## Map of sites and plots with elevation
 g2 = 
@@
 ## melt data.table
-dataC_all <- melt(dataC, id.vars = c("N_PLOT", "ID_VILLAGE", "ID_AF", "village"), 
+dataC_all <- melt(dataC, id.vars = c("N_PLOT", "ID_VILLAGE", "ID_AF", "site"), 
                   measure.vars = c("meanAGC", "meanBGC", "meanSOC"))
 
-dataC_all <- dataC_all[, .(value = mean(value, na.rm = TRUE)), .(village, ID_AF, variable)]
+dataC_all <- dataC_all[, .(value = mean(value, na.rm = TRUE)), .(site, ID_AF, variable)]
 
@@
 levels(dataC_all$variable) <- c("AGC", "BGC", "SOC")
-dataC_all[, village := factor(village)]
-levels(dataC_all$village) <- c("Gariuai", "Osso Luga")
+dataC_all[, site := factor(site)]
+levels(dataC_all$site) <- c("Gariuai", "Osso Luga")
 
@@
   geom_bar(position="stack", stat="identity") + 
-  facet_wrap(~village, nrow = 1, strip.position = "bottom") +
+  facet_wrap(~site, nrow = 1, strip.position = "bottom") +
   scale_fill_manual(values = c("forestgreen", "darkgoldenrod1", "coral4")) +
@@
 ## mixed effect linear regression
-lmeReg <- lmerTest::lmer(log(meanSOC) ~ log(meanAGC) + (1|village), data = dataC)
+lmeReg <- lmerTest::lmer(log(meanSOC) ~ log(meanAGC) + (1|site), data = dataC)
 # Nakagawa’s R2 for mixed models
@@
 # same for topsoil layer (0-10 cm)
-lmeReg10 <- lmerTest::lmer(log(SC10) ~ log(meanAGC) + (1|village), data = dataC)
+lmeReg10 <- lmerTest::lmer(log(SC10) ~ log(meanAGC) + (1|site), data = dataC)
 # Nakagawa’s R2 for mixed models
@@
 
-# change village names
-dataC[, village := factor(village)]
-levels(dataC$village) <- c("Gariuai", "Osso Luga")
+# change site names
+dataC[, site := factor(site)]
+levels(dataC$site) <- c("Gariuai", "Osso Luga")
 
 # stack data for figure
-dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSOC", "lwrSOC", "uprSOC")]
+dataC_30 = dataC[, c("N_PLOT", "site", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSOC", "lwrSOC", "uprSOC")]
 dataC_30$depth = "Soil depth: 0-30 cm"
-dataC_10 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSOC10", "uprSOC10")]
+dataC_10 = dataC[, c("N_PLOT", "site", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSOC10", "uprSOC10")]
 dataC_10$depth = "Soil depth: 0-10 cm"
@@
 ggplot(dataC_fig, aes(color = ID_AF)) +
-  geom_point(aes(x = meanAGC, y = meanSOC, shape = village), size = 2) +
+  geom_point(aes(x = meanAGC, y = meanSOC, shape = site), size = 2) +
   geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
@@
 
+![(#fig:design) Plot sampling design.
+(a) Typical plot < 1 ha, in all AFS except SP.
+(b) Typical 1-ha plot in SP systems.
+Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.](figures/plot-design.jpg)
+
 ```{r afs-typology}
@@
             aes(x = CF*100-5, y = 15,
-                label = paste("Village:", ID_VILLAGE, "/ SAF:", ID_TYPE, "/ Plot:", N_PLOT))) +
+                label = paste("Site:", ID_VILLAGE, "/ SAF:", ID_TYPE, "/ Plot:", N_PLOT))) +
   labs(x = "Elements grossiers (% du volume)",
@@
 ```{r sfig-clay-corg, fig.height = 4, fig.width = 6}
-data_soil[, village := substr(ID_VILLAGE, 1, 1)]
-data_soil[village == "S", village := "Osso Luga"]
-data_soil[village == "G", village := "Gariuai"]
+data_soil[, site := substr(ID_VILLAGE, 1, 1)]
+data_soil[site == "S", site := "Osso Luga"]
+data_soil[site == "G", site := "Gariuai"]
 
@@
   log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
-    (1|village)+ 
+    (1|site)+ 
     (1|SAMPLE_DEPTH), 
@@
 
-ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, col = SAMPLE_DEPTH, shape = village)) +
+ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, col = SAMPLE_DEPTH, shape = site)) +
   geom_point() +
@@
 dataCelev = merge(dataC[, c("N_PLOT", "meanSOC", "uprSOC", "lwrSOC")], 
-                  data_plot[, c("N_PLOT", "elev", "village")], by = "N_PLOT")
+                  data_plot[, c("N_PLOT", "elev", "site")], by = "N_PLOT")
 
-lmerTest::lmer(log(meanSOC) ~ log(elev) + (1|village), data = dataCelev) |> 
+lmerTest::lmer(log(meanSOC) ~ log(elev) + (1|site), data = dataCelev) |> 
   summary()
 
-ggplot(dataCelev, aes(x = elev, y = meanSOC, col = village)) +
+ggplot(dataCelev, aes(x = elev, y = meanSOC, col = site)) +
   geom_pointrange(aes(ymin=lwrSOC, ymax = uprSOC)) + 
@@
 vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "CA", "Mg", "K", "Na", "CEC")
-data_soil_plot = melt(data_soil, id.vars = c("village", "N_PLOT", "SAMPLE_DEPTH"), 
+data_soil_plot = melt(data_soil, id.vars = c("site", "N_PLOT", "SAMPLE_DEPTH"), 
                       measure.vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n"))
@@
     value = rnorm(100, value, sd = 0.12*value), n = 1:100), 
-    .(village, N_PLOT, SAMPLE_DEPTH, variable)]
+    .(site, N_PLOT, SAMPLE_DEPTH, variable)]
 
@@
     value = mean(sample(value, length(value), replace = TRUE))),
-    .(village, variable, n)]
+    .(site, variable, n)]
 
@@
   )), 
-  .(village, variable)]
+  .(site, variable)]
 
 # change table shape to have one column per variable
-data_soil_site = dcast(data_soil_site, village ~ variable, value.var = "value")
-data_soil_site = data_soil_site[, c("village", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "Ca", "K", "Mg", "Na", "CEC")]
+data_soil_site = dcast(data_soil_site, site ~ variable, value.var = "value")
+data_soil_site = data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "Ca", "K", "Mg", "Na", "CEC")]
 colnames(data_soil_site) = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)","N", "C/N", "pH", "Ca", "K", "Mg", "Na", "CEC")
@@
 
-```{r tab-tetum-names}
+```{r tetum-names}
 sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)

### 2024-03-21 manuscript.Rmd
correct palm biomass + add results and discussion

@@
 date: "2024-03-19"
-output: 
-  bookdown::word_document2:
-  fig_caption: yes
+output: officedown::rdocx_document
 bibliography: references.bib
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide", tinytex.clean = FALSE)
-packages_needed = c("ggplot2", "readxl", "data.table", "knitr", "BIOMASS", "terra", "utils", "car")
+knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide")
+packages_needed = c(
+  "ggplot2",
+  "readxl",
+  "data.table",
+  "knitr",
+  "BIOMASS",
+  "terra",
+  "utils",
+  "car",
+  "lmerTest",
+  "performance"
+)
 
@@
 
-**Word count**: xxx; **Figure and table count**: 5
+**Word count**: 7200 (with SI); **Figure and table count**: 5 + 6 in SI
 
@@
 In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastoral systems, young agroforests, home gardens and forest gardens.
-Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 255 MgC/ha, similar to an old-growth tropical forest.
+Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, similar to an old-growth tropical forest.
 Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
@@
 
-Although AFS can make agriculture more resilient while improving its carbon footprint, poor adaptation to the local context can explain many failures to introduce new agroforestry techniques [@Coe2014].
+<!-- Although AFS can make agriculture more resilient while improving its carbon footprint, poor adaptation to the local context can explain many failures to introduce new agroforestry techniques [@Coe2014].
 Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
@@
 A better understanding of traditional agroforestry practices and their ecological characteristics could provide better support for the local development of agroforestry.
+-->
 
@@
 In this context, productive, resilient and sustainable agriculture is critical for the country, and agroforestry has been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022;@Cogne2024].
-Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
+Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024] and can be key elements for biodiversity conservation in the tropics [@Perfecto2008;@Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
 Better quantification of these carbon stocks would improve the understanding of the ecological importance of these AFS at both local and global scales, and help to integrate them into national and international climate change mitigation initiatives.
@@
 The research was carried out in the center of the Baucau district, on the north-eastern coast of Timor-Leste.
-Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso Luga", elev]` m (Fig.
-\@ref(fig:map-sites)).
-The average soil properties differed between sites (Table \@ref(tab:stab-site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
+Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso Luga", elev]` m (Figure\ \@ref(fig:map-sites)).
+The average soil properties differed between sites (Table S\@ref(tab:site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
 The two sites also have had different histories.
@@
 We used the typology described in @Cogne2024, developed in the same study area.
-This typology was based on the one proposed by @Nair1993;@Nair2022, and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
+This typology was based on the one proposed by @Nair1993;@Nair2021, and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
 The five AFS identified were: crop system including a wooded fallow phase (CF), Sylvopastoral (SP), Young agroforest (YA), Home garden (HG) and Forest garden (FG).
@@
 
-The AFS are presented in more detail in the supplemental information (Table \@ref(tab:afs-typology)) and in @Cogne2024.
+The AFS are presented in more detail in the supplemental information (Table S\@ref(tab:afs-typology)) and in @Cogne2024.
 
@@
 
-We randomly selected 3 plots from each of the 5 traditional AFS identified in each of the two sites (30 plots in total; Fig.
-\@ref(fig:map-sites)).
+We randomly selected 3 plots from each of the 5 traditional AFS identified in each of the two sites (30 plots in total; Figure\ \@ref(fig:map-sites)).
 Sampling plots were defined as the entire agricultural plot (area between 0.06-1.25 ha) for all AFS except SP plots, which were generally much larger than 1 ha.
@@
 Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
-Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Fig.
-\@ref(fig:design)): within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
+Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Figure\ S\@ref(fig:design)): within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
 Height measurements were done with a dendrometer.
@@
 Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using iNaturalist and Pl\@ntnet application, Geoffrey Hull’s monograph [@Hull2006], the data bases of “the useful tropical plants” (https://tropical.theferns.info/) as well as the native plants websites of the NT herbarium from the Australian government (https://nt.gov.au/environment/native-plants/native-plants-and-nt-herbarium).
-The correspondence between tetum names and scientific names is provided in the supplemental material (Table \@ref(tab:tetum-names)).
+The correspondence between tetum names and scientific names is provided in the supplemental material (Table S\@ref(tab:tetum-names)).
 
@@
 
-In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (Fig.
-\@ref(fig:design)).
+In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (Figure\ S\@ref(fig:design)).
 Soils were sampled with an auger at 0-10 cm and 20-30 cm.
@@
 
+# estimate mean AGC
+data_tot$mean <- BIOMASS::computeAGB(
+  D = data_tot$DIAM_130_CM, 
+  WD = data_wd$meanWD, 
+  H = data_tot$HEIGHT
+) * 0.4713 # conversion factor from biomass to carbon
+
+### estimate AGC with uncertainty propagation
+
 # 15% coefficient of variation on height
@@
 
-### estimate AGC
 data_sim_agc <- data.table(
@@
   melt(
-    measure.vars = paste0("V", 1:1000),
+    measure.vars = c("mean", paste0("V", 1:1000)),
     variable.name = "iter",
@@
 data_sim_agc[GENUS %in% palm_genera, 
-             HEIGHT := rnorm(1, HEIGHT, HEIGHT*cvH)*
-               runif(1, .8, 1)* 100]
+             HEIGHT := ifelse(
+               iter == "mean", 
+               HEIGHT*.9* 100,  ## height in cm
+               rnorm(length(HEIGHT), HEIGHT, HEIGHT*cvH)*
+                 runif(length(HEIGHT), .8, 1)* 100
+             )]
 
@@
 data_wd_palms = unique(data_wd[data_wd$genus %in% palm_genera, c("genus", "meanWD", "sdWD")])
-data_sim_agc = merge(data_sim_agc, data_wd_palms, all.x = TRUE, by.x = "GENUS", by.y = "genus")
-data_sim_agc[GENUS %in% palm_genera, WD := rnorm(1, meanWD,sdWD)]
+data_sim_agc = merge(
+  data_sim_agc,
+  data_wd_palms,
+  all.x = TRUE,
+  by.x = "GENUS",
+  by.y = "genus"
+)
+data_sim_agc[GENUS %in% palm_genera, 
+             WD := ifelse(iter == "mean", 
+                          meanWD, 
+                          rnorm(length(meanWD), meanWD,sdWD))]
 
@@
                0.89 * # volumetric shrinkage
-               1e-6  # convert from g to tons
+               1e-6 * # convert from g to tons
+               0.4713 # conversion factor from biomass to carbon
 ]
@@
 The height of the palms was measured above the leaves and was therefore multiplied by a conversion factor of 0.9 to obtain the trunk height.
+Biomass was then converted into carbon with a factor of 0.4713 [@Thomas2012].
 
@@
 # melt data frame
-data_sim <- melt(data_sim_agc, id.vars = c("N_PLOT", "N_Q_ID", "ID_VILLAGE", "ID_AF", "iter", "area"), measure.vars = c("AGC", "BGC"))
+data_sim <-
+  melt(
+    data_sim_agc,
+    id.vars = c("N_PLOT", "N_Q_ID", "ID_VILLAGE", "ID_AF", "iter", "area"),
+    measure.vars = c("AGC", "BGC")
+  )
 
 # estimate biomass per quadrat and plot
-data_sim <- data_sim[, .(value = sum(value/area)), .(N_PLOT, N_Q_ID, ID_VILLAGE, ID_AF, iter, variable)]
-
-## bootstrap AGC and BGC values in small plots
+data_sim <-
+  data_sim[, .(value = sum(value / area)), .(N_PLOT, N_Q_ID, ID_VILLAGE, ID_AF, iter, variable)]
+
+## bootstrap AGC and BGC values in small plots (except for iter = mean)
 data_sim = 
@@
     data_sim[!grepl("Q", N_Q_ID)], 
-    data_sim[grepl("Q", N_Q_ID), .(
+    rbind(
+      data_sim[grepl("Q", N_Q_ID) & iter != "mean", .(
         valueSmall = ifelse(
@@
         )), .(N_PLOT, variable, iter)], 
-    by = c("N_PLOT", "variable", "iter")
+      data_sim[grepl("Q", N_Q_ID) & iter == "mean", 
+               .(valueSmall = sum(value / 5)), .(N_PLOT, variable, iter)]
+    ),
+    by = c("N_PLOT", "variable", "iter"), all = TRUE
   )
 
+# replace NAs is valueSmall (ie no trees in quadrats) with 0
+data_sim[, valueSmall := ifelse(is.na(valueSmall), 0, valueSmall)]
+
+## sum mean biomass in quadrats and entire plot
 data_sim[, value := value + valueSmall]
@@
 
-$$SOC_i = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10-6) \cdot (Splot \cdot \Delta d)}{Vcyl}$$
+$$SOC_i = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl}$$
 
@@
 ## error propagation using a Monte Carlo method -------------
-data_sim_soil = data_soil[, .(mfine = rnorm(1000, Sample_DRY, Sample_DRY*0.1), # 10% error on fine soil mass
-                              Corg = rnorm(1000, Corg, Corg*0.12), # 12% error on C content measurement
-                              iter = paste0("V", 1:1000)), 
+data_sim_soil = data_soil[, .(mfine = c(Sample_DRY, rnorm(1000, Sample_DRY, Sample_DRY*0.1)), # 10% error on fine soil mass
+                              Corg = c(Corg, rnorm(1000, Corg, Corg*0.12)), # 12% error on C content measurement
+                              iter = c("mean", paste0("V", 1:1000))), 
                           .(N_PLOT, SAMPLE_DEPTH, h1, h0)]
@@
 
-The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo() function from the BIOMASS package, using a (rather conservative) coefficient of variation of 15% on height measurements (Hunter et al 2013); large and small errors on 5 and 95% of trees respectively (Chave et al 2004; Réjou-Méchain et al., 2017).
-To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (Cirad laboratory in Montpellier, France), 10% for the mass of the fine fraction (Page-Dumroese et al.
-1999).
+The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo() function from the BIOMASS package, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013]; large and small errors on 5 and 95% of trees respectively [@Chave2004;@Rejou-Mechain2017].
+To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (Cirad laboratory in Montpellier, France), 10% for the mass of the fine fraction [@Page1999].
 
@@
 dataC_plot = data_sim[, .(
-  mean = mean(value),
+  mean = value[iter=="mean"],
   lwr = quantile(value, 0.025),
-  med = quantile(value, 0.5),
   upr= quantile(value, 0.975)
@@
 ), .(afs = ID_AF, variable, iter)][, .(
-  mean = mean(mean),        
+  mean = mean[iter=="mean"],        
   lwr_mean = quantile(meanBs, 0.025),   
@@
 ), .(site, variable, iter)][, .(
-  mean = mean(mean),        
+  mean = mean[iter=="mean"],    
   lwr_mean = quantile(meanBs, 0.025),   
@@
 ), .(site, variable)]
+
+# AFS + site level statistics (with bootstrapping)
+dataC_afs_site = data_sim[, .(
+  mean = mean(value),
+  meanBs = mean(sample(value, replace = TRUE))
+), .(afs = ID_AF, site, variable, iter)][, .(
+  mean = mean[iter=="mean"],        
+  lwr_mean = quantile(meanBs, 0.025),   
+  upr_mean = quantile(meanBs, 0.975)
+), .(afs, site, variable)]
 ```
@@
 
+```{r lmereg}
+data_reg = melt(dataC_plot, measure.vars = c("mean", "lwr", "upr"), variable.name = "stat") |> 
+  dcast(plot+ site+ afs ~ variable + stat)
+
+## mixed effect linear regression
+lmeReg <- lmerTest::lmer(log(SOC_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
+# Nakagawa’s R2 for mixed models
+R2 = performance::r2_nakagawa(lmeReg)
+summary(lmeReg)
+
+# same for topsoil layer (0-10 cm)
+lmeReg10 <- lmerTest::lmer(log(SOC10_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
+# Nakagawa’s R2 for mixed models
+R210 = performance::r2_nakagawa(lmeReg10)
+summary(lmeReg10)
+
+dfreg = 
+  data.table(
+    depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
+    intercept = c(summary(lmeReg)$coefficients[1,1], summary(lmeReg10)$coefficients[1,1]), 
+    slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
+    pvalue = c(summary(lmeReg)$coefficients[2,5], summary(lmeReg10)$coefficients[2,5]),
+    R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
+    R2_marg = c(R2$R2_marginal, R210$R2_marginal)
+  )
+```
+
 # Results
 
-```{r summary, results = "show"}
+## Total carbon stocks
+
+Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+
+## Aboveground carbon
+
+AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)), but not between site (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
+The AFS with the lowest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+
+## Soil organic carbon
+
+SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)).
+However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
+The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$) than those in Osso Luga (`r dataC_site[site=="Osso Luga" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
+
+## Relationship between AGC and SOC
+
+The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure\ \@ref(fig:plot-lmereg)).
+This relationship remained insignificant (yet slightly stronger; p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
+The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure\ \@ref(fig:plot-lmereg)).
+Overall, AGC stocks were a poor predictor of SOC stocks.
+
+# Discussion
+
+## Large C stocks in traditional AFS in Timor Leste
+
+Compared to values reported in the literature for other AFS, the carbon stocks in this study were particularly high.
+The value of SOC stocks was similar to other reported values [@Feliciano2018], but some of the AFS studied here have particularly high stocks in their biomass [compared to the average of 74\ Mg\ C\ ha$^{-1}$ in @Ma2020].
+The AFS with the highest carbon stocks, forest gardens, had stocks close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$; @Lal2005].
+These high carbon stocks could be explained by the fact that forest gardens have generally grown for more than 50 years, with little human intervention (Table S1).
+In terms of structure and carbon stocks, forest gardens should therefore resemble old secondary forests, which have been shown to recover more than 60% of the AGC stocks and almost all the SOC stocks of an old forest [@Poorter2021].
+In addition, FGs are generally located close to villages, where soils are enriched by organic waste and animal manure, which can enhance tree growth [@Madejon2016].
+
+Although agroforestry has been identified as one of the possible levers to increase carbon storage in the Timorese agricultural landscape [@UNFCCC2022;@Paudel2022], to our knowledge no study has quantified carbon stocks in these systems.
+Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste's climate change mitigation commitments [@UNFCCC2022], and the potential carbon gains from wider adoption of these techniques.
+Further studies should help to increase the representativeness of the AFS sampled by including a greater diversity of AFS and environmental conditions.
+Another critical aspect is the quantification of temporal changes in carbon stocks, which could be assessed in future studies either by having chronosequences of plots where the time since installation is known, or by remeasuring plots through time.
+
+## AFS differ in C stored in their biomass
+
+The AFS studied here vary considerably in terms of AGC stocks, in agreement with our hypothesis H1a.
+This reflects the density and size of trees in the plot, which in turn depends on the importance of trees in the production system.
+In some systems, such as CF and SP, trees do not contribute to food production but have other functions, such as restoring soil fertility (e.g.
+during the fallow period) or providing shade for livestock.
+This more limited role for trees may explain why their total biomass is lower in these systems: in the CF system, the plot is cleared and annual crops are planted in full sun, while in the SP system, the growth of grazing grasses is favored by reducing the number of trees that shade them.
+
+It is also interesting to note that, in line with previous results, the age of the trees in the AFS largely explains the carbon stored in their biomass [@Ma2020].
+Indeed, the systems described in this study reflect different stages of transition in land use, which follow the growth of their trees.
+For example, a farmer may plant trees in a CF field, which will then become a YA with trees aged between one and ten years, where the main production still comes from annual crops.
+When the trees are mature and producing fruit, the YA becomes a HG or a FG.
+In the latter systems, the trees become the main source of production, both for food and wood.
+
+The location of AFS in the landscape is not random and could also explain the observed differences in AGC stocks.
+For example, most HGs are located close to houses, in alluvial valleys that are well supplied with water and where the soil is enriched with organic waste and animal manure, which is likely to increase soil fertility [@Palm2001].
+The same applies to FGs, which are former HGs that have been more or less abandoned, and YAs, where the farmer has chosen the site with the intention of turning it into a HG.
+We therefore expect HGs, FGs and YAs to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuter2018].
+Conversely, CF and SP are mainly located on limestone plateaus where water is deep and soils are shallow, which could reduce tree growth and biomass [@Imada2008].
+
+## Importance of historical and environmental factors on SOC
+
+Our results revealed that SOC was mainly sensitive to the differences in location and villages rather than AFS systems in each location.
+This was also reported by the meta-analysis of Feliciano et al.
+(2018) who highlighted that only 3.2% of the variance of the soil carbon sequestration was explained by agroforestry system choices.
+The two villages studied were different in terms of inherent soil properties (Table S3; Fig S2), altitudes (Fig.
+1; Fig.
+S3) and also history.
+All those driving factors independent of the AFS management could not be disentangled in this study but their cumulative effects were clearly observed.
+The difference in terms of soil texture and especially clay content was reported to be a key driving process in terms of SOC storage in the tropics (Don et al., 2010; Matus, 2021).
+In this study, the fine fractions of soil texture were however not linked to the SOC stocks (Fig S2).
+In addition to soil inherent conditions, other key driving factors for SOC changes, independent to the AFS management practices, may have led to those differences between.
+Gariuai was inhabited whereas Osso-Luga was not for a long time.
+The management and care of the plot in Gariuai such as pruning, livestock visits, optimizing the different stratas in terms of C-input could have led to increased soil organic carbon compared to abandoned systems (Lorenz & Lal, 2014).
+
+The AGC stocks differed from one system to another but we could not detect any significant difference in SOC between AFSs.
+This observation was in contradiction with our H1b.
+Our systems were rather contrasted and previous studies reported significant differences in SOC sequestration between different AFS (Ma et al., 2020, Kimaro et al., 2024).
+Local conditions and the rapid SOC mineralisation in tropical areas may explain this lack of sensitivity against AFs management practices (Sayer et al., 2019).
+Our experimental design with a limited number of replicates within each village could have also limited the significance of the tests.
+Some trends were highlighted, such as higher SOC in SP in Gariuai.
+The design was based on 3 true replicates for each of the systems in each village with the objective to limit as much as possible the possible influence of external co-factor in each village (e.g.
+farmer management practices variability, micro-topography, etc).
+However, much variability cannot be controlled in farmers’ plots and increasing the number of replicates for each AFS may have strengthened the power of the tests and may have led to more differences between systems.
+
+## Lack of relationship between soil and biomass carbon [Vincent]
+
+- Low correlation between NPP and SOC in tropical soils [ref: Sayer et al. (2019).  Tropical forest soil carbon stocks do not increase despite 15 years of doubled litter inputs.]
+
+- Functioning of AFS is similar to that of a tropical forest
+
+Some AFS have high SOC for reasons other than the presence of trees; e.g. effect of cattle in SP systems.
+
+## Perspectives: quantify biodiversity in these AFS [Marguerite]
+
+importance of ecological continuity - a range of biodiversity (plant/animal/micro-organisms) that can develop in these traditional AF systems: plots of less than 1ha, density + distribution of trees (intra- and inter-plot), multi-strata.
+
+perspectives: quantify other indicators, such as biodiversity and composition, not only for plants but also for other groups such as animals and micro-organisms; These other organisms are as important as trees in carbon storage [ref?]
+
+## How to conserve and improve carbon stocks in traditional AFS in Timor Leste?
+
+Although agroforestry can make agriculture more resilient while improving its carbon footprint and biodiversity conservation value (Cardinael et al., 2021), poor adaptation to the local context can explain many failures to adopt new agroforestry techniques (Coe et al., 2014).
+Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it (Gouyon et al., 1993, Aumeeruddy-Thomas, 1994; Peltier, 1996).
+This is the case of the traditional AFS studied here: their heritage and socio-economic values (as shown by Cogné & Lescuyer, 2024), in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally..
+
+Conservation of these traditional AFS should avoid their abandonment as the trees wither or as their production value declines.
+Clear-cutting old agroforests for full-sun crops may then become an attractive short-term strategy, providing a large cash inflow from the sale of wood products, followed by 4 or 5 years of agricultural production on fertile land.
+Unfortunately, this results in a significant loss of biodiversity and carbon stored in the biomass, and can lead to an irreversible loss of fertility on the plot through soil degradation and erosion.
+This phenomenon has been documented in a similar context in Sumatra (Indonesia), where Gouyon et al.
+(1993) highlighted the fact that old agroforests, which they called "jungle-rubber" and which are quite similar in structure to "forest gardens" (except for the rubber tree component, which has been replaced by other species in Timor), were being cleared in favor of more intensive production systems.
+This trend towards clearing jungle rubber was subsequently confirmed by Ekadinata & Vincent.
+(2011), and its impact on carbon stock decline was quantified by Villamor et al.
+(2014).
+
+Strategies to avoid clearing of these traditional AFS could be to develop new ways of intensifying production while maintaining high carbon and biodiversity values, and gradually renewing the tree population in the plots.
+For example, it would be interesting to develop renewal through clearings of about one hundred square meters (corresponding to the felling of 1 to 10 trees, depending on their crown area).
+In these clearings, young trees of species considered interesting for their future production would be associated, first with light crops (cereals, etc.) and then with crops of shady species (taro, etc.).
+Additional mechanisms to encourage the renewal of traditional AFS could be the provision of subsidies (e.g.
+payment of carbon credits) or low-interest loans to farmers wishing to renew their agroforestry plots.
+In addition, a "sustainable agriculture" type label could be awarded to these AFS products, and their value chains could be supported by governments and international certification bodies, as is already the case for coffee in many countries (Bertrand et al., 2019).
+
+## Conclusion
+
+Agroforestry systems in the Baucau district of Timor had variable but overall high carbon stocks.
+These systems have only recently been described in the scientific literature, and much remains to be done to improve our knowledge of their diversity and functioning.
+It is therefore important that all stakeholders including local authorities, universities and NGOs, take into account their existence, diversity and environmental and social value, in order to propose possible innovations to maintain or improve their functioning.
+
+# Tables
+
+```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (between parentheses) of total, aboveground, belowground, and soil organic carbon stocks in each agroforestry system."}
 # calculate median and 95% confidence interval at site level
-dataC_afs[, value := paste0(
+table_afs = dataC_afs[, value := paste0(
   signif(mean, 3), " (",
@@
   subset(variable != "SOC10") |> 
-  dcast(afs ~ variable, value.var = "value") |> 
-  knitr::kable(col.names = c("Agroforestry system", 
+  dcast(afs ~ variable, value.var = "value")
+
+ knitr::kable(table_afs[, c("afs", "TOT", "AGC", "BGC", "SOC")],
+    col.names = c("Agroforestry system", 
+                  "Total carbon (MgC/ha)", 
                   "Aboveground carbon (MgC/ha)", 
                   "Belowground carbon (MgC/ha)", 
-                             "Soil organic carbon (MgC/ha)", 
-                             "Total carbon (MgC/ha)")) 
-```
-
-```{r anova, results = "show"}
-knitr::kable(anova_table[, c("variable", "factor", "pvalue")])
-```
-
-```{r map-sites, fig.height = 6, fig.width = 7}
-
+                  "Soil organic carbon (MgC/ha)")) 
+```
+
+```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAS with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
+The p-value gives the probability of the means of the categories of each independent variable being the same.
+Bold p-values are lower than 0.05, and the mean difference between groups is thus considered significantly different than zero."}
+knitr::kable(anova_table[, c("variable", "factor", "pvalue")], 
+             col.names = c("Dependent variable", "Independent variable", "p-value"))
+```
+
+# Figures
+
+```{r fig.id = "map-sites", fig.height = 5, fig.width = 7, fig.cap="Plot location.
+(a) Map of Timor-Leste with the location of the two study sites (Gariuai and Osso-Luga) in the Bacau district.
+(b) Location of the 16 plots in Gariuai; (c) Location of the 15 plots in Osso-Luga.
+Panels (b) and (c) share the same spatial scale.
+In (b) and (c), the plots are represented by dots whose color represents the agroforestry system it belongs to.
+The background color represents the elevation (m), from the Shuttle Radar Topography Mission (SRTM), specifically the hole-filled CGIAR-SRTM (90 m resolution)."}
 ## site location: create box around plots
@@
 ggpubr::ggarrange(ggpubr::ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))
-
-ggsave("figures/mapPlotsSRTM.png", height =4.5, width = 6)
-```
-
-```{r fig-2-barplot, fig.height = 5, fig.width = 6}
-## melt data.table
-dataC_all <- melt(dataC, id.vars = c("N_PLOT", "ID_VILLAGE", "ID_AF", "site"), 
-                  measure.vars = c("meanAGC", "meanBGC", "meanSOC"))
-
-dataC_all <- dataC_all[, .(value = mean(value, na.rm = TRUE)), .(site, ID_AF, variable)]
-
+```
+
+```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Barplot of the amount of carbon in each AFS (CF: crop and fallow, SP: silvopastoral, YA: young agroforest, HG: home garden, FG: forest garden), per site.
+AGC: Aboveground (biomass) carbon; BGC: belowground (biomass) carbon; SOC: soil organic carbon."}
 ## reorder factors 
-
-dataC_all$ID_AF <- factor(dataC_all$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
-
-# uncertainty fig 2: bootstrap by MC iteration?
-
-dataC_all[, variable := factor(variable, levels = c("meanAGC", "meanBGC", "meanSOC"))]
-levels(dataC_all$variable) <- c("AGC", "BGC", "SOC")
-dataC_all[, site := factor(site)]
-levels(dataC_all$site) <- c("Gariuai", "Osso Luga")
-
-ggplot(dataC_all, aes(x = ID_AF, y = value, fill = variable)) +
+dataC_afs_site[, afs := factor(afs, levels = c("CF", "SP", "YA", "HG", "FG"))]
+
+ggplot(dataC_afs_site[variable %in% c("AGC", "BGC", "SOC")], 
+       aes(x = afs, y = mean, fill = variable)) +
   geom_bar(position="stack", stat="identity") + 
@@
   scale_y_continuous(expand = c(0, 0))
-ggsave("figures/fig-carbon-stocks-stack.png", height = 5, width = 6)
-```
-
-```{r tab-3-reg, results = "show"}
-## mixed effect linear regression
-lmeReg <- lmerTest::lmer(log(meanSOC) ~ log(meanAGC) + (1|site), data = dataC)
-# Nakagawa’s R2 for mixed models
-R2 = performance::r2_nakagawa(lmeReg)
-summary(lmeReg)
-
-# same for topsoil layer (0-10 cm)
-lmeReg10 <- lmerTest::lmer(log(SC10) ~ log(meanAGC) + (1|site), data = dataC)
-# Nakagawa’s R2 for mixed models
-R210 = performance::r2_nakagawa(lmeReg10)
-summary(lmeReg10)
-
-dfreg = 
-  data.table(
-    depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
-    intercept = c(summary(lmeReg)$coefficients[1,1], summary(lmeReg10)$coefficients[1,1]), 
-    slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
-    pvalue = c(summary(lmeReg)$coefficients[2,5], summary(lmeReg10)$coefficients[2,5]),
-    R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
-    R2_marg = c(R2$R2_marginal, R210$R2_marginal)
-  )
-
-```
-
-```{r fig-3-reg, fig.height = 4, fig.width = 8}
-dfpred = expand.grid(meanAGC = exp(seq(min(log(dataC$meanAGC), na.rm = TRUE),
-                                       max(log(dataC$meanAGC), na.rm = TRUE),
+```
+
+```{r fig.id = "plot-lmereg", fig.height = 3.5, fig.width = 7, fig.cap = "Scatter plot of aboveground carbon (AGC; x-axis) and soil organic carbon (SOC; y-axis) stocks.
+SOC stocks were estimated at the topsoil layer (0-10 cm; left panel) and at the total measured depth (0-30 cm; right panel).
+Each point represents the median C stock value in a sample plot, and the vertical and horizontal error bars are the 95% confidence interval.
+The colors correspond to the different agroforestry systems (CF: crop and fallow, SP: silvopastoral, YA: young agroforest, HG: home garden, FG: forest garden); the different shapes represent the two sites.
+The dotted lines represent the (non-significant) linear mixed models between log-transformed AGC and SOC stocks.
+For each linear mixed model, we report the marginal R2 (proportion of variance explained by fixed effects only) and the conditional R2 (proportion of variance explained by both fixed and random effects).
+Both x- and y-axes are log-transformed."}
+dfpred = expand.grid(AGC_mean = exp(seq(min(log(data_reg$AGC_mean), na.rm = TRUE),
+                                        max(log(data_reg$AGC_mean), na.rm = TRUE),
                                         length.out = 20)),
@@
 setDT(dfpred)
-dfpred[, meanSOC := exp(intercept + slope * log(meanAGC))]
-
-# change site names
-dataC[, site := factor(site)]
-levels(dataC$site) <- c("Gariuai", "Osso Luga")
+dfpred[, SOC_mean := exp(intercept + slope * log(AGC_mean))]
 
 # stack data for figure
-dataC_30 = dataC[, c("N_PLOT", "site", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSOC", "lwrSOC", "uprSOC")]
+dataC_30 = data_reg[, c("plot", "site", "afs",  
+                        "AGC_mean", "AGC_lwr", "AGC_upr",
+                        "SOC_mean", "SOC_lwr", "SOC_upr")]
 dataC_30$depth = "Soil depth: 0-30 cm"
-dataC_10 = dataC[, c("N_PLOT", "site", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSOC10", "uprSOC10")]
+dataC_10 = data_reg[, c("plot", "site", "afs",   
+                        "AGC_mean", "AGC_lwr", "AGC_upr",
+                        "SOC10_mean", "SOC10_lwr", "SOC10_upr")]
 dataC_10$depth = "Soil depth: 0-10 cm"
@@
 ## figure
-ggplot(dataC_fig, aes(color = ID_AF)) +
-  geom_point(aes(x = meanAGC, y = meanSOC, shape = site), size = 2) +
+ggplot(dataC_fig, aes(color = afs)) +
+  geom_point(aes(x = AGC_mean, y = SOC_mean, shape = site), size = 2) +
   geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
@@
             x = -Inf, y = Inf, hjust = -0.1, vjust = 2.2, parse = TRUE, color = 1) +
-  geom_linerange(aes(x = meanAGC, y = meanSOC, ymin = lwrSOC, ymax = uprSOC)) +
-  geom_errorbarh(aes(y = meanSOC, xmin = lwrAGC, xmax = uprAGC)) +
-  # geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
-  geom_line(data = dfpred, aes(x = meanAGC, y = meanSOC), lty = 2, col = 1) +
+  geom_linerange(aes(x = AGC_mean, y = SOC_mean, ymin = SOC_lwr, ymax = SOC_upr)) +
+  geom_errorbarh(aes(y = SOC_mean, xmin = AGC_lwr, xmax = AGC_upr)) +
+  geom_line(data = dfpred, aes(x = AGC_mean, y = SOC_mean), lty = 2, col = 1) +
   labs(color = "AFS", x = "AGC stocks (Mg C/ha)", y = "SOC stocks (Mg C/ha)", shape = "Site") +
@@
   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
-  # scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
   scale_x_log10() +
@@
   theme_classic()
-
-ggsave("figures/fig-scatterplot-carbon.png", height = 4, width = 8)
-```
-
-## Total carbon stocks
-
-Total carbon stocks varied between `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(min(V1))]` and `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(max(V1))]` Mg C/ha, with an average of `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(mean(V1))]` Mg C/ha and a coefficient of variation of `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT)][,round(sd(V1)/mean(V1)*100)]`%  (Table 1; Figure 2).
-The AFS with the lowest total carbon stocks was the `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT,ID_AF)][, ID_AF[which.min(V1)]]` with a median of `r data_boot_tot[variable=="TOT", mean(value), .(N_PLOT,ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest total carbon stocks was the `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT,ID_AF)][, ID_AF[which.max(V1)]]` with a median of `r data_boot_tot[variable=="TOT", median(value), .(N_PLOT,ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).
-
-## Aboveground carbon
-
-AGC stocks varied between `r dataC[, round(min(meanAGC, na.rm = TRUE))]` and `r dataC[, round(max(meanAGC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanAGC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanAGC, na.rm = TRUE)/mean(meanAGC, na.rm = TRUE))]`%  (Table 1; Figure 2).
-In line with our hypothesis (H1a), AGC stocks varied significantly between AFS (ANOVA p-value: `r tab_anov$p[tab_anov$response=="AGC" & tab_anov$ind_var == "AF"]`), but not between site (ANOVA p-value = `r tab_anov$p[tab_anov$response=="AGC" & tab_anov$ind_var == "SUCO"]`, Table 2).
-The AFS with the lowest AGC stocks was the `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.min(V1)]]` with an average of `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest AGC stocks was the `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.max(V1)]]` with an average of `r dataC[, mean(meanAGC , na.rm = TRUE), .(ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).
-
-## Soil organic carbon
-
-SOC stocks varied between `r dataC[, round(min(meanSOC, na.rm = TRUE))]` and `r dataC[, round(max(meanSOC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanSOC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanSOC, na.rm = TRUE)/mean(meanSOC, na.rm = TRUE))]`%  (Table 1; Figure 2).
-In disagreement with our hypothesis (H1b), no significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "AF"]`).
-However, the measured SOC stocks differed from suco to suco (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "SUCO"]`; Table 2).
-The SOC stocks in Gariuai were on average higher (`r dataC[ID_SUCO=="G", round(mean(meanSOC, na.rm = TRUE))]` Mg C/ha) than those in Osso Luga (`r dataC[ID_SUCO=="S", round(mean(meanSOC, na.rm = TRUE))]` Mg C/ha; Figure 2).
-
-## Relationship between AGC and SOC
-
-The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure 3), as hypothesized (H2b).
-Contrary to our hypothesis (H2a), this relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
-The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure 3).
-Overall, AGC stocks were a poor predictor of SOC stocks.
+```
 
@@
 
-![(#fig:design) Plot sampling design.
-(a) Typical plot < 1 ha, in all AFS except SP.
-(b) Typical 1-ha plot in SP systems.
-Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.](figures/plot-design.jpg)
-
-```{r afs-typology}
+## Tables
+
+```{r tab.id = "afs-typology", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
 typ = readxl::read_excel("tables/afs-typology.xlsx")
@@
 
-```{r sfig-distr-sol, fig.height = 5, fig.width = 15}
-## diagnostic données sol pour marguerite
-g1 <- ggplot(data_soil, aes(x = CF*100, col = SAMPLE_DEPTH)) +
-  geom_histogram(bins = 30) +
-  geom_text(data = subset(data_soil, CF > 0.3), angle = -60,
-            aes(x = CF*100-5, y = 15,
-                label = paste("Site:", ID_VILLAGE, "/ SAF:", ID_TYPE, "/ Plot:", N_PLOT))) +
-  labs(x = "Elements grossiers (% du volume)",
-       y = "Nombre d'échantillons (site x profondeur)",
-       col = "Profondeur (cm)") +
-  scale_y_continuous(expand = expand_scale(0, 0)) +
-  theme_classic()  +
-  theme(legend.position = c(0.8, 0.9))
-
-g2 <- ggplot(data_soil, aes(x = Corg, col = SAMPLE_DEPTH)) +
-  geom_histogram(bins = 20) +
-  labs(x = "Teneur en carbone (%)",
-       y = "Nombre d'échantillons (site x profondeur)",
-       col = "Profondeur (cm)") +
-  scale_y_continuous(expand = expand_scale(0, 0)) +
-  theme_classic() +
-  theme(legend.position = c(0.8, 0.9))
-
-g3 <- ggplot(data_soil, aes(x = BD, col = SAMPLE_DEPTH)) +
-  geom_histogram(bins = 20) +
-  labs(x = "Densité apparente",
-       y = "Nombre d'échantillons (site x profondeur)",
-       col = "Profondeur (cm)") +
-  scale_y_continuous(expand = expand_scale(0, 0)) +
-  theme_classic()+
-  theme(legend.position = c(0.8, 0.9))
-
-ggpubr::ggarrange(g1, g2, g3, labels = "auto", nrow = 1)
-ggsave("figures/fig-distr-sol.png", height = 5, width = 15)
-```
-
-```{r sfig-clay-corg, fig.height = 4, fig.width = 6}
-data_soil[, site := substr(ID_VILLAGE, 1, 1)]
-data_soil[site == "S", site := "Osso Luga"]
-data_soil[site == "G", site := "Gariuai"]
-
-ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY), y = Corg, col = SAMPLE_DEPTH)) +
-  geom_point() +
-  labs(col = "Soil depth (cm)", x = "Clay content (%)", y = "SOC content (%)") +
-  scale_color_brewer(type = "qual") +
-  scale_x_log10() + 
-  scale_y_log10() +
-  geom_smooth(method = "lm") +
-  theme_classic()
-ggsave("figures/fig-clay-corg.pdf", height = 4, width = 6)
-
-lmerTest::lmer(
-  log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
-    (1|site)+ 
-    (1|SAMPLE_DEPTH), 
-  data = data_soil[SAMPLE_DEPTH != "10-20"]) |> 
-  summary()
-
-ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, col = SAMPLE_DEPTH, shape = site)) +
-  geom_point() +
-  labs(col = "Soil depth (cm)", x = "Clay and fine silt content (%)", y = "SOC content (%)", shape = "Site") +
-  scale_color_brewer(type = "qual") +
-  # scale_x_log10() +
-  # scale_y_log10() +
-  # geom_smooth(method = "lm") +
-  theme_classic()
-ggsave("figures/fig-clay-silt-corg.png", height = 4, width = 6)
-```
-
-```{r test-ac-soc-by-depth}
-data_soil_agc <- merge(data_soil[, c("N_PLOT", "SAMPLE_DEPTH", "Cstock")],
-                       dataC[, c("N_PLOT","ID_AF", "ID_VILLAGE",  "meanAGC")], by = "N_PLOT")
-
-ggplot(data_soil_agc, aes(x = meanAGC, y = Cstock, col = SAMPLE_DEPTH)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-lm(data = data_soil_agc, Cstock ~ meanAGC * SAMPLE_DEPTH) |> summary()
-lm(data = data_soil_agc[SAMPLE_DEPTH == "0-10"], Cstock ~ meanAGC) |> summary()
-lm(data = data_soil_agc[SAMPLE_DEPTH == "10-20"], Cstock ~ meanAGC ) |> summary()
-lm(data = data_soil_agc[SAMPLE_DEPTH == "20-30"], Cstock ~ meanAGC ) |> summary()
-```
-
-```{r compare-lit-values}
-litval <- read_excel("data/literature-values.xlsx")
-colnames(litval) <- c("system", "pool", "value", "location", "ref")
-
-litval$main_crop <- NA
-litval$main_crop[grepl("palm", litval$system)] <- "oil palm"
-litval$main_crop[grepl("offee", litval$system)] <- "coffee"
-unique(litval$system[is.na(litval$main_crop)])
-
-litval$main_per <- NA
-litval$main_per[grepl("nset", litval$system)] <- "enset"
-litval$main_per[grepl("Inga", litval$system)] <- "inga"
-litval$main_per[grepl("Eucalyptus", litval$system)] <- "eucalyptus"
-litval$main_per[grepl("Pinus", litval$system)] <- "pinus"
-unique(litval$system[is.na(litval$main_per)])
-
-litval$ID_AF <- NA
-litval$ID_AF[grepl("orests|orest |orest$", litval$system)] <- "forest"
-litval$ID_AF[grepl("onoculture", litval$system)] <- "monoculture"
-litval$ID_AF[!is.na(litval$main_crop) & !is.na(litval$main_per)] <- "two-crops"
-litval$ID_AF[grepl("ome garden", litval$system)] <- "HG"
-unique(litval$system[is.na(litval$ID_AF)])
-
-litval$var <- as.numeric(gsub(" ", "", tstrsplit(litval$value, "±")[[2]]))
-litval$value <- gsub(" ", "", tstrsplit(litval$value, "±")[[1]])
-
-mult_values <- which(sapply(strsplit(litval$value, ",|and"), length)>1)
-
-new_rows = lapply(mult_values, function(i){
-  vals = as.numeric(strsplit(litval$value[i], ",|and")[[1]])
-  vals = vals[!is.na(vals)]
-
-  x = matrix(litval[i,], ncol = ncol(litval), nrow = length(vals), byrow = TRUE)
-  x[, colnames(litval) == "value"] = vals
-
-  x = as.data.frame(x)
-  colnames(x) = colnames(litval)
-  return(x)
-}
-)
-
-new_rows = do.call(rbind, new_rows)
-
-litval2 = rbind(new_rows, litval[-mult_values,])
-litval = data.frame(system = unlist(litval2$system),
-                    pool = unlist(litval2$pool),
-                    value = unlist(litval2$value),
-                    location = unlist(litval2$location),
-                    ref = unlist(litval2$ref),
-                    main_crop = unlist(litval2$main_crop),
-                    main_per = unlist(litval2$main_per),
-                    ID_AF = unlist(litval2$ID_AF),
-                    var = unlist(litval2$var))
-
-```
-
-### Effect of altitude on SOC stocks
-
-```{r sfig-altitude}
-dataCelev = merge(dataC[, c("N_PLOT", "meanSOC", "uprSOC", "lwrSOC")], 
-                  data_plot[, c("N_PLOT", "elev", "site")], by = "N_PLOT")
-
-lmerTest::lmer(log(meanSOC) ~ log(elev) + (1|site), data = dataCelev) |> 
-  summary()
-
-ggplot(dataCelev, aes(x = elev, y = meanSOC, col = site)) +
-  geom_pointrange(aes(ymin=lwrSOC, ymax = uprSOC)) + 
-  theme_classic() +
-  labs(x = "Elevation (m)", y = "SOC stocks (Mg C/ha)", col = "Site")
-ggsave("figures/soc-elev.png", height = 4, width = 6)
-```
-
-### Site soil characteristics
-
-```{r stab-site-soil}
+```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show"}
 data_soil$silt <- as.numeric(data_soil$`FINE SILT`) + as.numeric(data_soil$`COARSE SILT`)
@@
 
-vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "CA", "Mg", "K", "Na", "CEC")
-data_soil_plot = melt(data_soil, id.vars = c("site", "N_PLOT", "SAMPLE_DEPTH"), 
-                      measure.vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n"))
+# add site info to data_soil 
+data_soil <- merge(data_soil, data_plot[, c("N_PLOT", "site")])
+
+# choose variables to show
+vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n")
+
+data_soil_plot = melt(data_soil, 
+                      id.vars = c("N_PLOT", "SAMPLE_DEPTH", "site"), 
+                      measure.vars = vars)
 
@@
 
-# simulation 100 values per plot, soil depth and variable with a 12% coefficient
+# simulation 1000 values per plot, soil depth and variable with a 12% coefficient
 # of variation (CV value from lab)
@@
   data_soil_plot[, .(
-    value = rnorm(100, value, sd = 0.12*value), n = 1:100), 
+    value = rnorm(1000, value, sd = 0.12*value), n = 1:1000), 
     .(site, N_PLOT, SAMPLE_DEPTH, variable)]
@@
 data_soil_site = data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "Ca", "K", "Mg", "Na", "CEC")]
-colnames(data_soil_site) = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)","N", "C/N", "pH", "Ca", "K", "Mg", "Na", "CEC")
-
-write.csv(data_soil_site, "data/soil_site.csv", row.names = FALSE)
-```
-
-### Botanical identification from Tetum names
-
-```{r tetum-names}
+
+kable(data_soil_site, col.names = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)","N", "C/N", "pH", "Ca", "K", "Mg", "Na", "CEC"))
+```
+
+```{r tab.id = "tetum-names", tab.lp="supp-tab", results = "show"}
 sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
@@
 vern = vern[, .(latin = paste(latin, collapse = ", ")), .(tetum)]
-write.csv(vern, "data/species-correspondance.csv", row.names = FALSE)
-```
-
-```{r timer-leste-c-footprint-lulcc}
-df_eluc <- read_excel("C:/Users/piponiot-laroche/Downloads/National_LandUseChange_Carbon_Emissions_2023v1.0.xlsx", sheet = "OSCAR") 
-colnames(df_eluc) <- df_eluc[7,]
-df_eluc <- df_eluc[-(1:8),colnames(df_eluc) %in% c("unit: Tg C/year", "Timor-Leste")]
+kable(vern, col.names = c("Tetum name", "Scientific name"))
+```
+
+## Figures
+
+```{r fig.id="design", fig.cap = "Plot sampling design.
+(a) Typical plot < 1 ha, in all AFS except SP.
+(b) Typical 1-ha plot in SP systems.
+Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
+knitr::include_graphics("figures/plot-design.jpg") 
+```
+
+```{r fig.id = "clay-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap="Distribution of SOC content (y-axis; in %) values as a function of fine silt and clay content (x-axis; in %), soil depth (point color; in cm), and site (point shape).", fig.cap.pre="Figure S"}
+data_soil[, site := substr(ID_VILLAGE, 1, 1)]
+data_soil[site == "S", site := "Osso Luga"]
+data_soil[site == "G", site := "Gariuai"]
+
+lmerTest::lmer(
+  log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
+    (1|site)+ 
+    (1|SAMPLE_DEPTH), 
+  data = data_soil[SAMPLE_DEPTH != "10-20"]) |> 
+  summary()
+
+ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, shape = SAMPLE_DEPTH, col = site)) +
+  geom_point() +
+  scale_x_log10() + scale_y_log10() +
+  labs(shape = "Soil depth (cm)", 
+       x = "Clay and fine silt content (%)", 
+       y = "SOC content (%)", 
+       col = "Site") +
+  theme_classic()
+```
+
+```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in MgC/ha).
+Points are coloured by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.", fig.cap.pre="Figure S"}
+dataCelev = merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")], 
+                  data_plot[, c("N_PLOT", "long", "lat")], by.x = "plot", by.y = "N_PLOT")
+
+# extract elevation
+dataCelev[, elev := terra::extract(
+  terra::rast("data/srtm_62_14.tif"), 
+  cbind(long, lat)
+)]
+
+lmerTest::lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |> 
+  summary()
+
+ggplot(dataCelev, aes(x = elev, y = mean, col = site)) +
+  geom_pointrange(aes(ymin=lwr, ymax = upr)) + 
+  theme_classic() +
+  labs(x = "Elevation (m)", y = "SOC stocks (Mg C/ha)", col = "Site")
 ```

### 2024-03-22 manuscript.Rmd
add table referencing

@@
 date: "2024-03-19"
-output: officedown::rdocx_document
+output:
+  officedown::rdocx_document:
+    number_sections: true
 bibliography: references.bib
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide")
 packages_needed = c(
-  "ggplot2",
-  "readxl",
-  "data.table",
-  "knitr",
-  "BIOMASS",
-  "terra",
-  "utils",
-  "car",
-  "lmerTest",
-  "performance"
+  "ggplot2",      # version 3.4.4
+  "ggrepel",      # version 0.9.4
+  "ggspatial",    # version 1.1.9
+  "ggpubr",       # version 0.6.0
+  "readxl",       # version 1.4.3
+  "data.table",   # version 1.14.8
+  "knitr",        # version 1.45
+  "BIOMASS",      # version 2.1.11
+  "terra",        # version 1.7-55
+  "car",          # version 3.1-2
+  "lmerTest",     # version 3.1-3
+  "performance",  # version 0.10.8
+  "flextable"     # version 0.9.5
 )
@@
 lapply(packages_needed, require, character.only = TRUE)
+
+opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide")
 ```
@@
 In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastoral systems, young agroforests, home gardens and forest gardens.
-Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, similar to an old-growth tropical forest.
+Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 210\ MgC\ ha$^{-1}$, similar to an old-growth tropical forest.
 Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
@@
 
-Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO2e/yr over 2010-2014; @Pendrill2019].
+Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO$_2$e/yr over 2010-2014; @Pendrill2019].
 The demand for tropical agricultural land is expected to continue to increase as the population grows [@UN2022].
@@
 if (!dir.exists("data")) create.dir("data")
-# utils::unzip("data/dataverse_files.zip")
+unzip("data/dataverse_files.zip")
 ```
@@
 # get plot number
-data_coord[, N_PLOT := as.numeric(data.table::tstrsplit(N_Q_ID, "_")[[1]])]
+data_coord[, N_PLOT := as.numeric(tstrsplit(N_Q_ID, "_")[[1]])]
 
@@
 ## Load plot general information
-data_plot <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")
+data_plot <- read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")
 
@@
   destfile <- "data/CHELSA_bio12_1981-2010_V.2.1.tif"
-  utils::download.file(url, destfile, method = "curl")
+  download.file(url, destfile, method = "curl")
 
@@
   rst <- terra::crop(rst, terra::ext(box))
-  terra::writeRaster(rst, filename = destfile, overwrite = TRUE) 
+  writeRaster(rst, filename = destfile, overwrite = TRUE) 
 }
@@
   destfile <- "data/srtm_62_14.zip"
-  utils::download.file(url, destfile, method = "curl")
-  utils::unzip(destfile, exdir = "data", overwrite = TRUE)
+  download.file(url, destfile, method = "curl")
+  unzip(destfile, exdir = "data", overwrite = TRUE)
 }
@@
   destfile <- "data/CWD.tif.zip"
-  utils::download.file(url, destfile, method = "curl")
-  utils::unzip(destfile, exdir = "data", overwrite = TRUE)
+  download.file(url, destfile, method = "curl")
+  unzip(destfile, exdir = "data", overwrite = TRUE)
 }
@@
 
-# TODO add code to download prec and elev
+# get precipitation data for the two sites
 site_env[, prec := terra::extract(
@@
 
+# get elevation data for the two sites
 site_env[, elev := terra::extract(
@@
 Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso Luga", elev]` m (Figure\ \@ref(fig:map-sites)).
-The average soil properties differed between sites (Table S\@ref(tab:site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
+The average soil properties differed between sites (Table\ S\@ref(tab:site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
 The two sites also have had different histories.
@@
 
-The AFS are presented in more detail in the supplemental information (Table S\@ref(tab:afs-typology)) and in @Cogne2024.
+The AFS are presented in more detail in the supplemental information (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
 
@@
 
-Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using iNaturalist and Pl\@ntnet application, Geoffrey Hull’s monograph [@Hull2006], the data bases of “the useful tropical plants” (https://tropical.theferns.info/) as well as the native plants websites of the NT herbarium from the Australian government (https://nt.gov.au/environment/native-plants/native-plants-and-nt-herbarium).
+Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using iNaturalist and Pl\@ntnet application, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory government's Native Plants Herbarium [@NT2013].
 The correspondence between tetum names and scientific names is provided in the supplemental material (Table S\@ref(tab:tetum-names)).
@@
 
-## Biomass estimation
+## Carbon stocks estimation
+
+### Biomass estimation
 
@@
 # Open large tree data (> 100 cm circumference)
-data_large <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV >100CM")
+data_large <- read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV >100CM")
 
 # Open small tree data (30-100 cm circumference)
-data_small <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV 30-100CM")
+data_small <- read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV 30-100CM")
 
@@
 # convert data_tot into data.table 
-data.table::setDT(data_tot)
+setDT(data_tot)
 
@@
 # split scientific names into genus and species
-names <- data.table::tstrsplit(data_tot$SCIENTIFIC_NAME, " ")
+names <- tstrsplit(data_tot$SCIENTIFIC_NAME, " ")
 
@@
 # get wood density
-data_wd <- BIOMASS::getWoodDensity(genus = data_tot$GENUS, species = data_tot$SPECIES)
+data_wd <- getWoodDensity(genus = data_tot$GENUS, species = data_tot$SPECIES)
 
 # estimate mean AGC
-data_tot$mean <- BIOMASS::computeAGB(
+data_tot$mean <- computeAGB(
   D = data_tot$DIAM_130_CM, 
@@
   data_tot,
-  BIOMASS::AGBmonteCarlo(
+  AGBmonteCarlo(
     D = data_tot$DIAM_130_CM, 
@@
 The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density.
-The volume of the palm trunks was estimated as the volume of a cylinder, using the height and diameter measured in situ, with a volumetric shrinkage of 11% [@WD2023].
+The volume of the palm trunks was estimated as the volume of a cylinder, using the height and diameter measured in situ, with a volumetric shrinkage of 11% [@WD2007].
 The height of the palms was measured above the leaves and was therefore multiplied by a conversion factor of 0.9 to obtain the trunk height.
@@
 
-## Soil organic carbon estimation
+### Soil organic carbon estimation
 
@@
 # open soil lab data
-data_soil_lab <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")
+data_soil_lab <- read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")
 
 # open soil field measurements
-data_soil_field <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")
+data_soil_field <- read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")
 
@@
 data_soil <- merge(data_soil_lab, data_soil_field, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
-data.table::setDT(data_soil)
+setDT(data_soil)
 ```
@@
 ```{r soc-error-propagation}
-data_soil[, c("h0", "h1") := lapply(data.table::tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
+data_soil[, c("h0", "h1") := lapply(tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
 
@@
 data_sim = rbind(data_sim, data_sim[, .(value = sum(value[variable != "SOC10"]), variable = "TOT"), .(N_PLOT, site, ID_AF, iter)])
+# group total C stocks in plots 6 and 8 (otherwise will underestimate CF total C
+# stocks); this will not be used for analysis, only to display average C stocks
+# per AFS
+data_sim <- rbind(
+  data_sim[variable != "TOT" | !N_PLOT %in% c(6, 8)], 
+  data_sim[variable == "TOT" & N_PLOT %in% c(6, 8), .(value = sum(value), N_PLOT = 6), .(site, ID_AF, variable, iter)]
+)
 ```
@@
 ## mixed effect linear regression
-lmeReg <- lmerTest::lmer(log(SOC_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
+lmeReg <- lmer(log(SOC_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
 # Nakagawa’s R2 for mixed models
-R2 = performance::r2_nakagawa(lmeReg)
+R2 = r2_nakagawa(lmeReg)
 summary(lmeReg)
@@
 # same for topsoil layer (0-10 cm)
-lmeReg10 <- lmerTest::lmer(log(SOC10_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
+lmeReg10 <- lmer(log(SOC10_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
 # Nakagawa’s R2 for mixed models
-R210 = performance::r2_nakagawa(lmeReg10)
+R210 = r2_nakagawa(lmeReg10)
 summary(lmeReg10)
@@
 
-Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ MgC\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ MgC\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ MgC\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ MgC\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 
@@
 
-AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ MgC\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ MgC\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)), but not between site (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
-The AFS with the lowest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+The AFS with the lowest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ MgC\ ha$^{-1}$; the AFS with the highest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ MgC\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 
@@
 
-SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ MgC\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ MgC\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)).
 However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
-The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$) than those in Osso Luga (`r dataC_site[site=="Osso Luga" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
+The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ MgC\ ha$^{-1}$) than those in Osso Luga (`r dataC_site[site=="Osso Luga" & variable=="SOC", round(mean)]`\ MgC\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
 
@@
 The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure\ \@ref(fig:plot-lmereg)).
-This relationship remained insignificant (yet slightly stronger; p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
+This relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
 The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure\ \@ref(fig:plot-lmereg)).
@@
 Compared to values reported in the literature for other AFS, the carbon stocks in this study were particularly high.
-The value of SOC stocks was similar to other reported values [@Feliciano2018], but some of the AFS studied here have particularly high stocks in their biomass [compared to the average of 74\ Mg\ C\ ha$^{-1}$ in @Ma2020].
-The AFS with the highest carbon stocks, forest gardens, had stocks close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$; @Lal2005].
+The value of SOC stocks was similar to other reported values [@Feliciano2018], but some of the AFS studied here have particularly high stocks in their biomass [compared to the average of 74\ MgC\ ha$^{-1}$ in @Ma2020].
+The AFS with the highest carbon stocks, forest gardens, had stocks close to those of old-growth tropical forests [270\ MgC\ ha$^{-1}$; @Lal2005].
 These high carbon stocks could be explained by the fact that forest gardens have generally grown for more than 50 years, with little human intervention (Table S1).
@@
 The same applies to FGs, which are former HGs that have been more or less abandoned, and YAs, where the farmer has chosen the site with the intention of turning it into a HG.
-We therefore expect HGs, FGs and YAs to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuter2018].
+We therefore expect HGs, FGs and YAs to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuer2018].
 Conversely, CF and SP are mainly located on limestone plateaus where water is deep and soils are shallow, which could reduce tree growth and biomass [@Imada2008].
@@
 Our results revealed that SOC was mainly sensitive to the differences in location and villages rather than AFS systems in each location.
-This was also reported by the meta-analysis of Feliciano et al.
-(2018) who highlighted that only 3.2% of the variance of the soil carbon sequestration was explained by agroforestry system choices.
-The two villages studied were different in terms of inherent soil properties (Table S3; Fig S2), altitudes (Fig.
-1; Fig.
-S3) and also history.
+This was also reported by the meta-analysis of @Feliciano2018 who highlighted that only 3.2% of the variance of the soil carbon sequestration was explained by agroforestry system choices.
+The two villages studied were different in terms of inherent soil properties (Table S\@ref(tab:site-soil); Figure S\@ref(fig:clay-corg)), altitudes (Figure \@ref(fig:map-sites); Figure S\@ref(fig:altitude-corg)) and also history.
 All those driving factors independent of the AFS management could not be disentangled in this study but their cumulative effects were clearly observed.
-The difference in terms of soil texture and especially clay content was reported to be a key driving process in terms of SOC storage in the tropics (Don et al., 2010; Matus, 2021).
-In this study, the fine fractions of soil texture were however not linked to the SOC stocks (Fig S2).
+The difference in terms of soil texture and especially clay content was reported to be a key driving process in terms of SOC storage in the tropics [@Don2010;@Matus2021].
+In this study, the fine fractions of soil texture were however not linked to the SOC stocks (Figure S\@ref(fig:clay-corg)).
 In addition to soil inherent conditions, other key driving factors for SOC changes, independent to the AFS management practices, may have led to those differences between.
 Gariuai was inhabited whereas Osso-Luga was not for a long time.
-The management and care of the plot in Gariuai such as pruning, livestock visits, optimizing the different stratas in terms of C-input could have led to increased soil organic carbon compared to abandoned systems (Lorenz & Lal, 2014).
+The management and care of the plot in Gariuai such as pruning, livestock visits, optimizing the different stratas in terms of carbon input could have led to increased soil organic carbon compared to abandoned systems [@Lorenz2014].
 
@@
 This observation was in contradiction with our H1b.
-Our systems were rather contrasted and previous studies reported significant differences in SOC sequestration between different AFS (Ma et al., 2020, Kimaro et al., 2024).
-Local conditions and the rapid SOC mineralisation in tropical areas may explain this lack of sensitivity against AFs management practices (Sayer et al., 2019).
+Our systems were rather contrasted and previous studies reported significant differences in SOC sequestration between different AFS [@Ma2020;@Kimaro2024].
+Local conditions and the rapid SOC mineralisation in tropical areas may explain this lack of sensitivity against AFs management practices [@Sayer2019].
 Our experimental design with a limited number of replicates within each village could have also limited the significance of the tests.
@@
 
-Although agroforestry can make agriculture more resilient while improving its carbon footprint and biodiversity conservation value (Cardinael et al., 2021), poor adaptation to the local context can explain many failures to adopt new agroforestry techniques (Coe et al., 2014).
-Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it (Gouyon et al., 1993, Aumeeruddy-Thomas, 1994; Peltier, 1996).
-This is the case of the traditional AFS studied here: their heritage and socio-economic values (as shown by Cogné & Lescuyer, 2024), in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally..
+Although agroforestry can make agriculture more resilient while improving its carbon footprint and biodiversity conservation value [@Cardinael2021], poor adaptation to the local context can explain many failures to adopt new agroforestry techniques [@Coe2014].
+Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
+This is the case of the traditional AFS studied here: their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally..
 
@@
 Unfortunately, this results in a significant loss of biodiversity and carbon stored in the biomass, and can lead to an irreversible loss of fertility on the plot through soil degradation and erosion.
-This phenomenon has been documented in a similar context in Sumatra (Indonesia), where Gouyon et al.
-(1993) highlighted the fact that old agroforests, which they called "jungle-rubber" and which are quite similar in structure to "forest gardens" (except for the rubber tree component, which has been replaced by other species in Timor), were being cleared in favor of more intensive production systems.
-This trend towards clearing jungle rubber was subsequently confirmed by Ekadinata & Vincent.
-(2011), and its impact on carbon stock decline was quantified by Villamor et al.
-(2014).
+This phenomenon has been documented in a similar context in Sumatra (Indonesia), where @Gouyon1993 highlighted the fact that old agroforests, which they called "jungle-rubber" and which are quite similar in structure to "forest gardens" (except for the rubber tree component, which has been replaced by other species in Timor), were being cleared in favor of more intensive production systems.
+This trend towards clearing jungle rubber was subsequently confirmed by @Ekadinata2011, and its impact on carbon stock decline was quantified by @Villamor2014.
 
@@
 payment of carbon credits) or low-interest loans to farmers wishing to renew their agroforestry plots.
-In addition, a "sustainable agriculture" type label could be awarded to these AFS products, and their value chains could be supported by governments and international certification bodies, as is already the case for coffee in many countries (Bertrand et al., 2019).
-
-## Conclusion
-
-Agroforestry systems in the Baucau district of Timor had variable but overall high carbon stocks.
+In addition, a "sustainable agriculture" type label could be awarded to these AFS products, and their value chains could be supported by governments and international certification bodies, as is already the case for coffee in many countries [@Bertrand2019].
+
+# Conclusion
+
+Agroforestry systems in the Baucau district of Timor-Leste had variable but overall high carbon stocks.
 These systems have only recently been described in the scientific literature, and much remains to be done to improve our knowledge of their diversity and functioning.
@@
 
-```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (between parentheses) of total, aboveground, belowground, and soil organic carbon stocks in each agroforestry system."}
+```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (between parentheses) of total, aboveground carbon (AGC), belowground carbon (BGC), and soil organic carbon (SOC) stocks in each agroforestry system."}
 # calculate median and 95% confidence interval at site level
-table_afs = dataC_afs[, value := paste0(
+dataC_afs[, value := paste0(
   signif(mean, 3), " (",
@@
   subset(variable != "SOC10") |> 
-  dcast(afs ~ variable, value.var = "value")
-
- knitr::kable(table_afs[, c("afs", "TOT", "AGC", "BGC", "SOC")],
-    col.names = c("Agroforestry system", 
-                  "Total carbon (MgC/ha)", 
-                  "Aboveground carbon (MgC/ha)", 
-                  "Belowground carbon (MgC/ha)", 
-                  "Soil organic carbon (MgC/ha)")) 
+  dcast(afs ~ variable, value.var = "value") |> 
+  flextable(col_keys = c("afs", "TOT", "AGC", "BGC", "SOC"), 
+            cwidth = c(0.9, 1.2, 1.3, 1.3, 1.2)) |> 
+  compose(i = 1, j = "afs", part = "header", value = as_paragraph("Agroforestry system")) |>
+  compose(i = 1, j = "TOT", part = "header", value = as_paragraph( "Total C stocks\n(MgC ha", as_sup("-1"), ")"))|> 
+  compose(i = 1, j = "AGC", part = "header", value = as_paragraph( "AGC stocks\n(MgC ha", as_sup("-1"), ")"))|> 
+  compose(i = 1, j = "BGC", part = "header", value = as_paragraph( "BGC stocks\n(MgC ha", as_sup("-1"), ")"))|> 
+  compose(i = 1, j = "SOC", part = "header", value = as_paragraph( "SOC stocks\n(MgC ha", as_sup("-1"), ")"))
 ```
@@
 Bold p-values are lower than 0.05, and the mean difference between groups is thus considered significantly different than zero."}
-knitr::kable(anova_table[, c("variable", "factor", "pvalue")], 
-             col.names = c("Dependent variable", "Independent variable", "p-value"))
+anova_table$factor = factor(anova_table$factor)
+levels(anova_table$factor) = c("AFS", "site")
+flextable(anova_table, col_keys = c("variable", "factor", "pvalue"), 
+          cwidth = c(1, 1, 0.8)) |> 
+  bold(i = anova_table$pvalue < 0.05, j = 2:3) |> 
+  set_header_labels(values = c("Dependent variable", "Independent variable", "p-value")) |> 
+  merge_at(i = 1:2, j = 1) |> merge_at(i = 3:4, j = 1) |> 
+  fix_border_issues() |> hline(i=2)
 ```
@@
   geom_tile(data = data_box, aes(x = xc, y = yc, width = w, height = w), fill = NA, color = 1, linewidth = 0.5) +
-  ggrepel::geom_text_repel(data = data_box, aes(x = xc, y = yc, label = site)) + 
+  geom_text_repel(data = data_box, aes(x = xc, y = yc, label = site)) + 
   labs(x = "", y = "") +
@@
              aes(x = long, y = lat), shape = 1, size = 2) + 
-  # coord_equal() +
-  # viridis::scale_fill_viridis(option = "D",
-  #                             limits = range(c(elevS$srtm_62_14, elevG$srtm_62_14))
-  # ) +
   scale_fill_gradientn(colours = terrain.colors(10)) +
-  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
+  scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
   labs(x = NULL, y = NULL, color = NULL, fill="Elevation (m)") +
@@
         axis.title = element_blank(),
-        # axis.ticks = element_blank(), axis.text = element_blank(), 
         plot.title = element_text(hjust = 0.5), 
@@
   scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
-  ggspatial::annotation_scale() +
+  annotation_scale() +
   coord_sf(crs = 'WGS84')
@@
   guides(color=guide_legend(ncol=2)) +
-  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
+  scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
   theme_classic()
-l1 = ggpubr::get_legend(g3)
-
-ggpubr::ggarrange(ggpubr::ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))
+l1 = get_legend(g3)
+
+ggarrange(ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))
 ```
@@
   theme(strip.placement = "outside") +
-  labs(x = "", y = "Carbon stocks (Mg C/ha)", fill = "Carbon\npool") +
+  labs(x = "", y = expression("Carbon stocks (MgC"*" "*ha^{-1}*")"), fill = "Carbon\npool") +
   scale_y_continuous(expand = c(0, 0))
@@
 
+## reorder factors 
+dataC_fig[, afs := factor(afs, levels = c("CF", "SP", "YA", "HG", "FG"))]
+
 ## figure
@@
   geom_line(data = dfpred, aes(x = AGC_mean, y = SOC_mean), lty = 2, col = 1) +
-  labs(color = "AFS", x = "AGC stocks (Mg C/ha)", y = "SOC stocks (Mg C/ha)", shape = "Site") +
+  labs(color = "AFS", x = expression("AGC stocks (MgC"*" "*ha^{-1}*")"), 
+       y = expression("SOC stocks (MgC"*" "*ha^{-1}*")"), shape = "Site") +
   expand_limits(x = 0, y = 0) +
-  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
+  scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
   scale_x_log10() +
@@
 
-# Supplementary material
+# Supplementary information
 
@@
 
-```{r tab.id = "afs-typology", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
-typ = readxl::read_excel("tables/afs-typology.xlsx")
-kable(typ)
-```
-
-```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show"}
+```{r site-soil-char}
 data_soil$silt <- as.numeric(data_soil$`FINE SILT`) + as.numeric(data_soil$`COARSE SILT`)
@@
 # add site info to data_soil 
-data_soil <- merge(data_soil, data_plot[, c("N_PLOT", "site")])
+data_soil_plot <- merge(data_soil, data_plot[, c("N_PLOT", "site")])
 
@@
 
-data_soil_plot = melt(data_soil, 
+data_soil_plot = melt(data_soil_plot, 
                       id.vars = c("N_PLOT", "SAMPLE_DEPTH", "site"), 
@@
 data_soil_site = dcast(data_soil_site, site ~ variable, value.var = "value")
-data_soil_site = data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "Ca", "K", "Mg", "Na", "CEC")]
-
-kable(data_soil_site, col.names = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)","N", "C/N", "pH", "Ca", "K", "Mg", "Na", "CEC"))
-```
-
-```{r tab.id = "tetum-names", tab.lp="supp-tab", results = "show"}
-sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
-sheets2 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)
+```
+
+```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (average over depths 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
+The confidence intervals were calculated by propagating a 12% error on the lab-provided values and obtaining site-level estimates through bootstrapping plot values (1000 simulations)."}
+flextable(
+  data_soil_site,
+  col_keys = c("site", "CLAY", "silt", "sand", "Corg",
+               "N", "Corg_N", "pH", "CEC", "Ca", "K", "Mg", "Na"), 
+  cwidth = 0.5
+) |> set_header_labels(
+  values = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)", 
+             "N", "C/N", "pH", "CEC", "Ca", "K", "Mg", "Na"))|> 
+  compose(i = 1, j = "N", part = "header", value = as_paragraph("N (%", as_sub("o"), ")")) |>
+  compose(i = 1, j = "CEC", part = "header", value = as_paragraph( "CEC (cmolc kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "Ca", part = "header", value = as_paragraph( "Ca (cmolc kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "K", part = "header", value = as_paragraph( "K (cmolc kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "Mg", part = "header", value = as_paragraph( "Mg (cmolc kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "Na", part = "header", value = as_paragraph( "Na (cmolc kg", as_sup("-1"), ")"))
+```
+
+```{r tab.id = "afs-typology", tab.cap.pre="Table S", tab.cap = "Description of 5 agroforestry systems in Central Timor [adaptation of the name given by Nair, name in the Tetum language, position in the landscape, surface area, age since creation, previous cultivation, main crops, associated livestock, planted and spontaneous tree species.", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
+read_excel("tables/afs-typology.xlsx") |> 
+  flextable(col_keys = c(
+    "Typology adapted from Nair (1993)",
+    "Usual location", 
+    "Typical plot size (ha)", 
+    "Time since installation", 
+    "Previous land use"), 
+    cwidth = c(1.5, 1.5, 1, 1, 1.5)) 
+```
+
+```{r tetum-scientific-correspondance}
+sheets1 = grep("INV", excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
+sheets2 = grep("INV", excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)
 
 vern1 <- lapply(sheets1, \(x){
-  df = readxl::read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
+  df = read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
   df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
@@
 vern2 <- lapply(sheets2, \(x){
-  df = readxl::read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
+  df = read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
   df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
@@
 vern[species == "purpuroides", species := NA]
-corr = BIOMASS::correctTaxo(genus = vern$genus, species = vern$species)
+corr = correctTaxo(genus = vern$genus, species = vern$species)
 vern = cbind(vern, corr)
@@
 vern = vern[, .(latin = paste(latin, collapse = ", ")), .(tetum)]
-kable(vern, col.names = c("Tetum name", "Scientific name"))
+```
+
+```{r tab.id = "tetum-names", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Correspondence between tetum names and scientific names used in this study."}
+flextable(vern) |> 
+  set_header_labels(values =  c("Tetum name", "Scientific name")) |> 
+  autofit()
 ```
@@
 (b) Typical 1-ha plot in SP systems.
-Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
-knitr::include_graphics("figures/plot-design.jpg") 
+Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 5, fig.height = 3}
+include_graphics("figures/plot-design.jpg") 
 ```
@@
 
-lmerTest::lmer(
+lmer(
   log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
@@
 
-```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in MgC/ha).
-Points are coloured by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.", fig.cap.pre="Figure S"}
+```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in MgC ha^-1^).
+Points are colored by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.", fig.cap.pre="Figure S"}
 dataCelev = merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")], 
@@
 
-lmerTest::lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |> 
+lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |> 
   summary()
@@
   theme_classic() +
-  labs(x = "Elevation (m)", y = "SOC stocks (Mg C/ha)", col = "Site")
+  labs(x = "Elevation (m)", y = expression("SOC stocks (MgC"*" "*ha^{-1}*")"), col = "Site")
 ```

### 2024-03-27 manuscript.Rmd
update with Marguerite's info on methodology

@@
   officedown::rdocx_document:
-    number_sections: true
+    reference_docx: word-styles-ref-02.docx
 bibliography: references.bib
@@
 ```{r setup, include=FALSE}
+# Run with R version 4.3.2
+
 packages_needed = c(
@@
   "ggspatial",    # version 1.1.9
+  "maps",         # version 3.4.2
   "ggpubr",       # version 0.6.0
@@
   "lmerTest",     # version 3.1-3
+  "lme4",         # version 1.1-35.1
+  "Matrix",       # version 1.6
   "performance",  # version 0.10.8
@@
 In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastoral systems, young agroforests, home gardens and forest gardens.
-Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 210\ MgC\ ha$^{-1}$, similar to an old-growth tropical forest.
+Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, similar to an old-growth tropical forest.
 Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
@@
 ## Load plot general information
-data_plot <- read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")
+data_plot <- read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO") |> 
+  subset(N_PLOT != 8)  # remove plot 8: no soil data
 
@@
   ## crop to reduce memory usage
-  box = c(124.5, 127.5, -9.2, -8.3) # Timor Leste
+  box = c(124.5, 127.5, -9.2, -8.3) # Timor-Leste
   rst <- terra::rast(destfile)
@@
 The two sites also have had different histories.
-Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2000.
+Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2003.
 During the period of abandonment, the various agroforestry systems in Osso Luga lay fallow.
@@
 This typology was based on the one proposed by @Nair1993;@Nair2021, and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
-The five AFS identified were: crop system including a wooded fallow phase (CF), Sylvopastoral (SP), Young agroforest (YA), Home garden (HG) and Forest garden (FG).
-CF are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
+The five AFS identified were: crop system including a wooded fallow phase (CF), Silvopastoral (SP), Young agroforest (YA), Home garden (HG) and Forest garden (FG).
+CF systems are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
 SP land is a vast area (generally 200 to 500 ha) where cattle are grazed under a layer of trees.
-YA are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
-HG are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
+YAs are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
+HGs are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
 FGs are generally former HGs (over 50 years old), which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
@@
 
-Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using iNaturalist and Pl\@ntnet application, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory government's Native Plants Herbarium [@NT2013].
+Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory government's Native Plants Herbarium [@NT2013].
 The correspondence between tetum names and scientific names is provided in the supplemental material (Table S\@ref(tab:tetum-names)).
@@
 # stack tree tables
-data_tot <- rbind(data_large[, col_keep], data_small[, col_keep])
+data_tree <- rbind(data_large[, col_keep], data_small[, col_keep])
 
 # merge with plot information
-data_tot <- merge(data_tot, data_plot, by = "N_PLOT")
-
-# convert data_tot into data.table 
-setDT(data_tot)
+data_tree <- merge(data_tree, data_plot, by = "N_PLOT")
+
+# convert data_tree into data.table 
+setDT(data_tree)
 
@@
 # remove NA or <30 cm circumference values or < 100 cm outside subplots
-data_tot <- subset(data_tot, 
+data_tree <- subset(data_tree, 
                    !is.na(DIAM_130_CM) & 
@@
 # split scientific names into genus and species
-names <- tstrsplit(data_tot$SCIENTIFIC_NAME, " ")
+names <- tstrsplit(data_tree$SCIENTIFIC_NAME, " ")
 
 # get genera
-data_tot$GENUS <- names[[1]]
+data_tree$GENUS <- names[[1]]
 
 # get species
-data_tot$SPECIES <- names[[2]]
+data_tree$SPECIES <- names[[2]]
 
 # Timoneus should be Timonius
-data_tot$GENUS[!is.na(data_tot$GENUS) & data_tot$GENUS == "Timoneus"] <- "Timonius"
+data_tree$GENUS[!is.na(data_tree$GENUS) & data_tree$GENUS == "Timoneus"] <- "Timonius"
 
 # remove one tree with DBH = 0
-data_tot <- subset(data_tot, DIAM_130_CM > 0)
+data_tree <- subset(data_tree, DIAM_130_CM > 0)
 
@@
 # get wood density
-data_wd <- getWoodDensity(genus = data_tot$GENUS, species = data_tot$SPECIES)
+data_wd <- getWoodDensity(genus = data_tree$GENUS, species = data_tree$SPECIES)
 
 # estimate mean AGC
-data_tot$mean <- computeAGB(
-  D = data_tot$DIAM_130_CM, 
+data_tree$mean <- computeAGB(
+  D = data_tree$DIAM_130_CM, 
   WD = data_wd$meanWD, 
-  H = data_tot$HEIGHT
+  H = data_tree$HEIGHT
 ) * 0.4713 # conversion factor from biomass to carbon
@@
 data_sim_agc <- data.table(
-  data_tot,
+  data_tree,
   AGBmonteCarlo(
-    D = data_tot$DIAM_130_CM, 
+    D = data_tree$DIAM_130_CM, 
     WD = data_wd$meanWD, 
     errWD = data_wd$sdWD, 
-    H = data_tot$HEIGHT, 
-    errH = data_tot$HEIGHT*cvH, 
+    H = data_tree$HEIGHT, 
+    errH = data_tree$HEIGHT*cvH, 
     Dpropag = "chave2004", 
@@
 data_sim[, value := value + valueSmall]
+
+## add zeros for plots with no tree above 10 cm DBH
+# deal with non detectable values: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9216349/
+# use uniform distribution between zero and limit of detectability
+data_sim_empty = expand.grid(N_PLOT = setdiff(data_plot$N_PLOT, data_sim$N_PLOT), 
+                             iter = unique(data_sim$iter), 
+                             variable = c("AGC", "BGC"))|> 
+  setDT() |> merge(data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")]) 
+data_sim_empty$value = 0
+
+# limit of detectability: tree of 30 cm CBH, mean WD and height (for this DBH)
+# LOD = computeAGB(D = 30/pi, 
+#                  WD = unique(data_wd$meanWD[data_wd$genus=="NR"]), 
+#                  H = data_tree[DIAM_130_CM == 30/pi, mean(HEIGHT)])
+# data_sim_empty[iter != "mean", AGC := runif(sum(iter != "mean"), 0, LOD)]
+# data_sim_empty[iter == "mean", AGC := LOD/2]
+# RS = mean(data_sim_agc$RS[data_sim_agc$DIAM_130_CM==30/pi])
+# data_sim_empty[, BGC := AGC * RS]
+
+# data_sim_empty = melt(data_sim_empty, measure.vars = c("AGC", "BGC"))
+
+data_sim = rbind(data_sim[, colnames(data_sim_empty), with = FALSE], data_sim_empty)
 ```
@@
 
-Soil organic carbon stocks (in Mg.ha$^{-1}$) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:
+Soil organic carbon stocks (in Mg\ C\ ha$^{-1}$) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:
 
@@
 # per AFS
-data_sim <- rbind(
-  data_sim[variable != "TOT" | !N_PLOT %in% c(6, 8)], 
-  data_sim[variable == "TOT" & N_PLOT %in% c(6, 8), .(value = sum(value), N_PLOT = 6), .(site, ID_AF, variable, iter)]
-)
+# data_sim <- rbind(
+#   data_sim[variable != "TOT" | !N_PLOT %in% c(6, 8)], 
+#   data_sim[variable == "TOT" & N_PLOT %in% c(6, 8), .(value = sum(value), N_PLOT = 6), .(site, ID_AF, variable, iter)]
+# )
 ```
@@
 The estimated AGC and SOC stocks were log-transformed for all statistical tests, because we expect them to be log-normally distributed as they only take positive values.
+We removed one CF plot at Gariuai from the following statistical tests, as it did not contain any trees exceeding the minimum measurement size (CBH = 30 cm) and its AGC stocks were therefore estimated at 0\ Mg\ C\ ha$^{-1}$ (even if there were trees < 30 cm CBH) and could not be log-transformed.
 After testing for normality (Shapiro-Wilks test) and homoscedasticity (Levene’s test), we performed a two-way ANOVA on the log-transformed estimates of AGC and SOC stocks, grouped by AFS and site.
@@
 # shapiro wilk test - normality
-dataC_plot[variable %in% c("AGC", "SOC"), .(
+dataC_plot[variable %in% c("AGC", "SOC") & plot != 6, .(
   shap = shapiro.test(log(mean))$p.value
@@
 # levene test of homoscedasticity
-dataC_plot[variable %in% c("AGC", "SOC"), 
+dataC_plot[variable %in% c("AGC", "SOC") & plot != 6, 
            leveneTest(log(mean) ~ afs*site)$`Pr(>F)`[1], .(variable)]
@@
 anova_table = 
-  dataC_plot[variable %in% c("AGC", "SOC"), 
+  dataC_plot[variable %in% c("AGC", "SOC") & plot != 6, 
              .(pvalue = round(anova(lm(log(mean) ~ site + afs))$`Pr(>F)`[1:2], 3)), 
@@
 ```{r lmereg}
-data_reg = melt(dataC_plot, measure.vars = c("mean", "lwr", "upr"), variable.name = "stat") |> 
-  dcast(plot+ site+ afs ~ variable + stat)
+data_reg = melt(dataC_plot[plot != 6], measure.vars = c("mean", "lwr", "upr"), variable.name = "stat") |> 
+  dcast(plot + site+ afs ~ variable + stat)
 
@@
 
-Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ MgC\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ MgC\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ MgC\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ MgC\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 
@@
 
-AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ MgC\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ MgC\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)), but not between site (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
-The AFS with the lowest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ MgC\ ha$^{-1}$; the AFS with the highest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ MgC\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+The AFS with the lowest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 
@@
 
-SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ MgC\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ MgC\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)).
 However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
-The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ MgC\ ha$^{-1}$) than those in Osso Luga (`r dataC_site[site=="Osso Luga" & variable=="SOC", round(mean)]`\ MgC\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
+The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$) than those in Osso Luga (`r dataC_site[site=="Osso Luga" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
 
@@
 
-## Large C stocks in traditional AFS in Timor Leste
+## Large C stocks in traditional AFS in Timor-Leste
 
 Compared to values reported in the literature for other AFS, the carbon stocks in this study were particularly high.
-The value of SOC stocks was similar to other reported values [@Feliciano2018], but some of the AFS studied here have particularly high stocks in their biomass [compared to the average of 74\ MgC\ ha$^{-1}$ in @Ma2020].
-The AFS with the highest carbon stocks, forest gardens, had stocks close to those of old-growth tropical forests [270\ MgC\ ha$^{-1}$; @Lal2005].
+The value of SOC stocks was similar to other reported values [@Feliciano2018], but some of the AFS studied here have particularly high stocks in their biomass [compared to the average of 74\ Mg\ C\ ha$^{-1}$ in @Ma2020].
+The AFS with the highest carbon stocks, forest gardens, had stocks close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$; @Lal2005].
 These high carbon stocks could be explained by the fact that forest gardens have generally grown for more than 50 years, with little human intervention (Table S1).
@@
 We therefore expect HGs, FGs and YAs to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuer2018].
-Conversely, CF and SP are mainly located on limestone plateaus where water is deep and soils are shallow, which could reduce tree growth and biomass [@Imada2008].
+Conversely, CF and SP fields are mainly located on limestone plateaus where water is deep and soils are shallow, which could reduce tree growth and biomass [@Imada2008].
 
@@
 
-## How to conserve and improve carbon stocks in traditional AFS in Timor Leste?
+## How to conserve and improve carbon stocks in traditional AFS in Timor-Leste?
 
@@
 Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
-This is the case of the traditional AFS studied here: their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally..
+This is the case of the traditional AFS studied here: their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally.
 
@@
 It is therefore important that all stakeholders including local authorities, universities and NGOs, take into account their existence, diversity and environmental and social value, in order to propose possible innovations to maintain or improve their functioning.
+
+# Authors' contribution
+
+All authors contributed to the design of the sampling protocol.
+MC carried out the fieldwork with the help of RP, AT and VF [mention students?].
+CP analyzed the data with the help of VF and MC.
+CP prepared the manuscript and all authors contributed to the drafting and revision of the manuscript.
+
+# Acknowledgements
+
+# Data availability
 
@@
   compose(i = 1, j = "afs", part = "header", value = as_paragraph("Agroforestry system")) |>
-  compose(i = 1, j = "TOT", part = "header", value = as_paragraph( "Total C stocks\n(MgC ha", as_sup("-1"), ")"))|> 
-  compose(i = 1, j = "AGC", part = "header", value = as_paragraph( "AGC stocks\n(MgC ha", as_sup("-1"), ")"))|> 
-  compose(i = 1, j = "BGC", part = "header", value = as_paragraph( "BGC stocks\n(MgC ha", as_sup("-1"), ")"))|> 
-  compose(i = 1, j = "SOC", part = "header", value = as_paragraph( "SOC stocks\n(MgC ha", as_sup("-1"), ")"))
+  compose(i = 1, j = "TOT", part = "header", value = as_paragraph( "Total C stocks\n(Mg C ha", as_sup("-1"), ")"))|> 
+  compose(i = 1, j = "AGC", part = "header", value = as_paragraph( "AGC stocks\n(Mg C ha", as_sup("-1"), ")"))|> 
+  compose(i = 1, j = "BGC", part = "header", value = as_paragraph( "BGC stocks\n(Mg C ha", as_sup("-1"), ")"))|> 
+  compose(i = 1, j = "SOC", part = "header", value = as_paragraph( "SOC stocks\n(Mg C ha", as_sup("-1"), ")"))
 ```
@@
 (a) Map of Timor-Leste with the location of the two study sites (Gariuai and Osso-Luga) in the Bacau district.
-(b) Location of the 16 plots in Gariuai; (c) Location of the 15 plots in Osso-Luga.
+(b) Location of the 15 plots in Gariuai; (c) Location of the 15 plots in Osso-Luga.
 Panels (b) and (c) share the same spatial scale.
@@
   theme(strip.placement = "outside") +
-  labs(x = "", y = expression("Carbon stocks (MgC"*" "*ha^{-1}*")"), fill = "Carbon\npool") +
+  labs(x = "", y = expression("Carbon stocks (Mg C"*" "*ha^{-1}*")"), fill = "Carbon\npool") +
   scale_y_continuous(expand = c(0, 0))
@@
 For each linear mixed model, we report the marginal R2 (proportion of variance explained by fixed effects only) and the conditional R2 (proportion of variance explained by both fixed and random effects).
-Both x- and y-axes are log-transformed."}
+Both x- and y-axes are log-transformed.
+One plot (AFS: CF, site: Gariuai) with no trees over 30 cm in circumference at breast height (minimum size for measurement) had AGC stocks estimated at 0 Mg C ha^-1^ and was therefore removed from the linear model and the figure."}
 dfpred = expand.grid(AGC_mean = exp(seq(min(log(data_reg$AGC_mean), na.rm = TRUE),
@@
   geom_line(data = dfpred, aes(x = AGC_mean, y = SOC_mean), lty = 2, col = 1) +
-  labs(color = "AFS", x = expression("AGC stocks (MgC"*" "*ha^{-1}*")"), 
-       y = expression("SOC stocks (MgC"*" "*ha^{-1}*")"), shape = "Site") +
+  labs(color = "AFS", x = expression("AGC stocks (Mg C"*" "*ha^{-1}*")"), 
+       y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), shape = "Site") +
   expand_limits(x = 0, y = 0) +
@@
 
-```{r tab.id = "afs-typology", tab.cap.pre="Table S", tab.cap = "Description of 5 agroforestry systems in Central Timor [adaptation of the name given by Nair, name in the Tetum language, position in the landscape, surface area, age since creation, previous cultivation, main crops, associated livestock, planted and spontaneous tree species.", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
+```{r tab.id = "afs-typology", tab.cap.pre="Table S", tab.cap = "Description of 5 agroforestry systems in Central Timor-Leste [adaptation of the name given by Nair, position in the landscape, surface area, age since creation, previous cultivation", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
 read_excel("tables/afs-typology.xlsx") |> 
@@
 
-```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in MgC ha^-1^).
+```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in\ Mg\ C\ ha^-1^).
 Points are colored by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.", fig.cap.pre="Figure S"}
@@
   theme_classic() +
-  labs(x = "Elevation (m)", y = expression("SOC stocks (MgC"*" "*ha^{-1}*")"), col = "Site")
+  labs(x = "Elevation (m)", y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), col = "Site")
 ```

### 2024-04-10 manuscript.Rmd
add diagram of AFS sequence

@@
 Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their potential inclusion in climate change mitigation strategies.
-In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastoral systems, young agroforests, home gardens and forest gardens.
+In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
 Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, similar to an old-growth tropical forest.
@@
 
-Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO$_2$e/yr over 2010-2014; @Pendrill2019].
+Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO~2~e/yr over 2010-2014; @Pendrill2019].
 The demand for tropical agricultural land is expected to continue to increase as the population grows [@UN2022].
@@
 Although AFS appear to increase SOC stocks overall, some AFS such as woodlots (i.e.
-planting trees with intercropping during the establishment phase) may decrease SOC compared to non-AFS, while AFS such as silvopastoral systems have been shown to strongly increase SOC stocks [@Feliciano2018].
+planting trees with intercropping during the establishment phase) may decrease SOC compared to non-AFS, while AFS such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
 Other factors such as climate and previous land use type have also been shown to strongly influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018;@Ma2020;@Martin2020].
 Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS could only be demonstrated for the topsoil layer [< 10 cm; @Ma2020].
-
-<!-- Although AFS can make agriculture more resilient while improving its carbon footprint, poor adaptation to the local context can explain many failures to introduce new agroforestry techniques [@Coe2014].
-Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
-Moreover, traditional small-scale AFS typically retain high levels of biodiversity, making them key elements for biodiversity conservation in the tropics [@Perfecto2008;@Torquebiau1992].
-A better understanding of traditional agroforestry practices and their ecological characteristics could provide better support for the local development of agroforestry.
--->
 
@@
 
+Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural fields and diversifying agricultural landscapes.
+Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their potential inclusion in climate change mitigation strategies.
+In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
+Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 255 MgC/ha, similar to an old-growth tropical forest.
+Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
+We found no relationship between the amount of carbon stored in biomass and soil.
+Our results highlight the high diversity of traditional agroforestry systems in Timor-Leste and their high capacity to store carbon.
+
 # Materials and methods
@@
 ## download data from dataverse - once published
-dataverse_url <- "https://dataverse.cirad.fr/privateurl.xhtml?token=f8eabb2a-3635-499b-aee5-80fd2b7495a0"
-dataverse_doi <- "10.18167/DVN1/QCXWIY"
-
-# for now : download manually from temporary link
-if (!dir.exists("data")) create.dir("data")
-unzip("data/dataverse_files.zip")
+# dataverse_url <- "https://dataverse.cirad.fr/privateurl.xhtml?token=f8eabb2a-3635-499b-aee5-80fd2b7495a0"
+# 
+# # for now : download manually from temporary link
+# if (!dir.exists("data")) create.dir("data")
+# unzip("data/dataverse_files.zip")
+# 
+# 
+# library(dataverse)
+# server<- "dataverse.cirad.fr"
+# doi_dataset <- "doi:10.18167/DVN1/QCXWIY"
+# # key<- "xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx"
+# 
+# Sys.setenv(DATAVERSE_SERVER = server)
+# # Sys.setenv(DATAVERSE_KEY = key)
+# 
+# # get the dataset
+# ds1<-get_dataset(doi_dataset,version = ":latest")
+# #lister les fichiers
+# ds1$files$filename
+# 
+# #recupérer le fichier par le nom
+# data <- get_dataframe_by_name("DATA_BIOMASS_TL_LARGE.xlsx", doi_dataset)
 ```
@@
 Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso Luga", elev]` m (Figure\ \@ref(fig:map-sites)).
-The average soil properties differed between sites (Table\ S\@ref(tab:site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
-The two sites also have had different histories.
+The soils of the Baucau district are classified using the WRB soil classification [@Mantel2023] as Lithosols and Calcic Luvisols [@FAO2012].
+The average soil properties differed between villages (Table\ S\@ref(tab:site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
+The two villages also have had different histories.
 Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2003.
@@
 
-We used the typology described in @Cogne2024, developed in the same study area.
+We used the typology described in@Cogne2024, developed in the same study area.
 This typology was based on the one proposed by @Nair1993;@Nair2021, and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
-The five AFS identified were: crop system including a wooded fallow phase (CF), Silvopastoral (SP), Young agroforest (YA), Home garden (HG) and Forest garden (FG).
-CF systems are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
+The five AFS identified were: Crop system including a wooded Fallow phase (CF), Silvopasture (SP), Young Agroforest (YA), Home Garden (HG) and Forest Garden (FG).
+CF are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
 SP land is a vast area (generally 200 to 500 ha) where cattle are grazed under a layer of trees.
-YAs are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
-HGs are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
+YA are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
+HG are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
 FGs are generally former HGs (over 50 years old), which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
@@
 
-In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (Figure\ S\@ref(fig:design)).
+In each of the 30 sample plots, three composite samples were taken from three of the five 10 x 10 m subplots, following a transect covering the within plot variability and at a minimum distance of 1 m from any tree (Figure\ S\@ref(fig:design)).
 Soils were sampled with an auger at 0-10 cm and 20-30 cm.
 Soils samples were then air-dried and sieved at 2 mm.
-They were the subject of standard physical and chemical analyses at Cirad laboratories in Montpellier, France.
+They were subjected to standard physical and chemical analyses at the Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil-char)).
 Among these analyses, total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
@@
 
-Soil organic carbon stocks (in Mg\ C\ ha$^{-1}$) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:
+Soil organic carbon stocks (in Mg\ C\ ha$^{-1}$) were then calculated as follows:
 
@@
 
-Where $Corg_i$ is the organic carbon content at depth $i$ (either 0-10, 10-20 or 20-30 cm); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth $i$ (in g), converted into Mg with a factor of 10$^{-6}$; Splot is the surface of a 1-ha plot (10$^8$ cm$^2$); $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
+Where $Corg_i$ is the organic carbon content at depth $i$ (either 0-10, 10-20 or 20-30 cm); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth $i$ (in g), converted into Mg with a factor of 10$^{-6}$; $Splot$ is the surface of a 1-ha plot (10$^8$ cm$^2$); $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
+This calculation corresponds to the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil.
 At depth $i$ = 10-20 cm, where the carbon content was not measured (see "Soil measurements" section), we used the average of the carbon content values at 0-10 and 20-30 cm.
@@
 
-## Lack of relationship between soil and biomass carbon [Vincent]
-
-- Low correlation between NPP and SOC in tropical soils [ref: Sayer et al. (2019).  Tropical forest soil carbon stocks do not increase despite 15 years of doubled litter inputs.]
-
-- Functioning of AFS is similar to that of a tropical forest
-
-Some AFS have high SOC for reasons other than the presence of trees; e.g. effect of cattle in SP systems.
-
-## Perspectives: quantify biodiversity in these AFS [Marguerite]
-
-importance of ecological continuity - a range of biodiversity (plant/animal/micro-organisms) that can develop in these traditional AF systems: plots of less than 1ha, density + distribution of trees (intra- and inter-plot), multi-strata.
-
-perspectives: quantify other indicators, such as biodiversity and composition, not only for plants but also for other groups such as animals and micro-organisms; These other organisms are as important as trees in carbon storage [ref?]
+## Lack of relationship between soil and biomass carbon
+
+We found no relationship between AGC and SOC stocks both in the topsoil layer 0-10 cm depth (H2a refuted) and in 0-30 cm depth (H2b confirmed).
+
+These results could be explained by previous findings that SOC stocks saturate much faster than AGC stocks in tropical AFS.
+For example, in a global meta-analysis, @Ma2020 estimated that it took only 5 years for SOC stocks in tropical AFS to reach an equilibrium value, but more than three decades for AGC stocks.
+This desynchronisation of SOC and AGC stocks recovery has also been observed in naturally regenerating tropical forests.
+In a systematic review, @Martin2013 found that SOC in tropical secondary forests changed little with time since disturbance, while AGB increased.
+@Ojoatre2024 found similar results in a tropical montane secondary forest.
+These results imply that there is little relationship between AGC and SOC once agroforestry systems are more than 5 years old, which is the case in our study for FGs, HGs (> 5 years old) and partially for YAs (2 to 10 years; Table S\@ref(tab:afs-typology)).
+This saturation of SOC even with increasing AGC is counterintuitive, as it might be expected that greater AGC would lead to greater litter production and decomposition, some of which would increase SOC.
+However, this was not shown by @Sayer2019, who found that SOC did not increase in a lowland tropical forest in Panama after 15 years of an experimentally doubled aboveground litter input.
+Two processes could explain this.
+On the one hand, the addition of easily decomposable litter could stimulate microbial decomposition and turnover of old stored SOC [i.e.
+priming effect; @Sayer2019].
+On the other hand, the capacity of the soil to store SOC through organo-mineral associations would be limited by its fine element content (clay + fine silt) (i.e.
+saturation concept).
 
@@
 elevAll = rbind(elevS, elevG)
+
+data_plot[, ID_AF := factor(ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))]
 
@@
 
-```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Barplot of the amount of carbon in each AFS (CF: crop and fallow, SP: silvopastoral, YA: young agroforest, HG: home garden, FG: forest garden), per site.
+```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Barplot of the amount of carbon in each AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden), per site.
 AGC: Aboveground (biomass) carbon; BGC: belowground (biomass) carbon; SOC: soil organic carbon."}
@@
 Each point represents the median C stock value in a sample plot, and the vertical and horizontal error bars are the 95% confidence interval.
-The colors correspond to the different agroforestry systems (CF: crop and fallow, SP: silvopastoral, YA: young agroforest, HG: home garden, FG: forest garden); the different shapes represent the two sites.
+The colors correspond to the different agroforestry systems (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden); the different shapes represent the two sites.
 The dotted lines represent the (non-significant) linear mixed models between log-transformed AGC and SOC stocks.
@@
 
+```{r fig.id="seq-diagram", fig.cap = "Typical sequence of AFS in the sites included in this study.
+Transitions between two systems are represented by arrows; dashed arrows represent transitions that usually include tree removal, whereas solid arrows represent transitions where trees are planted and/or the AFS is left to regenerate naturally.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 8, fig.height = 2.2}
+# save_png <- function(plot, path){
+#   DiagrammeRsvg::export_svg(plot) %>%
+#     charToRaw() %>%
+#     rsvg::rsvg() %>%
+#     png::writePNG(path)
+# }
+# 
+# DiagrammeR::grViz("
+#       digraph {
+#       graph [layout = dot, rankdir = LR]
+# 
+#       node [shape = ellipse]
+#       CF[color = '#1B9E77'];
+#       SP[color = '#D95F02'];
+#       YA[color = '#7570B3'];
+#       HG[color = '#E7298A'];
+#       FG[color = '#E6AB02'];
+# 
+#       NE[label = 'Forest or\nsavanna', shape = rectangle, fillcolor ='black'];
+# 
+#       # Connect nodes with edges and labels
+#       CF -> YA -> HG -> FG;
+#       edge[style = 'dashed']
+#       SP -> CF;
+#       FG -> CF[label = 'tree removal']
+#       FG -> YA
+#       NE -> SP
+#       NE -> CF
+#        }
+# 
+#       ") %>%
+#   save_png("figures/seq_diagram.png")
+include_graphics("figures/seq_diagram.png") 
+```
+
 # References

### 2024-04-16 manuscript.Rmd
version of the manuscript that will be sent to Grace

@@
 title: "Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass"
-author: "Marguerite Cogné, Régis Peltier, Vincent Freycon, Alexis Toumazeau, Camille Piponiot"
+author: "Camille Piponiot*, Marguerite Cogné*, Vincent Freycon, Alexis Toumazeau, Régis Peltier"
 date: "2024-03-19"
@@
 bibliography: references.bib
+nocite: |
+  @Orsini1976, @Nair1993, @Nair2021
+csl: misc/springer-basic-author-date.csl
 ---
@@
 
-**Max words**: The total length of Original Articles, including figures, tables and references, should not exceed **10,000 words**. The number of figures and tables should not exceed 10 in total.
-
-**Word count**: 7200 (with SI); **Figure and table count**: 5 + 6 in SI
+**Word count**: 8365; **Figure and table count**: 5
 
@@
 The soils of the Baucau district are classified using the WRB soil classification [@Mantel2023] as Lithosols and Calcic Luvisols [@FAO2012].
-The average soil properties differed between villages (Table\ S\@ref(tab:site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture.
+The average soil properties differed between villages (Table\ S\@ref(tab:site-soil)).
+Gariuai had a clay loam texture whereas Osso Luga had a clay texture.
 The two villages also have had different histories.
@@
 
-We used the typology described in@Cogne2024, developed in the same study area.
-This typology was based on the one proposed by @Nair1993;@Nair2021, and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
-The five AFS identified were: Crop system including a wooded Fallow phase (CF), Silvopasture (SP), Young Agroforest (YA), Home Garden (HG) and Forest Garden (FG).
-CF are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
-SP land is a vast area (generally 200 to 500 ha) where cattle are grazed under a layer of trees.
-YA are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
-HG are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
+We used the typology described in @Cogne2024, developed in the same study area.
+This typology was based on the one proposed by Nair et al.
+(1993, 2021) and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
+The five AFSs identified were: Crop system including a wooded Fallow phase (CF), Silvopasture (SP), Young Agroforest (YA), Home Garden (HG) and Forest Garden (FG).
+CFs are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
+SPs consist in a vast area (generally 200 to 500 ha) where cattle are grazed under a layer of trees.
+YAs are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
+HGs are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
 FGs are generally former HGs (over 50 years old), which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
 
-The AFS are presented in more detail in the supplemental information (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
+The AFSs are presented in more detail in the supplemental information (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
 
@@
 
-We randomly selected 3 plots from each of the 5 traditional AFS identified in each of the two sites (30 plots in total; Figure\ \@ref(fig:map-sites)).
-Sampling plots were defined as the entire agricultural plot (area between 0.06-1.25 ha) for all AFS except SP plots, which were generally much larger than 1 ha.
+We randomly selected 3 plots from each of the 5 traditional AFSs identified in each of the two sites (30 plots in total; Figure\ \@ref(fig:map-sites)).
+Sampling plots were defined as the entire agricultural plot (area between 0.06-1.25 ha) for all AFSs except SP plots, which were generally much larger than 1 ha.
 In SP plots, a 1 ha sampling plot was delineated within the agricultural plot.
@@
 Soils samples were then air-dried and sieved at 2 mm.
-They were subjected to standard physical and chemical analyses at the Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil-char)).
+They were subjected to standard physical and chemical analyses at the Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil)).
 Among these analyses, total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
@@
 In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm$^3$ at 2-7 cm, 12-17 cm and 22-27 cm.
-Each sample was dried at 105°C in a classic oven until reaching a stable dry mass, sieved to 2 mm, and then the fine fraction (< 2 mm) was weighted.
+Each sample was dried at 105°C in a classic oven until reaching a stable dry mass, sieved to 2 mm, and then the fine fraction (< 2 mm) was weighed.
 
@@
 data_sim = rbind(data_sim, data_sim[, .(value = sum(value[variable != "SOC10"]), variable = "TOT"), .(N_PLOT, site, ID_AF, iter)])
-# group total C stocks in plots 6 and 8 (otherwise will underestimate CF total C
-# stocks); this will not be used for analysis, only to display average C stocks
-# per AFS
-# data_sim <- rbind(
-#   data_sim[variable != "TOT" | !N_PLOT %in% c(6, 8)], 
-#   data_sim[variable == "TOT" & N_PLOT %in% c(6, 8), .(value = sum(value), N_PLOT = 6), .(site, ID_AF, variable, iter)]
-# )
 ```
@@
 ## Statistical analysis
+
+All statistical analyses were carried out using R [@R2023].
 
@@
 
-Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
@@
 
-AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)), but not between site (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
+AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table\ \@ref(tab:anova)), but not between site (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
 The AFS with the lowest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
@@
 
-SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)).
-However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
+SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table\ \@ref(tab:anova)).
+However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
 The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$) than those in Osso Luga (`r dataC_site[site=="Osso Luga" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
@@
 
-## Large C stocks in traditional AFS in Timor-Leste
-
-Compared to values reported in the literature for other AFS, the carbon stocks in this study were particularly high.
-The value of SOC stocks was similar to other reported values [@Feliciano2018], but some of the AFS studied here have particularly high stocks in their biomass [compared to the average of 74\ Mg\ C\ ha$^{-1}$ in @Ma2020].
-The AFS with the highest carbon stocks, forest gardens, had stocks close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$; @Lal2005].
-These high carbon stocks could be explained by the fact that forest gardens have generally grown for more than 50 years, with little human intervention (Table S1).
-In terms of structure and carbon stocks, forest gardens should therefore resemble old secondary forests, which have been shown to recover more than 60% of the AGC stocks and almost all the SOC stocks of an old forest [@Poorter2021].
-In addition, FGs are generally located close to villages, where soils are enriched by organic waste and animal manure, which can enhance tree growth [@Madejon2016].
+## Traditional AFS in Timor Leste can store large C stocks
+
+In comparison to values reported in the literature for other AFS, the carbon stocks in this study were relatively high.
+Specifically, average SOC stocks (`r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$) were higher than other reported values: in a meta-analysis by @Shi2018, it was estimated that AFS stored around 50\ Mg\ C\ ha$^{-1}$ at 0-30 cm depth.
+These large SOC stocks could be due to the high soil pH associated with high exchangeable Ca content at our sites (Table\ \@ref(tab:site-soil)), as Ca is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
+Average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`\ Mg\ C\ ha$^{-1}$), in contrast, were more similar to the average of 74\ Mg\ C\ ha$^{-1}$ estimated by @Ma2020, but some AFS had particularly high biomass values.
+The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`s, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$) similar to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$, @Lal2005].
+The high carbon stocks in FGs can be attributed to their long period since establishment, which is generally more than 50 years, with little human intervention (Table\ S\@ref(tab:afs-typology)).
+In terms of structure and carbon stocks, FGs should therefore resemble old secondary forests, which have been shown to recover more than 60% of the AGC stocks and almost all the SOC stocks of an old forest [@Poorter2021].
+In addition, FGs contain a large number of trees of the same age that were planted when the AFS was established and can survive to reach significant sizes, similar to forest plantations that can grow to old-growth-forest levels of AGC stocks after only a few decades [@Brown2020;@Njoukam1996].
 
@@
 
-The AFS studied here vary considerably in terms of AGC stocks, in agreement with our hypothesis H1a.
+The AFS studied here vary significantly in terms of AGC stocks, in agreement with our hypothesis H1a.
 This reflects the density and size of trees in the plot, which in turn depends on the importance of trees in the production system.
@@
 during the fallow period) or providing shade for livestock.
-This more limited role for trees may explain why their total biomass is lower in these systems: in the CF system, the plot is cleared and annual crops are planted in full sun, while in the SP system, the growth of grazing grasses is favored by reducing the number of trees that shade them.
+This more limited role for trees may explain why their total biomass is lower in these systems: in the CF system, the plot is regularly cleared and annual crops are planted in full sun, while in the SP system, the growth of grazing grasses is favored by reducing the number of trees that shade them.
 
 It is also interesting to note that, in line with previous results, the age of the trees in the AFS largely explains the carbon stored in their biomass [@Ma2020].
-Indeed, the systems described in this study reflect different stages of transition in land use, which follow the growth of their trees.
-For example, a farmer may plant trees in a CF field, which will then become a YA with trees aged between one and ten years, where the main production still comes from annual crops.
-When the trees are mature and producing fruit, the YA becomes a HG or a FG.
-In the latter systems, the trees become the main source of production, both for food and wood.
+Indeed, the systems described in this study reflect different stages of transition in land use, which follow the growth of their trees (Figure\ S\@ref(fig:seq-diagram)).
+For example, a farmer might plant trees in a CF field, which will then become a YA that can still produce annual crops for the first 2 to 4 years, before the shade from the trees becomes too great.
+The trees are then used as property markers and, when they reach maturity, to produce wood and fruit: the YA then becomes a HG or a FG.
+Likewise, our results show that AGC stocks increase from CF to YA, HG and FG (Table\ \@ref(tab:summary)).
 
@@
 We therefore expect HGs, FGs and YAs to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuer2018].
-Conversely, CF and SP fields are mainly located on limestone plateaus where water is deep and soils are shallow, which could reduce tree growth and biomass [@Imada2008].
-
-## Importance of historical and environmental factors on SOC
-
-Our results revealed that SOC was mainly sensitive to the differences in location and villages rather than AFS systems in each location.
-This was also reported by the meta-analysis of @Feliciano2018 who highlighted that only 3.2% of the variance of the soil carbon sequestration was explained by agroforestry system choices.
-The two villages studied were different in terms of inherent soil properties (Table S\@ref(tab:site-soil); Figure S\@ref(fig:clay-corg)), altitudes (Figure \@ref(fig:map-sites); Figure S\@ref(fig:altitude-corg)) and also history.
-All those driving factors independent of the AFS management could not be disentangled in this study but their cumulative effects were clearly observed.
-The difference in terms of soil texture and especially clay content was reported to be a key driving process in terms of SOC storage in the tropics [@Don2010;@Matus2021].
-In this study, the fine fractions of soil texture were however not linked to the SOC stocks (Figure S\@ref(fig:clay-corg)).
-In addition to soil inherent conditions, other key driving factors for SOC changes, independent to the AFS management practices, may have led to those differences between.
-Gariuai was inhabited whereas Osso-Luga was not for a long time.
-The management and care of the plot in Gariuai such as pruning, livestock visits, optimizing the different stratas in terms of carbon input could have led to increased soil organic carbon compared to abandoned systems [@Lorenz2014].
-
-The AGC stocks differed from one system to another but we could not detect any significant difference in SOC between AFSs.
-This observation was in contradiction with our H1b.
-Our systems were rather contrasted and previous studies reported significant differences in SOC sequestration between different AFS [@Ma2020;@Kimaro2024].
-Local conditions and the rapid SOC mineralisation in tropical areas may explain this lack of sensitivity against AFs management practices [@Sayer2019].
-Our experimental design with a limited number of replicates within each village could have also limited the significance of the tests.
-Some trends were highlighted, such as higher SOC in SP in Gariuai.
-The design was based on 3 true replicates for each of the systems in each village with the objective to limit as much as possible the possible influence of external co-factor in each village (e.g.
-farmer management practices variability, micro-topography, etc).
-However, much variability cannot be controlled in farmers’ plots and increasing the number of replicates for each AFS may have strengthened the power of the tests and may have led to more differences between systems.
-
-## Lack of relationship between soil and biomass carbon
-
-We found no relationship between AGC and SOC stocks both in the topsoil layer 0-10 cm depth (H2a refuted) and in 0-30 cm depth (H2b confirmed).
-
-These results could be explained by previous findings that SOC stocks saturate much faster than AGC stocks in tropical AFS.
+Conversely, CF and SP are mainly located on limestone plateaus where water is deep and soils are shallow, which could reduce tree growth and biomass [@Imada2008].
+
+## Historical and environmental factors explain differences in SOC stocks
+
+We found no significant difference in SOC between AFS (H1b rejected).
+This result is inconsistent with our expectation that agroforestry practices would influence SOC stocks [@Minasny2017;@Wiesmeier2019] and with previous studies showing significant differences in SOC stocks between different AFSs [e.g., @Ramesh2015].
+However, it is worth noting that some single-site studies [e.g.
+@Betemariyam2020;@GamaRodrigues2010] and at least two meta-analyses failed to find a strong effect of AFS on SOC stocks: @Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by AFS selection, and @Shi2018 found no significant differences in SOC stocks between 4 different AFSs in 354 previous studies.
+We offer three main explanations to reconcile these differences between studies.
+First, in a global meta-analysis, @Ma2020 estimated that it takes only 5 years for SOC stocks in tropical AFS to reach an equilibrium value.
+In our study, at least three of the five AFS studied (namely YAs, HGs and FGs) represent different stages of an AFS transition depending on the age of the trees (Figure\ S\@ref(fig:seq-diagram).
+Most of these AFSs are older than 5 years and may have already reached their equilibrium SOC stocks, which could explain why there is little difference between them.
+Second, we had a limited number of replicates of each AFS within each site, which certainly reduced the power of statistical tests.
+We were only able to identify trends, such as higher SOC in SPs in Gariuai, which were similar to previous studies [@Feliciano2018].
+Increasing the number of replicates for each AFS could have increased the power of the tests and potentially led to more significant differences between AFSs.
+Third, other pedoclimatic and historical factors influence carbon stocks, sometimes to a greater extent than AFS, e.g.
+within-site variability in climate, soil texture (Figures\ S\@ref(fig:clay-corg), S\@ref(fig:altitude-corg)) or previous land use may have significantly influenced SOC stocks [@Wiesmeier2019;@Ma2020].
+Our experimental design aimed to limit as much as possible the influence of external co-factors at each site (e.g.
+variability in farmers' management practices, micro-topography, etc.), but much of the variability cannot be controlled in farmers' plots.
+
+We found a significant difference in SOC between the two sites: SOC stocks were higher in Gariuai than in Osso Luga.
+A priori, three factors that differentiate the sites could explain these differences in SOC stocks: climate (through elevation; Figure\ \@ref(fig:map-sites); Figure\ S\@ref(fig:altitude-corg)), soil texture (Table\ S\@ref(fig:site-soil); Figure\ S\@ref(fig:clay-corg)) and history.
+First, we found that SOC stocks were positively correlated with elevation, both between and within sites (Figure\ S\@ref(fig:altitude-corg)).
+This result is consistent with those of @Raich2006, who found higher SOC at higher elevations in tropical evergreen forests in different countries (e.g.
+Congo, Thailand).
+As changes in elevation are associated with variations in climate (precipitation, temperature), our result is also consistent with previous studies that found higher SOC in areas with higher precipitation [@Schuur2001] and lower temperature [@Raich2006;@MarinSpiotta2013].
+Second, soil texture is a key driving process for SOC storage in the tropics [@Don2010;@Wiesmeier2019;@Matus2021].
+In particular, many studies have found a strong positive correlation between SOC stocks and the fine mineral fractions of soils [i.e.
+clay or clay + fine silt content; @Zinn2005].
+In our study, we found no correlation between SOC stocks and clay + fine silt content when site and depth were included in the model (Figure\ S\@ref(fig:clay-corg)).
+Furthermore, the Gariuai soils, which had the highest stocks of soil organic carbon (SOC), were also the least clayey.
+This contradicts the typical (positive) relationship between the fine mineral fraction and SOC stocks.
+It is therefore unlikely that the difference in soil organic carbon (SOC) between the two sites can be attributed to soil texture.
+Thirdly, history could be a significant driving factor.
+Gariuai was continuously inhabited for over 50 years, whereas Osso-Luga was abandoned for almost three decades.
+The management and care of the AFSs in Gariuai, such as pruning and livestock visits, as well as optimizing the different strata in terms of C-input, could have led to an increase in soil organic carbon compared to abandoned systems [@Lorenz2014].
+
+## SOC stocks are not explained by variation in AGC stocks
+
+We found no significant relationship between AGC and SOC stocks both in the topsoil layer 0-10 cm depth (H2a refuted) and in 0-30 cm depth (H2b confirmed).
+These results may relate to previous findings that SOC stocks saturate much faster than AGC stocks in tropical AFS.
 For example, in a global meta-analysis,  @Ma2020 estimated that it took only 5 years for SOC stocks in tropical AFS to reach an equilibrium value, but more than three decades for AGC stocks.
 This desynchronisation of SOC and AGC stocks recovery has also been observed in naturally regenerating tropical forests.
-In a systematic review, @Martin2013 found that SOC in tropical secondary forests changed little with time since disturbance, while AGB increased.
+In a systematic review, @Martin2013 found that SOC in tropical secondary forests changed little with time after disturbance, while AGB increased.
 @Ojoatre2024 found similar results in a tropical montane secondary forest.
-These results imply that there is little relationship between AGC and SOC once agroforestry systems are more than 5 years old, which is the case in our study for FGs, HGs (> 5 years old) and partially for YAs (2 to 10 years; Table S\@ref(tab:afs-typology)).
+These results imply that there is little relationship between AGC and SOC once agroforestry systems are more than 5 years old, which is the case in our study for FGs, HGs (> 5 years old) and partially for YAs (2 to 10 years; Table\ S\@ref(tab:afs-typology)).
 This saturation of SOC even with increasing AGC is counterintuitive, as it might be expected that greater AGC would lead to greater litter production and decomposition, some of which would increase SOC.
-However, this was not shown by @Sayer2019, who found that SOC did not increase in a lowland tropical forest in Panama after 15 years of an experimentally doubled aboveground litter input.
+However, @Sayer2019, found that SOC did not increase in a lowland tropical forest in Panama after 15 years of an experimentally doubled aboveground litter input.
 Two processes could explain this.
@@
 
-## How to conserve and improve carbon stocks in traditional AFS in Timor-Leste?
-
-Although agroforestry can make agriculture more resilient while improving its carbon footprint and biodiversity conservation value [@Cardinael2021], poor adaptation to the local context can explain many failures to adopt new agroforestry techniques [@Coe2014].
+In addition to biomass stocks, other vegetation characteristics may account for variations in soil carbon stocks between AFSs.
+For instance, previous studies have shown that tree diversity can enhance SOC sequestration in natural ecosystems [@Spohn2023;@Chen2018] and in AFSs [@Manaye2021;@Islam2015].
+Moreover, total soil organic carbon may be less sensitive to management practices such as biomass inputs than specific fractions of soil organic carbon [@Haynes2005].
+Therefore, it would be interesting to continue with further studies that focus not only on the quantity but also the quality of soil organic matter and its relationship with biodiversity (flora and fauna) in traditional Timorese AFSs.
+
+## How to conserve and improve carbon stocks in traditional AFS in Timor Leste?
+
+Although agroforestry can make agriculture more resilient to climate change while improving its carbon footprint and biodiversity conservation value [@Cardinael2021], poor adaptation to the local context can explain many failures to adopt new agroforestry techniques [@Coe2014].
 Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
@@
 
-Conservation of these traditional AFS should avoid their abandonment as the trees wither or as their production value declines.
+Conservation of these traditional AFSs should avoid their abandonment as the trees wither or as their production value declines.
 Clear-cutting old agroforests for full-sun crops may then become an attractive short-term strategy, providing a large cash inflow from the sale of wood products, followed by 4 or 5 years of agricultural production on fertile land.
@@
 payment of carbon credits) or low-interest loans to farmers wishing to renew their agroforestry plots.
-In addition, a "sustainable agriculture" type label could be awarded to these AFS products, and their value chains could be supported by governments and international certification bodies, as is already the case for coffee in many countries [@Bertrand2019].
+Another strategy, which has been successfully implemented in other tropical countries (e.g.
+for agroforestry coffee), is to improve the incomes of AFS farmers by creating 'sustainable agriculture'-type labels [@Bertrand2019], although this is primarily applicable to the production of export goods, whereas the AFSs we have studied are currently oriented more towards local consumption and domestic markets.
 
@@
 
-Agroforestry systems in the Baucau district of Timor-Leste had variable but overall high carbon stocks.
+Agroforestry systems in the Baucau district of Timor had variable but overall high carbon stocks.
 These systems have only recently been described in the scientific literature, and much remains to be done to improve our knowledge of their diversity and functioning.
@@
 All authors contributed to the design of the sampling protocol.
-MC carried out the fieldwork with the help of RP, AT and VF [mention students?].
-CP analyzed the data with the help of VF and MC.
+MC carried out the fieldwork.
+CP analyzed the data with the help of MC and VF.
 CP prepared the manuscript and all authors contributed to the drafting and revision of the manuscript.
@@
 
+We are grateful to Marcal Gusmao (National University of Timor-Leste) and Alain Rival (Cirad) for their facilitation of the field study, and to the trainees who helped with field data collection: Juvencio Dos Santos (vegetation) and Teofilio Gomes (soil).
+We would also like to thank the 31 farmers who gave up their time to take part in the interviews and field measurements.
+
+This project was co-funded by the European Union and the German Federal Ministry for Economic Cooperation and Development (BMZ) as part of the Partnership for Sustainable Agroforestry (PSAF) - Ai ba Futuru project; the Deutsche Gesellschaft für Internationale Zusammenarbeit (GIZ); the French Embassy in Jakarta; and the French Institute of Timor-Leste.
+
 # Data availability
+
+All data collected for this study is available in the following Dataverse repository ______.
+
+The R codes used to analyse the data and produce the manuscript (including figures and tables) are available on ______.
+
+\newpage
 
@@
 
+</br>
+
 ```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAS with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
@@
 ```
+
+\newpage
 
@@
 
+</br>
+
 ```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Barplot of the amount of carbon in each AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden), per site.
@@
 
+</br>
+
 ```{r fig.id = "plot-lmereg", fig.height = 3.5, fig.width = 7, fig.cap = "Scatter plot of aboveground carbon (AGC; x-axis) and soil organic carbon (SOC; y-axis) stocks.
@@
 The dotted lines represent the (non-significant) linear mixed models between log-transformed AGC and SOC stocks.
-For each linear mixed model, we report the marginal R2 (proportion of variance explained by fixed effects only) and the conditional R2 (proportion of variance explained by both fixed and random effects).
+For each linear mixed model, we report the marginal R^2^ (proportion of variance explained by fixed effects only) and the conditional R^2^ (proportion of variance explained by both fixed and random effects).
 Both x- and y-axes are log-transformed.
@@
 
+# References
+
+<div id="refs"></div>
+
+\newpage
+
 # Supplementary information
@@
 
-```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (average over depths 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
-The confidence intervals were calculated by propagating a 12% error on the lab-provided values and obtaining site-level estimates through bootstrapping plot values (1000 simulations)."}
+```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (depth of 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
+The confidence intervals were calculated by propagating a 12% error on the lab-provided values and obtaining site-level estimates through bootstrapping plot values (1000 simulations).
+Total organic carbon content (dry combustion method) ; total nitrogen content (dry combustion method) ; pH in water (soil:deionized water ratio of 1:2.5) ; CEC - cation exchangeable capacity - and exchangeable cations (Ca, K, Mg and Na) at soil pH [Cobaltihexamine method; Orsini and Remy, 1976]."}
 flextable(
@@
 
-```{r tab.id = "afs-typology", tab.cap.pre="Table S", tab.cap = "Description of 5 agroforestry systems in Central Timor-Leste [adaptation of the name given by Nair, position in the landscape, surface area, age since creation, previous cultivation", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
+</br>
+
+```{r tab.id = "afs-typology", tab.cap.pre="Table S", tab.cap = "Description of 5 agroforestry systems in Central Timor-Leste.", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
 read_excel("tables/afs-typology.xlsx") |> 
@@
 ```
+
+</br>
 
@@
 
+## correct typos in vernacular names
+vern[, gsub("kamii", "kami", TETUM_NAME)]
+vern[, gsub("ai-haek", "ai-hanek", TETUM_NAME)]
+vern[, gsub("ai-bobur", "ai-bubur", TETUM_NAME)]
+
 ## correct scientific names
@@
 vern = vern[!(tetum %in% dupl_tetum & is.na(speciesCorrected))]
-setorder(vern, tetum)
-
-vern = vern[, .(latin = paste(latin, collapse = ", ")), .(tetum)]
+setorder(vern, latin)
+
+vern = vern[, .(tetum = paste(tetum, collapse = ", ")), .(latin)]
 ```
@@
 flextable(vern) |> 
-  set_header_labels(values =  c("Tetum name", "Scientific name")) |> 
+  set_header_labels(values =  c("Scientific name", "Tetum name")) |> 
   autofit()
@@
 ```{r fig.id="design", fig.cap = "Plot sampling design.
-(a) Typical plot < 1 ha, in all AFS except SP.
-(b) Typical 1-ha plot in SP systems.
+(a) Typical plot < 1 ha, in all AFS except SPs.
+(b) Typical 1-ha plot in SPs.
 Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 5, fig.height = 3}
@@
 
-```{r fig.id = "clay-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap="Distribution of SOC content (y-axis; in %) values as a function of fine silt and clay content (x-axis; in %), soil depth (point color; in cm), and site (point shape).", fig.cap.pre="Figure S"}
-data_soil[, site := substr(ID_VILLAGE, 1, 1)]
-data_soil[site == "S", site := "Osso Luga"]
-data_soil[site == "G", site := "Gariuai"]
-
-lmer(
-  log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
-    (1|site)+ 
-    (1|SAMPLE_DEPTH), 
-  data = data_soil[SAMPLE_DEPTH != "10-20"]) |> 
-  summary()
-
-ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, shape = SAMPLE_DEPTH, col = site)) +
-  geom_point() +
-  scale_x_log10() + scale_y_log10() +
-  labs(shape = "Soil depth (cm)", 
-       x = "Clay and fine silt content (%)", 
-       y = "SOC content (%)", 
-       col = "Site") +
-  theme_classic()
-```
-
-```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in\ Mg\ C\ ha^-1^).
-Points are colored by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.", fig.cap.pre="Figure S"}
-dataCelev = merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")], 
-                  data_plot[, c("N_PLOT", "long", "lat")], by.x = "plot", by.y = "N_PLOT")
-
-# extract elevation
-dataCelev[, elev := terra::extract(
-  terra::rast("data/srtm_62_14.tif"), 
-  cbind(long, lat)
-)]
-
-lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |> 
-  summary()
-
-ggplot(dataCelev, aes(x = elev, y = mean, col = site)) +
-  geom_pointrange(aes(ymin=lwr, ymax = upr)) + 
-  theme_classic() +
-  labs(x = "Elevation (m)", y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), col = "Site")
-```
+</br>
 
 ```{r fig.id="seq-diagram", fig.cap = "Typical sequence of AFS in the sites included in this study.
-Transitions between two systems are represented by arrows; dashed arrows represent transitions that usually include tree removal, whereas solid arrows represent transitions where trees are planted and/or the AFS is left to regenerate naturally.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 8, fig.height = 2.2}
+Transitions between two systems are represented by arrows: dashed arrows represent transitions that usually include tree removal, whereas solid arrows represent transitions where trees are planted and/or the AFS is left to regenerate naturally.
+These transitions may vary according to the farmer's strategy and available resources (Cogné and Lescuyer, 2024).", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 8, fig.height = 2.2}
 # save_png <- function(plot, path){
@@
 #       CF -> YA -> HG -> FG;
+#       YA -> FG;
 #       edge[style = 'dashed']
 #       SP -> CF;
-#       FG -> CF[label = 'tree removal']
+#       FG -> CF
 #       FG -> YA
@@
 
-# References
+</br>
+
+```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in\ Mg\ C\ ha^-1^).
+Points are colored by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.
+We performed a linear mixed effects regression with log-transformed SOC stocks as the dependent variable, log-transformed elevation as the independent variable, and site as random effect.
+Results are presented as regression lines and the p-value of the relationship between SOC stocks and elevation.", fig.cap.pre="Figure S"}
+dataCelev = merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")], 
+                  data_plot[, c("N_PLOT", "long", "lat")], by.x = "plot", by.y = "N_PLOT")
+
+# extract elevation
+dataCelev[, elev := terra::extract(
+  terra::rast("data/srtm_62_14.tif"), 
+  cbind(long, lat)
+)]
+
+regElev = lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |> 
+  summary()
+pvalueE = signif(regElev$coefficients[2, 5], 2)
+slp = regElev$coefficients[2, 1]
+
+dfElev = dataCelev[, .(
+  int = mean(log(mean)) - slp * mean(log(elev)), 
+  elev = min(elev), elevMax = max(elev),
+  slp = slp), .(site)] 
+dfElev[, mean := exp(slp*log(elev) + int)]
+dfElev[, predMax := exp(slp*log(elevMax) + int)]
+
+ggplot(dataCelev, aes(x = elev, y = mean, col = site)) +
+  geom_pointrange(aes(ymin=lwr, ymax = upr)) + 
+  geom_segment(data = dfElev, aes(xend = elevMax, yend = predMax)) +
+  theme_classic() +
+  annotate(geom = "text", label = paste("p-value =", pvalueE), x = 500, y = 32) +
+  labs(x = "Elevation (m)", y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), col = "Site") +
+  scale_x_log10() + scale_y_log10()
+```
+
+</br>
+
+```{r fig.id = "clay-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap="Distribution of SOC content (y-axis; in %) values as a function of fine silt and clay content (x-axis; in %), soil depth (point color; in cm), and site (point shape).
+We performed a linear mixed effects regression with log-transformed SOC content as the dependent variable, log-transformed fine soil texture as the independent variable, and soil depth and site as random effects.
+Results are presented as regression lines and the p-value of the relationship between SOC content and fine soil texture.", fig.cap.pre="Figure S"}
+
+data_soil[, site := substr(ID_VILLAGE, 1, 1)]
+data_soil[site == "S", site := "Osso Luga"]
+data_soil[site == "G", site := "Gariuai"]
+
+regText = lmer(
+  log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
+    (1|site)+ 
+    (1|SAMPLE_DEPTH), 
+  data = data_soil[SAMPLE_DEPTH != "10-20"]) |> 
+  summary()
+pvalueT = signif(regText$coefficients[2, 5], 3)
+slp = regText$coefficients[2, 1]
+
+dfText = data_soil[SAMPLE_DEPTH != "10-20" & !is.na(CLAY), .(
+  int = mean(log(Corg)) - slp * mean(log(as.numeric(CLAY) + as.numeric(`FINE SILT`))), 
+  xmin = min(as.numeric(CLAY) + as.numeric(`FINE SILT`)), 
+  xmax = max(as.numeric(CLAY) + as.numeric(`FINE SILT`)),
+  slp = slp), .(site, SAMPLE_DEPTH)] 
+dfText[, predMin := exp(slp*log(xmin) + int)]
+dfText[, predMax := exp(slp*log(xmax) + int)]
+
+ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, shape = SAMPLE_DEPTH, col = site)) +
+  geom_point() +
+  geom_segment(data = dfText, aes(x = xmin, y = predMin, xend = xmax, yend = predMax, lty = SAMPLE_DEPTH)) +
+  scale_x_log10() + scale_y_log10() +
+  scale_linetype_manual(values=c("twodash", "dotted")) +
+  labs(shape = "Soil depth (cm)", lty = "Soil depth (cm)", 
+       x = "Clay and fine silt content (%)", 
+       y = "SOC content (%)", 
+       col = "Site") +
+  annotate(geom = "text", label = paste("p-value =", pvalueT), x = 45, y = 1) +
+  theme_classic()
+```

### 2024-04-17 manuscript.Rmd
change units in site-soil table: from % into g/kg

@@
 
+# convert sand, silt, clay and C content from % into g/kg
+data_soil_site_simu[ variable %in% c("Corg", "CLAY", "sand", "silt"), value := value*10]
+
 # calculate mean and 95% confidence interval at site level
@@
                "N", "Corg_N", "pH", "CEC", "Ca", "K", "Mg", "Na"), 
-  cwidth = 0.5
-) |> set_header_labels(
-  values = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)", 
-             "N", "C/N", "pH", "CEC", "Ca", "K", "Mg", "Na"))|> 
-  compose(i = 1, j = "N", part = "header", value = as_paragraph("N (%", as_sub("o"), ")")) |>
+  cwidth = c(0.5, rep(0.6, 5), rep(0.5, 2), rep(0.6, 5)) 
+) |> 
+  compose(i = 1, j = "CLAY", part = "header", value = as_paragraph("Clay\n(g kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "silt", part = "header", value = as_paragraph("Silt (g kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "sand", part = "header", value = as_paragraph("Sand (g kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "Corg", part = "header", value = as_paragraph("C content (g kg", as_sup("-1"), ")")) |> 
+  compose(i = 1, j = "N", part = "header", value = as_paragraph("N content (g kg", as_sup("-1"), ")")) |>
+  compose(i = 1, j = "Corg_N", part = "header", value = as_paragraph("C/N")) |> 
   compose(i = 1, j = "CEC", part = "header", value = as_paragraph( "CEC (cmolc kg", as_sup("-1"), ")")) |> 

### 2024-04-23 manuscript.Rmd
add suggestions from alain and marcal

@@
 title: "Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass"
-author: "Camille Piponiot*, Marguerite Cogné*, Vincent Freycon, Alexis Toumazeau, Régis Peltier"
+author: "Camille Piponiot*, Marguerite Cogné*, Vincent Freycon, Alexis Thoumazeau, Marçal Gusmão, Régis Peltier"
 date: "2024-03-19"
@@
   officedown::rdocx_document:
+    base_format: "bookdown::word_document2"
     reference_docx: word-styles-ref-02.docx
@@
 
-Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural fields and diversifying agricultural landscapes.
-Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their potential inclusion in climate change mitigation strategies.
-In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
-Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, similar to an old-growth tropical forest.
+Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural land and diversifying landscapes.
+Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their inclusion in climate change mitigation strategies.
+In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFSs) in Timor-Leste, namely: cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
+Our results show that these traditional AFSs can store large amounts of carbon: they stored an average of 156\ Mg\ C\ ha$^{-1}$, and the AFSs with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, close to the values of old-growth tropical forests.
 Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
 We found no relationship between the amount of carbon stored in biomass and soil.
-Our results highlight the high diversity of traditional agroforestry systems in Timor-Leste and their high capacity to store carbon.
+Our results highlight the high diversity of traditional AFSs in Timor-Leste and their high carbon sequestration capacity.
+These results could provide an important baseline for the inclusion of AFS in Timor-Leste's climate change mitigation strategy, and could serve as a reference for future AFS studies in different agro-climates of Timor-Leste.
 
@@
 Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO~2~e/yr over 2010-2014; @Pendrill2019].
-The demand for tropical agricultural land is expected to continue to increase as the population grows [@UN2022].
+The demand for tropical agricultural land is expected to continue to grow as population increases [@UN2022].
 In addition, tropical areas are expected to be disproportionately affected by climate change, putting their agriculture and food systems at risk [@Lawrence2015].
 There is therefore an urgent need to promote tropical agricultural landscapes that are both more resilient to climate change, while maintaining high productivity and ecological value [@Harvey2014].
-In particular, increasing the carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of the food production system [@Lal2004;@Minasny2017].
+In particular, increasing the amount of carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of the food production system [@Lal2004;@Minasny2017].
 
@@
 
-Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation and soil, and this carbon storage capacity depends on the characteristics of AFS considered [@Albrecht2003;@Feliciano2018].
-As trees store large amounts of carbon, the carbon stored in the vegetation of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
-However, the effect of AFS on soil organic carbon (SOC) has been relatively little studied [@Beillouin2021], is more variable and depends on a number of factors.
-Although AFS appear to increase SOC stocks overall, some AFS such as woodlots (i.e.
-planting trees with intercropping during the establishment phase) may decrease SOC compared to non-AFS, while AFS such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
-Other factors such as climate and previous land use type have also been shown to strongly influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018;@Ma2020;@Martin2020].
-Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS could only be demonstrated for the topsoil layer [< 10 cm; @Ma2020].
-
-In Timor-Leste, a large proportion of the population depends on agriculture for their livelihoods, but food security is threatened by population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime [@WB2009;@Molyneux2012].
+Agroforestry systems (AFSs) have the potential to store large amounts of carbon in the vegetation and soil, and this carbon storage capacity depends on the characteristics of AFS considered [@Albrecht2003;@Feliciano2018].
+As trees can store large amounts of carbon, the carbon stored in the vegetation of AFSs is largely dependent on the number and size of trees in the system [@Ma2020].
+However, the effect of AFSs on soil organic carbon (SOC) has been relatively little studied [@Beillouin2021], is more variable and depends on a number of factors.
+Although AFSs appear to increase SOC stocks overall, some AFS such as woodlots (i.e.
+planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFSs, while AFSs such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
+Other factors such as climate and previous land use type can also influence SOC in AFSs, which may explain the variability of observations reported in the literature [@Feliciano2018;@Ma2020;@Martin2020].
+Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS could only be evidenced for the topsoil layer [< 10 cm; @Ma2020].
+
+In Timor-Leste, a large proportion of the population depends on agriculture for their livelihoods, and food security is threatened by population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime [@WB2009;@Molyneux2012].
 In this context, productive, resilient and sustainable agriculture is critical for the country, and agroforestry has been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022;@Cogne2024].
-Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024] and can be key elements for biodiversity conservation in the tropics [@Perfecto2008;@Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
-Better quantification of these carbon stocks would improve the understanding of the ecological importance of these AFS at both local and global scales, and help to integrate them into national and international climate change mitigation initiatives.
-
-In this study, we aim to better understand traditional agroforestry systems in Timor-Leste, and in particular their carbon storage both in their biomass and in the soil.
-We hypothesize that different types of traditional agroforestry systems will exhibit measurable differences in aboveground carbon (AGC; H1a) and SOC (H1b) stocks.
-We also assume, following @Ma2020, that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; H2a), but no relationship with SOC below 10 cm (H2b).
-
-Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural fields and diversifying agricultural landscapes.
-Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their potential inclusion in climate change mitigation strategies.
-In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste:  cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
-Our results show that these traditional AFS store large amounts of carbon: the AFS with the highest carbon stocks (forest gardens) stored an average of 255 MgC/ha, similar to an old-growth tropical forest.
-Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
-We found no relationship between the amount of carbon stored in biomass and soil.
-Our results highlight the high diversity of traditional agroforestry systems in Timor-Leste and their high capacity to store carbon.
+Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024] and can be key elements for biodiversity conservation [@Perfecto2008;@Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFSs.
+A precise quantification of these carbon stocks would improve the understanding of the ecological importance of these AFSs at both local and global scales, and help to integrate them into national and international climate change mitigation initiatives.
+
+In this study, we aim to better understand traditional AFSs in Timor-Leste by quantifying the carbon storage in both biomass and soil of five traditional AFSs: cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
+We hypothesize that different types of traditional AFSs will exhibit measurable differences in aboveground carbon (AGC; H1a) and SOC (H1b) stocks.
+We anticipate, following @Ma2020, that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; H2a), but no relationship with SOC below 10 cm (H2b).
 
@@
 data_plot[, site := substr(ID_VILLAGE, 1, 1)]
-data_plot[site == "S", site := "Osso Luga"]
+data_plot[site == "S", site := "Osso-Luga"]
 data_plot[site == "G", site := "Gariuai"]
@@
 
-The research was carried out in the center of the Baucau district, on the north-eastern coast of Timor-Leste.
-Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso Luga", elev]` m (Figure\ \@ref(fig:map-sites)).
-The soils of the Baucau district are classified using the WRB soil classification [@Mantel2023] as Lithosols and Calcic Luvisols [@FAO2012].
-The average soil properties differed between villages (Table\ S\@ref(tab:site-soil)).
-Gariuai had a clay loam texture whereas Osso Luga had a clay texture.
-The two villages also have had different histories.
-Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2003.
-During the period of abandonment, the various agroforestry systems in Osso Luga lay fallow.
+The research was carried out in the center of the Baucau municipality, on the north-eastern coast of Timor-Leste.
+Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m above sea level (asl); and (ii) the village of Osso-Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso-Luga", coord]`) located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso-Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso-Luga", elev]` m asl (Figure\ \@ref(fig:map-sites)).
+The soils of the Baucau municipality are classified using the WRB soil classification [@Mantel2023] as Lithosols and Calcic Luvisols [@FAO2012].
+The average soil properties differed between villages (Table\ S\@ref(tab:site-soil)): Gariuai had a clay loam texture whereas Osso-Luga had a clay texture.
+The two villages also have different recent histories: Gariuai has been inhabited for at least 50 years, while Osso-Luga was abandoned during the Indonesian occupation and was only repopulated after Timor-Leste gained independence in 2002.
+During the period of abandonment, the AFSs in Osso-Luga lay fallow.
 
@@
 
-We used the typology described in @Cogne2024, developed in the same study area.
+We used the typology described in @Cogne2024, which was developed in the same study area.
 This typology was based on the one proposed by Nair et al.
@@
 SPs consist in a vast area (generally 200 to 500 ha) where cattle are grazed under a layer of trees.
-YAs are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years, usually with the help of projects.
-HGs are plots of land close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
+YAs are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years.
+HGs are plots of land located close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
 FGs are generally former HGs (over 50 years old), which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
 
-The AFSs are presented in more detail in the supplemental information (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
+The AFSs are presented in more detail as supplemental information (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
 
@@
 We randomly selected 3 plots from each of the 5 traditional AFSs identified in each of the two sites (30 plots in total; Figure\ \@ref(fig:map-sites)).
-Sampling plots were defined as the entire agricultural plot (area between 0.06-1.25 ha) for all AFSs except SP plots, which were generally much larger than 1 ha.
+Sampling plots were defined as the entire agricultural plot (area between 0.06 and 1.25 ha) for all AFSs except SP plots, which were generally much larger than 1 ha.
 In SP plots, a 1 ha sampling plot was delineated within the agricultural plot.
@@
 
-Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory government's Native Plants Herbarium [@NT2013].
-The correspondence between tetum names and scientific names is provided in the supplemental material (Table S\@ref(tab:tetum-names)).
+Taxonomic identification was carried out using Tetum  (a national official language) vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory government's Native Plants Herbarium [@NT2013].
+The correspondence between Tetum names and scientific names is provided in the supplemental material (Table S\@ref(tab:tetum-names)).
 
@@
 
-In each of the 30 sample plots, three composite samples were taken from three of the five 10 x 10 m subplots, following a transect covering the within plot variability and at a minimum distance of 1 m from any tree (Figure\ S\@ref(fig:design)).
-Soils were sampled with an auger at 0-10 cm and 20-30 cm.
+In each of the 30 sample plots, three composite samples were taken from three of the five 10 x 10 m subplots, following a transect covering the intra-plot variability and at a minimum distance of 1 m from any tree (Figure\ S\@ref(fig:design)).
+Soils were sampled with a 15cm auger at 0-10 cm and 20-30 cm depth; due to the size of the auger we couldn't accurately sample the intermediate soil layer.
 Soils samples were then air-dried and sieved at 2 mm.
-They were subjected to standard physical and chemical analyses at the Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil)).
+They were subjected to standard physical and chemical analyses at Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil)).
 Among these analyses, total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
@@
 In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm$^3$ at 2-7 cm, 12-17 cm and 22-27 cm.
-Each sample was dried at 105°C in a classic oven until reaching a stable dry mass, sieved to 2 mm, and then the fine fraction (< 2 mm) was weighed.
+Each sample was dried at 105°C in a classic oven until reaching a stable dry mass, sieved to 2 mm, and then the fine fraction (< 2 mm) was weighed to be used in SOC stocks calculation (equation \@ref(eq:soc)).
 
@@
 The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density.
-The volume of the palm trunks was estimated as the volume of a cylinder, using the height and diameter measured in situ, with a volumetric shrinkage of 11% [@WD2007].
+The volume of the palm trunks was estimated as a cylinder, using the height and diameter measured *in situ*, with a volumetric shrinkage of 11% [@WD2007].
 The height of the palms was measured above the leaves and was therefore multiplied by a conversion factor of 0.9 to obtain the trunk height.
@@
 
-Tree-level belowground carbon (BGC) stocks were estimated using the root-to-shoot ratio equation from @Ledo2018, which is based on tree DBH and climatic water deficit from @Chave2014.
-This ratio was then multiplied by the estimated aboveground carbon stock to obtain tree-by-tree belowground carbon stock.
+Belowground carbon (BGC) stocks were estimated at tree level using the root-to-shoot ratio equation from @Ledo2018, which is based on tree DBH and climatic water deficit from @Chave2014.
+This ratio was then multiplied by the estimated AGC stock to obtain tree-by-tree BGC stock.
 
@@
 
-$$SOC_i = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl}$$
+\begin{equation}
+SOC_i = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl} (\#eq:soc)
+\end{equation}
 
@@
 
-All statistical analyses were carried out using R [@R2023].
+All statistical analyses were carried out using the R software [@R2023].
 
@@
 
-The estimated AGC and SOC stocks were log-transformed for all statistical tests, because we expect them to be log-normally distributed as they only take positive values.
+Estimated AGC and SOC stocks were log-transformed for all statistical tests, because we expected them to be log-normally distributed as they only take positive values.
 We removed one CF plot at Gariuai from the following statistical tests, as it did not contain any trees exceeding the minimum measurement size (CBH = 30 cm) and its AGC stocks were therefore estimated at 0\ Mg\ C\ ha$^{-1}$ (even if there were trees < 30 cm CBH) and could not be log-transformed.
@@
 
-The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo() function from the BIOMASS package, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013]; large and small errors on 5 and 95% of trees respectively [@Chave2004;@Rejou-Mechain2017].
-To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (Cirad laboratory in Montpellier, France), 10% for the mass of the fine fraction [@Page1999].
+The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo function from the R package BIOMASS, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013]; large and small errors on 5 and 95% of trees respectively [@Chave2004;@Rejou-Mechain2017].
+To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (as recommended by Cirad Laboratories in Montpellier, France), 10% for the mass of the fine fraction [@Page1999].
 
@@
 
-Total carbon stocks varied between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+Total carbon stocks ranged between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
@@
 
-AGC stocks varied between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table\ \@ref(tab:anova)), but not between site (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
@@
 However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
-The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$) than those in Osso Luga (`r dataC_site[site=="Osso Luga" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
+The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$) than those in Osso-Luga (`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
 
@@
 The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure\ \@ref(fig:plot-lmereg)).
-Overall, AGC stocks were a poor predictor of SOC stocks.
 
@@
 
-In comparison to values reported in the literature for other AFS, the carbon stocks in this study were relatively high.
+Compared to values reported in the literature for other AFS, the carbon stocks in this study were relatively high.
 Specifically, average SOC stocks (`r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$) were higher than other reported values: in a meta-analysis by @Shi2018, it was estimated that AFS stored around 50\ Mg\ C\ ha$^{-1}$ at 0-30 cm depth.
-These large SOC stocks could be due to the high soil pH associated with high exchangeable Ca content at our sites (Table\ \@ref(tab:site-soil)), as Ca is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
-Average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`\ Mg\ C\ ha$^{-1}$), in contrast, were more similar to the average of 74\ Mg\ C\ ha$^{-1}$ estimated by @Ma2020, but some AFS had particularly high biomass values.
-The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`s, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$) similar to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$, @Lal2005].
-The high carbon stocks in FGs can be attributed to their long period since establishment, which is generally more than 50 years, with little human intervention (Table\ S\@ref(tab:afs-typology)).
+Such large SOC stocks could be due to the high soil pH associated with high exchangeable Ca content at our sites (Table\ \@ref(tab:site-soil)), as Ca is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
+Average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`\ Mg\ C\ ha$^{-1}$), in contrast, were similar to the average of 74\ Mg\ C\ ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
+The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`s, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$) close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$, @Lal2005].
+The high carbon stocks in FGs can be attributed to their long period since establishment, which was generally longer than 50 years, with little human intervention (Table\ S\@ref(tab:afs-typology)).
 In terms of structure and carbon stocks, FGs should therefore resemble old secondary forests, which have been shown to recover more than 60% of the AGC stocks and almost all the SOC stocks of an old forest [@Poorter2021].
@@
 
-Although agroforestry has been identified as one of the possible levers to increase carbon storage in the Timorese agricultural landscape [@UNFCCC2022;@Paudel2022], to our knowledge no study has quantified carbon stocks in these systems.
-Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste's climate change mitigation commitments [@UNFCCC2022], and the potential carbon gains from wider adoption of these techniques.
-Further studies should help to increase the representativeness of the AFS sampled by including a greater diversity of AFS and environmental conditions.
-Another critical aspect is the quantification of temporal changes in carbon stocks, which could be assessed in future studies either by having chronosequences of plots where the time since installation is known, or by remeasuring plots through time.
+Although agroforestry has been identified as one of the possible levers to increase carbon storage in the Timorese agricultural landscape [@UNFCCC2020;@Paudel2022], to our knowledge no study has quantified carbon stocks in these systems.
+Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste's climate change mitigation commitments [@UNFCCC2020], and the potential carbon gains from wider adoption of these techniques.
+Further studies should help to increase the representativeness of the AFS sampled by including a greater diversity of AFSs and environmental conditions.
+Another critical aspect is the quantification of temporal changes in carbon stocks, which could be assessed in future studies either by monitoring chronosequences of plots where the time since installation is known, or by remeasuring plots through time.
 
@@
 
-The AFS studied here vary significantly in terms of AGC stocks, in agreement with our hypothesis H1a.
-This reflects the density and size of trees in the plot, which in turn depends on the importance of trees in the production system.
-In some systems, such as CF and SP, trees do not contribute to food production but have other functions, such as restoring soil fertility (e.g.
+The studied AFSs were found to differ significantly in terms of AGC stocks, in agreement with our hypothesis H1a.
+This result reflects the density and size of trees in the plot, which in turn depends on the importance of trees in the production system.
+In some systems, such as CF and SP, trees do not contribute directly to food production but have other functions, such as restoring soil fertility (e.g.
 during the fallow period) or providing shade for livestock.
-This more limited role for trees may explain why their total biomass is lower in these systems: in the CF system, the plot is regularly cleared and annual crops are planted in full sun, while in the SP system, the growth of grazing grasses is favored by reducing the number of trees that shade them.
+This more limited role for trees may explain why their total biomass is lower in these systems: in the CF system, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun, while in the SP system, the growth of grazing grasses is favored by reducing the number of trees.
 
@@
 For example, a farmer might plant trees in a CF field, which will then become a YA that can still produce annual crops for the first 2 to 4 years, before the shade from the trees becomes too great.
-The trees are then used as property markers and, when they reach maturity, to produce wood and fruit: the YA then becomes a HG or a FG.
-Likewise, our results show that AGC stocks increase from CF to YA, HG and FG (Table\ \@ref(tab:summary)).
+The trees are then used as property markers and, as they mature, produce wood and fruits: the YA then becomes HG or FG.
+Similarly, our results show that AGC stocks increase from CF to YA, HG and FG (Table\ \@ref(tab:summary)).
 
 The location of AFS in the landscape is not random and could also explain the observed differences in AGC stocks.
-For example, most HGs are located close to houses, in alluvial valleys that are well supplied with water and where the soil is enriched with organic waste and animal manure, which is likely to increase soil fertility [@Palm2001].
+For example, most HGs are located close to houses, in alluvial valleys that are well supplied with water and where soils are enriched with organic waste and animal manure [@Palm2001].
 The same applies to FGs, which are former HGs that have been more or less abandoned, and YAs, where the farmer has chosen the site with the intention of turning it into a HG.
 We therefore expect HGs, FGs and YAs to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuer2018].
-Conversely, CF and SP are mainly located on limestone plateaus where water is deep and soils are shallow, which could reduce tree growth and biomass [@Imada2008].
+Conversely, CF and SP are mainly located on limestone plateaus where water tables are deeper and soils are shallower, which could reduce tree growth and biomass [@Imada2008].
 
@@
 We found no significant difference in SOC between AFS (H1b rejected).
-This result is inconsistent with our expectation that agroforestry practices would influence SOC stocks [@Minasny2017;@Wiesmeier2019] and with previous studies showing significant differences in SOC stocks between different AFSs [e.g., @Ramesh2015].
+This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017;@Wiesmeier2019], and with previous studies that show significant differences in SOC stocks between AFSs [e.g., @Ramesh2015].
 However, it is worth noting that some single-site studies [e.g.
-@Betemariyam2020;@GamaRodrigues2010] and at least two meta-analyses failed to find a strong effect of AFS on SOC stocks: @Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by AFS selection, and @Shi2018 found no significant differences in SOC stocks between 4 different AFSs in 354 previous studies.
-We offer three main explanations to reconcile these differences between studies.
+@Betemariyam2020;@GamaRodrigues2010] and at least two meta-analyses failed to find any sizeable effect of AFS on SOC stocks: @Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by the type of AFS, and @Shi2018 found no significant differences in SOC stocks between 4 different AFSs in 354 previous studies.
+We propose three main explanations to reconcile these different findings.
 First, in a global meta-analysis, @Ma2020 estimated that it takes only 5 years for SOC stocks in tropical AFS to reach an equilibrium value.
-In our study, at least three of the five AFS studied (namely YAs, HGs and FGs) represent different stages of an AFS transition depending on the age of the trees (Figure\ S\@ref(fig:seq-diagram).
-Most of these AFSs are older than 5 years and may have already reached their equilibrium SOC stocks, which could explain why there is little difference between them.
+In our study, at least three of the five AFSs (namely YAs, HGs and FGs) represent different stages of an AFS transition depending on the age of the trees (Figure\ S\@ref(fig:seq-diagram)).
+Most of these AFSs are older than 5 years and may have already reached their equilibrium SOC stocks, which could explain why there was little difference between them.
 Second, we had a limited number of replicates of each AFS within each site, which certainly reduced the power of statistical tests.
 We were only able to identify trends, such as higher SOC in SPs in Gariuai, which were similar to previous studies [@Feliciano2018].
-Increasing the number of replicates for each AFS could have increased the power of the tests and potentially led to more significant differences between AFSs.
-Third, other pedoclimatic and historical factors influence carbon stocks, sometimes to a greater extent than AFS, e.g.
+Increasing the number of replicates for each AFS could have increased the power of the tests and potentially resulted in more significant differences between AFSs.
+Third, other pedoclimatic and historical factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g.
 within-site variability in climate, soil texture (Figures\ S\@ref(fig:clay-corg), S\@ref(fig:altitude-corg)) or previous land use may have significantly influenced SOC stocks [@Wiesmeier2019;@Ma2020].
@@
 
-We found a significant difference in SOC between the two sites: SOC stocks were higher in Gariuai than in Osso Luga.
+We found a significant difference in SOC between the two sites, with higher SOC stocks at Gariuai than at Osso-Luga.
 A priori, three factors that differentiate the sites could explain these differences in SOC stocks: climate (through elevation; Figure\ \@ref(fig:map-sites); Figure\ S\@ref(fig:altitude-corg)), soil texture (Table\ S\@ref(fig:site-soil); Figure\ S\@ref(fig:clay-corg)) and history.
 First, we found that SOC stocks were positively correlated with elevation, both between and within sites (Figure\ S\@ref(fig:altitude-corg)).
-This result is consistent with those of @Raich2006, who found higher SOC at higher elevations in tropical evergreen forests in different countries (e.g.
+This result is consistent with those of @Raich2006, who found higher SOC at higher elevations in tropical evergreen forests in various countries (e.g.
 Congo, Thailand).
@@
 clay or clay + fine silt content; @Zinn2005].
-In our study, we found no correlation between SOC stocks and clay + fine silt content when site and depth were included in the model (Figure\ S\@ref(fig:clay-corg)).
-Furthermore, the Gariuai soils, which had the highest stocks of soil organic carbon (SOC), were also the least clayey.
+In our study, we found no significant correlation between SOC stocks and clay + fine silt content when site and depth were included in the model (Figure\ S\@ref(fig:clay-corg)).
+Furthermore, the Gariuai soils, which had the highest SOC stocks, were also the least clayey.
 This contradicts the typical (positive) relationship between the fine mineral fraction and SOC stocks.
-It is therefore unlikely that the difference in soil organic carbon (SOC) between the two sites can be attributed to soil texture.
+It is therefore unlikely that the difference in SOC stocks between the two sites can be attributed to soil texture.
 Thirdly, history could be a significant driving factor.
 Gariuai was continuously inhabited for over 50 years, whereas Osso-Luga was abandoned for almost three decades.
-The management and care of the AFSs in Gariuai, such as pruning and livestock visits, as well as optimizing the different strata in terms of C-input, could have led to an increase in soil organic carbon compared to abandoned systems [@Lorenz2014].
+The management and care of the AFSs in Gariuai, such as pruning, organic inputs related to the presence of livestock, as well as optimization of the different strata in terms of C input, may have led to an increase in soil organic carbon compared to abandoned systems [@Lorenz2014].
 
@@
 
-We found no significant relationship between AGC and SOC stocks both in the topsoil layer 0-10 cm depth (H2a refuted) and in 0-30 cm depth (H2b confirmed).
+We found no significant relationship between AGC and SOC stocks both in the topsoil layer 0-10 cm depth (H2a refuted) and at 0-30 cm depth (H2b confirmed).
+Overall, AGC stocks were a poor predictor of SOC stocks.
 These results may relate to previous findings that SOC stocks saturate much faster than AGC stocks in tropical AFS.
 For example, in a global meta-analysis,  @Ma2020 estimated that it took only 5 years for SOC stocks in tropical AFS to reach an equilibrium value, but more than three decades for AGC stocks.
-This desynchronisation of SOC and AGC stocks recovery has also been observed in naturally regenerating tropical forests.
+This desynchronization of SOC and AGC stocks recovery has also been observed in naturally regenerating tropical forests.
 In a systematic review, @Martin2013 found that SOC in tropical secondary forests changed little with time after disturbance, while AGB increased.
-@Ojoatre2024 found similar results in a tropical montane secondary forest.
-These results imply that there is little relationship between AGC and SOC once agroforestry systems are more than 5 years old, which is the case in our study for FGs, HGs (> 5 years old) and partially for YAs (2 to 10 years; Table\ S\@ref(tab:afs-typology)).
-This saturation of SOC even with increasing AGC is counterintuitive, as it might be expected that greater AGC would lead to greater litter production and decomposition, some of which would increase SOC.
-However, @Sayer2019, found that SOC did not increase in a lowland tropical forest in Panama after 15 years of an experimentally doubled aboveground litter input.
-Two processes could explain this.
+@Ojoatre2024 found similar results in a secondary tropical mountain forest.
+These results suggest that there is little correlation between AGC and SOC stocks once AFSs are more than 5 years old, which is the case in our study for FGs, HGs (> 5 years old) and partially for YAs (2 to 10 years; Table\ S\@ref(tab:afs-typology)).
+This saturation of SOC stocks even with increasing AGC stocks is counterintuitive, as it might be expected that higher AGC stocks would lead to more litter production and decomposition, some of which would increase SOC stocks.
+However, @Sayer2019, found that SOC stocks did not increase in a lowland tropical forest in Panama after 15 years of an experimentally doubled aboveground litter input.
+Two processes could explain these findings.
 On the one hand, the addition of easily decomposable litter could stimulate microbial decomposition and turnover of old stored SOC [i.e.
@@
 For instance, previous studies have shown that tree diversity can enhance SOC sequestration in natural ecosystems [@Spohn2023;@Chen2018] and in AFSs [@Manaye2021;@Islam2015].
-Moreover, total soil organic carbon may be less sensitive to management practices such as biomass inputs than specific fractions of soil organic carbon [@Haynes2005].
+Moreover, total SOC may be less sensitive to management practices (such as biomass inputs) than specific fractions of SOC [@Haynes2005].
 Therefore, it would be interesting to continue with further studies that focus not only on the quantity but also the quality of soil organic matter and its relationship with biodiversity (flora and fauna) in traditional Timorese AFSs.
@@
 Although agroforestry can make agriculture more resilient to climate change while improving its carbon footprint and biodiversity conservation value [@Cardinael2021], poor adaptation to the local context can explain many failures to adopt new agroforestry techniques [@Coe2014].
-Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and are therefore adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
-This is the case of the traditional AFS studied here: their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally.
-
-Conservation of these traditional AFSs should avoid their abandonment as the trees wither or as their production value declines.
+Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and became more adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
+This is the case of the traditional AFSs studied here in Timor-Leste: their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally.
+
+Conservation of these traditional AFSs should avoid their abandonment as the trees wither or as their productivity declines.
 Clear-cutting old agroforests for full-sun crops may then become an attractive short-term strategy, providing a large cash inflow from the sale of wood products, followed by 4 or 5 years of agricultural production on fertile land.
@@
 
-Strategies to avoid clearing of these traditional AFS could be to develop new ways of intensifying production while maintaining high carbon and biodiversity values, and gradually renewing the tree population in the plots.
-For example, it would be interesting to develop renewal through clearings of about one hundred square meters (corresponding to the felling of 1 to 10 trees, depending on their crown area).
-In these clearings, young trees of species considered interesting for their future production would be associated, first with light crops (cereals, etc.) and then with crops of shady species (taro, etc.).
-Additional mechanisms to encourage the renewal of traditional AFS could be the provision of subsidies (e.g.
+Strategies to avoid clearing of traditional AFSs could involve new ways of intensifying production while maintaining high carbon and biodiversity values, and gradually renewing the tree population in the plots.
+For example, it would be interesting to improve tree regeneration through clearings of about one hundred square meters (corresponding to the felling of 1 to 10 trees, depending on their crown area).
+In these clearings, young trees of species considered interesting for their future production would be associated, first with heliophile crops (e.g.
+cereals) and then with shade-tolerant crops (e.g.
+taro).
+Additional mechanisms to encourage the renewal of traditional AFSs could include the provision of subsidies (e.g.
 payment of carbon credits) or low-interest loans to farmers wishing to renew their agroforestry plots.
 Another strategy, which has been successfully implemented in other tropical countries (e.g.
-for agroforestry coffee), is to improve the incomes of AFS farmers by creating 'sustainable agriculture'-type labels [@Bertrand2019], although this is primarily applicable to the production of export goods, whereas the AFSs we have studied are currently oriented more towards local consumption and domestic markets.
+for agroforestry coffee), is to improve the incomes of AFS farmers through 'sustainable agriculture' certification [@Bertrand2019], although this is primarily applicable to the production of export goods, whereas the AFSs studied here are currently oriented towards local consumption and domestic markets.
 
@@
 
-Agroforestry systems in the Baucau district of Timor had variable but overall high carbon stocks.
-These systems have only recently been described in the scientific literature, and much remains to be done to improve our knowledge of their diversity and functioning.
-It is therefore important that all stakeholders including local authorities, universities and NGOs, take into account their existence, diversity and environmental and social value, in order to propose possible innovations to maintain or improve their functioning.
+Agroforestry systems in the Baucau municipality of Timor-Leste had variable but overall high carbon stocks.
+These systems have only recently been described in the scientific literature, and more research is needed to improve our knowledge of their diversity and functioning.
+It is therefore important that all stakeholders, including local authorities, universities and NGOs, are aware of their existence, diversity and environmental and social value, in order to propose possible innovations to maintain or improve their functioning.
 
@@
 
-We are grateful to Marcal Gusmao (National University of Timor-Leste) and Alain Rival (Cirad) for their facilitation of the field study, and to the trainees who helped with field data collection: Juvencio Dos Santos (vegetation) and Teofilio Gomes (soil).
+We are grateful to Alain Rival (Cirad) for his facilitation of the field studies and critical revision of the paper, and to the trainees who helped with field data collection: Juvencio Dos Santos (vegetation) and Teofilio Gomes (soil).
 We would also like to thank the 31 farmers who gave up their time to take part in the interviews and field measurements.
@@
 
-```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (between parentheses) of total, aboveground carbon (AGC), belowground carbon (BGC), and soil organic carbon (SOC) stocks in each agroforestry system."}
+```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (in parentheses) of total, aboveground carbon (AGC), belowground carbon (BGC), and soil organic carbon (SOC) stocks in each agroforestry system."}
 # calculate median and 95% confidence interval at site level
@@
 ```{r fig.id = "map-sites", fig.height = 5, fig.width = 7, fig.cap="Plot location.
-(a) Map of Timor-Leste with the location of the two study sites (Gariuai and Osso-Luga) in the Bacau district.
-(b) Location of the 15 plots in Gariuai; (c) Location of the 15 plots in Osso-Luga.
+(a) Map of Timor-Leste showing the location of the two study sites (Gariuai and Osso-Luga) in the Baucau municipality (b) Location of the 15 plots in Gariuai; (c) Location of the 15 plots in Osso-Luga.
 Panels (b) and (c) share the same spatial scale.
@@
 data_box[site == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
-data_box[site == "Osso Luga", w := 0.05]
+data_box[site == "Osso-Luga", w := 0.05]
 data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
@@
 elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
-elevS <- subset(elev, x > data_box[site == "Osso Luga"]$x1 & 
-                  x < data_box[site == "Osso Luga"]$x2 & 
-                  y > data_box[site == "Osso Luga"]$y1 & 
-                  y < data_box[site == "Osso Luga"]$y2)
-elevS$site = "Osso Luga"
+elevS <- subset(elev, x > data_box[site == "Osso-Luga"]$x1 & 
+                  x < data_box[site == "Osso-Luga"]$x2 & 
+                  y > data_box[site == "Osso-Luga"]$y1 & 
+                  y < data_box[site == "Osso-Luga"]$y2)
+elevS$site = "Osso-Luga"
 elevG <- subset(elev, x > data_box[site == "Gariuai"]$x1 & 
@@
 
-```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Barplot of the amount of carbon in each AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden), per site.
+```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Distribution of carbon stocks in each type of AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden), per site.
 AGC: Aboveground (biomass) carbon; BGC: belowground (biomass) carbon; SOC: soil organic carbon."}
@@
 
-```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (depth of 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
+```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (average over 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
 The confidence intervals were calculated by propagating a 12% error on the lab-provided values and obtaining site-level estimates through bootstrapping plot values (1000 simulations).
@@
 
-```{r tab.id = "tetum-names", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Correspondence between tetum names and scientific names used in this study."}
+```{r tab.id = "tetum-names", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Correspondence between Tetum names and scientific names used in this study."}
 flextable(vern) |> 
@@
 data_soil[, site := substr(ID_VILLAGE, 1, 1)]
-data_soil[site == "S", site := "Osso Luga"]
+data_soil[site == "S", site := "Osso-Luga"]
 data_soil[site == "G", site := "Gariuai"]

### 2024-05-13 manuscript.Rmd
add gdl's corrections

@@
 title: "Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass"
-author: "Camille Piponiot*, Marguerite Cogné*, Vincent Freycon, Alexis Thoumazeau, Marçal Gusmão, Régis Peltier"
-date: "2024-03-19"
+author: "Camille Piponiot^*^, Marguerite Cogné^*^, Vincent Freycon, Alexis Thoumazeau, Marçal Gusmão, Régis Peltier"
+date: ""
 output: 
@@
 Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their inclusion in climate change mitigation strategies.
-In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFSs) in Timor-Leste, namely: cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
-Our results show that these traditional AFSs can store large amounts of carbon: they stored an average of 156\ Mg\ C\ ha$^{-1}$, and the AFSs with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, close to the values of old-growth tropical forests.
+In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste, namely cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
+Our results show that these traditional AFS can store large amounts of carbon, with the average being 156\ Mg\ C\ ha$^{-1}$.
+The AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, close to the values of old-growth tropical forests.
 Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
 We found no relationship between the amount of carbon stored in biomass and soil.
-Our results highlight the high diversity of traditional AFSs in Timor-Leste and their high carbon sequestration capacity.
-These results could provide an important baseline for the inclusion of AFS in Timor-Leste's climate change mitigation strategy, and could serve as a reference for future AFS studies in different agro-climates of Timor-Leste.
+Our results highlight the high diversity of traditional AFS in Timor-Leste and their high carbon sequestration capacity.
+These results could provide an important baseline for the inclusion of AFS in Timor-Leste’s climate change mitigation strategy, and could serve as a reference for future AFS studies in different agro-climates of Timor-Leste.
 
@@
 Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO~2~e/yr over 2010-2014; @Pendrill2019].
-The demand for tropical agricultural land is expected to continue to grow as population increases [@UN2022].
+The demand for tropical agricultural land is expected to continue to grow as the world’s population increases [@UN2022].
 In addition, tropical areas are expected to be disproportionately affected by climate change, putting their agriculture and food systems at risk [@Lawrence2015].
-There is therefore an urgent need to promote tropical agricultural landscapes that are both more resilient to climate change, while maintaining high productivity and ecological value [@Harvey2014].
-In particular, increasing the amount of carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of the food production system [@Lal2004;@Minasny2017].
+There is therefore an urgent need to promote tropical agricultural landscapes that are more resilient to climate change while maintaining high productivity and ecological value [@Harvey2014].
+In particular, increasing the amount of carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of food production systems [@Lal2004;@Minasny2017].
 
@@
 Agroforestry is a collective name for land use systems and technologies where woody perennials (trees, shrubs, bamboos, creepers, etc.) are deliberately used on the same land management units as agricultural crops and/or animals, in some form of spatial arrangement or temporal sequence [@Lundgren1982].
-Agroforestry allows for diversification of agricultural production, making agricultural systems more resilient [@Duffy2021;@Terasaki2023].
-
-Agroforestry systems (AFSs) have the potential to store large amounts of carbon in the vegetation and soil, and this carbon storage capacity depends on the characteristics of AFS considered [@Albrecht2003;@Feliciano2018].
-As trees can store large amounts of carbon, the carbon stored in the vegetation of AFSs is largely dependent on the number and size of trees in the system [@Ma2020].
-However, the effect of AFSs on soil organic carbon (SOC) has been relatively little studied [@Beillouin2021], is more variable and depends on a number of factors.
-Although AFSs appear to increase SOC stocks overall, some AFS such as woodlots (i.e.
-planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFSs, while AFSs such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
-Other factors such as climate and previous land use type can also influence SOC in AFSs, which may explain the variability of observations reported in the literature [@Feliciano2018;@Ma2020;@Martin2020].
+Agroforestry allows for the diversification of agricultural production, making agricultural systems more resilient [@Duffy2021;@Terasaki2023].
+
+Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation and soil, and this carbon storage capacity depends on the characteristics of the AFS considered [@Albrecht2003;@Feliciano2018].
+As trees can store large amounts of carbon, the carbon stored in the vegetation of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
+However, the effect of AFS on soil organic carbon (SOC) has been relatively little studied [@Beillouin2021], is more variable and depends on several factors.
+Although AFS appear to increase SOC stocks overall, some AFS such as woodlots (i.e., planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFS, while AFS such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
+Other factors such as climate and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018;@Ma2020;@Martin2020].
 Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS could only be evidenced for the topsoil layer [< 10 cm; @Ma2020].
 
-In Timor-Leste, a large proportion of the population depends on agriculture for their livelihoods, and food security is threatened by population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime [@WB2009;@Molyneux2012].
-In this context, productive, resilient and sustainable agriculture is critical for the country, and agroforestry has been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022;@Cogne2024].
-Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024] and can be key elements for biodiversity conservation [@Perfecto2008;@Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFSs.
-A precise quantification of these carbon stocks would improve the understanding of the ecological importance of these AFSs at both local and global scales, and help to integrate them into national and international climate change mitigation initiatives.
-
-In this study, we aim to better understand traditional AFSs in Timor-Leste by quantifying the carbon storage in both biomass and soil of five traditional AFSs: cropping systems with fallow, silvopastures, young agroforests, home gardens and forest gardens.
-We hypothesize that different types of traditional AFSs will exhibit measurable differences in aboveground carbon (AGC; H1a) and SOC (H1b) stocks.
-We anticipate, following @Ma2020, that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; H2a), but no relationship with SOC below 10 cm (H2b).
+In Timor-Leste, a large proportion of the population depends on agriculture for their livelihoods.
+However, population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime are threatening food security [@WB2009;@Molyneux2012].
+In this context, productive, resilient, and sustainable agriculture is critical for the country, and agroforestry has been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022;@Cogne2024].
+Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024] and can be key elements for biodiversity conservation [@Perfecto2008;@Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
+A precise quantification of these carbon stocks would improve understanding of the ecological importance of these AFS at both local and global scales, and help integrate them into national and international climate change mitigation initiatives.
+
+In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying the carbon storage in both biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
+We hypothesize that there are measurable differences in aboveground carbon (AGC; H1a) and SOC (H1b) stocks between these different types of traditional AFS.
+In line with Ma et al.
+(2020), we anticipate that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; H2a), but no relationship with SOC below 10 cm (H2b).
 
@@
   prec = round(mean(prec)), 
-  elev = paste(round(quantile(elev, c(0.025, 0.975))), collapse = " - ")),
+  elev = paste(round(quantile(elev, c(0.025, 0.975))/10)*10, collapse = " and ")),
   .(site)]
@@
 
-The research was carried out in the center of the Baucau municipality, on the north-eastern coast of Timor-Leste.
-Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m above sea level (asl); and (ii) the village of Osso-Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso-Luga", coord]`) located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso-Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso-Luga", elev]` m asl (Figure\ \@ref(fig:map-sites)).
+The research was carried out in the center of the Baucau municipality, located on the north-eastern coast of Timor-Leste.
+Two communities, or *suco* (a group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases.
+The first, Gariuai *suco* (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`), is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging between `r site_summary[site=="Gariuai", elev]` m above sea level (asl).
+The second, the village of Osso-Luga in Samalari *suco* (long-lat coordinates: `r site_summary[site=="Osso-Luga", coord]`), is located on the Matebian foothills based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso-Luga", prec]` mm and an altitude ranging between `r site_summary[site=="Osso-Luga", elev]` m asl (Figure\ \@ref(fig:map-sites)).
 The soils of the Baucau municipality are classified using the WRB soil classification [@Mantel2023] as Lithosols and Calcic Luvisols [@FAO2012].
-The average soil properties differed between villages (Table\ S\@ref(tab:site-soil)): Gariuai had a clay loam texture whereas Osso-Luga had a clay texture.
-The two villages also have different recent histories: Gariuai has been inhabited for at least 50 years, while Osso-Luga was abandoned during the Indonesian occupation and was only repopulated after Timor-Leste gained independence in 2002.
-During the period of abandonment, the AFSs in Osso-Luga lay fallow.
+The average soil properties differ between the two sites (Table\ S\@ref(tab:site-soil)); Gariuai has a clay loam texture whereas Osso-Luga has a clay texture.
+The two villages also have different recent histories.
+Gariuai has been inhabited for at least 50 years, while Osso-Luga was abandoned during the Indonesian occupation and was only repopulated after Timor-Leste gained independence in 2002.
+The AFS in Osso-Luga lay fallow during the period of abandonment.
 
@@
 (1993, 2021) and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
-The five AFSs identified were: Crop system including a wooded Fallow phase (CF), Silvopasture (SP), Young Agroforest (YA), Home Garden (HG) and Forest Garden (FG).
-CFs are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts and squash) which are left fallow after 1 to 3 years.
-SPs consist in a vast area (generally 200 to 500 ha) where cattle are grazed under a layer of trees.
-YAs are agricultural plots where the farmer has been planting rows of trees, spaced at around 5x10m, for two to 10 years.
-HGs are plots of land located close to houses, where crops and livestock (poultry, pigs, cattle, etc.) have been grown in the shade of numerous trees and palms for multiple uses (wood, fruit, fibre, etc.) for more than five years.
-FGs are generally former HGs (over 50 years old), which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
-
-The AFSs are presented in more detail as supplemental information (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
+The five AFS identified were as follows: Crop systems including a wooded Fallow phase (CF), Silvopastures (SP), Young Agroforests (YA), Home Gardens (HG), and Forest Gardens (FG).
+CF are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts, and squash) which are left fallow after one to three years.
+SP are vast areas (generally 200 to 500 ha) where cattle are grazed under the shade of trees.
+YA are agricultural plots where farmers plant rows of trees spaced approximately 5 x 10 m apart, which they maintain for two to ten years.
+HG are plots of land located close to houses where crops and livestock (poultry, pigs, cattle, etc.) are kept in the shade of numerous trees and palms for multiple uses (wood, fruit, fiber, etc.).
+FG are generally former HG (over 50 years old) which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
+
+The AFS are presented in more detail in the supplemental information section (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
 
@@
 
-We randomly selected 3 plots from each of the 5 traditional AFSs identified in each of the two sites (30 plots in total; Figure\ \@ref(fig:map-sites)).
-Sampling plots were defined as the entire agricultural plot (area between 0.06 and 1.25 ha) for all AFSs except SP plots, which were generally much larger than 1 ha.
+We randomly selected three plots from each of the five traditional AFS identified in each of the two sites (30 plots in total; Figure\ \@ref(fig:map-sites)).
+Sampling plots were defined as the entire agricultural plot (area between 0.06 and 1.25 ha) for all AFS except SP plots, which were generally much larger than 1 ha.
 In SP plots, a 1 ha sampling plot was delineated within the agricultural plot.
@@
 Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
-Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Figure\ S\@ref(fig:design)): within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
-Height measurements were done with a dendrometer.
-
-Taxonomic identification was carried out using Tetum  (a national official language) vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory government's Native Plants Herbarium [@NT2013].
+Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Figure\ S\@ref(fig:design)).
+Within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
+Height measurements were taken with a dendrometer.
+
+Taxonomic identification was carried out using Tetum  (a national official language) vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory Government's Native Plants Herbarium [@NT2013].
 The correspondence between Tetum names and scientific names is provided in the supplemental material (Table S\@ref(tab:tetum-names)).
@@
 In each of the 30 sample plots, three composite samples were taken from three of the five 10 x 10 m subplots, following a transect covering the intra-plot variability and at a minimum distance of 1 m from any tree (Figure\ S\@ref(fig:design)).
-Soils were sampled with a 15cm auger at 0-10 cm and 20-30 cm depth; due to the size of the auger we couldn't accurately sample the intermediate soil layer.
+Soils were sampled with a 15 cm auger at 0-10 cm and 20-30 cm depth; due to the size of the auger we could not accurately sample the intermediate soil layer.
 Soils samples were then air-dried and sieved at 2 mm.
-They were subjected to standard physical and chemical analyses at Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil)).
-Among these analyses, total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
-
-In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm$^3$ at 2-7 cm, 12-17 cm and 22-27 cm.
-Each sample was dried at 105°C in a classic oven until reaching a stable dry mass, sieved to 2 mm, and then the fine fraction (< 2 mm) was weighed to be used in SOC stocks calculation (equation \@ref(eq:soc)).
+They underwent standard physical and chemical analyses at Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil)).
+Total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
+
+In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm3 at 2-7 cm, 12-17 cm and 22-27 cm.
+Each sample was dried at 105°C in a classic oven until a stable dry mass was reached and then sieved to 2 mm.
+The fine fraction (< 2 mm) was weighed to be used in the SOC stocks calculation (equation \@ref(eq:soc)).
 
@@
 
-We estimated tree-level AGC stocks from the measured tree CBH and height using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017].
+We estimated tree-level AGC stocks from the tree CBH and height measurements using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017].
 Each tree was assigned a mean and standard deviation of wood density at the species or genus level when at least one value was available from the Global Wood Density Database [@Zanne2009].
@@
 
-Estimated AGC and SOC stocks were log-transformed for all statistical tests, because we expected them to be log-normally distributed as they only take positive values.
+The estimated AGC and SOC stocks were log-transformed for all statistical tests because, as they only take positive values, we expected them to be log-normally distributed.
 We removed one CF plot at Gariuai from the following statistical tests, as it did not contain any trees exceeding the minimum measurement size (CBH = 30 cm) and its AGC stocks were therefore estimated at 0\ Mg\ C\ ha$^{-1}$ (even if there were trees < 30 cm CBH) and could not be log-transformed.
@@
 
-The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo function from the R package BIOMASS, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013]; large and small errors on 5 and 95% of trees respectively [@Chave2004;@Rejou-Mechain2017].
-To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (as recommended by Cirad Laboratories in Montpellier, France), 10% for the mass of the fine fraction [@Page1999].
+The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo function from the R package BIOMASS, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013] and large and small errors for 5 and 95% of trees, respectively [@Chave2004;@Rejou-Mechain2017].
+To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (as recommended by Cirad laboratories in Montpellier, France), and 10% for the mass of the fine fraction [@Page1999].
 
@@
 Total carbon stocks ranged between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-The AFS with the lowest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks was the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+The AFS with the lowest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 
@@
 AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table\ \@ref(tab:anova)), but not between site (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
-The AFS with the lowest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest AGC stocks was the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table\ \@ref(tab:anova)), but not between sites (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
+The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
 
@@
 This relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
-The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure\ \@ref(fig:plot-lmereg)).
+The conditional R^2^ (i.e., the proportion of variance explained by the linear mixed models) were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths, respectively, but the marginal R^2^ (i.e., when considering only the fixed effect) were only `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`%, respectively (Figure\ \@ref(fig:plot-lmereg)).
 
@@
 
-## Traditional AFS in Timor Leste can store large C stocks
+## Traditional AFS in Timor-Leste can store large C stocks
 
 Compared to values reported in the literature for other AFS, the carbon stocks in this study were relatively high.
-Specifically, average SOC stocks (`r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$) were higher than other reported values: in a meta-analysis by @Shi2018, it was estimated that AFS stored around 50\ Mg\ C\ ha$^{-1}$ at 0-30 cm depth.
+Specifically, the average SOC stocks (`r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$) were higher than other reported values.
+In a meta-analysis by @Shi2018, it was estimated that AFS stored around 50\ Mg\ C\ ha$^{-1}$ at a 0-30 cm depth.
 Such large SOC stocks could be due to the high soil pH associated with high exchangeable Ca content at our sites (Table\ \@ref(tab:site-soil)), as Ca is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
-Average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`\ Mg\ C\ ha$^{-1}$), in contrast, were similar to the average of 74\ Mg\ C\ ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
-The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`s, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$) close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$, @Lal2005].
-The high carbon stocks in FGs can be attributed to their long period since establishment, which was generally longer than 50 years, with little human intervention (Table\ S\@ref(tab:afs-typology)).
-In terms of structure and carbon stocks, FGs should therefore resemble old secondary forests, which have been shown to recover more than 60% of the AGC stocks and almost all the SOC stocks of an old forest [@Poorter2021].
-In addition, FGs contain a large number of trees of the same age that were planted when the AFS was established and can survive to reach significant sizes, similar to forest plantations that can grow to old-growth-forest levels of AGC stocks after only a few decades [@Brown2020;@Njoukam1996].
+In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`\ Mg\ C\ ha$^{-1}$) were similar to the average of 74\ Mg\ C\ ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
+The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$) close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$, @Lal2005].
+The high carbon stocks in FG can be attributed to the length of time they had been established, which was generally longer than 50 years, with little human intervention (Table\ S\@ref(tab:afs-typology)).
+In terms of structure and carbon stocks, FG should therefore resemble old secondary forests, which have been shown to recover more than 60% of the AGC stocks and almost all the SOC stocks of an old forest [@Poorter2021].
+In addition, FG contain a large number of trees of the same age that were planted when the AFS was established and can survive to reach significant sizes, similar to forest plantations that can grow to old-growth-forest levels of AGC stocks after only a few decades [@Brown2020;@Njoukam1996].
 
 Although agroforestry has been identified as one of the possible levers to increase carbon storage in the Timorese agricultural landscape [@UNFCCC2020;@Paudel2022], to our knowledge no study has quantified carbon stocks in these systems.
-Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste's climate change mitigation commitments [@UNFCCC2020], and the potential carbon gains from wider adoption of these techniques.
-Further studies should help to increase the representativeness of the AFS sampled by including a greater diversity of AFSs and environmental conditions.
-Another critical aspect is the quantification of temporal changes in carbon stocks, which could be assessed in future studies either by monitoring chronosequences of plots where the time since installation is known, or by remeasuring plots through time.
+Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste's climate change mitigation commitments [@UNFCCC2020] and the potential carbon gains from a wider adoption of these techniques.
+Further studies should help to increase the representativeness of the AFS sampled by including a greater diversity of AFS and environmental conditions.
+Another critical aspect is the quantification of temporal changes in carbon stocks, which could be assessed in future studies either by monitoring chronosequences of plots where the time since installation is known, or by remeasuring plots over time.
 
@@
 
-The studied AFSs were found to differ significantly in terms of AGC stocks, in agreement with our hypothesis H1a.
-This result reflects the density and size of trees in the plot, which in turn depends on the importance of trees in the production system.
-In some systems, such as CF and SP, trees do not contribute directly to food production but have other functions, such as restoring soil fertility (e.g.
-during the fallow period) or providing shade for livestock.
-This more limited role for trees may explain why their total biomass is lower in these systems: in the CF system, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun, while in the SP system, the growth of grazing grasses is favored by reducing the number of trees.
+The AFS studied were found to differ significantly in terms of AGC stocks, in agreement with our hypothesis H1a.
+This result reflects the density and size of trees in the plot, which in turn depend on the importance of trees in the production system.
+In some systems, such as CF and SP, trees do not contribute directly to food production but have other functions, such as restoring soil fertility (e.g., during the fallow period) or providing shade for livestock.
+This more limited role for trees may explain why their total biomass is lower in these systems.
+In CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun, while in SP systems, the growth of grazing grasses is favored by reducing the number of trees.
 
 It is also interesting to note that, in line with previous results, the age of the trees in the AFS largely explains the carbon stored in their biomass [@Ma2020].
-Indeed, the systems described in this study reflect different stages of transition in land use, which follow the growth of their trees (Figure\ S\@ref(fig:seq-diagram)).
+The AFS described in this study reflect different stages of transition in land use, which follow the growth of their trees (Figure\ S\@ref(fig:seq-diagram)).
 For example, a farmer might plant trees in a CF field, which will then become a YA that can still produce annual crops for the first 2 to 4 years, before the shade from the trees becomes too great.
-The trees are then used as property markers and, as they mature, produce wood and fruits: the YA then becomes HG or FG.
+The trees are then used as property markers and, as they mature, produce wood and fruits; the YA then becomes an HG or FG.
 Similarly, our results show that AGC stocks increase from CF to YA, HG and FG (Table\ \@ref(tab:summary)).
 
-The location of AFS in the landscape is not random and could also explain the observed differences in AGC stocks.
-For example, most HGs are located close to houses, in alluvial valleys that are well supplied with water and where soils are enriched with organic waste and animal manure [@Palm2001].
-The same applies to FGs, which are former HGs that have been more or less abandoned, and YAs, where the farmer has chosen the site with the intention of turning it into a HG.
-We therefore expect HGs, FGs and YAs to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuer2018].
+The location of AFS in the landscape is not random, and could also explain the observed differences in AGC stocks.
+For example, most HG are located close to houses, in alluvial valleys that are well supplied with water and where soils are enriched with organic waste and animal manure [@Palm2001].
+The same applies to FG, which are former HG that have been more or less abandoned, and YA, where the farmer has chosen the site with the intention of turning it into a HG.
+We therefore expect HG, FG and YA to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuer2018].
 Conversely, CF and SP are mainly located on limestone plateaus where water tables are deeper and soils are shallower, which could reduce tree growth and biomass [@Imada2008].
@@
 
-We found no significant difference in SOC between AFS (H1b rejected).
-This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017;@Wiesmeier2019], and with previous studies that show significant differences in SOC stocks between AFSs [e.g., @Ramesh2015].
-However, it is worth noting that some single-site studies [e.g.
-@Betemariyam2020;@GamaRodrigues2010] and at least two meta-analyses failed to find any sizeable effect of AFS on SOC stocks: @Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by the type of AFS, and @Shi2018 found no significant differences in SOC stocks between 4 different AFSs in 354 previous studies.
+We found no significant difference in SOC between the AFS (H1b rejected).
+This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017;@Wiesmeier2019], and with previous studies that show significant differences in SOC stocks between AFS [e.g., @Ramesh2015].
+However, it is worth noting that some single-site studies [e.g., @Betemariyam2020;@GamaRodrigues2010] and at least two meta-analyses failed to find any sizable effect of AFS on SOC stocks.
+@Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by the type of AFS, and @Shi2018 found no significant differences in SOC stocks between four different AFS in 354 previous studies.
 We propose three main explanations to reconcile these different findings.
-First, in a global meta-analysis, @Ma2020 estimated that it takes only 5 years for SOC stocks in tropical AFS to reach an equilibrium value.
-In our study, at least three of the five AFSs (namely YAs, HGs and FGs) represent different stages of an AFS transition depending on the age of the trees (Figure\ S\@ref(fig:seq-diagram)).
-Most of these AFSs are older than 5 years and may have already reached their equilibrium SOC stocks, which could explain why there was little difference between them.
-Second, we had a limited number of replicates of each AFS within each site, which certainly reduced the power of statistical tests.
-We were only able to identify trends, such as higher SOC in SPs in Gariuai, which were similar to previous studies [@Feliciano2018].
-Increasing the number of replicates for each AFS could have increased the power of the tests and potentially resulted in more significant differences between AFSs.
-Third, other pedoclimatic and historical factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g.
-within-site variability in climate, soil texture (Figures\ S\@ref(fig:clay-corg), S\@ref(fig:altitude-corg)) or previous land use may have significantly influenced SOC stocks [@Wiesmeier2019;@Ma2020].
-Our experimental design aimed to limit as much as possible the influence of external co-factors at each site (e.g.
-variability in farmers' management practices, micro-topography, etc.), but much of the variability cannot be controlled in farmers' plots.
+First, in a global meta-analysis, @Ma2020 estimated that it takes only five years for SOC stocks in tropical AFS to reach an equilibrium value.
+In our study, at least three of the five AFS (namely YA, HG and FG) represent different stages of an AFS transition depending on the age of the trees (Figure\ S\@ref(fig:seq-diagram)).
+Most of these AFS are more than five years old and may have already reached their equilibrium SOC stocks, which could explain why there was little difference between them.
+Second, we had a limited number of replicates of each AFS within each site, which certainly reduced the power of the statistical tests.
+We were only able to identify trends, such as higher SOC in SP in Gariuai, which were similar to previous studies [@Feliciano2018].
+Increasing the number of replicates for each AFS could have increased the power of the tests, and potentially resulted in more significant differences between AFS.
+Third, other pedoclimatic and historical factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g., within-site variability in climate, soil texture (Figures\ S\@ref(fig:clay-corg), S\@ref(fig:altitude-corg)) or previous land use may have significantly influenced SOC stocks [@Wiesmeier2019;@Ma2020].
+Our experimental design aimed to limit as much as possible the influence of external co-factors at each site (e.g., variability in farmers' management practices, micro-topography, etc.), but much of the variability cannot be controlled in farmers' plots.
 
@@
 First, we found that SOC stocks were positively correlated with elevation, both between and within sites (Figure\ S\@ref(fig:altitude-corg)).
-This result is consistent with those of @Raich2006, who found higher SOC at higher elevations in tropical evergreen forests in various countries (e.g.
-Congo, Thailand).
+This result is consistent with those of @Raich2006, who found higher SOC at higher elevations in tropical evergreen forests in various countries (e.g., Congo, Thailand).
 As changes in elevation are associated with variations in climate (precipitation, temperature), our result is also consistent with previous studies that found higher SOC in areas with higher precipitation [@Schuur2001] and lower temperature [@Raich2006;@MarinSpiotta2013].
 Second, soil texture is a key driving process for SOC storage in the tropics [@Don2010;@Wiesmeier2019;@Matus2021].
-In particular, many studies have found a strong positive correlation between SOC stocks and the fine mineral fractions of soils [i.e.
-clay or clay + fine silt content; @Zinn2005].
+In particular, many studies have found a strong positive correlation between SOC stocks and the fine mineral fractions of soils [i.e., clay or clay + fine silt content; @Zinn2005].
 In our study, we found no significant correlation between SOC stocks and clay + fine silt content when site and depth were included in the model (Figure\ S\@ref(fig:clay-corg)).
@@
 Gariuai was continuously inhabited for over 50 years, whereas Osso-Luga was abandoned for almost three decades.
-The management and care of the AFSs in Gariuai, such as pruning, organic inputs related to the presence of livestock, as well as optimization of the different strata in terms of C input, may have led to an increase in soil organic carbon compared to abandoned systems [@Lorenz2014].
+The management and care of the AFS in Gariuai, such as pruning, organic inputs related to the presence of livestock, as well as optimization of the different strata in terms of C input, may have led to an increase in soil organic carbon compared to abandoned systems [@Lorenz2014].
 
@@
 
-We found no significant relationship between AGC and SOC stocks both in the topsoil layer 0-10 cm depth (H2a refuted) and at 0-30 cm depth (H2b confirmed).
+We found no significant relationship between AGC and SOC stocks in either the topsoil layer 0-10 cm depth (H2a refuted) or at 0-30 cm depth (H2b confirmed).
 Overall, AGC stocks were a poor predictor of SOC stocks.
 These results may relate to previous findings that SOC stocks saturate much faster than AGC stocks in tropical AFS.
-For example, in a global meta-analysis,  @Ma2020 estimated that it took only 5 years for SOC stocks in tropical AFS to reach an equilibrium value, but more than three decades for AGC stocks.
+For example, in a global meta-analysis, @Ma2020 estimated that it took only five years for SOC stocks in tropical AFS to reach an equilibrium value, but more than three decades for AGC stocks.
 This desynchronization of SOC and AGC stocks recovery has also been observed in naturally regenerating tropical forests.
-In a systematic review, @Martin2013 found that SOC in tropical secondary forests changed little with time after disturbance, while AGB increased.
+In a systematic review, @Martin2013 found that SOC in tropical secondary forests changed little with time after disturbance, while AGC increased.
 @Ojoatre2024 found similar results in a secondary tropical mountain forest.
-These results suggest that there is little correlation between AGC and SOC stocks once AFSs are more than 5 years old, which is the case in our study for FGs, HGs (> 5 years old) and partially for YAs (2 to 10 years; Table\ S\@ref(tab:afs-typology)).
+These results suggest that there is little correlation between AGC and SOC stocks once AFS are more than five years old, which is the case in our study for FG, HG (> 5 years old) and partially for YA (2 to 10 years; Table\ S\@ref(tab:afs-typology)).
 This saturation of SOC stocks even with increasing AGC stocks is counterintuitive, as it might be expected that higher AGC stocks would lead to more litter production and decomposition, some of which would increase SOC stocks.
@@
 Two processes could explain these findings.
-On the one hand, the addition of easily decomposable litter could stimulate microbial decomposition and turnover of old stored SOC [i.e.
-priming effect; @Sayer2019].
-On the other hand, the capacity of the soil to store SOC through organo-mineral associations would be limited by its fine element content (clay + fine silt) (i.e.
-saturation concept).
-
-In addition to biomass stocks, other vegetation characteristics may account for variations in soil carbon stocks between AFSs.
-For instance, previous studies have shown that tree diversity can enhance SOC sequestration in natural ecosystems [@Spohn2023;@Chen2018] and in AFSs [@Manaye2021;@Islam2015].
+On the one hand, the addition of easily decomposable litter could stimulate microbial decomposition and turnover of old stored SOC [i.e., priming effect; @Sayer2019].
+On the other hand, the capacity of the soil to store SOC through organo-mineral associations would be limited by its fine element content (clay + fine silt) (i.e., saturation concept).
+
+In addition to biomass stocks, other vegetation characteristics may account for variations in soil carbon stocks between AFS.
+For instance, previous studies have shown that tree diversity can enhance SOC sequestration in natural ecosystems [@Spohn2023;@Chen2018] and in AFS [@Manaye2021;@Islam2015].
 Moreover, total SOC may be less sensitive to management practices (such as biomass inputs) than specific fractions of SOC [@Haynes2005].
-Therefore, it would be interesting to continue with further studies that focus not only on the quantity but also the quality of soil organic matter and its relationship with biodiversity (flora and fauna) in traditional Timorese AFSs.
-
-## How to conserve and improve carbon stocks in traditional AFS in Timor Leste?
+Therefore, it would be interesting to continue with further studies that focus not only on the quantity but also the quality of soil organic matter and its relationship with biodiversity (flora and fauna) in traditional Timorese AFS.
+
+## How to conserve and improve carbon stocks in traditional AFS in Timor-Leste?
 
 Although agroforestry can make agriculture more resilient to climate change while improving its carbon footprint and biodiversity conservation value [@Cardinael2021], poor adaptation to the local context can explain many failures to adopt new agroforestry techniques [@Coe2014].
-Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and became more adapted to it [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
-This is the case of the traditional AFSs studied here in Timor-Leste: their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally.
-
-Conservation of these traditional AFSs should avoid their abandonment as the trees wither or as their productivity declines.
-Clear-cutting old agroforests for full-sun crops may then become an attractive short-term strategy, providing a large cash inflow from the sale of wood products, followed by 4 or 5 years of agricultural production on fertile land.
-Unfortunately, this results in a significant loss of biodiversity and carbon stored in the biomass, and can lead to an irreversible loss of fertility on the plot through soil degradation and erosion.
-This phenomenon has been documented in a similar context in Sumatra (Indonesia), where @Gouyon1993 highlighted the fact that old agroforests, which they called "jungle-rubber" and which are quite similar in structure to "forest gardens" (except for the rubber tree component, which has been replaced by other species in Timor), were being cleared in favor of more intensive production systems.
-This trend towards clearing jungle rubber was subsequently confirmed by @Ekadinata2011, and its impact on carbon stock decline was quantified by @Villamor2014.
-
-Strategies to avoid clearing of traditional AFSs could involve new ways of intensifying production while maintaining high carbon and biodiversity values, and gradually renewing the tree population in the plots.
-For example, it would be interesting to improve tree regeneration through clearings of about one hundred square meters (corresponding to the felling of 1 to 10 trees, depending on their crown area).
-In these clearings, young trees of species considered interesting for their future production would be associated, first with heliophile crops (e.g.
-cereals) and then with shade-tolerant crops (e.g.
-taro).
-Additional mechanisms to encourage the renewal of traditional AFSs could include the provision of subsidies (e.g.
-payment of carbon credits) or low-interest loans to farmers wishing to renew their agroforestry plots.
-Another strategy, which has been successfully implemented in other tropical countries (e.g.
-for agroforestry coffee), is to improve the incomes of AFS farmers through 'sustainable agriculture' certification [@Bertrand2019], although this is primarily applicable to the production of export goods, whereas the AFSs studied here are currently oriented towards local consumption and domestic markets.
+Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and have adapted to this context [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
+This is the case of the traditional AFS studied here in Timor-Leste.
+Their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally.
+
+These traditional AFS should be conserved to prevent them from being abandoned as the trees wither or their productivity declines.
+Clear-cutting old agroforests for full-sun crops can appear to be an attractive short-term strategy, providing a large cash inflow from the sale of wood products, followed by four or five years of agricultural production on fertile land.
+However, this practice unfortunately results in a significant loss of biodiversity and carbon stored in the biomass, and can lead to an irreversible loss of fertility on the plot through soil degradation and erosion.
+This phenomenon has been documented in a similar context in Sumatra (Indonesia), where @Gouyon1993 highlighted the fact that old agroforests, which they called “jungle-rubber”, and which are quite similar in structure to “forest gardens” (except for the rubber tree component, which has been replaced by other species in Timor-Leste), were being cleared in favor of more intensive production systems.
+This trend toward clearing jungle-rubber was subsequently confirmed by @Ekadinata2011, and its impact on carbon stock decline was quantified by @Villamor2014.
+
+Strategies to avoid the clearing of traditional AFS could involve new ways of intensifying production while maintaining high carbon and biodiversity values, and gradually renewing the tree population in the plots.
+For example, it would be interesting to improve tree regeneration by clearing about one hundred square meters (corresponding to the felling of 1 to 10 trees, depending on their crown area).
+In these clearings, young trees of species considered interesting for their future production could be associated with first heliophile crops (e.g., cereals), and then shade-tolerant crops (e.g., taro).
+Additional mechanisms to encourage the renewal of traditional AFS could include the provision of subsidies (e.g., payment of carbon credits) or low-interest loans to farmers wishing to renew their agroforestry plots.
+Another strategy, which has been successfully implemented in other tropical countries (e.g., for agroforestry coffee), is to improve the incomes of AFS farmers through ‘sustainable agriculture’ certification [@Bertrand2019], although this is primarily applicable to the production of export goods, whereas the AFS studied here are currently oriented toward local consumption and domestic markets.
 
@@
 These systems have only recently been described in the scientific literature, and more research is needed to improve our knowledge of their diversity and functioning.
-It is therefore important that all stakeholders, including local authorities, universities and NGOs, are aware of their existence, diversity and environmental and social value, in order to propose possible innovations to maintain or improve their functioning.
+It is therefore important that all stakeholders, including local authorities, universities, and NGOs, are aware of their existence, diversity, and environmental and social value, in order to propose possible innovations to maintain or improve their functioning.
 
@@
 We are grateful to Alain Rival (Cirad) for his facilitation of the field studies and critical revision of the paper, and to the trainees who helped with field data collection: Juvencio Dos Santos (vegetation) and Teofilio Gomes (soil).
-We would also like to thank the 31 farmers who gave up their time to take part in the interviews and field measurements.
-
-This project was co-funded by the European Union and the German Federal Ministry for Economic Cooperation and Development (BMZ) as part of the Partnership for Sustainable Agroforestry (PSAF) - Ai ba Futuru project; the Deutsche Gesellschaft für Internationale Zusammenarbeit (GIZ); the French Embassy in Jakarta; and the French Institute of Timor-Leste.
+We would also like to thank the 31 farmers who gave their time to take part in the interviews and field measurements.
+A first English version of this paper was proofread by Grace Delobel.
+
+This project was co-funded by the European Union and the German Federal Ministry for Economic Cooperation and Development (BMZ) as part of the Partnership for Sustainable Agroforestry (PSAF) - Ai ba Futuru project, implemented by the Deutsche Gesellschaft für Internationale Zusammenarbeit (GIZ).
+Additional funding for this study was provided by the French Embassy in Jakarta and the French Institute of Timor-Leste.
 
@@
 
-The R codes used to analyse the data and produce the manuscript (including figures and tables) are available on ______.
+The R codes used to analyze the data and produce the manuscript (including figures and tables) are available on ______.
 
@@
 
-```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAS with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
+```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAs with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
 The p-value gives the probability of the means of the categories of each independent variable being the same.
@@
 ## correct typos in vernacular names
-vern[, gsub("kamii", "kami", TETUM_NAME)]
-vern[, gsub("ai-haek", "ai-hanek", TETUM_NAME)]
-vern[, gsub("ai-bobur", "ai-bubur", TETUM_NAME)]
+vern[, TETUM_NAME := tolower(gsub(" oan| kain| restu| hun restu", "", TETUM_NAME))]
+
+vern[,grep("tali", TETUM_NAME, value=T)]
+
+vern[grep("kami", TETUM_NAME), TETUM_NAME := "ai-kamii"]
+vern[, TETUM_NAME := gsub("ai-haek", "ai-hanek", TETUM_NAME)]
+vern[, TETUM_NAME := gsub("ai-bobur", "ai-bubur", TETUM_NAME)]
+vern[, TETUM_NAME := gsub("duut", "du'ut", TETUM_NAME)]
+vern[, TETUM_NAME := gsub("aifarina", "ai-farina", TETUM_NAME)]
+vern[, TETUM_NAME := gsub("aitarak", "ai-tarak", TETUM_NAME)]
+vern[, TETUM_NAME := gsub("ainanas", "ai-nanas", TETUM_NAME)]
+vern[, TETUM_NAME := gsub("fore tali", "fore-tali", TETUM_NAME)]
+vern[, TETUM_NAME := gsub("kromolaina", "kromolaena", TETUM_NAME)]
 

### 2024-05-16 manuscript.Rmd
pre-submission manuscript

@@
 title: "Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass"
-author: "Camille Piponiot^*^, Marguerite Cogné^*^, Vincent Freycon, Alexis Thoumazeau, Marçal Gusmão, Régis Peltier"
+author: "Camille Piponiot, Marguerite Cogné, Vincent Freycon, Alexis Thoumazeau, Marçal Gusmão, Régis Peltier"
+
+abstract: |
+  Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural land and diversifying landscapes.
+Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their inclusion in climate change mitigation strategies.
+In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste, namely cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
+Our results show that these traditional AFS can store large amounts of carbon, with the average being 156\ Mg\ C\ ha$^{-1}$.
+The AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, close to the values of old-growth tropical forests.
+Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
+We found no relationship between the amount of carbon stored in biomass and soil.
+Our results highlight the high diversity of traditional AFS in Timor-Leste and their high carbon sequestration capacity.
+These results could provide an important baseline for the inclusion of AFS in Timor-Leste’s climate change mitigation strategy, and could serve as a reference for future AFS studies in different agro-climates of Timor-Leste.
+
+journal: "Agroforestry systems"
 date: ""
+
 output:
@@
     base_format: "bookdown::word_document2"
-    reference_docx: word-styles-ref-02.docx
-bibliography: references.bib
+    reference_docx: misc/word-styles-ref-02.docx
+bibliography: misc/references.bib
 nocite: |
-  @Orsini1976, @Nair1993, @Nair2021
+  @Nair1993, @Nair2021, @Orsini1976
 csl: misc/springer-basic-author-date.csl
+keywords: "Traditional agroforestry, carbon stocks, soil, biomass"
 ---
+
+Camille Piponiot(*) - Marguerite Cogné - Vincent Freycon - Régis Peltier
+
+UR Forests and Societies, CIRAD, Univ Montpellier, Montpellier, France
+
+(*) email: camille.piponiot-laroche@cirad.fr
+
+Alexis Thoumazeau
+
+UR ABSys, CIRAD, Univ Montpellier, CIHEAM-IAMM, INRAE, Institut Agro, Montpellier, France
+
+HRPP, Kasetsart University, Bangkok, Thailand
+
+Marçal Gusmão
+
+National University of Timor Lorosa'e, Dili, Timor-Leste
 
@@
 ```
-
-**Journal**: Agroforestry Systems
-
-**Word count**: 8365; **Figure and table count**: 5
-
-# Abstract
-
-Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural land and diversifying landscapes.
-Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their inclusion in climate change mitigation strategies.
-In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste, namely cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
-Our results show that these traditional AFS can store large amounts of carbon, with the average being 156\ Mg\ C\ ha$^{-1}$.
-The AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, close to the values of old-growth tropical forests.
-Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
-We found no relationship between the amount of carbon stored in biomass and soil.
-Our results highlight the high diversity of traditional AFS in Timor-Leste and their high carbon sequestration capacity.
-These results could provide an important baseline for the inclusion of AFS in Timor-Leste’s climate change mitigation strategy, and could serve as a reference for future AFS studies in different agro-climates of Timor-Leste.
 
@@
 
-In each of the 30 sample plots, three composite samples were taken from three of the five 10 x 10 m subplots, following a transect covering the intra-plot variability and at a minimum distance of 1 m from any tree (Figure\ S\@ref(fig:design)).
+In each of the 30 sample plots, a composite of three samples was taken from three of the five 10 x 10 m subplots, following a transect covering the intra-plot variability and at a minimum distance of 1 m from any tree (Figure\ S\@ref(fig:design)).
 Soils were sampled with a 15 cm auger at 0-10 cm and 20-30 cm depth; due to the size of the auger we could not accurately sample the intermediate soil layer.
@@
 
-In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm3 at 2-7 cm, 12-17 cm and 22-27 cm.
+In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm^3^ at depths of 2-7 cm, 12-17 cm and 22-27 cm.
 Each sample was dried at 105°C in a classic oven until a stable dry mass was reached and then sieved to 2 mm.
@@
 
-Soil organic carbon stocks (in Mg\ C\ ha$^{-1}$) were then calculated as follows:
+Soil organic carbon stocks (in Mg\ C\ ha$^{-1}$) were calculated as follows:
 
@@
 ), .(plot = N_PLOT, site, afs = ID_AF, variable)]
+# save for supp info doc 
+write.csv(dataC_plot, "dataC_plot.csv", row.names = FALSE)
 
@@
 In a meta-analysis by @Shi2018, it was estimated that AFS stored around 50\ Mg\ C\ ha$^{-1}$ at a 0-30 cm depth.
-Such large SOC stocks could be due to the high soil pH associated with high exchangeable Ca content at our sites (Table\ \@ref(tab:site-soil)), as Ca is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
+Such large SOC stocks could be due to the high exchangeable Ca^2+^ content at the sites included in this study (Table\ \@ref(tab:site-soil)), as Ca^2+^ is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
 In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`\ Mg\ C\ ha$^{-1}$) were similar to the average of 74\ Mg\ C\ ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
@@
 This more limited role for trees may explain why their total biomass is lower in these systems.
-In CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun, while in SP systems, the growth of grazing grasses is favored by reducing the number of trees.
+Indeed, in CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun, while in SP systems, the growth of grazing grasses is favored by reducing the number of trees.
 
@@
 These results may relate to previous findings that SOC stocks saturate much faster than AGC stocks in tropical AFS.
-For example, in a global meta-analysis, @Ma2020 estimated that it took only five years for SOC stocks in tropical AFS to reach an equilibrium value, but more than three decades for AGC stocks.
+For example, in their global meta-analysis, @Ma2020 estimated that it took only five years for SOC stocks in tropical AFS to reach an equilibrium value, but more than three decades for AGC stocks.
 This desynchronization of SOC and AGC stocks recovery has also been observed in naturally regenerating tropical forests.
@@
 dataC_afs[, value := paste0(
-  signif(mean, 3), " (",
-  signif(lwr_mean, 3), "-", 
-  signif(upr_mean, 3), ")")] |> 
+  round(mean), " (",
+  round(lwr_mean), "-", 
+  round(upr_mean), ")")] |> 
   subset(variable != "SOC10") |> 
@@
             cwidth = c(0.9, 1.2, 1.3, 1.3, 1.2)) |> 
-  compose(i = 1, j = "afs", part = "header", value = as_paragraph("Agroforestry system")) |>
+  compose(i = 1, j = "afs", part = "header", value = as_paragraph("Agroforestry\nsystem")) |>
   compose(i = 1, j = "TOT", part = "header", value = as_paragraph( "Total C stocks\n(Mg C ha", as_sup("-1"), ")"))|> 
@@
 ggarrange(ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))
+ggsave("fig1.pdf")
 ```
@@
   scale_y_continuous(expand = c(0, 0))
+ggsave("fig2.pdf")
 ```
@@
   theme_classic()
+ggsave("fig3.pdf")
 ```
@@
 
-\newpage
-
 # Supplementary information
@@
 ```{r site-soil-char}
+data_soil[, Corg := `Corg\n`]
+data_soil[, muCorg := mean(Corg, na.rm = TRUE), .(N_PLOT)]
+data_soil[is.na(Corg), Corg := muCorg]
+
+data_soil[, c("h0", "h1") := lapply(tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
+
 data_soil$silt <- as.numeric(data_soil$`FINE SILT`) + as.numeric(data_soil$`COARSE SILT`)
@@
 data_soil_site  = data_soil_site_simu[, .(
-  value = paste0(
-    signif(mean(value), 3), " (",
-    signif(quantile(value, 0.025), 3), "-", 
-    signif(quantile(value, 0.975), 3), ")"
-  )), 
+  mean = signif(mean(value), 3), 
+  lwr = signif(quantile(value, 0.025), 3),
+  upr = signif(quantile(value, 0.975), 3)
+  ), 
   .(site, variable)]
+data_soil_site[, value := paste0(round(mean, 1), " (", round(lwr, 1), "-", round(upr, 1), ")")]
+data_soil_site[variable %in% c("K", "Mg", "Na"), value := paste0(round(mean, 2), " (", round(lwr, 2), "-", round(upr, 2), ")")]
 
@@
 data_soil_site = dcast(data_soil_site, site ~ variable, value.var = "value")
+
+## add unit row
+data_soil_site = data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "CEC", "Ca", "K", "Mg", "Na")]
+colnames(data_soil_site) = c(" ", "Clay", "Silt", "Sand", "C content", "N content", "C:N", "pH", "CEC", "Ca", "K", "Mg", "Na")
 ```
@@
 ```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (average over 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
+Units are: in g/kg for clay, silt, sand, C and N contents; in cmolc/kg for CEC, Ca, K, Mg and Na; and no units for C:N and pH.
 The confidence intervals were calculated by propagating a 12% error on the lab-provided values and obtaining site-level estimates through bootstrapping plot values (1000 simulations).
 Total organic carbon content (dry combustion method) ; total nitrogen content (dry combustion method) ; pH in water (soil:deionized water ratio of 1:2.5) ; CEC - cation exchangeable capacity - and exchangeable cations (Ca, K, Mg and Na) at soil pH [Cobaltihexamine method; Orsini and Remy, 1976]."}
-flextable(
-  data_soil_site,
-  col_keys = c("site", "CLAY", "silt", "sand", "Corg",
-               "N", "Corg_N", "pH", "CEC", "Ca", "K", "Mg", "Na"), 
-  cwidth = c(0.5, rep(0.6, 5), rep(0.5, 2), rep(0.6, 5)) 
-) |> 
-  compose(i = 1, j = "CLAY", part = "header", value = as_paragraph("Clay\n(g kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "silt", part = "header", value = as_paragraph("Silt (g kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "sand", part = "header", value = as_paragraph("Sand (g kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "Corg", part = "header", value = as_paragraph("C content (g kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "N", part = "header", value = as_paragraph("N content (g kg", as_sup("-1"), ")")) |>
-  compose(i = 1, j = "Corg_N", part = "header", value = as_paragraph("C/N")) |> 
-  compose(i = 1, j = "CEC", part = "header", value = as_paragraph( "CEC (cmolc kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "Ca", part = "header", value = as_paragraph( "Ca (cmolc kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "K", part = "header", value = as_paragraph( "K (cmolc kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "Mg", part = "header", value = as_paragraph( "Mg (cmolc kg", as_sup("-1"), ")")) |> 
-  compose(i = 1, j = "Na", part = "header", value = as_paragraph( "Na (cmolc kg", as_sup("-1"), ")"))
+flextable(data_soil_site, cwidth = 0.5)
 ```
@@
 vern[, TETUM_NAME := tolower(gsub(" oan| kain| restu| hun restu", "", TETUM_NAME))]
-
-vern[,grep("tali", TETUM_NAME, value=T)]
-
 vern[grep("kami", TETUM_NAME), TETUM_NAME := "ai-kamii"]
@@
   set_header_labels(values =  c("Scientific name", "Tetum name")) |> 
-  autofit()
+  autofit() |> 
+  italic(j=1)
 ```
@@
 Transitions between two systems are represented by arrows: dashed arrows represent transitions that usually include tree removal, whereas solid arrows represent transitions where trees are planted and/or the AFS is left to regenerate naturally.
-These transitions may vary according to the farmer's strategy and available resources (Cogné and Lescuyer, 2024).", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 8, fig.height = 2.2}
+These transitions may vary according to the farmer's strategy and available resources (Cogné and Lescuyer, 2024).", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 8, fig.height = 2.2, out.width="50%"}
 # save_png <- function(plot, path){
@@
 Results are presented as regression lines and the p-value of the relationship between SOC stocks and elevation.", fig.cap.pre="Figure S"}
-dataCelev = merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")], 
+dataC_plot <- read.csv("dataC_plot.csv") |> setDT()
+dataCelev <- merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")], 
                   data_plot[, c("N_PLOT", "long", "lat")], by.x = "plot", by.y = "N_PLOT")
@@
 Results are presented as regression lines and the p-value of the relationship between SOC content and fine soil texture.", fig.cap.pre="Figure S"}
-
 data_soil[, site := substr(ID_VILLAGE, 1, 1)]

### 2024-11-29 manuscript.Rmd
use bayesian approach

@@
   "ggpubr",       # version 0.6.0
+  "rdryad",       # version 1.0.0
   "readxl",       # version 1.4.3
@@
 
+# get carbon content 
+GLOWCAD_files = dryad_download("10.5061/dryad.18931zcxk")
+GLOWCAD = read.csv(GLOWCAD_files[[1]][3]) |> 
+  subset(grepl("tree", growth.form)) |> 
+  subset(tissue %in% c("stem", "root", "coarseroot")) |> 
+  subset(biome == "tropical") 
+data_ccontent <-
+
   # estimate mean AGC
@@
   ) * 0.4713 # conversion factor from biomass to carbon
+
+# carbon content
 
@@
 
-Soil organic carbon stocks (in Mg\ C\ ha$^{-1}$) were calculated as follows:
+SOC stocks (in Mg\ C\ ha$^{-1}$) in plot $p$ were calculated as follows:
 
 \begin{equation}
-SOC_i = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl} (\#eq:soc)
+SOC_p = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl} (\#eq:soc)
 \end{equation}
 
-Where $Corg_i$ is the organic carbon content at depth $i$ (either 0-10, 10-20 or 20-30 cm); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth $i$ (in g), converted into Mg with a factor of 10$^{-6}$; $Splot$ is the surface of a 1-ha plot (10$^8$ cm$^2$); $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
+$$SOC_p = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl}$$
+
+Where $Corg_i$ is the SOC content at depth $i$ (either 0-10, 10-20 or 20-30 cm); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth $i$ (in g), converted into Mg with a factor of 10$^{-6}$; $Splot$ is the surface of a 1-ha plot (10$^8$ cm$^2$); $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
 This calculation corresponds to the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil.
-At depth $i$ = 10-20 cm, where the carbon content was not measured (see "Soil measurements" section), we used the average of the carbon content values at 0-10 and 20-30 cm.
+At depth $i$ = 10-20 cm, where SOC content was not measured (see "Soil measurements" section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a negative exponential model of SOC content decrease with soil depth [@Mishra2009;@Murphy2019]:
+
+\begin{equation}
+Corg_{10,20} = Corg_{0,10} \cdot \sqrt{\frac{Corg_{20,30}}{Corg_{0,10}}} (\#eq:intersoc)
+\end{equation}
+
+where $Corg_{a,b}$ is the average SOC content between depths a and b (see rationale for equation \@ref(eq:intersoc) in Supplementary Information xx).
 
@@
 ```{r add-missing-values}
-## add intermediate soil carbon in 10-20 cm depth
 data_soil[, Corg := `Corg\n`]
-data_soil[, muCorg := mean(Corg, na.rm = TRUE), .(N_PLOT)]
-data_soil[is.na(Corg), Corg := muCorg]
+## add intermediate soil carbon in 10-20 cm depth under the assumption of an
+# organic carbon content that decreases as a negative exponential of depth, we can deduce the carbon content at 10-20 cm with the following function: SOC[10-20] = SOC[0-10]*sqrt(SOC[20-30]/SOC[0-10])
+data_soil = merge(
+  data_soil, 
+  data_soil[, .(
+    Corg1020 = Corg[SAMPLE_DEPTH=="0-10"]*
+      sqrt(Corg[SAMPLE_DEPTH=="20-30"]/Corg[SAMPLE_DEPTH=="0-10"])), 
+    .(N_PLOT)])
+
+data_soil[SAMPLE_DEPTH=="10-20", Corg := Corg1020]
 ```
@@
 
+```{r add-other-soil-properties}
+data_soil_summ <- data_soil |> 
+  subset(!is.na(TC)) |> 
+  melt(id.vars = c("N_PLOT", "SAMPLE_DEPTH"), 
+       measure.vars= c("CLAY", "FINE SILT", "pH\n", "Ca", "CEC\n"))  |> 
+  setDT()
+data_soil_summ[, variable := gsub("\n", "", variable)]
+data_soil_summ <- 
+  data_soil_summ[, .(value = mean(as.numeric(value), na.rm = TRUE)), .(N_PLOT, variable)] |> 
+  dcast(N_PLOT ~ variable)
+data_soil_summ[, fine_soil := CLAY + `FINE SILT`] 
+data_sim <- merge(data_sim, data_soil_summ)  
+```
+
 ## Statistical analysis
@@
 For multiplot estimates of carbon stocks, values within each grouping factor were additionally bootstrapped before calculating quantiles (median and 95% confidence intervals).
+
+```{r}
+stan_model <- "
+data {
+  int<lower=0> N;
+  int<lower=1> AF;
+  int<lower=1> S;
+  int<lower=1,upper=AF> afs[N];
+  int<lower=1,upper=S> site[N];
+  real clay_sc[N];
+  real agc_sc[N];
+  real<lower=0> agc[N];
+  real<lower=0> soc[N];
+  real<lower=0> soc10[N];
+}
+
+parameters {
+  real mu_a;
+  real delta_afs_a[AF];
+  real delta_site_a[S];
+  real<lower=0> sigma_a;
+
+  real mu_s;
+  real delta_afs_s[AF];
+  real delta_site_s[S];
+  real theta_agc_s;
+  real theta_clay_s;
+  real<lower=0> sigma_s;
+
+  real mu_s10;
+  real delta_afs_s10[AF];
+  real delta_site_s10[S];
+  real theta_agc_s10;
+  real theta_clay_s10;
+  real<lower=0> sigma_s10;
+
+  real<lower=0> sigma_site_a;
+  real<lower=0> sigma_afs_a;
+  real<lower=0> sigma_site_s;
+  real<lower=0> sigma_afs_s;
+  real<lower=0> sigma_site_s10;
+  real<lower=0> sigma_afs_s10;
+
+}
+
+model {
+  real pred_a[N];
+  real pred_s[N];
+  real pred_s10[N];
+
+  for (i in 1:N) {
+    pred_a[i] = mu_a + delta_afs_a[afs[i]] + delta_site_a[site[i]];
+    pred_s[i] = mu_s + theta_agc_s*agc_sc[i] + theta_clay_s*clay_sc[i] + delta_afs_s[afs[i]] + delta_site_s[site[i]];
+    pred_s10[i] = mu_s10 + theta_agc_s10*agc_sc[i] + theta_clay_s10*clay_sc[i] + delta_afs_s10[afs[i]] + delta_site_s10[site[i]];
+  }
+
+  agc ~ normal(pred_a, sigma_a);
+  soc ~ normal(pred_s, sigma_s);
+  soc10 ~ normal(pred_s10, sigma_s);
+
+  // random effects
+  delta_afs_a ~ normal(0, sigma_afs_a);
+  delta_afs_s ~ normal(0, sigma_afs_s);
+  delta_afs_s10 ~ normal(0, sigma_afs_s10);
+  delta_site_s ~ normal(0, sigma_site_s);
+  delta_site_s10 ~ normal(0, sigma_site_s10);
+  delta_site_a ~ normal(0, sigma_site_a);
+
+  //  priors
+  theta_agc_s ~ normal(0,1);
+  theta_agc_s10 ~ normal(0,1);
+  theta_clay_s ~ normal(0,1);
+  theta_clay_s10 ~ normal(0,1);
+  sigma_a ~ normal(0,1);
+  sigma_s ~ normal(0,1);
+  sigma_s10 ~ normal(0,1);
+  sigma_site_a ~ normal(0,1);
+  sigma_afs_a ~ normal(0,1);
+  sigma_site_s ~ normal(0,1);
+  sigma_afs_s ~ normal(0,1);
+  sigma_site_s10 ~ normal(0,1);
+  sigma_afs_s10 ~ normal(0,1);
+}
+
+"
+```
+
+```{r stan}
+if (!dir.exists("outputs")) dir.create("outputs")
+
+if (!file.exists("outputs/data_par.csv")) {
+  library(parallel)
+
+  cl <- makeCluster(detectCores()-1)
+  clusterEvalQ(cl, {library(rstan); library(data.table)})
+  clusterExport(cl, "data_sim"); clusterEvalQ(cl, data_sim)
+  clusterExport(cl, "stan_model"); clusterEvalQ(cl, stan_model)
+
+  data_par <- parLapply(cl, unique(data_sim$iter), function(i) {
+    rstan_options(auto_write = TRUE)
+
+    data_stan <- data_sim |> 
+      subset(iter==i) |> 
+      dcast(N_PLOT + site + ID_AF + fine_soil ~ variable) |> 
+      with(
+        list(
+          N = length(N_PLOT),
+          AF = length(unique(ID_AF)),
+          S = length(unique(site)),
+          afs = as.numeric(as.factor(ID_AF)), 
+          site = as.numeric(as.factor(site)), 
+          clay_sc = (fine_soil-mean(fine_soil))/sd(fine_soil),
+          agc_sc = (AGC-mean(AGC))/sd(AGC),
+          agc = AGC,
+          soc = SOC,
+          soc10 = SOC10
+        ))
+
+    fit <- stan(model_code = stan_model, data = data_stan, 
+                iter = 60, warmup = 50, chains = 1)
+    pars <- extract(fit) |> 
+      lapply(function(x) {
+        df <- data.table(iter_stan = 1:dim(x)[1], x) |> 
+          melt(id.vars = "iter_stan", variable.name = "level_num")  
+        return(df)
+      }) |> 
+      rbindlist(idcol = "param")
+    return(pars)
+  }) |> rbindlist(idcol = "iter")
+
+  write.csv(data_par, file = "outputs/data_par.csv", row.names = FALSE)
+
+  stopCluster(cl)
+
+} else data_par <- read.csv("outputs/data_par.csv") |> setDT()
+
+data_par[, effect := tstrsplit(param, "_")[[2]]]
+data_par[! effect %in% c("afs", "site", "agc", "clay"), effect := NA]
+data_par[, stock := tstrsplit(param, "_")[[3]]]
+data_par[is.na(stock) & param != "lp__", stock := tstrsplit(param, "_")[[2]]]
+data_par[stock == "a", stock := "agc"]
+data_par[stock == "s", stock := "soc"]
+data_par[stock == "s10", stock := "soc10"]
+data_par[, param := tstrsplit(param, "_")[[1]]]
+
+data_par[, level_num := as.numeric(gsub("V", "", level_num))]
+data_par[effect == "afs" & !is.na(level_num), 
+         level := levels(as.factor(data_sim$ID_AF))[level_num]]
+data_par[effect == "site" & !is.na(level_num),
+         level := levels(as.factor(data_sim$site))[level_num]]
+
+df_mu_delta <- 
+  data_par[param == "delta" & !is.na(effect)] |> 
+  dcast(iter + iter_stan + stock + effect + level ~ param) |> 
+  merge(dcast(data_par[param == "mu"], iter + iter_stan + stock ~ param))
+df_mu_delta[, `:=`(value = mu + delta, param = "pred_med")]
+data_par <- rbind(
+  data_par[, c("param", "iter", "iter_stan", "stock", "effect", "level", "value")], 
+  df_mu_delta[, c("param", "iter", "iter_stan",  "stock", "effect", "level", "value")]
+)
+
+itermaxlp = data_par[param == "lp", iter[which.max(value)]]
+pars_summary <- data_par[, .(
+  mean = mean(value), 
+  maxlp = value[iter==itermaxlp],
+  low95 = quantile(value, 0.025), 
+  low80 = quantile(value, 0.1), 
+  med = median(value), 
+  upp80 = quantile(value, 0.9),
+  upp95 = quantile(value, 0.975)
+), .(
+  param, stock, effect, level
+)]
+```
 
@@
 ggarrange(ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))
-ggsave("fig1.pdf")
 ```
@@
   scale_y_continuous(expand = c(0, 0))
-ggsave("fig2.pdf")
 ```
@@
 </br>
+
+```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3}
+labels_stock = function(x) paste(toupper(x), "(Mg/ha)")
+labels_effect = function(x) ifelse(x %in% c("afs", "agc"), toupper(x), x)
+
+pars_summary |> 
+  subset(param == "pred_med" & stock != "soc10") |> 
+  ggplot(aes(y = level)) + 
+  geom_segment(aes(x = low95, xend = upp95)) + 
+  geom_segment(aes(x = low80, xend = upp80), col = 2, lwd = 3) + 
+  geom_point(aes(x = mean), size = 2) +
+  facet_grid(
+    effect ~ stock, scales = "free", 
+    labeller = labeller(stock = labels_stock, effect = labels_effect)) +
+  labs(x=NULL, y = NULL) +
+  theme_bw()
+```
+
+```{r fig.id = "effect-agc-clay", fig.width = 7, fig.height = 2}
+labels_effect = function(x) ifelse(x=="agc", "AGC", "clay + fine silt")
+pars_summary |> 
+  subset(param == "theta")  |> 
+  ggplot(aes(y = ifelse(stock=="soc10", "0-10 cm", "0-30 cm"))) + 
+  geom_vline(xintercept = 0, lty = 2) +
+  geom_segment(aes(x = low95, xend = upp95)) + 
+  geom_segment(aes(x = low80, xend = upp80), col = 2, lwd = 3) + 
+  geom_point(aes(x = mean), size = 2) +
+  facet_wrap(~effect, labeller = labeller(effect = labels_effect)) + 
+  labs(x="Effect on SOC", y = NULL) +
+  theme_bw()
+```
 
@@
 One plot (AFS: CF, site: Gariuai) with no trees over 30 cm in circumference at breast height (minimum size for measurement) had AGC stocks estimated at 0 Mg C ha^-1^ and was therefore removed from the linear model and the figure."}
-dfpred = expand.grid(AGC_mean = exp(seq(min(log(data_reg$AGC_mean), na.rm = TRUE),
-                                        max(log(data_reg$AGC_mean), na.rm = TRUE),
-                                        length.out = 20)),
-                     depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm")) |>
-  merge(dfreg[, c("depth", "intercept", "slope")], by = "depth")
-setDT(dfpred)
-dfpred[, SOC_mean := exp(intercept + slope * log(AGC_mean))]
+
+dfpred <- expand.grid(agc = seq(0, 170, length.out = 20), 
+  iter = unique(data_par$iter), 
+  iter_stan = unique(data_par$iter_stan)
+) |> 
+  merge(
+    data_par[grepl("soc", stock) & (param =="mu" | effect == "agc")] |> 
+      dcast(iter + iter_stan + stock ~ param)
+  ) |> 
+  merge(
+    data_sim[variable=="AGC", .(mua = mean(value), sda = sd(value)), 
+             .(iter = as.numeric(as.factor(iter)))]
+  ) |> setDT()
+dfpred[, pred := (mu - mua*theta/sda) + (theta/sda) * agc]
+dfpred <- dfpred[, .(
+  mean = mean(pred), 
+  lower = quantile(pred, 0.025), 
+  upper = quantile(pred, 0.975)
+  ), .(agc, stock)]
+dfpred[, depth := ifelse(grepl(10, stock), "Soil depth: 0-10 cm", "Soil depth: 0-30 cm")]
 
@@
 ggplot(dataC_fig, aes(color = afs)) +
+  geom_ribbon(data = dfpred, aes(x = agc, ymin = lower, ymax = upper), 
+              col = NA, alpha = 0.1) +
   geom_point(aes(x = AGC_mean, y = SOC_mean, shape = site), size = 2) +
-  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
-            x = -Inf, y = Inf, hjust = -0.1, vjust = 1.2, parse = TRUE, color = 1) +
-  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
-            x = -Inf, y = Inf, hjust = -0.1, vjust = 2.2, parse = TRUE, color = 1) +
+  # geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
+  #           x = -Inf, y = Inf, hjust = -0.1, vjust = 1.2, parse = TRUE, color = 1) +
+  # geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
+  #           x = -Inf, y = Inf, hjust = -0.1, vjust = 2.2, parse = TRUE, color = 1) +
   geom_linerange(aes(x = AGC_mean, y = SOC_mean, ymin = SOC_lwr, ymax = SOC_upr)) +
   geom_errorbarh(aes(y = SOC_mean, xmin = AGC_lwr, xmax = AGC_upr)) +
-  geom_line(data = dfpred, aes(x = AGC_mean, y = SOC_mean), lty = 2, col = 1) +
+  geom_line(data = dfpred, aes(x = agc, y = mean), lty = 2, col = 1) +
   labs(color = "AFS", x = expression("AGC stocks (Mg C"*" "*ha^{-1}*")"), 
@@
   scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
-  scale_x_log10() +
-  scale_y_log10() +
+  # scale_x_log10() +
+  # scale_y_log10() +
   facet_wrap(~depth) +
   theme_classic()
-ggsave("fig3.pdf")
 ```
@@
 # Supplementary information
+
+## Supplementary methods
+
+### Calculation of 10-20 cm average carbon content
+
+During field data collection, because the auger length (15 cm) didn't match the target soil sampling depths (10 cm each), we could only accurately sample and analyse the 0-10 cm and 20-30 cm soil depths for SOC content.
+The SOC content of the intermediate 10-20 cm soil depth had to be estimated based on these first two measurements.
+
+  To reduce the bias in this estimation, we used the standard assumption that SOC content $c(x)$ decreases as an exponential function of soil depth $x$:
+
+  $$c(x) = C_0 \cdot exp(-\lambda\cdot x) $$ 
+  where $C_0$ is the carbon content at the surface (depth $x= 0$) and $\lambda$ is the rate at which the carbon content decreases with soil depth. 
+  Under this assumption, we can calculate the mean SOC content between depths $x$ and $x+\Delta$ as:
+
+  $$\bar c(x, x+\Delta) = \frac{1}{\Delta}\int_{x}^{x+\Delta}  C_0 \cdot exp(-\lambda\cdot x) dx$$
+
+  $$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{-\lambda \cdot \Delta} \left[ exp(-\lambda\cdot (x+\Delta)) - exp(-\lambda\cdot x) \right]$$
+
+  $$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{\lambda \cdot \Delta} exp(-\lambda\cdot x)\left[1- exp(-\lambda\cdot\Delta) \right]$$
+  The average SOC contents between 0-10 cm, 10-20 cm and 20-30 cm can respectively be written as:
+
+  $$C_{0,10} = \frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)$$
+
+  $$C_{10,20} = \frac{C_0}{10\lambda} exp(-10\lambda)\left(1- exp(-10\lambda) \right)$$
+
+  $$C_{20,30} = \frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)$$
+  We can thus write:
+
+  $$\frac{C_{20,30}}{C_{0,10}} = \frac{\frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)}{\frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)} = exp(-20\lambda)$$
+  $$ \Rightarrow \lambda = \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20}$$
+  We can also write:
+
+  $$ \frac{C_{10,20}}{C_{0,10}} = exp(-10\lambda) $$
+  $$\Rightarrow C_{10,20} = C_{0,10} \cdot exp(-10 \cdot \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20})$$
+
+  $$ \Rightarrow C_{10,20} = C_{0,10} \cdot \left( \frac{C_{20,30}}{C_{0,10}}\right)^{\frac{1}{2}}$$
 

### 2024-12-03 manuscript.Rmd
uppdate statistical section in methods

@@
 author: "Camille Piponiot, Marguerite Cogné, Vincent Freycon, Alexis Thoumazeau, Marçal Gusmão, Régis Peltier"
-
 abstract: |
-Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and other woody perennials into agricultural land and diversifying landscapes.
+Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and/or other woody perennials into agricultural land thereby  diversifying landscapes.
 Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their inclusion in climate change mitigation strategies.
-In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste, namely cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
-Our results show that these traditional AFS can store large amounts of carbon, with the average being 156\ Mg\ C\ ha$^{-1}$.
-The AFS with the highest carbon stocks (forest gardens) stored an average of 210\ Mg\ C\ ha$^{-1}$, close to the values of old-growth tropical forests.
-Biomass carbon was strongly dependent on the type of AFS (which differed in tree cover), while soil carbon was less variable between AFS but more dependent on site.
+In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste, namely cropping systems with fallow (CF), silvopastures (SP), young agroforests (YA), home gardens (HG), and forest gardens (FG).
+Our results show that these traditional AFS can store large amounts of carbon, with the average being 155\ Mg\ C\ ha$^{-1}$.
+The AFS with the highest carbon stocks (FG) stored an average of 210\ Mg\ C\ ha$^{-1}$,  close to the values of old-growth tropical forests.
+Other AFS stored on average: 107\ Mg\ C\ ha$^{-1}$ in CF, 158\ Mg\ C\ ha$^{-1}$ in SP, 132\ Mg\ C\ ha$^{-1}$ in YA, and 167\ Mg\ C\ ha$^{-1}$ in HG.
+Biomass carbon stocks varied significantly between AFS (which differed in tree cover), while soil carbon stocks were less variable between AFS but more dependent on site.
 We found no relationship between the amount of carbon stored in biomass and soil.
@@
 date: ""
-
 output:
@@
 nocite: |
-@Nair1993, @Nair2021, @Orsini1976
 csl: misc/springer-basic-author-date.csl
 keywords: "Traditional agroforestry, carbon stocks, soil, biomass"
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
 
-Camille Piponiot(*) - Marguerite Cogné - Vincent Freycon - Régis Peltier
+Camille Piponiot(\*) - Marguerite Cogné - Vincent Freycon - Régis Peltier
 
@@
 
-(*) email: camille.piponiot-laroche@cirad.fr
+(\*) email: [camille.piponiot-laroche\@cirad.fr](mailto:camille.piponiot-laroche@cirad.fr){.email}
 
@@
 
-Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO~2~e/yr over 2010-2014; @Pendrill2019].
+Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022; @Gibbs2010; @Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO~2~e/yr over 2010-2014; @Pendrill2019].
 The demand for tropical agricultural land is expected to continue to grow as the world’s population increases [@UN2022].
-In addition, tropical areas are expected to be disproportionately affected by climate change, putting their agriculture and food systems at risk [@Lawrence2015].
-There is therefore an urgent need to promote tropical agricultural landscapes that are more resilient to climate change while maintaining high productivity and ecological value [@Harvey2014].
-In particular, increasing the amount of carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of food production systems [@Lal2004;@Minasny2017].
-
-Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Albrecht2003;@Verchot2007;@Lasco2014;@Cardinael2021].
+In addition, tropical areas are expected to be disproportionately affected by climate change, putting their agriculture and food production systems at risk [@Lawrence2015].
+Therefore, there is an urgent need to promote tropical agricultural landscapes that are more resilient to climate change while maintaining high productivity and ecological value [@Harvey2014].
+Increasing the amount of carbon stored in the soil and vegetation biomass of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of food production systems [@Lal2004; @Minasny2017].
+
+Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Albrecht2003; @Verchot2007; @Lasco2014; @Cardinael2021].
 Agroforestry is a collective name for land use systems and technologies where woody perennials (trees, shrubs, bamboos, creepers, etc.) are deliberately used on the same land management units as agricultural crops and/or animals, in some form of spatial arrangement or temporal sequence [@Lundgren1982].
-Agroforestry allows for the diversification of agricultural production, making agricultural systems more resilient [@Duffy2021;@Terasaki2023].
-
-Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation and soil, and this carbon storage capacity depends on the characteristics of the AFS considered [@Albrecht2003;@Feliciano2018].
-As trees can store large amounts of carbon, the carbon stored in the vegetation of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
-However, the effect of AFS on soil organic carbon (SOC) has been relatively little studied [@Beillouin2021], is more variable and depends on several factors.
-Although AFS appear to increase SOC stocks overall, some AFS such as woodlots (i.e., planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFS, while AFS such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
-Other factors such as climate and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018;@Ma2020;@Martin2020].
-Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS could only be evidenced for the topsoil layer [< 10 cm; @Ma2020].
+Therefore, agroforestry practices allow for the diversification of agricultural production, making agricultural systems more resilient to climate change [@Duffy2021; @Terasaki2023].
+
+Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation biomass and soil, and this carbon storage capacity depends on the type of the AFS in addition to environmental and edaphic conditions [@Albrecht2003; @Feliciano2018].
+Since trees can store large amounts of carbon, the carbon stored in vegetation aboveground biomass of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
+However, the effect of AFS on soil organic carbon (SOC) is more variable: although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e., planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFS, while AFS such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
+Other factors such as climate and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018; @Ma2020; @Martin2020].
+Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS was only evidenced to a depth of 10 cm [@Ma2020].
 
 In Timor-Leste, a large proportion of the population depends on agriculture for their livelihoods.
-However, population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime are threatening food security [@WB2009;@Molyneux2012].
-In this context, productive, resilient, and sustainable agriculture is critical for the country, and agroforestry has been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022;@Cogne2024].
-Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022;@Cogne2024] and can be key elements for biodiversity conservation [@Perfecto2008;@Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
-A precise quantification of these carbon stocks would improve understanding of the ecological importance of these AFS at both local and global scales, and help integrate them into national and international climate change mitigation initiatives.
-
-In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying the carbon storage in both biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
-We hypothesize that there are measurable differences in aboveground carbon (AGC; H1a) and SOC (H1b) stocks between these different types of traditional AFS.
-In line with Ma et al.
-(2020), we anticipate that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; H2a), but no relationship with SOC below 10 cm (H2b).
-
-# Materials and methods
+However, population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime are threatening food security [@WB2009; @Molyneux2012].
+In this context, productive, resilient, and sustainable agriculture is critical for the country and the livelihood of its people.
+To address this, AFS have been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022; @Cogne2024].
+Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022; @Cogne2024] and can be key elements for biodiversity conservation [@Perfecto2008; @Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
+A precise quantification of these carbon stocks will improve understanding of the ecological importance of AFS at both local and global scales, and help integrate them into national and international climate change mitigation initiatives.
+
+In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying the carbon storage in both vegetation biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
+We hypothesize that there are measurable differences in carbon in aboveground vegetation biomass (hereinafter referred to as above ground carbon or AGC; Hypothesis 1a) and SOC (Hypothesis 1b) stocks between these different types of traditional AFS.
+In line with @Ma2020, we anticipate that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; Hypothesis 2a), but no relationship with SOC below 10 cm (Hypothesis 2b).
+
+# Materials and Methods
 
@@
 site_summary = site_env[, .(
-  coord = paste(round(mean(long),2), round(mean(lat),2), sep = ", "), 
+  coord = paste(round(mean(lat),2), round(mean(long),2), sep = ", "), 
   prec = round(mean(prec)), 
@@
 Two communities, or *suco* (a group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases.
-The first, Gariuai *suco* (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`), is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging between `r site_summary[site=="Gariuai", elev]` m above sea level (asl).
-The second, the village of Osso-Luga in Samalari *suco* (long-lat coordinates: `r site_summary[site=="Osso-Luga", coord]`), is located on the Matebian foothills based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso-Luga", prec]` mm and an altitude ranging between `r site_summary[site=="Osso-Luga", elev]` m asl (Figure\ \@ref(fig:map-sites)).
+The first, Gariuai *suco* (Lat-Lon coordinates: `r site_summary[site=="Gariuai", coord]`), is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging between `r site_summary[site=="Gariuai", elev]` m above sea level (masl).
+The second, the village of Osso-Luga in Samalari *suco* (Lat-Lon coordinates: `r site_summary[site=="Osso-Luga", coord]`), is located on the Matebian foothills based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso-Luga", prec]` mm and an altitude ranging between `r site_summary[site=="Osso-Luga", elev]` masl (Figure \@ref(fig:map-sites)).
 The soils of the Baucau municipality are classified using the WRB soil classification [@Mantel2023] as Lithosols and Calcic Luvisols [@FAO2012].
-The average soil properties differ between the two sites (Table\ S\@ref(tab:site-soil)); Gariuai has a clay loam texture whereas Osso-Luga has a clay texture.
-The two villages also have different recent histories.
+The average soil properties differ between the two sites (Table S\@ref(tab:site-soil)); Gariuai has a clay loam texture whereas Osso-Luga has a clay texture.
+The two villages also have different histories.
 Gariuai has been inhabited for at least 50 years, while Osso-Luga was abandoned during the Indonesian occupation and was only repopulated after Timor-Leste gained independence in 2002.
 The AFS in Osso-Luga lay fallow during the period of abandonment.
+The land use of these 2 villages is typical of the Baucau district.
+Land use in these two villages is typical of the Baucau district.
+The 5 AFS described in this article [see detailed description in @Cogne2024] are found in the valleys and high plateaux, while non-AFS such as rainfed rice and horticultural plots are more likely to be found in the lowlands.
+Other non-agricultural land uses are also present, such as limestone and limestone quarrying.
+Some natural forest parks are also preserved around mountain peaks, such as Matebian or Mundo Perdido.
+Villages are concentrated near paved roads and on hilltops, more rarely near rivers because of the risk of flooding.
 
@@
 
-We used the typology described in @Cogne2024, which was developed in the same study area.
-This typology was based on the one proposed by Nair et al.
-(1993, 2021) and adapted to the local context with semi-structured interviews conducted in the same study area [@Cogne2024].
+We used the typology of AFS described in @Cogne2024, which was developed in the same study area.
+This typology was based on the one proposed by Nair *et al.* (1993, 2021) and adapted to the local context with semi-structured interviews previously conducted in another study at the same sites [@Cogne2024].
 The five AFS identified were as follows: Crop systems including a wooded Fallow phase (CF), Silvopastures (SP), Young Agroforests (YA), Home Gardens (HG), and Forest Gardens (FG).
-CF are plantations of a few non-perennial crops (such as maize, sweet potatoes, groundnuts, and squash) which are left fallow after one to three years.
-SP are vast areas (generally 200 to 500 ha) where cattle are grazed under the shade of trees.
-YA are agricultural plots where farmers plant rows of trees spaced approximately 5 x 10 m apart, which they maintain for two to ten years.
-HG are plots of land located close to houses where crops and livestock (poultry, pigs, cattle, etc.) are kept in the shade of numerous trees and palms for multiple uses (wood, fruit, fiber, etc.).
-FG are generally former HG (over 50 years old) which have been abandoned, for example as a result of the war, and are now used mainly for gathering with limited human influence.
-
-The AFS are presented in more detail in the supplemental information section (Table\ S\@ref(tab:afs-typology)) and in @Cogne2024.
+The AFS are presented in more detail in the supplemental information section (Table S\@ref(tab:afs-typology)) and in @Cogne2024.
+
+Briefly, the CF are plantations of a few non-perennial crops (e.g., maize [*Zea mays*], sweet potatoes [*Ipomoea batatas*], groundnuts [*Arachis hypogaea*], and squash gourd [*Cucurbita moschata*]) which are left fallow after one to three years of cropping.
+SP are vast areas (generally 200 to 500 ha) where cattle (*Bos taurus*, *Bos domesticus*, *Bubalus bubalis*) are grazed under the shade of palms, like coconut (*Cocos nucifera*), and trees, like tamarind (*Tamarindus indica*) or eucalyptus (*Eucalyptus alba*) and jujube (*Ziziphus sp.*).
+YA are agricultural plots where farmers plant rows of trees, like teak (*Tectona grandis*) and mahogany (*Swietenia mahagoni*), spaced approximately 5 x 10 m apart, which are maintained for two to ten years and then either cut for wood or left to grow to become HG or FG (see Figure S\@ref(fig:seq-diagram)).
+HG are plots of land located close to houses where crops and livestock (poultry [*Gallus gallus domesticus*], pigs [*Sus domesticus*], cattle, etc.) are kept in the shade of numerous trees, like Breadfruit (*Artocarpus altilis*), candlenut (*Aleurites moluccanus*) and palms, like areca palm (*Areca catechu*), for multiple uses (wood, fruit, fiber, etc.).
+FG are generally former HG (over 50 years old) which have been abandoned, for example as a result of the war, and are now used mainly for gathering fruits (e.g.
+mango [*Mangifera indica*]) and non-timber forest products (e.g.
+candlenut) with limited human influence.
 
@@
 
-We randomly selected three plots from each of the five traditional AFS identified in each of the two sites (30 plots in total; Figure\ \@ref(fig:map-sites)).
+We randomly selected three plots from each of the five traditional AFS identified in each of the two sites (30 plots in total; Figure \@ref(fig:map-sites)).
 Sampling plots were defined as the entire agricultural plot (area between 0.06 and 1.25 ha) for all AFS except SP plots, which were generally much larger than 1 ha.
@@
 Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
-Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Figure\ S\@ref(fig:design)).
+Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Figure S\@ref(fig:design)).
 Within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
@@
 
-In each of the 30 sample plots, a composite of three samples was taken from three of the five 10 x 10 m subplots, following a transect covering the intra-plot variability and at a minimum distance of 1 m from any tree (Figure\ S\@ref(fig:design)).
+In each of the 30 sample plots, a composite of three samples was taken from three of the five 10 x 10 m subplots, following a transect covering the intra-plot variability and at a minimum distance of 1 m from any tree (Figure S\@ref(fig:design)).
 Soils were sampled with a 15 cm auger at 0-10 cm and 20-30 cm depth; due to the size of the auger we could not accurately sample the intermediate soil layer.
 Soils samples were then air-dried and sieved at 2 mm.
-They underwent standard physical and chemical analyses at Cirad laboratories in Montpellier, France (Table\ S\@ref(tab:site-soil)).
+They underwent standard physical and chemical analyses at Cirad laboratories in Montpellier, France (Table S\@ref(tab:site-soil)).
 Total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
@@
 Each sample was dried at 105°C in a classic oven until a stable dry mass was reached and then sieved to 2 mm.
-The fine fraction (< 2 mm) was weighed to be used in the SOC stocks calculation (equation \@ref(eq:soc)).
+The fine fraction (\< 2 mm) was weighed to be used in the SOC stocks calculation (equation \@ref(eq:soc)).
 
@@
 
-SOC stocks (in Mg\ C\ ha$^{-1}$) in plot $p$ were calculated as follows:
+SOC stocks (in Mg C ha$^{-1}$) in plot $p$ were calculated as follows:
 
@@
 This calculation corresponds to the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil.
-At depth $i$ = 10-20 cm, where SOC content was not measured (see "Soil measurements" section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a negative exponential model of SOC content decrease with soil depth [@Mishra2009;@Murphy2019]:
+At depth $i$ = 10-20 cm, where SOC content was not measured (see "Soil measurements" section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a negative exponential model of SOC content decrease with soil depth [@Mishra2009; @Murphy2019]:
 
@@
 
-### Statistical tests
-
-The estimated AGC and SOC stocks were log-transformed for all statistical tests because, as they only take positive values, we expected them to be log-normally distributed.
-We removed one CF plot at Gariuai from the following statistical tests, as it did not contain any trees exceeding the minimum measurement size (CBH = 30 cm) and its AGC stocks were therefore estimated at 0\ Mg\ C\ ha$^{-1}$ (even if there were trees < 30 cm CBH) and could not be log-transformed.
-After testing for normality (Shapiro-Wilks test) and homoscedasticity (Levene’s test), we performed a two-way ANOVA on the log-transformed estimates of AGC and SOC stocks, grouped by AFS and site.
-To test the relationship between log-transformed SOC and AGC stocks, we fitted two linear mixed models.
-We predicted SOC stocks at two different depths (either 0-10 or 0-30 cm) as a function of AGC stocks (fixed effect) and site (Samalari or Osso-Luga; random effect).
-
 ### Uncertainty propagation
 
-The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo function from the R package BIOMASS, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013] and large and small errors for 5 and 95% of trees, respectively [@Chave2004;@Rejou-Mechain2017].
+The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo function from the R package BIOMASS, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013] and large and small errors for 5 and 95% of trees, respectively [@Chave2004; @Rejou-Mechain2017].
 To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (as recommended by Cirad laboratories in Montpellier, France), and 10% for the mass of the fine fraction [@Page1999].
@@
 For multiplot estimates of carbon stocks, values within each grouping factor were additionally bootstrapped before calculating quantiles (median and 95% confidence intervals).
+
+### Model calibration
+
+We calibrated the following models for AGC stocks of plot $i$ in site $k$ and AFS type $j$:
+
+$$AGC_{i,j,k}  = \mu_A + \alpha_{A,j} + \beta_{A,k} + \epsilon_{A,i}$$
+with $\mu_A$ the average AGC stocks in our study area, $\alpha_{A,j}$ and  $\beta_{A,k}$ the random AFS and site effects, respectively distributed as:
+
+$$\alpha_{A,j} \sim \mathcal{N}(0,\sigma_{\alpha A}^2)$$
+
+$$\beta_{A,k} \sim \mathcal{N}(0,\sigma_{\beta A}^2)$$
+And $\epsilon_{A,i}$ the random error, distributed as:
+
+$$\epsilon_{A,i} \sim \mathcal{N}(0,\sigma_A^2)$$
+$\sigma_A$, $\sigma_{\alpha A}$, and $\sigma_{\beta A}$ are the standard deviation of the random error, the AFS effect and the site effect respectively.
+
+Similarly, we modelled SOC stocks of plot $i$ in site $k$ and AFS type $j$ as:
+
+$$SOC_{i,j,k} = \mu_S +  \alpha_{Sj} + \beta_{Sk} + \theta_1 agc_i + \theta_2 fine_i+ \epsilon_{S,i}$$
+with $\mu_S$ the average SOC stocks in our study area; $\theta_1$ the effect of AGC stocks on SOC stocks and $agc_i$ the standardised AGC stocks of plot $i$; $\theta_2$ the effect of fine soil fraction (clay + fine silt) on SOC stocks and $fine_i$ the standardised fine soil fraction of plot $i$.
+We standardised AGC stocks and fine soil fraction by scaling and centring values.
+
+$\alpha_{S,j}$ and $\beta_{S,k}$ are the random AFS and site effects, respectively distributed as:
+
+$$\alpha_{S,j} \sim \mathcal{N}(0,\sigma_{\alpha S}^2)$$
+
+$$\beta_{S,k} \sim \mathcal{N}(0,\sigma_{\beta S}^2)$$
+And $\epsilon_{S,i}$ the random error, distributed as:
+
+$$\epsilon_{S,i} \sim \mathcal{N}(0,\sigma_S^2)$$
+$\sigma_S$, $\sigma_{\alpha S}$, and $\sigma_{\beta S}$ are the standard deviation of the random error, the AFS effect and the site effect respectively.
+
+Calibration was carried out using an adapted form of the Hamiltonian Monte Carlo using Stan’s programming language [@Carpenter2017].
+We used the standard half-normal distribution as the prior for all standard deviation parameters, and the standard normal distribution for both $\theta_1$ and $\theta_2$ parameters.
+To propagate uncertainties on the model outputs, we calibrated the model for each of the 1000 iterations and sampled one set of parameter values in the posterior distributions for each iteration.
 
@@
 
-Total carbon stocks ranged between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-The AFS with the lowest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+Total carbon stocks ranged between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+The AFS with the lowest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
 
@@
 
-AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table\ \@ref(tab:anova)), but not between sites (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
-The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]`\ Mg\ C\ ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]`\ Mg\ C\ ha$^{-1}$ (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
+AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)), but not between sites (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
+The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
 
@@
 
-SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]`\ Mg\ C\ ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`%  (Table\ \@ref(tab:summary); Figure\ \@ref(fig:barplot)).
-No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table\ \@ref(tab:anova)).
-However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table\ \@ref(tab:anova)).
-The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$) than those in Osso-Luga (`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]`\ Mg\ C\ ha$^{-1}$; Figure\ \@ref(fig:barplot)).
+SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)).
+However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
+The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$) than those in Osso-Luga (`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$; Figure \@ref(fig:barplot)).
 
@@
 
-The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure\ \@ref(fig:plot-lmereg)).
+The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure \@ref(fig:plot-lmereg)).
 This relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
-The conditional R^2^ (i.e., the proportion of variance explained by the linear mixed models) were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths, respectively, but the marginal R^2^ (i.e., when considering only the fixed effect) were only `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`%, respectively (Figure\ \@ref(fig:plot-lmereg)).
+The conditional R^2^ (i.e., the proportion of variance explained by the linear mixed models) were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths, respectively, but the marginal R^2^ (i.e., when considering only the fixed effect) were only `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`%, respectively (Figure \@ref(fig:plot-lmereg)).
 
@@
 Compared to values reported in the literature for other AFS, the carbon stocks in this study were relatively high.
-Specifically, the average SOC stocks (`r dataC_plot[variable=="SOC", round(mean(mean))]`\ Mg\ C\ ha$^{-1}$) were higher than other reported values.
-In a meta-analysis by @Shi2018, it was estimated that AFS stored around 50\ Mg\ C\ ha$^{-1}$ at a 0-30 cm depth.
-Such large SOC stocks could be due to the high exchangeable Ca^2+^ content at the sites included in this study (Table\ \@ref(tab:site-soil)), as Ca^2+^ is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
-In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`\ Mg\ C\ ha$^{-1}$) were similar to the average of 74\ Mg\ C\ ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
-The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]`\ Mg\ C\ ha$^{-1}$) close to those of old-growth tropical forests [270\ Mg\ C\ ha$^{-1}$, @Lal2005].
-The high carbon stocks in FG can be attributed to the length of time they had been established, which was generally longer than 50 years, with little human intervention (Table\ S\@ref(tab:afs-typology)).
+Specifically, the average SOC stocks (`r dataC_plot[variable=="SOC", round(mean(mean))]` Mg C ha$^{-1}$) were higher than other reported values.
+In a meta-analysis by @Shi2018, it was estimated that AFS stored around 50 Mg C ha$^{-1}$ at a 0-30 cm depth.
+Such large SOC stocks could be due to the high exchangeable Ca^2+^ content at the sites included in this study (Table \@ref(tab:site-soil)), as Ca^2+^ is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
+In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]` Mg C ha$^{-1}$) were similar to the average of 74 Mg C ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
+The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]` Mg C ha$^{-1}$) close to those of old-growth tropical forests [270 Mg C ha$^{-1}$, @Lal2005].
+The high carbon stocks in FG can be attributed to the length of time they had been established, which was generally longer than 50 years, with little human intervention (Table S\@ref(tab:afs-typology)).
 In terms of structure and carbon stocks, FG should therefore resemble old secondary forests, which have been shown to recover more than 60% of the AGC stocks and almost all the SOC stocks of an old forest [@Poorter2021].
-In addition, FG contain a large number of trees of the same age that were planted when the AFS was established and can survive to reach significant sizes, similar to forest plantations that can grow to old-growth-forest levels of AGC stocks after only a few decades [@Brown2020;@Njoukam1996].
-
-Although agroforestry has been identified as one of the possible levers to increase carbon storage in the Timorese agricultural landscape [@UNFCCC2020;@Paudel2022], to our knowledge no study has quantified carbon stocks in these systems.
+In addition, FG contain a large number of trees of the same age that were planted when the AFS was established and can survive to reach significant sizes, similar to forest plantations that can grow to old-growth-forest levels of AGC stocks after only a few decades [@Brown2020; @Njoukam1996].
+
+Although agroforestry has been identified as one of the possible levers to increase carbon storage in the Timorese agricultural landscape [@UNFCCC2020; @Paudel2022], to our knowledge no study has quantified carbon stocks in these systems.
 Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste's climate change mitigation commitments [@UNFCCC2020] and the potential carbon gains from a wider adoption of these techniques.
@@
 It is also interesting to note that, in line with previous results, the age of the trees in the AFS largely explains the carbon stored in their biomass [@Ma2020].
-The AFS described in this study reflect different stages of transition in land use, which follow the growth of their trees (Figure\ S\@ref(fig:seq-diagram)).
+The AFS described in this study reflect different stages of transition in land use, which follow the growth of their trees (Figure S\@ref(fig:seq-diagram)).
 For example, a farmer might plant trees in a CF field, which will then become a YA that can still produce annual crops for the first 2 to 4 years, before the shade from the trees becomes too great.
 The trees are then used as property markers and, as they mature, produce wood and fruits; the YA then becomes an HG or FG.
-Similarly, our results show that AGC stocks increase from CF to YA, HG and FG (Table\ \@ref(tab:summary)).
+Similarly, our results show that AGC stocks increase from CF to YA, HG and FG (Table \@ref(tab:summary)).
 
@@
 The same applies to FG, which are former HG that have been more or less abandoned, and YA, where the farmer has chosen the site with the intention of turning it into a HG.
-We therefore expect HG, FG and YA to be located in areas that are more favorable for tree growth [@Wagner2012;@Madejon2016;@Treuer2018].
+We therefore expect HG, FG and YA to be located in areas that are more favorable for tree growth [@Wagner2012; @Madejon2016; @Treuer2018].
 Conversely, CF and SP are mainly located on limestone plateaus where water tables are deeper and soils are shallower, which could reduce tree growth and biomass [@Imada2008].
@@
 We found no significant difference in SOC between the AFS (H1b rejected).
-This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017;@Wiesmeier2019], and with previous studies that show significant differences in SOC stocks between AFS [e.g., @Ramesh2015].
-However, it is worth noting that some single-site studies [e.g., @Betemariyam2020;@GamaRodrigues2010] and at least two meta-analyses failed to find any sizable effect of AFS on SOC stocks.
+This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017; @Wiesmeier2019], and with previous studies that show significant differences in SOC stocks between AFS [e.g., @Ramesh2015].
+However, it is worth noting that some single-site studies [e.g., @Betemariyam2020; @GamaRodrigues2010] and at least two meta-analyses failed to find any sizable effect of AFS on SOC stocks.
 @Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by the type of AFS, and @Shi2018 found no significant differences in SOC stocks between four different AFS in 354 previous studies.
@@
 First, in a global meta-analysis, @Ma2020 estimated that it takes only five years for SOC stocks in tropical AFS to reach an equilibrium value.
-In our study, at least three of the five AFS (namely YA, HG and FG) represent different stages of an AFS transition depending on the age of the trees (Figure\ S\@ref(fig:seq-diagram)).
+In our study, at least three of the five AFS (namely YA, HG and FG) represent different stages of an AFS transition depending on the age of the trees (Figure S\@ref(fig:seq-diagram)).
 Most of these AFS are more than five years old and may have already reached their equilibrium SOC stocks, which could explain why there was little difference between them.
@@
 Increasing the number of replicates for each AFS could have increased the power of the tests, and potentially resulted in more significant differences between AFS.
-Third, other pedoclimatic and historical factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g., within-site variability in climate, soil texture (Figures\ S\@ref(fig:clay-corg), S\@ref(fig:altitude-corg)) or previous land use may have significantly influenced SOC stocks [@Wiesmeier2019;@Ma2020].
+Third, other pedoclimatic and historical factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g., within-site variability in climate, soil texture (Figures S\@ref(fig:clay-corg), S\@ref(fig:altitude-corg)) or previous land use may have significantly influenced SOC stocks [@Wiesmeier2019; @Ma2020].
 Our experimental design aimed to limit as much as possible the influence of external co-factors at each site (e.g., variability in farmers' management practices, micro-topography, etc.), but much of the variability cannot be controlled in farmers' plots.
@@
 We found a significant difference in SOC between the two sites, with higher SOC stocks at Gariuai than at Osso-Luga.
-A priori, three factors that differentiate the sites could explain these differences in SOC stocks: climate (through elevation; Figure\ \@ref(fig:map-sites); Figure\ S\@ref(fig:altitude-corg)), soil texture (Table\ S\@ref(fig:site-soil); Figure\ S\@ref(fig:clay-corg)) and history.
-First, we found that SOC stocks were positively correlated with elevation, both between and within sites (Figure\ S\@ref(fig:altitude-corg)).
+A priori, three factors that differentiate the sites could explain these differences in SOC stocks: climate (through elevation; Figure \@ref(fig:map-sites); Figure S\@ref(fig:altitude-corg)), soil texture (Table S\@ref(fig:site-soil); Figure S\@ref(fig:clay-corg)) and history.
+First, we found that SOC stocks were positively correlated with elevation, both between and within sites (Figure S\@ref(fig:altitude-corg)).
 This result is consistent with those of @Raich2006, who found higher SOC at higher elevations in tropical evergreen forests in various countries (e.g., Congo, Thailand).
-As changes in elevation are associated with variations in climate (precipitation, temperature), our result is also consistent with previous studies that found higher SOC in areas with higher precipitation [@Schuur2001] and lower temperature [@Raich2006;@MarinSpiotta2013].
-Second, soil texture is a key driving process for SOC storage in the tropics [@Don2010;@Wiesmeier2019;@Matus2021].
+As changes in elevation are associated with variations in climate (precipitation, temperature), our result is also consistent with previous studies that found higher SOC in areas with higher precipitation [@Schuur2001] and lower temperature [@Raich2006; @MarinSpiotta2013].
+Second, soil texture is a key driving process for SOC storage in the tropics [@Don2010; @Wiesmeier2019; @Matus2021].
 In particular, many studies have found a strong positive correlation between SOC stocks and the fine mineral fractions of soils [i.e., clay or clay + fine silt content; @Zinn2005].
-In our study, we found no significant correlation between SOC stocks and clay + fine silt content when site and depth were included in the model (Figure\ S\@ref(fig:clay-corg)).
+In our study, we found no significant correlation between SOC stocks and clay + fine silt content when site and depth were included in the model (Figure S\@ref(fig:clay-corg)).
 Furthermore, the Gariuai soils, which had the highest SOC stocks, were also the least clayey.
@@
 @Ojoatre2024 found similar results in a secondary tropical mountain forest.
-These results suggest that there is little correlation between AGC and SOC stocks once AFS are more than five years old, which is the case in our study for FG, HG (> 5 years old) and partially for YA (2 to 10 years; Table\ S\@ref(tab:afs-typology)).
+These results suggest that there is little correlation between AGC and SOC stocks once AFS are more than five years old, which is the case in our study for FG, HG (\> 5 years old) and partially for YA (2 to 10 years; Table S\@ref(tab:afs-typology)).
 This saturation of SOC stocks even with increasing AGC stocks is counterintuitive, as it might be expected that higher AGC stocks would lead to more litter production and decomposition, some of which would increase SOC stocks.
@@
 In addition to biomass stocks, other vegetation characteristics may account for variations in soil carbon stocks between AFS.
-For instance, previous studies have shown that tree diversity can enhance SOC sequestration in natural ecosystems [@Spohn2023;@Chen2018] and in AFS [@Manaye2021;@Islam2015].
+For instance, previous studies have shown that tree diversity can enhance SOC sequestration in natural ecosystems [@Spohn2023; @Chen2018] and in AFS [@Manaye2021; @Islam2015].
 Moreover, total SOC may be less sensitive to management practices (such as biomass inputs) than specific fractions of SOC [@Haynes2005].
@@
 Although agroforestry can make agriculture more resilient to climate change while improving its carbon footprint and biodiversity conservation value [@Cardinael2021], poor adaptation to the local context can explain many failures to adopt new agroforestry techniques [@Coe2014].
-Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and have adapted to this context [@Gouyon1993;@Aumeeruddy1994;@Peltier1996].
+Conversely, some traditional agroforestry techniques have evolved in a particular context (biophysical, social and economic) and have adapted to this context [@Gouyon1993; @Aumeeruddy1994; @Peltier1996].
 This is the case of the traditional AFS studied here in Timor-Leste.
@@
 
-All data collected for this study is available in the following Dataverse repository ______.
-
-The R codes used to analyze the data and produce the manuscript (including figures and tables) are available on ______.
+All data collected for this study is available in the following Dataverse repository \_\_\_\_\_\_.
+
+The R codes used to analyze the data and produce the manuscript (including figures and tables) are available on \_\_\_\_\_\_.
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
-  $$c(x) = C_0 \cdot exp(-\lambda\cdot x) $$ 
-  where $C_0$ is the carbon content at the surface (depth $x= 0$) and $\lambda$ is the rate at which the carbon content decreases with soil depth. 
+$$c(x) = C_0 \cdot exp(-\lambda\cdot x) $$ where $C_0$ is the carbon content at the surface (depth $x= 0$) and $\lambda$ is the rate at which the carbon content decreases with soil depth.
 Under this assumption, we can calculate the mean SOC content between depths $x$ and $x+\Delta$ as:
@@
 
-  $$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{\lambda \cdot \Delta} exp(-\lambda\cdot x)\left[1- exp(-\lambda\cdot\Delta) \right]$$
-  The average SOC contents between 0-10 cm, 10-20 cm and 20-30 cm can respectively be written as:
+$$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{\lambda \cdot \Delta} exp(-\lambda\cdot x)\left[1- exp(-\lambda\cdot\Delta) \right]$$ The average SOC contents between 0-10 cm, 10-20 cm and 20-30 cm can respectively be written as:
 
@@
 
-  $$C_{20,30} = \frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)$$
-  We can thus write:
-
-  $$\frac{C_{20,30}}{C_{0,10}} = \frac{\frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)}{\frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)} = exp(-20\lambda)$$
-  $$ \Rightarrow \lambda = \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20}$$
-  We can also write:
-
-  $$ \frac{C_{10,20}}{C_{0,10}} = exp(-10\lambda) $$
-  $$\Rightarrow C_{10,20} = C_{0,10} \cdot exp(-10 \cdot \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20})$$
+$$C_{20,30} = \frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)$$ We can thus write:
+
+$$\frac{C_{20,30}}{C_{0,10}} = \frac{\frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)}{\frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)} = exp(-20\lambda)$$ $$ \Rightarrow \lambda = \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20}$$ We can also write:
+
+$$ \frac{C_{10,20}}{C_{0,10}} = exp(-10\lambda) $$ $$\Rightarrow C_{10,20} = C_{0,10} \cdot exp(-10 \cdot \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20})$$
 
@@
 
-  ## Tables
+\## Tables
 
@@
 
-  ## Figures
+\## Figures
 

### 2024-12-20 manuscript.Rmd
restructure manuscript with new model and elevation gradient

@@
 
+[paragraph on possible drivers]
+C stocks (in biomass and soil) of AFS could be driven by a number of factors: 
+- agricultural practices (ref)
+- climate (ref)
+- soil properties (especially for SOC) > in particular, fine content
+
 In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying the carbon storage in both vegetation biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
-We hypothesize that there are measurable differences in carbon in aboveground vegetation biomass (hereinafter referred to as above ground carbon or AGC; Hypothesis 1a) and SOC (Hypothesis 1b) stocks between these different types of traditional AFS.
+[xxx 2 sites] We hypothesize that there are measurable differences in carbon in aboveground vegetation biomass (hereinafter referred to as above ground carbon or AGC; Hypothesis 1a) and SOC (Hypothesis 1b) stocks between these different types of traditional AFS.
 In line with @Ma2020, we anticipate that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; Hypothesis 2a), but no relationship with SOC below 10 cm (Hypothesis 2b).
@@
 # get coordinates
-data_coord <- data_coord[, .(long = mean(as.numeric(LONG)), lat = mean(as.numeric(LAT))), .(N_PLOT)]
+data_coord <- data_coord[, .(
+  long = mean(as.numeric(LONG)), 
+  lat = mean(as.numeric(LAT))
+), .(N_PLOT)]
 
@@
 ```{r get-env-data}
-# precipitation from CHELSA
-if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
-  url <- "https://os.zhdk.cloud.switch.ch/envicloud/chelsa/chelsa_V2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio12_1981-2010_V.2.1.tif"
-  destfile <- "data/CHELSA_bio12_1981-2010_V.2.1.tif"
+# temperature and precipitation from CHELSA
+if (!file.exists("data/CHELSA_bio1_1981-2010_V.2.1.tif")) {
+  url <- "https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio1_1981-2010_V.2.1.tif"
+  destfile <- "data/CHELSA_bio1_1981-2010_V.2.1.tif"
   download.file(url, destfile, method = "curl")
@@
 
+if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
+  url <- "https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio12_1981-2010_V.2.1.tif"
+  destfile <- "data/CHELSA_bio12_1981-2010_V.2.1.tif"
+  download.file(url, destfile, method = "curl")
+
+  ## crop to reduce memory usage
+  box = c(124.5, 127.5, -9.2, -8.3) # Timor-Leste
+  rst <- terra::rast(destfile)
+  rst <- terra::crop(rst, terra::ext(box))
+  writeRaster(rst, filename = destfile, overwrite = TRUE) 
+}
+
 # 30m resolution SRTM
-if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
+if (!file.exists("data/srtm_62_14.zip")) {
   url <- "https://srtm.csi.cgiar.org/wp-content/uploads/files/srtm_5x5/TIFF/srtm_62_14.zip"
@@
 ```{r extract-env-data}
+# extract elevation, precipitation and temperature for all plots
+data_plot[, `:=`(
+  elev = terra::extract(terra::rast("data/srtm_62_14.tif"), 
+                        cbind(long, lat))[[1]], 
+  prec = terra::extract(terra::rast("data/CHELSA_bio12_1981-2010_V.2.1.tif"), 
+                        cbind(long, lat))[[1]], 
+  temp = terra::extract(terra::rast("data/CHELSA_bio1_1981-2010_V.2.1.tif"), 
+                        cbind(long, lat))[[1]]
+)]
+
+cor(data_plot[, c("elev", "prec", "temp")])
+
 # extract elevation and rainfall for the 2 sites
@@
 SP are vast areas (generally 200 to 500 ha) where cattle (*Bos taurus*, *Bos domesticus*, *Bubalus bubalis*) are grazed under the shade of palms, like coconut (*Cocos nucifera*), and trees, like tamarind (*Tamarindus indica*) or eucalyptus (*Eucalyptus alba*) and jujube (*Ziziphus sp.*).
-YA are agricultural plots where farmers plant rows of trees, like teak (*Tectona grandis*) and mahogany (*Swietenia mahagoni*), spaced approximately 5 x 10 m apart, which are maintained for two to ten years and then either cut for wood or left to grow to become HG or FG (see Figure S\@ref(fig:seq-diagram)).
+YA are agricultural plots where farmers plant rows of trees, like teak (*Tectona grandis*) and mahogany (*Swietenia mahagoni*), spaced approximately 5 x 10 m apart, which are generally left to grow to become HG or FG (see Figure S\@ref(fig:seq-diagram)).
 HG are plots of land located close to houses where crops and livestock (poultry [*Gallus gallus domesticus*], pigs [*Sus domesticus*], cattle, etc.) are kept in the shade of numerous trees, like Breadfruit (*Artocarpus altilis*), candlenut (*Aleurites moluccanus*) and palms, like areca palm (*Areca catechu*), for multiple uses (wood, fruit, fiber, etc.).
@@
 
-### Tree measurements
-
 Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
@@
 
-### Soil measurements
-
-In each of the 30 sample plots, a composite of three samples was taken from three of the five 10 x 10 m subplots, following a transect covering the intra-plot variability and at a minimum distance of 1 m from any tree (Figure S\@ref(fig:design)).
-Soils were sampled with a 15 cm auger at 0-10 cm and 20-30 cm depth; due to the size of the auger we could not accurately sample the intermediate soil layer.
+In each of the 30 sample plots, a composite of three samples was taken from three of the five 10 x 10 m subplots (including the plot center and two opposite corners; Figure S\@ref(fig:design)), covering the intra-plot variability and at a minimum distance of 1 m from any tree.
+Soils were sampled with a soil auger ( Model Bor Tanah - Dormer Standard Soil Auger, Adiguna Karya Persada, Bogor, Indonesia) at 0-10 cm and 20-30 cm depth,.
+As the length of the head auger (15 cm) was longer than our target soil depths (10 cm), we could not accurately sample the 10-20 cm soil layer.
 Soils samples were then air-dried and sieved at 2 mm.
 They underwent standard physical and chemical analyses at Cirad laboratories in Montpellier, France (Table S\@ref(tab:site-soil)).
-Total organic carbon was determined by dry combustion and measured with a thermal conductive detector.
-
-In addition, at one location at the center of the plot, soils were sampled with a cylinder of 250 cm^3^ at depths of 2-7 cm, 12-17 cm and 22-27 cm.
+Total carbon content was determined by dry combustion, following the ISO 10694:1995 standard method, and measured with a thermal conductive detector (EA 1112 Elemental Analyser, Thermo Fischer Scientific, Netherlands).
+Carbonate content was determined with a calcimeter, following the ISO 10693 standard method.
+Soil organic carbon content was calculated as the difference of the total carbon and the carbonate contents.
+
+In addition, at one location in the center of the plot, three 250 cm$^3$ soil samples were taken with a soil sampling cylinder (5 cm in height, 7.98 cm inner diameter, and 8.4 cm outer diameter; Eijkelkamp Soil & Water, Giesbeek, Netherlands)  at depths of 2-7 cm, 12-17 cm and 22-27 cm.
+These depths were chosen to take account of the height of the cylinder (i.e.
+5 cm), the destruction of the soil below the cylinder during its extraction, and to be associated with depths 0-10 cm, 10-20 cm and 20-30 cm, the depths at which the organic carbon content was calculated or estimated.
 Each sample was dried at 105°C in a classic oven until a stable dry mass was reached and then sieved to 2 mm.
@@
 
-$$SOC_p = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl}$$
-
-Where $Corg_i$ is the SOC content at depth $i$ (either 0-10, 10-20 or 20-30 cm); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth $i$ (in g), converted into Mg with a factor of 10$^{-6}$; $Splot$ is the surface of a 1-ha plot (10$^8$ cm$^2$); $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
-This calculation corresponds to the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil.
-At depth $i$ = 10-20 cm, where SOC content was not measured (see "Soil measurements" section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a negative exponential model of SOC content decrease with soil depth [@Mishra2009; @Murphy2019]:
+Where $Corg_i$ is the SOC content at depth i (either 0-10, 10-20 or 20-30 cm) as measured in the laboratory from the soil sampled by the auger and estimated at depth 10-20 cm (see details below); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth i (in g), converted into Mg with a factor of 10$^-6$; $Surf\ =\ 10^8$ is the surface conversion factor from cm$^2$ to ha; $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
+This calculation corresponds to the method M4 recommended by @Poeplau2017, which correctly considers the coarse fraction of the soil.
+We considered that $mfine_i$ at depths of 2-7 cm, 12-17 cm and 22-27 cm was associated with $Corg_i$ at depths of 0-10, 10-20 cm and 20-30 cm, respectively.
+At depth $i$ = 10-20 cm, where the carbon content was not measured (see “Soil measurements” section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a standard negative exponential model of SOC content decrease with soil depth [@Mishra2009;@Murphy2019]:
 
 \begin{equation}
-Corg_{10,20} = Corg_{0,10} \cdot \sqrt{\frac{Corg_{20,30}}{Corg_{0,10}}} (\#eq:intersoc)
+Corg_{10,20}=Corg_{0,10}\cdot\sqrt{\frac{Corg_{20,30}}{Corg_{0,10}}}
+(\#eq:soc)
 \end{equation}
@@
 data_sim <- merge(data_sim, data_soil_summ)  
+# add elevation
+data_sim <- merge(data_sim, data_plot[, c("N_PLOT", "elev")])
 ```
@@
 
-We calibrated the following models for AGC stocks of plot $i$ in site $k$ and AFS type $j$:
-
-$$AGC_{i,j,k}  = \mu_A + \alpha_{A,j} + \beta_{A,k} + \epsilon_{A,i}$$
-with $\mu_A$ the average AGC stocks in our study area, $\alpha_{A,j}$ and  $\beta_{A,k}$ the random AFS and site effects, respectively distributed as:
-
-$$\alpha_{A,j} \sim \mathcal{N}(0,\sigma_{\alpha A}^2)$$
-
-$$\beta_{A,k} \sim \mathcal{N}(0,\sigma_{\beta A}^2)$$
-And $\epsilon_{A,i}$ the random error, distributed as:
-
-$$\epsilon_{A,i} \sim \mathcal{N}(0,\sigma_A^2)$$
-$\sigma_A$, $\sigma_{\alpha A}$, and $\sigma_{\beta A}$ are the standard deviation of the random error, the AFS effect and the site effect respectively.
-
-Similarly, we modelled SOC stocks of plot $i$ in site $k$ and AFS type $j$ as:
-
-$$SOC_{i,j,k} = \mu_S +  \alpha_{Sj} + \beta_{Sk} + \theta_1 agc_i + \theta_2 fine_i+ \epsilon_{S,i}$$
-with $\mu_S$ the average SOC stocks in our study area; $\theta_1$ the effect of AGC stocks on SOC stocks and $agc_i$ the standardised AGC stocks of plot $i$; $\theta_2$ the effect of fine soil fraction (clay + fine silt) on SOC stocks and $fine_i$ the standardised fine soil fraction of plot $i$.
-We standardised AGC stocks and fine soil fraction by scaling and centring values.
-
-$\alpha_{S,j}$ and $\beta_{S,k}$ are the random AFS and site effects, respectively distributed as:
-
-$$\alpha_{S,j} \sim \mathcal{N}(0,\sigma_{\alpha S}^2)$$
-
-$$\beta_{S,k} \sim \mathcal{N}(0,\sigma_{\beta S}^2)$$
-And $\epsilon_{S,i}$ the random error, distributed as:
-
-$$\epsilon_{S,i} \sim \mathcal{N}(0,\sigma_S^2)$$
-$\sigma_S$, $\sigma_{\alpha S}$, and $\sigma_{\beta S}$ are the standard deviation of the random error, the AFS effect and the site effect respectively.
+We modelled both AGC and SOC stocks as:
+
+$$Y_{i,j,k} = \mu +  \theta \cdot X_i + \alpha_{j} + \beta_{k} + \epsilon_{i}$$
+with $Y_{i,j,k}$ the stocks (either SOC or AGC) of plot $i$ in site $k$ and AFS type $j$
+
+$\mu$ is the average stocks in our study area.
+
+$\theta$ is the effect of the covariates on $Y$, and $X_i$ are the values of the covariates in plot $i$; they include elevation for AGC stocks [from NASA SRTM; @SRTM2024], or elevation [@SRTM2024], clay + fine soil content (from soil analysis; xxx methods) and AGC stocks for SOC stocks.
+All covariates were centred and scaled.
+Elevation was chosen as it was highly correlated (>95%) with both mean annual temperature and precipitation [@Karger2017]: it was thus a good indicator of a climate gradient in our dataset (see Supplementary information section xx).
+xxx justify these variables in the introduction
+
+$\alpha_{j}$ and $\beta_{k}$ are the random AFS and site effects, respectively distributed as:
+
+$$\alpha_{j} \sim \mathcal{N}(0,\sigma_{\alpha}^2)$$
+
+$$\beta_{k} \sim \mathcal{N}(0,\sigma_{\beta}^2)$$
+And $\epsilon_{i}$ the random error, distributed as:
+
+$$\epsilon_{i} \sim \mathcal{N}(0,\sigma^2)$$
+$\sigma$, $\sigma_{\alpha}$, and $\sigma_{\beta}$ are the standard deviation of the random error, the AFS effect and the site effect respectively.
 
 Calibration was carried out using an adapted form of the Hamiltonian Monte Carlo using Stan’s programming language [@Carpenter2017].
-We used the standard half-normal distribution as the prior for all standard deviation parameters, and the standard normal distribution for both $\theta_1$ and $\theta_2$ parameters.
+We used the standard half-normal distribution as the prior for all standard deviation parameters, and the standard normal distribution for all $\theta$ parameters.
 To propagate uncertainties on the model outputs, we calibrated the model for each of the 1000 iterations and sampled one set of parameter values in the posterior distributions for each iteration.
@@
   int<lower=1> S;
+  int<lower=1> K;
   int<lower=1,upper=AF> afs[N];
   int<lower=1,upper=S> site[N];
-  real clay_sc[N];
-  real agc_sc[N];
+  vector[N] elev;
+  matrix[N,K] covar;
   real<lower=0> agc[N];
@@
   real delta_site_a[S];
+  real theta_a;
   real<lower=0> sigma_a;
@@
   real delta_site_s[S];
-  real theta_agc_s;
-  real theta_clay_s;
+  vector[K] theta_s;
   real<lower=0> sigma_s;
@@
   real delta_site_s10[S];
-  real theta_agc_s10;
-  real theta_clay_s10;
+  vector[K] theta_s10;
   real<lower=0> sigma_s10;
@@
 model {
-  real pred_a[N];
-  real pred_s[N];
-  real pred_s10[N];
-
+  vector[N] pred_a;
+  vector[N] pred_s;
+  vector[N] pred_s10;
+
+  // predictions
   for (i in 1:N) {
-    pred_a[i] = mu_a + delta_afs_a[afs[i]] + delta_site_a[site[i]];
-    pred_s[i] = mu_s + theta_agc_s*agc_sc[i] + theta_clay_s*clay_sc[i] + delta_afs_s[afs[i]] + delta_site_s[site[i]];
-    pred_s10[i] = mu_s10 + theta_agc_s10*agc_sc[i] + theta_clay_s10*clay_sc[i] + delta_afs_s10[afs[i]] + delta_site_s10[site[i]];
+    pred_a[i] = mu_a + elev[i] * theta_a + delta_afs_a[afs[i]] + delta_site_a[site[i]];
+    pred_s[i] = mu_s + dot_product(row(covar, i), theta_s) + delta_afs_s[afs[i]] + delta_site_s[site[i]];
+    pred_s10[i] = mu_s10 + dot_product(row(covar, i), theta_s10) + delta_afs_s10[afs[i]] + delta_site_s10[site[i]];
   }
 
+  // likelihoods
   agc ~ normal(pred_a, sigma_a);
@@
   //  priors
-  theta_agc_s ~ normal(0,1);
-  theta_agc_s10 ~ normal(0,1);
-  theta_clay_s ~ normal(0,1);
-  theta_clay_s10 ~ normal(0,1);
+  theta_s ~ normal(0,1);
+  theta_s10 ~ normal(0,1);
   sigma_a ~ normal(0,1);
@@
 if (!file.exists("outputs/data_par.csv")) {
-  library(parallel)
-
-  cl <- makeCluster(detectCores()-1)
-  clusterEvalQ(cl, {library(rstan); library(data.table)})
-  clusterExport(cl, "data_sim"); clusterEvalQ(cl, data_sim)
-  clusterExport(cl, "stan_model"); clusterEvalQ(cl, stan_model)
-
-  data_par <- parLapply(cl, unique(data_sim$iter), function(i) {
+  library(rstan)
   rstan_options(auto_write = TRUE)
+
+  pars_prev <- list.files("outputs", patter = "data_par_") |> 
+    gsub(pattern = "data_par_|\\.csv", replacement = "") |> 
+    as.numeric()
+
+  for (n in setdiff(seq_len(100), pars_prev)) {
+    data_par <- lapply(paste0("V", seq((n-1)*10+1, n*10)), function(i) {
 
@@
         subset(iter==i) |> 
-      dcast(N_PLOT + site + ID_AF + fine_soil ~ variable) |> 
+        dcast(N_PLOT + site + ID_AF + fine_soil + elev ~ variable) |> 
         with(
@@
             S = length(unique(site)),
+            K = 3,
             afs = as.numeric(as.factor(ID_AF)), 
             site = as.numeric(as.factor(site)),
-          clay_sc = (fine_soil-mean(fine_soil))/sd(fine_soil),
-          agc_sc = (AGC-mean(AGC))/sd(AGC),
+            elev = (elev-mean(elev))/sd(elev),
+            covar = scale(cbind(elev, fine_soil, AGC)),
             agc = AGC,
@@
         rbindlist(idcol = "param")
-    return(pars)
+      return(pars[iter_stan==sample(iter_stan, 1), -"iter_stan"])
     }) |> rbindlist(idcol = "iter")
 
-  write.csv(data_par, file = "outputs/data_par.csv", row.names = FALSE)
-
-  stopCluster(cl)
-
-} else data_par <- read.csv("outputs/data_par.csv") |> setDT()
+    write.csv(data_par, 
+              file = paste0("outputs/data_par_", n, ".csv"), row.names = FALSE)
+  }
+
+  data_par <- list.files("outputs", pattern = "data_par_", full.names = TRUE) |> 
+    lapply(read.csv) |> rbindlist(idcol = "iter10") 
+  data_par[, `:=`(iter = (iter10-1)*10+iter, iter10 = NULL)]
+
+  write.csv(data_par, "outputs/data_par.csv", row.names = FALSE)
+
+  # list.files("outputs", pattern = "data_par_", full.names = TRUE) |>
+  #   file.remove()
+}
+
+data_par <- read.csv("outputs/data_par.csv") |> setDT()
 
 data_par[, effect := tstrsplit(param, "_")[[2]]]
-data_par[! effect %in% c("afs", "site", "agc", "clay"), effect := NA]
+data_par[! effect %in% c("afs", "site"), effect := NA]
 data_par[, stock := tstrsplit(param, "_")[[3]]]
@@
          level := levels(as.factor(data_sim$site))[level_num]]
-
+data_par[param == "theta",
+         level := c("elev", "fine", "agc")[level_num]]
+data_par[param == "theta" & stock == "agc", level := "elev"]
+
+# get levels of "delta" parameters
 df_mu_delta <- 
   data_par[param == "delta" & !is.na(effect)] |> 
-  dcast(iter + iter_stan + stock + effect + level ~ param) |> 
-  merge(dcast(data_par[param == "mu"], iter + iter_stan + stock ~ param))
+  dcast(iter + stock + effect + level ~ param) |> 
+  merge(dcast(data_par[param == "mu"], iter + stock ~ param))
 df_mu_delta[, `:=`(value = mu + delta, param = "pred_med")]
 data_par <- rbind(
-  data_par[, c("param", "iter", "iter_stan", "stock", "effect", "level", "value")], 
-  df_mu_delta[, c("param", "iter", "iter_stan",  "stock", "effect", "level", "value")]
+  data_par[, c("param", "iter", "stock", "effect", "level", "value")], 
+  df_mu_delta[, c("param", "iter", "stock", "effect", "level", "value")]
 )
@@
 
-```{r group-comparison}
-# shapiro wilk test - normality
-dataC_plot[variable %in% c("AGC", "SOC") & plot != 6, .(
-  shap = shapiro.test(log(mean))$p.value
-), .(afs, variable)]
-# levene test of homoscedasticity
-dataC_plot[variable %in% c("AGC", "SOC") & plot != 6, 
-           leveneTest(log(mean) ~ afs*site)$`Pr(>F)`[1], .(variable)]
-
-# 2-way anova 
-anova_table = 
-  dataC_plot[variable %in% c("AGC", "SOC") & plot != 6, 
-             .(pvalue = round(anova(lm(log(mean) ~ site + afs))$`Pr(>F)`[1:2], 3)), 
-             .(variable)]
-anova_table$factor = rep(c("site", "afs"), 2)
-```
-
-```{r lmereg}
-data_reg = melt(dataC_plot[plot != 6], measure.vars = c("mean", "lwr", "upr"), variable.name = "stat") |> 
-  dcast(plot + site+ afs ~ variable + stat)
-
-## mixed effect linear regression
-lmeReg <- lmer(log(SOC_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
-# Nakagawa’s R2 for mixed models
-R2 = r2_nakagawa(lmeReg)
-summary(lmeReg)
-
-# same for topsoil layer (0-10 cm)
-lmeReg10 <- lmer(log(SOC10_mean) ~ log(AGC_mean) + (1|site), data = data_reg)
-# Nakagawa’s R2 for mixed models
-R210 = r2_nakagawa(lmeReg10)
-summary(lmeReg10)
-
-dfreg = 
-  data.table(
-    depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
-    intercept = c(summary(lmeReg)$coefficients[1,1], summary(lmeReg10)$coefficients[1,1]), 
-    slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
-    pvalue = c(summary(lmeReg)$coefficients[2,5], summary(lmeReg10)$coefficients[2,5]),
-    R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
-    R2_marg = c(R2$R2_marginal, R210$R2_marginal)
-  )
-```
-
 # Results
 
-## Total carbon stocks
+AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+AGC stocks varied substantially between AFS, but not between sites (Figure \@ref(fig:effect-afs-site)).
+The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+
+SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+SOC stocks varied substantially between sites, but not between AFS (Figure \@ref(fig:effect-afs-site)).
+The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$) than those in Osso-Luga (`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$; Figure \@ref(fig:barplot)).
 
@@
 
-## Aboveground carbon
-
-AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-AGC stocks varied significantly between AFS (ANOVA p-value: `r subset(anova_table, variable=="AGC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)), but not between sites (ANOVA p-value = `r subset(anova_table, variable=="AGC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
-The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-
-## Soil organic carbon
-
-SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-No significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "afs")$pvalue`, Table \@ref(tab:anova)).
-However, the measured SOC stocks differed between sites (ANOVA p-value: `r subset(anova_table, variable=="SOC" & factor == "site")$pvalue`, Table \@ref(tab:anova)).
-The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$) than those in Osso-Luga (`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$; Figure \@ref(fig:barplot)).
-
-## Relationship between AGC and SOC
-
-The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure \@ref(fig:plot-lmereg)).
-This relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm).
-The conditional R^2^ (i.e., the proportion of variance explained by the linear mixed models) were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths, respectively, but the marginal R^2^ (i.e., when considering only the fixed effect) were only `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`%, respectively (Figure \@ref(fig:plot-lmereg)).
+## Effect of elevation and fine soil on AGC and SOC stocks
+
+## Relationship between AGC and SOC stocks
+
+The effect of AGC on SOC was close to zero (Figure \@ref(fig:plot-reg)).
 
@@
 
-```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAs with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
+<!-- ```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAs with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
 The p-value gives the probability of the means of the categories of each independent variable being the same.
-Bold p-values are lower than 0.05, and the mean difference between groups is thus considered significantly different than zero."}
-anova_table$factor = factor(anova_table$factor)
-levels(anova_table$factor) = c("AFS", "site")
-flextable(anova_table, col_keys = c("variable", "factor", "pvalue"), 
-          cwidth = c(1, 1, 0.8)) |> 
-  bold(i = anova_table$pvalue < 0.05, j = 2:3) |> 
-  set_header_labels(values = c("Dependent variable", "Independent variable", "p-value")) |> 
-  merge_at(i = 1:2, j = 1) |> merge_at(i = 3:4, j = 1) |> 
-  fix_border_issues() |> hline(i=2)
-```
+Bold p-values are lower than 0.05, and the mean difference between groups is thus considered significantly different than zero."} -->
+<!-- anova_table$factor = factor(anova_table$factor) -->
+<!-- levels(anova_table$factor) = c("AFS", "site") -->
+<!-- flextable(anova_table, col_keys = c("variable", "factor", "pvalue"),  -->
+<!--           cwidth = c(1, 1, 0.8)) |>  -->
+<!--   bold(i = anova_table$pvalue < 0.05, j = 2:3) |>  -->
+<!--   set_header_labels(values = c("Dependent variable", "Independent variable", "p-value")) |>  -->
+<!--   merge_at(i = 1:2, j = 1) |> merge_at(i = 3:4, j = 1) |>  -->
+<!--   fix_border_issues() |> hline(i=2) -->
+<!-- ``` -->
 
@@
 ```{r fig.id = "effect-agc-clay", fig.width = 7, fig.height = 2}
-labels_effect = function(x) ifelse(x=="agc", "AGC", "clay + fine silt")
+labels_level = function(x)
+  ifelse(x=="agc", "AGC", ifelse(x == "elev", "Elevation", "Clay + fine silt"))
 pars_summary |> 
   subset(param == "theta")  |> 
-  ggplot(aes(y = ifelse(stock=="soc10", "0-10 cm", "0-30 cm"))) + 
+  ggplot(aes(y = ifelse(stock=="soc10", "SOC : 0-10 cm", 
+                        ifelse(stock == "soc", "SOC : 0-30 cm", "AGC")))) + 
   geom_vline(xintercept = 0, lty = 2) +
@@
   geom_point(aes(x = mean), size = 2) +
-  facet_wrap(~effect, labeller = labeller(effect = labels_effect)) + 
+  facet_wrap(~level, labeller = labeller(level = labels_level)) + 
   labs(x="Effect on SOC", y = NULL) +
@@
 
-```{r fig.id = "plot-lmereg", fig.height = 3.5, fig.width = 7, fig.cap = "Scatter plot of aboveground carbon (AGC; x-axis) and soil organic carbon (SOC; y-axis) stocks.
+```{r fig.id = "plot-reg", fig.height = 3.5, fig.width = 7, fig.cap = "Scatter plot of aboveground carbon (AGC; x-axis) and soil organic carbon (SOC; y-axis) stocks.
 SOC stocks were estimated at the topsoil layer (0-10 cm; left panel) and at the total measured depth (0-30 cm; right panel).
@@
 dfpred <- expand.grid(agc = seq(0, 170, length.out = 20), 
-  iter = unique(data_par$iter), 
-  iter_stan = unique(data_par$iter_stan)
+                      iter = unique(data_par$iter)
 ) |> 
   merge(
-    data_par[grepl("soc", stock) & (param =="mu" | effect == "agc")] |> 
-      dcast(iter + iter_stan + stock ~ param)
+    data_par[grepl("soc", stock) & (param =="mu" | level == "agc")] |> 
+      dcast(iter + stock ~ param)
   ) |> 
@@
 ), .(agc, stock)]
-dfpred[, depth := ifelse(grepl(10, stock), "Soil depth: 0-10 cm", "Soil depth: 0-30 cm")]
-
-# stack data for figure
-dataC_30 = data_reg[, c("plot", "site", "afs",  
-                        "AGC_mean", "AGC_lwr", "AGC_upr",
-                        "SOC_mean", "SOC_lwr", "SOC_upr")]
-dataC_30$depth = "Soil depth: 0-30 cm"
-dataC_10 = data_reg[, c("plot", "site", "afs",   
-                        "AGC_mean", "AGC_lwr", "AGC_upr",
-                        "SOC10_mean", "SOC10_lwr", "SOC10_upr")]
-dataC_10$depth = "Soil depth: 0-10 cm"
-colnames(dataC_10) = colnames(dataC_30)
-dataC_fig = rbind(dataC_10, dataC_30)
+dfpred[, depth := ifelse(grepl(10, stock), 
+                         "Soil depth: 0-10 cm", "Soil depth: 0-30 cm")]
+
+# # stack data for figure
+dataC_30 <- dataC_plot[variable %in% c("AGC", "SOC")] |> 
+  dcast(plot + site + afs ~ variable, value.var = c("mean", "lwr", "upr"))
+dataC_30$depth <- "Soil depth: 0-30 cm"
+dataC_10 <- dataC_plot[variable %in% c("AGC", "SOC10")] |> 
+  dcast(plot + site + afs ~ variable, value.var = c("mean", "lwr", "upr"))
+colnames(dataC_10) <- gsub("SOC10", "SOC", colnames(dataC_10))
+dataC_10$depth <- "Soil depth: 0-10 cm"
+dataC_fig <- rbind(dataC_10, dataC_30)
 
@@
               col = NA, alpha = 0.1) +
-  geom_point(aes(x = AGC_mean, y = SOC_mean, shape = site), size = 2) +
-  # geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
-  #           x = -Inf, y = Inf, hjust = -0.1, vjust = 1.2, parse = TRUE, color = 1) +
-  # geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
-  #           x = -Inf, y = Inf, hjust = -0.1, vjust = 2.2, parse = TRUE, color = 1) +
-  geom_linerange(aes(x = AGC_mean, y = SOC_mean, ymin = SOC_lwr, ymax = SOC_upr)) +
-  geom_errorbarh(aes(y = SOC_mean, xmin = AGC_lwr, xmax = AGC_upr)) +
+  geom_point(aes(x = mean_AGC, y = mean_SOC, shape = site), size = 2) +
+  geom_linerange(aes(x = mean_AGC, y = mean_SOC, ymin = lwr_SOC, ymax = upr_SOC)) +
+  geom_errorbarh(aes(y = mean_SOC, xmin = lwr_AGC, xmax = upr_AGC)) +
   geom_line(data = dfpred, aes(x = agc, y = mean), lty = 2, col = 1) +
@@
   scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
-  # scale_x_log10() +
-  # scale_y_log10() +
   facet_wrap(~depth) +
@@
 \## Figures
+
+```{r dag, fig.cap="Direct acyclic plot of the hypothesised effect of fine soil (clay + fine silt content), climate (through elevation) and land use (here AFS) on AGC and SOC stocks."}
+# save_png <- function(plot, path){
+#   DiagrammeRsvg::export_svg(plot) %>%
+#     charToRaw() %>%
+#     rsvg::rsvg() %>%
+#     png::writePNG(path)
+# }
+# 
+# DiagrammeR::grViz("
+#       digraph {
+#       graph [layout = dot, rankdir = LR]
+# 
+#       node [shape = ellipse]
+#       AGC;
+#       SOC;
+#       node [shape = rectangle]
+#       elevation;
+#       climate;
+#       clay;
+#       AFS;
+# 
+#       # Connect nodes with edges and labels
+#       AFS -> AGC;
+#       climate -> AGC;
+#       AFS -> SOC;
+#       clay -> SOC;
+#       elevation -> climate;
+#       climate -> SOC;
+#       edge[style = 'dashed']
+#       AGC -> SOC;
+#        }
+# 
+#       ") %>%
+#   save_png("figures/dag.png")
+include_graphics("figures/dag.png") 
+```
 

### 2025-01-02 manuscript.Rmd
update results and figure to include new analysis

@@
 However, the effect of AFS on soil organic carbon (SOC) is more variable: although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e., planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFS, while AFS such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
-Other factors such as climate and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018; @Ma2020; @Martin2020].
+Other factors such as climate, soil characteristics and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018; @Ma2020; @Martin2020].
 Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS was only evidenced to a depth of 10 cm [@Ma2020].
@@
 A precise quantification of these carbon stocks will improve understanding of the ecological importance of AFS at both local and global scales, and help integrate them into national and international climate change mitigation initiatives.
-
-[paragraph on possible drivers]
-C stocks (in biomass and soil) of AFS could be driven by a number of factors: 
-- agricultural practices (ref)
-- climate (ref)
-- soil properties (especially for SOC) > in particular, fine content
 
@@
 ```{r get-env-data}
-# temperature and precipitation from CHELSA
+# mean annual temperature from CHELSA
 if (!file.exists("data/CHELSA_bio1_1981-2010_V.2.1.tif")) {
@@
 
+# mean annual precipitation from CHELSA
 if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
@@
   writeRaster(rst, filename = destfile, overwrite = TRUE) 
+}
+
+# precipitation variability from CHELSA
+if (!file.exists("data/CHELSA_bio15_1981-2010_V.2.1.tif")) {
+  url <- "https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio15_1981-2010_V.2.1.tif"
+  destfile <- "data/CHELSA_bio15_1981-2010_V.2.1.tif"
+  download.file(url, destfile, method = "curl")
+
+  ## crop to reduce memory usage
+  box = c(124.5, 127.5, -9.2, -8.3) # Timor-Leste
+  rst <- terra::rast(destfile)
+  rst <- terra::crop(rst, terra::ext(box))
+  writeRaster(rst, filename = destfile, overwrite = TRUE) 
+}
+
+# download cwd from Chave et al 2014
+if (!file.exists("data/CWD.tif")) {
+  url <- "http://chave.ups-tlse.fr/pantropical_allometry/CWD.tif.zip"
+  destfile <- "data/CWD.tif.zip"
+  download.file(url, destfile, method = "curl")
+  unzip(destfile, exdir = "data", overwrite = TRUE)
 }
@@
 } 
-
-# download cwd from Chave et al 2014
-if (!file.exists("data/CWD.tif")) {
-  url <- "http://chave.ups-tlse.fr/pantropical_allometry/CWD.tif.zip"
-  destfile <- "data/CWD.tif.zip"
-  download.file(url, destfile, method = "curl")
-  unzip(destfile, exdir = "data", overwrite = TRUE)
-}
 ```
@@
                         cbind(long, lat))[[1]], 
+  temp = terra::extract(terra::rast("data/CHELSA_bio1_1981-2010_V.2.1.tif"), 
+                        cbind(long, lat))[[1]], 
   prec = terra::extract(terra::rast("data/CHELSA_bio12_1981-2010_V.2.1.tif"), 
                         cbind(long, lat))[[1]], 
-  temp = terra::extract(terra::rast("data/CHELSA_bio1_1981-2010_V.2.1.tif"), 
+  varp = terra::extract(terra::rast("data/CHELSA_bio15_1981-2010_V.2.1.tif"), 
+                        cbind(long, lat))[[1]], 
+  cwd = terra::extract(terra::rast("data/CWD.tif"), 
                         cbind(long, lat))[[1]]
 )]
-
-cor(data_plot[, c("elev", "prec", "temp")])
 
@@
 The five AFS identified were as follows: Crop systems including a wooded Fallow phase (CF), Silvopastures (SP), Young Agroforests (YA), Home Gardens (HG), and Forest Gardens (FG).
-The AFS are presented in more detail in the supplemental information section (Table S\@ref(tab:afs-typology)) and in @Cogne2024.
+The AFS are presented in more detail in the Supplemental Information (Table S\@ref(tab:afs-typology)) and in @Cogne2024.
 
@@
 
-## Field inventories
+## Field inventories {#field-inventories}
 
@@
 Taxonomic identification was carried out using Tetum (a national official language) vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory Government's Native Plants Herbarium [@NT2013].
-The correspondence between Tetum names and scientific names is provided in the supplemental material (Table S\@ref(tab:tetum-names)).
+The correspondence between Tetum names and scientific names is provided in the Supplemental Information (Table S\@ref(tab:tetum-names)).
 
 In each of the 30 sample plots, a composite of three samples was taken from three of the five 10 x 10 m subplots (including the plot center and two opposite corners; Figure S\@ref(fig:design)), covering the intra-plot variability and at a minimum distance of 1 m from any tree.
-Soils were sampled with a soil auger ( Model Bor Tanah - Dormer Standard Soil Auger, Adiguna Karya Persada, Bogor, Indonesia) at 0-10 cm and 20-30 cm depth,.
+Soils were sampled with a soil auger (Model Bor Tanah - Dormer Standard Soil Auger, Adiguna Karya Persada, Bogor, Indonesia) at 0-10 cm and 20-30 cm depth.
 As the length of the head auger (15 cm) was longer than our target soil depths (10 cm), we could not accurately sample the 10-20 cm soil layer.
@@
 Soil organic carbon content was calculated as the difference of the total carbon and the carbonate contents.
+Soil texture (proportion of clay, silt and sand), pH, xxx and CEC were measured as xxx.
 
@@
 # use CWD values from Chave et al., 2014
-data_plot[, cwd := terra::extract(terra::rast("data/CWD.tif"), cbind(long, lat))]
-data_sim_agc <- merge(data_sim_agc, data_plot[, c("N_PLOT", "cwd")], by = "N_PLOT")
-
-# add uncertainty on R:S ratio? 
+
 data_sim_agc[, RS := exp(- 1.2312 - 0.0215 * DIAM_130_CM + 0.0002 * DIAM_130_CM^2 - 0.0007 * cwd)]
@@
 We considered that $mfine_i$ at depths of 2-7 cm, 12-17 cm and 22-27 cm was associated with $Corg_i$ at depths of 0-10, 10-20 cm and 20-30 cm, respectively.
-At depth $i$ = 10-20 cm, where the carbon content was not measured (see “Soil measurements” section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a standard negative exponential model of SOC content decrease with soil depth [@Mishra2009;@Murphy2019]:
+At depth $i$ = 10-20 cm, where the carbon content was not measured (see “Field inventories” section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a standard negative exponential model of SOC content decrease with soil depth [@Mishra2009;@Murphy2019]:
 
@@
 
-where $Corg_{a,b}$ is the average SOC content between depths a and b (see rationale for equation \@ref(eq:intersoc) in Supplementary Information xx).
+where $Corg_{a,b}$ is the average SOC content between depths a and b (see rationale for equation \@ref(eq:intersoc) in Supplemental Information).
 
@@
   dcast(N_PLOT ~ variable) 
-data_soil_summ[, fine_soil := CLAY + `FINE SILT`] 
 data_sim <- merge(data_sim, data_soil_summ)  
-# add elevation
-data_sim <- merge(data_sim, data_plot[, c("N_PLOT", "elev")])
+# add mean annual temperature
+data_sim <- merge(data_sim, data_plot[, c("N_PLOT", "temp")])
 ```
@@
 
-### Model calibration
-
-We modelled both AGC and SOC stocks as:
-
-$$Y_{i,j,k} = \mu +  \theta \cdot X_i + \alpha_{j} + \beta_{k} + \epsilon_{i}$$
-with $Y_{i,j,k}$ the stocks (either SOC or AGC) of plot $i$ in site $k$ and AFS type $j$
+### Carbon stocks model
+
+To test the relative effect of AFS type, climate and soil on carbon stocks, we modelled both AGC and SOC stocks as:
+
+$$Y_{i,j,k} = \mu +  \sum_{l} \left( \theta_l \cdot X_{l,i} \right) + \alpha_{j} + \beta_{k} + \epsilon_{i}$$
+with $Y_{i,j,k}$ the stocks (either AGC or SOC) of plot $i$ in site $k$ and AFS type $j$.
 
@@
 
-$\theta$ is the effect of the covariates on $Y$, and $X_i$ are the values of the covariates in plot $i$; they include elevation for AGC stocks [from NASA SRTM; @SRTM2024], or elevation [@SRTM2024], clay + fine soil content (from soil analysis; xxx methods) and AGC stocks for SOC stocks.
-All covariates were centred and scaled.
-Elevation was chosen as it was highly correlated (>95%) with both mean annual temperature and precipitation [@Karger2017]: it was thus a good indicator of a climate gradient in our dataset (see Supplementary information section xx).
-xxx justify these variables in the introduction
+$\theta_l$ is the effect of covariate $l$ on variable $Y$, and $X_{l,i}$ is the value of covariate $l$ in plot $i$.
+We included the following covariates: elevation [from NASA SRTM; @SRTM2024] to represent the climate gradient in our study area, and clay fraction (as measured in the Cirad laboratories; see “Field inventories” section) to represent the soil gradient in our study area.
+The rationale for the choice of these variables is given in the Supplemental Information.
+In addition, we included AGC stock as a covariate in the SOC stock model, as we wanted to quantify its effect on SOC stocks (hypotheses H2a and H2b).
+Covariates were all centred and scaled.
 
@@
 $\sigma$, $\sigma_{\alpha}$, and $\sigma_{\beta}$ are the standard deviation of the random error, the AFS effect and the site effect respectively.
+
+### Model calibration
 
@@
 
-```{r}
+```{r stan-model}
 stan_model <- "
@@
   int<lower=1> S;
-  int<lower=1> K;
+  int<lower=1> KA;
+  int<lower=1> KS;
   int<lower=1,upper=AF> afs[N];
   int<lower=1,upper=S> site[N];
-  vector[N] elev;
-  matrix[N,K] covar;
+  matrix[N,KA] covarA;
+  matrix[N,KS] covarS;
   real<lower=0> agc[N];
@@
   real delta_site_a[S];
-  real theta_a;
+  vector[KA] theta_a;
   real<lower=0> sigma_a;
@@
   real delta_site_s[S];
-  vector[K] theta_s;
+  vector[KS] theta_s;
   real<lower=0> sigma_s;
@@
   real delta_site_s10[S];
-  vector[K] theta_s10;
+  vector[KS] theta_s10;
   real<lower=0> sigma_s10;
@@
   for (i in 1:N) {
-    pred_a[i] = mu_a + elev[i] * theta_a + delta_afs_a[afs[i]] + delta_site_a[site[i]];
-    pred_s[i] = mu_s + dot_product(row(covar, i), theta_s) + delta_afs_s[afs[i]] + delta_site_s[site[i]];
-    pred_s10[i] = mu_s10 + dot_product(row(covar, i), theta_s10) + delta_afs_s10[afs[i]] + delta_site_s10[site[i]];
+    pred_a[i] = mu_a + dot_product(row(covarA, i), theta_a) + delta_afs_a[afs[i]] + delta_site_a[site[i]];
+    pred_s[i] = mu_s + dot_product(row(covarS, i), theta_s) + delta_afs_s[afs[i]] + delta_site_s[site[i]];
+    pred_s10[i] = mu_s10 + dot_product(row(covarS, i), theta_s10) + delta_afs_s10[afs[i]] + delta_site_s10[site[i]];
   }
@@
   soc ~ normal(pred_s, sigma_s);
-  soc10 ~ normal(pred_s10, sigma_s);
+  soc10 ~ normal(pred_s10, sigma_s10);
 
@@
   //  priors
-  theta_s ~ normal(0,1);
-  theta_s10 ~ normal(0,1);
+  theta_a ~ normal(0,1);
   sigma_a ~ normal(0,1);
-  sigma_s ~ normal(0,1);
-  sigma_s10 ~ normal(0,1);
   sigma_site_a ~ normal(0,1);
   sigma_afs_a ~ normal(0,1);
+  theta_s ~ normal(0,1);
+  sigma_s ~ normal(0,1);
   sigma_site_s ~ normal(0,1);
   sigma_afs_s ~ normal(0,1);
+  theta_s10 ~ normal(0,1);
+  sigma_s10 ~ normal(0,1);
   sigma_site_s10 ~ normal(0,1);
@@
 
-```{r stan}
+```{r stan-calibration}
 if (!dir.exists("outputs")) dir.create("outputs")
@@
         subset(iter==i) |> 
-        dcast(N_PLOT + site + ID_AF + fine_soil + elev ~ variable) |> 
+        dcast(N_PLOT + site + ID_AF + CLAY + temp ~ variable) |> 
         with(
@@
             S = length(unique(site)),
-            K = 3,
+            KA = 2,
+            KS = 3,
             afs = as.numeric(as.factor(ID_AF)), 
             site = as.numeric(as.factor(site)),
-            elev = (elev-mean(elev))/sd(elev),
-            covar = scale(cbind(elev, fine_soil, AGC)),
+            covarA = scale(cbind(temp, CLAY)),
+            covarS = scale(cbind(temp, CLAY, AGC)),
             agc = AGC,
@@
 data_par[param == "theta",
-         level := c("elev", "fine", "agc")[level_num]]
-data_par[param == "theta" & stock == "agc", level := "elev"]
+         level := c("mat", "clay", "agc")[level_num]]
 
@@
 
+## Distribution of carbon stocks by AFS and site
+
 AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
@@
 
-## Effect of elevation and fine soil on AGC and SOC stocks
+## Effect of mean annual temperature and clay fraction on AGC and SOC stocks
+
+Zero was included in the 80% credibility intervals for the effect of mean annual temperature and clay fraction on AGC, meaning that no strong effect was detected in our data (Figure \@ref(fig:effect-agc-clay)).
+Similarly, the clay fraction did not show a strong effect on SOC stocks, as zero was included in its 80% credible intervals (Figure \@ref(fig:effect-agc-clay)).
+
+Mean annual temperature, on the contrary, had a stronger negative effect on SOC stocks at both 0-10 cm and 0-30 cm depth (Figure \@ref(fig:effect-agc-clay)).
+Values of $\theta_{MAT}$ were negative in `r sum(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value<0)/10`% of cases for SOC at 0-30 cm depth, and in `r sum(subset(data_par, stock=="soc10" & param == "theta" & level == "mat")$value<0)/10`% of cases for SOC at 0-10 cm depth.
+They averaged `r round(mean(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value), 2)` at 0-30 cm depth, meaning that SOC stocks were estimated to decrease by `r abs(round(mean(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value), 2))` Mg C ha$^{-1}$ for an increase in mean annual temperature of one standard deviation, i.e.
+`r round(sd(data_plot$temp), 2)` °C.
 
@@
 
-The effect of AGC on SOC was close to zero (Figure \@ref(fig:plot-reg)).
+The relationship between AGC and SOC stocks was weak in our dataset (Figure  \@ref(fig:plot-reg)), as the 80% credibility interval of the associated parameter ($\theta_{AGC}$) included zero (Figure \@ref(fig:effect-agc-clay)).
 
@@
 
-```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3}
+```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3, fig.cap = "Effect of AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden) and site on carbon stocks (in MgC/ha).
+These values were obtained as the sum of the mean predicted stocks ($\\mu$) and the AFS or site random effects.
+The black dot represents the mean of the 1000 iterations; the red segment represents the 80% confidence interval and the black segment represents the 95% confidence interval.
+AGC: aboveground (biomass) carbon; SOC: soil organic carbon."}
 labels_stock = function(x) paste(toupper(x), "(Mg/ha)")
@@
 
-```{r fig.id = "effect-agc-clay", fig.width = 7, fig.height = 2}
-labels_level = function(x)
-  ifelse(x=="agc", "AGC", ifelse(x == "elev", "Elevation", "Clay + fine silt"))
+```{r fig.id = "effect-agc-clay", fig.width = 7, fig.height = 2, fig.cap = "Effect of covariates (MAT: mean annual temperature, Clay: clay fraction, AGC: aboveground carbon) on carbon stocks (SOC: soil organic carbon, at 0-10 cm or 0-30 cm depth).
+Covariates are all centred and scaled; positive values indicate that C stocks tend to be higher for higher values of the covariate.
+Black dots represent the mean of the 1000 iterations; red segments represent 80% confidence intervals and black segments represent 95% confidence intervals."}
+labels_stock = function(x)
+  ifelse(x=="agc", "AGC", ifelse(grepl("10", x), "SOC (0-10 cm)", "SOC (0-30 cm)"))
+
 pars_summary |> 
   subset(param == "theta")  |> 
-  ggplot(aes(y = ifelse(stock=="soc10", "SOC : 0-10 cm", 
-                        ifelse(stock == "soc", "SOC : 0-30 cm", "AGC")))) + 
+  ggplot(aes(y = level)) + 
   geom_vline(xintercept = 0, lty = 2) +
@@
   geom_point(aes(x = mean), size = 2) +
-  facet_wrap(~level, labeller = labeller(level = labels_level)) + 
-  labs(x="Effect on SOC", y = NULL) +
+  facet_wrap(~stock, labeller = labeller(stock = labels_stock)) + 
+  labs(x="Effect of covariates on C stocks", y = NULL) +
+  scale_y_discrete(labels= c("AGC", "Clay", "MAT")) +
   theme_bw()
@@
 SOC stocks were estimated at the topsoil layer (0-10 cm; left panel) and at the total measured depth (0-30 cm; right panel).
-Each point represents the median C stock value in a sample plot, and the vertical and horizontal error bars are the 95% confidence interval.
-The colors correspond to the different agroforestry systems (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden); the different shapes represent the two sites.
-The dotted lines represent the (non-significant) linear mixed models between log-transformed AGC and SOC stocks.
-For each linear mixed model, we report the marginal R^2^ (proportion of variance explained by fixed effects only) and the conditional R^2^ (proportion of variance explained by both fixed and random effects).
-Both x- and y-axes are log-transformed.
-One plot (AFS: CF, site: Gariuai) with no trees over 30 cm in circumference at breast height (minimum size for measurement) had AGC stocks estimated at 0 Mg C ha^-1^ and was therefore removed from the linear model and the figure."}
+Each point represents the mean SOC stocks in a sample plot, and the vertical and horizontal error bars are the 95% confidence intervals.
+The colours correspond to the different agroforestry systems (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden); the different shapes represent the two sites.
+The dotted lines represent the mean predicted SOC stocks for a given AGC stock value, and the grey areas represent the 95% confidence interval of this prediction."}
 
@@
 
-# Supplementary information
-
-## Supplementary methods
+# Supplemental Information
+
+## Supplemental methods
 
@@
 $$ \Rightarrow C_{10,20} = C_{0,10} \cdot \left( \frac{C_{20,30}}{C_{0,10}}\right)^{\frac{1}{2}}$$
+### Choice of climate and soil covariates
+
+As described in the "Carbon stocks model" section of the Materials and Methods, we hypothesized that C stocks (both AGC and SOC) are affected by climate and soil characteristics, together with land use (in this case, AFS type) and other spatial differences (in our case, a site effect).
+
+```{r pca-clim-soil}
+pca_clim <- data_plot[, c("temp", "prec", "varp")] |> 
+  scale() |> princomp()
+
+data_soil_melt <- 
+  data_soil |> 
+  melt(id.vars = c("N_PLOT", "SAMPLE_DEPTH"), 
+       measure.vars = c(
+         "pH\n", "CEC\n",  "CLAY", "FINE SILT", 
+         "COARSE SILT", "FINE SAND", "COARSE SAND"))
+
+# make values numeric; when detectability limit is reached, used limit value
+data_soil_melt[grepl("LD", value), value := gsub(",", ".", tstrsplit(value, "LD=")[[2]])]
+
+data_soil_melt <- 
+  data_soil_melt[, .(value = mean(as.numeric(value), na.rm = TRUE)), .(N_PLOT, variable)] |> 
+  dcast(N_PLOT ~ variable) 
+data_soil_melt[, silt := `COARSE SILT` +`FINE SILT`]
+data_soil_melt[, sand := `FINE SAND` + `COARSE SAND`]
+
+pca_soil <- data_soil_melt[, c("pH\n", "CEC\n",  "CLAY", "silt", "sand")] |> 
+  scale() |> princomp()
+```
+
+We selected the following set of variables to represent the effect of climate: mean annual temperature (BIO1), mean annual precipitation (BIO12), annual precipitation variability (BIO15) over the period 1981-2010.
+These bioclimatic variables were extracted from CHELSA 2.1 [@Krager2017].
+These variables were all highly correlated in our dataset (Figure S\@ref(fig:choice-clim-vars)a).
+We performed a principal component analysis after scaling and centering these three variables: the first axis explained `r round(pca_clim$sdev[1]^2/sum(pca_clim$sdev^2)*100)`% of the variance in our dataset (Figure S\@ref(fig:choice-clim-vars)b).
+We therefore decided to keep only one variable to represent this climatic gradient: mean annual temperature, which was most correlated with the first axis of the PCA (Figure S\@ref(fig:choice-clim-vars)b), and has been shown to influence C stocks in other contexts [@Wiesmeier2019;@Muller-Landau2021].
+
+We selected the following set of variables to represent the effect of soil properties on C stocks: clay fraction (<2 $\mu$m), silt fraction (2-50 $\mu$m), sand fraction (50-2000 $\mu$m), pH and cation exchange capacity xx.
+These edaphic variables were measured on our soil samples analysed in the Cirad laboratories (see section "Field inventories").
+After scaling and centering these variables, we performed a principal component analysis: the first axis explained `r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance in our data set (Figure S\@ref(fig:choice-soil-vars)b).
+We therefore decided to retain only one variable to represent this edaphic gradient: clay fraction, which was one of the most highly correlated with the first axis of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been shown to influence C stocks in other contexts [@Wiesmeier2019].
 
@@
 \## Figures
-
-```{r dag, fig.cap="Direct acyclic plot of the hypothesised effect of fine soil (clay + fine silt content), climate (through elevation) and land use (here AFS) on AGC and SOC stocks."}
-# save_png <- function(plot, path){
-#   DiagrammeRsvg::export_svg(plot) %>%
-#     charToRaw() %>%
-#     rsvg::rsvg() %>%
-#     png::writePNG(path)
-# }
-# 
-# DiagrammeR::grViz("
-#       digraph {
-#       graph [layout = dot, rankdir = LR]
-# 
-#       node [shape = ellipse]
-#       AGC;
-#       SOC;
-#       node [shape = rectangle]
-#       elevation;
-#       climate;
-#       clay;
-#       AFS;
-# 
-#       # Connect nodes with edges and labels
-#       AFS -> AGC;
-#       climate -> AGC;
-#       AFS -> SOC;
-#       clay -> SOC;
-#       elevation -> climate;
-#       climate -> SOC;
-#       edge[style = 'dashed']
-#       AGC -> SOC;
-#        }
-# 
-#       ") %>%
-#   save_png("figures/dag.png")
-include_graphics("figures/dag.png") 
-```
-
-```{r fig.id="design", fig.cap = "Plot sampling design.
-(a) Typical plot < 1 ha, in all AFS except SPs.
-(b) Typical 1-ha plot in SPs.
-Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 5, fig.height = 3}
-include_graphics("figures/plot-design.jpg") 
-```
-
-</br>
 
@@
 
-</br>
-
-```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in\ Mg\ C\ ha^-1^).
+```{r fig.id="design", fig.cap = "Plot sampling design.
+(a) Typical plot < 1 ha, in all AFS except SPs.
+(b) Typical 1-ha plot in SPs.
+Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 5, fig.height = 3}
+include_graphics("figures/plot-design.jpg") 
+```
+
+<!-- </br> -->
+
+<!-- ```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in\ Mg\ C\ ha^-1^).
 Points are colored by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.
 We performed a linear mixed effects regression with log-transformed SOC stocks as the dependent variable, log-transformed elevation as the independent variable, and site as random effect.
-Results are presented as regression lines and the p-value of the relationship between SOC stocks and elevation.", fig.cap.pre="Figure S"}
-dataC_plot <- read.csv("dataC_plot.csv") |> setDT()
-dataCelev <- merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")], 
-                   data_plot[, c("N_PLOT", "long", "lat")], by.x = "plot", by.y = "N_PLOT")
-
-# extract elevation
-dataCelev[, elev := terra::extract(
-  terra::rast("data/srtm_62_14.tif"), 
-  cbind(long, lat)
-)]
-
-regElev = lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |> 
-  summary()
-pvalueE = signif(regElev$coefficients[2, 5], 2)
-slp = regElev$coefficients[2, 1]
-
-dfElev = dataCelev[, .(
-  int = mean(log(mean)) - slp * mean(log(elev)), 
-  elev = min(elev), elevMax = max(elev),
-  slp = slp), .(site)] 
-dfElev[, mean := exp(slp*log(elev) + int)]
-dfElev[, predMax := exp(slp*log(elevMax) + int)]
-
-ggplot(dataCelev, aes(x = elev, y = mean, col = site)) +
-  geom_pointrange(aes(ymin=lwr, ymax = upr)) + 
-  geom_segment(data = dfElev, aes(xend = elevMax, yend = predMax)) +
-  theme_classic() +
-  annotate(geom = "text", label = paste("p-value =", pvalueE), x = 500, y = 32) +
-  labs(x = "Elevation (m)", y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), col = "Site") +
-  scale_x_log10() + scale_y_log10()
-```
+Results are presented as regression lines and the p-value of the relationship between SOC stocks and elevation.", fig.cap.pre="Figure S"} -->
+<!-- dataC_plot <- read.csv("dataC_plot.csv") |> setDT() -->
+<!-- dataCelev <- merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")],  -->
+<!--                    data_plot[, c("N_PLOT", "long", "lat")], by.x = "plot", by.y = "N_PLOT") -->
+
+<!-- # extract elevation -->
+<!-- dataCelev[, elev := terra::extract( -->
+<!--   terra::rast("data/srtm_62_14.tif"),  -->
+<!--   cbind(long, lat) -->
+<!-- )] -->
+
+<!-- regElev = lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |>  -->
+<!--   summary() -->
+<!-- pvalueE = signif(regElev$coefficients[2, 5], 2) -->
+<!-- slp = regElev$coefficients[2, 1] -->
+
+<!-- dfElev = dataCelev[, .( -->
+<!--   int = mean(log(mean)) - slp * mean(log(elev)),  -->
+<!--   elev = min(elev), elevMax = max(elev), -->
+<!--   slp = slp), .(site)]  -->
+<!-- dfElev[, mean := exp(slp*log(elev) + int)] -->
+<!-- dfElev[, predMax := exp(slp*log(elevMax) + int)] -->
+
+<!-- ggplot(dataCelev, aes(x = elev, y = mean, col = site)) + -->
+<!--   geom_pointrange(aes(ymin=lwr, ymax = upr)) +  -->
+<!--   geom_segment(data = dfElev, aes(xend = elevMax, yend = predMax)) + -->
+<!--   theme_classic() + -->
+<!--   annotate(geom = "text", label = paste("p-value =", pvalueE), x = 500, y = 32) + -->
+<!--   labs(x = "Elevation (m)", y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), col = "Site") + -->
+<!--   scale_x_log10() + scale_y_log10() -->
+<!-- ``` -->
+
+<!-- </br> -->
+
+<!-- ```{r fig.id = "clay-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap="Distribution of SOC content (y-axis; in %) values as a function of fine silt and clay content (x-axis; in %), soil depth (point color; in cm), and site (point shape).
+We performed a linear mixed effects regression with log-transformed SOC content as the dependent variable, log-transformed fine soil texture as the independent variable, and soil depth and site as random effects.
+Results are presented as regression lines and the p-value of the relationship between SOC content and fine soil texture.", fig.cap.pre="Figure S"} -->
+<!-- data_soil[, site := substr(ID_VILLAGE, 1, 1)] -->
+<!-- data_soil[site == "S", site := "Osso-Luga"] -->
+<!-- data_soil[site == "G", site := "Gariuai"] -->
+
+<!-- regText = lmer( -->
+<!--   log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) +  -->
+<!--     (1|site)+  -->
+<!--     (1|SAMPLE_DEPTH),  -->
+<!--   data = data_soil[SAMPLE_DEPTH != "10-20"]) |>  -->
+<!--   summary() -->
+<!-- pvalueT = signif(regText$coefficients[2, 5], 3) -->
+<!-- slp = regText$coefficients[2, 1] -->
+
+<!-- dfText = data_soil[SAMPLE_DEPTH != "10-20" & !is.na(CLAY), .( -->
+<!--   int = mean(log(Corg)) - slp * mean(log(as.numeric(CLAY) + as.numeric(`FINE SILT`))),  -->
+<!--   xmin = min(as.numeric(CLAY) + as.numeric(`FINE SILT`)),  -->
+<!--   xmax = max(as.numeric(CLAY) + as.numeric(`FINE SILT`)), -->
+<!--   slp = slp), .(site, SAMPLE_DEPTH)]  -->
+<!-- dfText[, predMin := exp(slp*log(xmin) + int)] -->
+<!-- dfText[, predMax := exp(slp*log(xmax) + int)] -->
+
+<!-- ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, shape = SAMPLE_DEPTH, col = site)) + -->
+<!--   geom_point() + -->
+<!--   geom_segment(data = dfText, aes(x = xmin, y = predMin, xend = xmax, yend = predMax, lty = SAMPLE_DEPTH)) + -->
+<!--   scale_x_log10() + scale_y_log10() + -->
+<!--   scale_linetype_manual(values=c("twodash", "dotted")) + -->
+<!--   labs(shape = "Soil depth (cm)", lty = "Soil depth (cm)",  -->
+<!--        x = "Clay and fine silt content (%)",  -->
+<!--        y = "SOC content (%)",  -->
+<!--        col = "Site") + -->
+<!--   annotate(geom = "text", label = paste("p-value =", pvalueT), x = 45, y = 1) + -->
+<!--   theme_classic() -->
+<!-- ``` -->
 
@@
 
-```{r fig.id = "clay-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap="Distribution of SOC content (y-axis; in %) values as a function of fine silt and clay content (x-axis; in %), soil depth (point color; in cm), and site (point shape).
-We performed a linear mixed effects regression with log-transformed SOC content as the dependent variable, log-transformed fine soil texture as the independent variable, and soil depth and site as random effects.
-Results are presented as regression lines and the p-value of the relationship between SOC content and fine soil texture.", fig.cap.pre="Figure S"}
-data_soil[, site := substr(ID_VILLAGE, 1, 1)]
-data_soil[site == "S", site := "Osso-Luga"]
-data_soil[site == "G", site := "Gariuai"]
-
-regText = lmer(
-  log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) + 
-    (1|site)+ 
-    (1|SAMPLE_DEPTH), 
-  data = data_soil[SAMPLE_DEPTH != "10-20"]) |> 
-  summary()
-pvalueT = signif(regText$coefficients[2, 5], 3)
-slp = regText$coefficients[2, 1]
-
-dfText = data_soil[SAMPLE_DEPTH != "10-20" & !is.na(CLAY), .(
-  int = mean(log(Corg)) - slp * mean(log(as.numeric(CLAY) + as.numeric(`FINE SILT`))), 
-  xmin = min(as.numeric(CLAY) + as.numeric(`FINE SILT`)), 
-  xmax = max(as.numeric(CLAY) + as.numeric(`FINE SILT`)),
-  slp = slp), .(site, SAMPLE_DEPTH)] 
-dfText[, predMin := exp(slp*log(xmin) + int)]
-dfText[, predMax := exp(slp*log(xmax) + int)]
-
-ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, shape = SAMPLE_DEPTH, col = site)) +
-  geom_point() +
-  geom_segment(data = dfText, aes(x = xmin, y = predMin, xend = xmax, yend = predMax, lty = SAMPLE_DEPTH)) +
-  scale_x_log10() + scale_y_log10() +
-  scale_linetype_manual(values=c("twodash", "dotted")) +
-  labs(shape = "Soil depth (cm)", lty = "Soil depth (cm)", 
-       x = "Clay and fine silt content (%)", 
-       y = "SOC content (%)", 
-       col = "Site") +
-  annotate(geom = "text", label = paste("p-value =", pvalueT), x = 45, y = 1) +
+```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesised effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks."}
+# save_png <- function(plot, path){
+#   DiagrammeRsvg::export_svg(plot) %>%
+#     charToRaw() %>%
+#     rsvg::rsvg() %>%
+#     png::writePNG(path)
+# }
+# 
+# DiagrammeR::grViz("
+#       digraph {
+#       graph [layout = dot, rankdir = LR]
+# 
+#       node [shape = ellipse]
+#       AGC;
+#       SOC;
+#       node [shape = rectangle]
+#       climate;
+#       soil;
+#       AFS;
+# 
+#       # Connect nodes with edges and labels
+#       AFS -> AGC;
+#       climate -> AGC;
+#       AFS -> SOC;
+#       soil -> AGC;
+#       soil -> SOC;
+#       climate -> SOC;
+#       edge[style = 'dashed']
+#       AGC -> SOC;
+#        }
+# 
+#       ") %>%
+#   save_png("figures/dag.png")
+include_graphics("figures/dag.png") 
+```
+
+</br>
+
+```{r fig.id="choice-clim-vars", fig.width = 10, fig.height=4, fig.cap=""}
+g1 <- data_plot[, c("temp", "prec", "varp")] |>
+  setnames(c("MAT", "MAP", "Pvar")) |>
+  cor() |> melt() |>
+  ggplot(aes(x = Var1, y = Var2, label = round(value, 2), fill = value)) +
+  geom_tile() +
+  geom_text() +
+  scale_fill_gradientn(colours = c("brown1", "#FFFFFF", "cornflowerblue")) +
+  theme_bw() +
+  labs(x = NULL, y = NULL, fill = "Pearson\nCorrelation\nCoefficient")
+
+g2 <- data_plot[, c("temp", "prec", "varp")] |>
+  setnames(c("MAT", "MAP", "Pvar")) |>
+  scale() |>
+  princomp() |>
+  autoplot(data = data_plot, colour = 'site', loadings.colour = 'black',
+         loadings.label = TRUE, loadings.label.col = 'black') +
   theme_classic()
-```
+ggpubr::ggarrange(g1, g2, labels = "auto")
+```
+
+```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4}
+g1 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n", "CEC\n")] |>
+  setnames(c("clay", "silt", "sand", "pH", "CEC")) |>
+  cor() |> melt() |>
+  ggplot(aes(x = Var1, y = Var2, label = round(value, 2), fill = value)) +
+  geom_tile() +
+  geom_text() +
+  scale_fill_gradientn(colours = c("brown1", "#FFFFFF", "cornflowerblue")) +
+  theme_bw() +
+  labs(x = NULL, y = NULL, fill = "Pearson\nCorrelation\nCoefficient")
+
+g2 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n", "CEC\n")] |>
+  setnames(c("clay", "silt", "sand", "pH", "CEC")) |>
+  scale() |>
+  princomp() |>
+  autoplot(data = data_plot, colour = 'site', loadings.colour = 'black',
+         loadings.label = TRUE, loadings.label.col = 'black') +
+  theme_classic()
+ggpubr::ggarrange(g1, g2, labels = "auto")
+```
+
+<!-- ```{r dag-code} -->
+<!-- # dag { -->
+<!-- # bb="-2.799,-2.922,2.216,3.074" -->
+<!-- # afs [exposure,pos="0.265,-1.155"] -->
+<!-- # agc [exposure,pos="-0.591,0.840"] -->
+<!-- # climate [adjusted,pos="-1.049,-1.155"] -->
+<!-- # soc [outcome,pos="0.137,1.233"] -->
+<!-- # soil [adjusted,pos="-0.426,-1.182"] -->
+<!-- # afs -> agc -->
+<!-- # afs -> soc -->
+<!-- # agc -> soc -->
+<!-- # climate -> agc -->
+<!-- # climate -> soc -->
+<!-- # soil -> agc -->
+<!-- # soil -> soc -->
+<!-- # } -->
+<!-- ``` -->

### 2025-01-03 manuscript.Rmd
add ggfortify package, figure captions and references

@@
   "ggspatial",    # version 1.1.9
+  "ggpubr",       # version 0.6.0
+  "ggfortify",    # version 0.4.17
   "maps",         # version 3.4.2
-  "ggpubr",       # version 0.6.0
   "rdryad",       # version 1.0.0
@@
 We selected the following set of variables to represent the effect of climate: mean annual temperature (BIO1), mean annual precipitation (BIO12), annual precipitation variability (BIO15) over the period 1981-2010.
-These bioclimatic variables were extracted from CHELSA 2.1 [@Krager2017].
+These bioclimatic variables were extracted from CHELSA V2.1 [@Krager2017].
 These variables were all highly correlated in our dataset (Figure S\@ref(fig:choice-clim-vars)a).
-We performed a principal component analysis after scaling and centering these three variables: the first axis explained `r round(pca_clim$sdev[1]^2/sum(pca_clim$sdev^2)*100)`% of the variance in our dataset (Figure S\@ref(fig:choice-clim-vars)b).
-We therefore decided to keep only one variable to represent this climatic gradient: mean annual temperature, which was most correlated with the first axis of the PCA (Figure S\@ref(fig:choice-clim-vars)b), and has been shown to influence C stocks in other contexts [@Wiesmeier2019;@Muller-Landau2021].
+We performed a principal component analysis after scaling and centering these three variables: the first component explained `r round(pca_clim$sdev[1]^2/sum(pca_clim$sdev^2)*100)`% of the variance in our dataset (Figure S\@ref(fig:choice-clim-vars)b).
+We therefore decided to keep only one variable to represent this climatic gradient: mean annual temperature, which was most correlated with the first component of the PCA (Figure S\@ref(fig:choice-clim-vars)b), and has been shown to influence C stocks in other contexts [@Wiesmeier2019;@Muller-Landau2021].
 
@@
 These edaphic variables were measured on our soil samples analysed in the Cirad laboratories (see section "Field inventories").
-After scaling and centering these variables, we performed a principal component analysis: the first axis explained `r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance in our data set (Figure S\@ref(fig:choice-soil-vars)b).
-We therefore decided to retain only one variable to represent this edaphic gradient: clay fraction, which was one of the most highly correlated with the first axis of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been shown to influence C stocks in other contexts [@Wiesmeier2019].
+After scaling and centering these variables, we performed a principal component analysis: the first component explained `r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance in our data set (Figure S\@ref(fig:choice-soil-vars)b).
+We therefore decided to retain only one variable to represent this edaphic gradient: clay fraction, which was one of the most highly correlated with the first component of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been shown to influence C stocks in other contexts [@Wiesmeier2019].
 
@@
 \## Figures
+
+```{r fig.id="design", fig.cap = "Plot sampling design.
+(a) Typical plot < 1 ha, in all AFS except SPs.
+(b) Typical 1-ha plot in SPs.
+Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 5, fig.height = 3}
+include_graphics("figures/plot-design.jpg") 
+```
+</br>
 
@@
 
-```{r fig.id="design", fig.cap = "Plot sampling design.
-(a) Typical plot < 1 ha, in all AFS except SPs.
-(b) Typical 1-ha plot in SPs.
-Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 5, fig.height = 3}
-include_graphics("figures/plot-design.jpg") 
-```
-
-<!-- </br> -->
-
 <!-- ```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in\ Mg\ C\ ha^-1^).
@@
 
-```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesised effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks."}
+```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesised effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
 # save_png <- function(plot, path){
@@
 
-```{r fig.id="choice-clim-vars", fig.width = 10, fig.height=4, fig.cap=""}
+```{r fig.id="choice-clim-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with three climate variables: mean annual temperature (MAT), mean annual precipitation (MAP) and precipitation variability (PVar).
+Values for the climate variables were extracted from CHELSA V2.1 (Karger et al., 2017) for each of the 30 study plots.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
 g1 <- data_plot[, c("temp", "prec", "varp")] |>
@@
   theme_classic()
+
 ggpubr::ggarrange(g1, g2, labels = "auto")
@@
 
-```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4}
+</br>
+
+```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with five soil variables: clay, silt and sand fraction, soil pH, and cation exchange capacity (CEC).
+Values for the soil variables were measured from in situ soil samples for each of the 30 study plots.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
 g1 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n", "CEC\n")] |>

### 2025-01-06 manuscript.Rmd
improve figures consistency and aesthetics

@@
 Soil organic carbon content was calculated as the difference of the total carbon and the carbonate contents.
-Soil texture (proportion of clay, silt and sand), pH, xxx and CEC were measured as xxx.
+Soil texture (proportion of clay, silt and sand), and pH were measured as xxx.
 
@@
   annotate(geom = "text", label = "a", x = -Inf, y = Inf, 
-           hjust = 0, vjust = 1, fontface = "bold", size = 6) + 
+           hjust = -3, vjust = 1, size = 5) + 
   geom_polygon(aes(x = long, y = lat, group = group), fill="lightgray", colour = "grey") +
@@
   annotate(geom = "text", label = "b", x = -Inf, y = Inf, 
-           hjust = -0.5, vjust = 1, fontface = "bold", size = 6) + 
+           hjust = -0.5, vjust = 1, size = 5) + 
   annotate(geom = "text", label = "c", x = Inf, y = -Inf, 
-           hjust = 12, vjust = -9.8, fontface = "bold", size = 6) + 
+           hjust = 13.7, vjust = -10.8, size = 5) + 
   guides(color = FALSE, fill=guide_colorbar(title.position = "top")) +
@@
 
-ggplot(dataC_afs_site[variable %in% c("AGC", "BGC", "SOC")], 
-       aes(x = afs, y = mean, fill = variable)) +
-  geom_bar(position="stack", stat="identity") + 
+# add facet letter
+data_letter <- data.frame( 
+  site = c("Gariuai", "Osso-Luga"), 
+  lab = letters[1:2]
+)
+
+dataC_afs_site |> 
+  subset(variable %in% c("AGC", "BGC", "SOC")) |> 
+ggplot(aes(x = afs, y = mean)) +
+  geom_bar(position="stack", stat="identity", aes(fill = variable)) + 
+  geom_text(data = data_letter, x = -Inf, y = Inf, hjust = -1, vjust = 1, 
+            aes(label = lab)) +
   facet_wrap(~site, nrow = 1, strip.position = "bottom") +
@@
   theme_classic() +
-  theme(strip.placement = "outside") +
   labs(x = "", y = expression("Carbon stocks (Mg C"*" "*ha^{-1}*")"), fill = "Carbon\npool") +
-  scale_y_continuous(expand = c(0, 0))
+  scale_y_continuous(expand = c(0, 0)) +
+  theme(strip.background = element_blank(), strip.placement = "outside")
 ```
@@
 
-```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3, fig.cap = "Effect of AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden) and site on carbon stocks (in MgC/ha).
+```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3, fig.cap = "Effect of (a, b) AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden) and (c, d) site (G: Gariuai, OL: Osso-Luga) on (a, c) aboveground carbon (AGC, in Mg C ha$^{-1}$) and (b, d) soil organic carbon (SOC, in Mg C ha$^{-1}$).
 These values were obtained as the sum of the mean predicted stocks ($\\mu$) and the AFS or site random effects.
-The black dot represents the mean of the 1000 iterations; the red segment represents the 80% confidence interval and the black segment represents the 95% confidence interval.
-AGC: aboveground (biomass) carbon; SOC: soil organic carbon."}
-labels_stock = function(x) paste(toupper(x), "(Mg/ha)")
-labels_effect = function(x) ifelse(x %in% c("afs", "agc"), toupper(x), x)
+The black dot represents the mean of the 1000 iterations; the red segment represents the 80% confidence interval and the black segment represents the 95% confidence interval."}
+labels_stock <- function(x) paste0(toupper(x), "~(Mg~C~ha^-1)")
+labels_effect <- function(x) ifelse(x=="afs", toupper(x), x)
+
+# add facet letter
+data_letter <- data.frame(
+  stock = rep(c("agc", "soc"), 2), 
+  effect = rep(c("afs", "site"), each = 2), 
+  lab = letters[1:4]
+)
+
+# keep only sites' initials to simplify figure
+initials <- function(x) gsub("(?<=[A-Z])[^A-Z]+", "", x, perl = TRUE)
 
@@
   subset(param == "pred_med" & stock != "soc10") |> 
-  ggplot(aes(y = level)) + 
+  ggplot(aes(y = initials(level))) + 
   geom_segment(aes(x = low95, xend = upp95)) + 
@@
   geom_point(aes(x = mean), size = 2) +
+  geom_text(data = data_letter, x = -Inf, y = Inf, hjust = -1, vjust = 1, col = 1, aes(label = lab)) +
   facet_grid(
-    effect ~ stock, scales = "free", 
-    labeller = labeller(stock = labels_stock, effect = labels_effect)) +
+    labels_effect(effect) ~ labels_stock(stock), 
+    scales = "free", switch = "both", labeller = label_parsed) +
   labs(x=NULL, y = NULL) +
-  theme_bw()
+  theme_bw() +
+  theme(strip.background = element_blank(), strip.placement = "outside")
 ```
@@
 labels_stock = function(x)
-  ifelse(x=="agc", "AGC", ifelse(grepl("10", x), "SOC (0-10 cm)", "SOC (0-30 cm)"))
+  ifelse(x=="agc", "a. AGC", ifelse(grepl("10", x), "c. SOC (0-10 cm)", "b. SOC (0-30 cm)"))
 
@@
   scale_y_discrete(labels= c("AGC", "Clay", "MAT")) +
-  theme_bw()
-```
-
-```{r fig.id = "plot-reg", fig.height = 3.5, fig.width = 7, fig.cap = "Scatter plot of aboveground carbon (AGC; x-axis) and soil organic carbon (SOC; y-axis) stocks.
-SOC stocks were estimated at the topsoil layer (0-10 cm; left panel) and at the total measured depth (0-30 cm; right panel).
+  theme_bw() +
+  theme(strip.text = element_text(hjust = 0, size = 10), strip.background = element_blank())
+```
+
+```{r fig.id = "plot-reg", fig.height = 3.5, fig.width = 7, fig.cap = "Scatter plot of aboveground carbon (AGC; x-axis) and soil organic carbon (SOC; y-axis) stocks, (a) at the total measured depth (0-30 cm) and (b) at the topsoil layer (0-10 cm).
 Each point represents the mean SOC stocks in a sample plot, and the vertical and horizontal error bars are the 95% confidence intervals.
-The colours correspond to the different agroforestry systems (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden); the different shapes represent the two sites.
+The colours correspond to the different agroforestry systems (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden); the different shapes represent the two sites (round: Gariuai, triangle: Osso-Luga).
 The dotted lines represent the mean predicted SOC stocks for a given AGC stock value, and the grey areas represent the 95% confidence interval of this prediction."}
@@
 ), .(agc, stock)]
-dfpred[, depth := ifelse(grepl(10, stock), 
-                         "Soil depth: 0-10 cm", "Soil depth: 0-30 cm")]
 
@@
   dcast(plot + site + afs ~ variable, value.var = c("mean", "lwr", "upr"))
-dataC_30$depth <- "Soil depth: 0-30 cm"
+dataC_30$stock <- "soc"
 dataC_10 <- dataC_plot[variable %in% c("AGC", "SOC10")] |> 
@@
 colnames(dataC_10) <- gsub("SOC10", "SOC", colnames(dataC_10))
-dataC_10$depth <- "Soil depth: 0-10 cm"
+dataC_10$stock <- "soc10"
 dataC_fig <- rbind(dataC_10, dataC_30)
@@
 dataC_fig[, afs := factor(afs, levels = c("CF", "SP", "YA", "HG", "FG"))]
+
+# create strip labels
+labels_depth <-
+  function(x) ifelse(grepl(10, x), "b. SOC (0-10 cm)", "a. SOC (0-30 cm)")
 
@@
   scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
-  facet_wrap(~depth) +
-  theme_classic()
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  facet_wrap(~stock, labeller = labeller(stock = labels_depth)) + 
+  theme_classic() +
+  theme(strip.text = element_text(hjust = 0, size = 11), strip.background = element_blank())
 ```
@@
 
-```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with five soil variables: clay, silt and sand fraction, soil pH, and cation exchange capacity (CEC).
+```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with four soil variables: clay, silt and sand fraction, and soil pH.
 Values for the soil variables were measured from in situ soil samples for each of the 30 study plots.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
-g1 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n", "CEC\n")] |>
-  setnames(c("clay", "silt", "sand", "pH", "CEC")) |>
+g1 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n")] |>
+  setnames(c("clay", "silt", "sand", "pH")) |>
   cor() |> melt() |>
@@
 
-g2 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n", "CEC\n")] |>
-  setnames(c("clay", "silt", "sand", "pH", "CEC")) |>
+g2 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n")] |>
+  setnames(c("clay", "silt", "sand", "pH")) |>
   scale() |>

### 2025-01-08 manuscript.Rmd
figure 2 with similarly spaced x axes

@@
 
-UR ABSys, CIRAD, Univ Montpellier, CIHEAM-IAMM, INRAE, Institut Agro, Montpellier, France
-
-HRPP, Kasetsart University, Bangkok, Thailand
+ABSys, CIRAD, INRAE, Institut Agro, Université de Montpellier, Montpellier, France
+
+CIRAD, UMR ABSys, Montpellier F-34398, France
+
+HRPP, Kasetsart University, Bangkok 10900, Thailand
 
@@
 
-Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation biomass and soil, and this carbon storage capacity depends on the type of the AFS in addition to environmental and edaphic conditions [@Albrecht2003; @Feliciano2018].
+Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation biomass and soil, and this carbon storage capacity depends on the type of the AFS in addition to environmental conditions [@Albrecht2003; @Feliciano2018].
 Since trees can store large amounts of carbon, the carbon stored in vegetation aboveground biomass of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
-However, the effect of AFS on soil organic carbon (SOC) is more variable: although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e., planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFS, while AFS such as silvopastures have been shown to strongly increase SOC stocks [@Feliciano2018].
-Other factors such as climate, soil characteristics and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Feliciano2018; @Ma2020; @Martin2020].
+However, the effect of AFS on soil organic carbon (SOC) is more variable.
+Although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e, planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFS [@Feliciano2018], while AFS such as alley cropping and home gardens have been shown to increase SOC stocks [@Oelbermann2004;@Shi2018], and other AFS such as silvopastures have mixed results [@Feliciano2018;@Shi2018].
+Other factors such as climate, soil characteristics and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Nair2009;@Feliciano2018;@Ma2020;@Martin2020].
 Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS was only evidenced to a depth of 10 cm [@Ma2020].
@@
 
-In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying the carbon storage in both vegetation biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
-[xxx 2 sites] We hypothesize that there are measurable differences in carbon in aboveground vegetation biomass (hereinafter referred to as above ground carbon or AGC; Hypothesis 1a) and SOC (Hypothesis 1b) stocks between these different types of traditional AFS.
-In line with @Ma2020, we anticipate that there is a positive relationship between AGC and SOC stocks in the topsoil layer (0-10 cm; Hypothesis 2a), but no relationship with SOC below 10 cm (Hypothesis 2b).
+In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying carbon storage in both vegetation biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
+To do this, we quantified carbon stored in biomass and soil in 30 agroforestry plots across 2 sites in the Baucau district of Timor-Leste.
+We developed a modeling framework to quantify the effect of agricultural practices (here, different types of traditional AFS) on vegetation aboveground carbon (AGC) and SOC stocks, while controlling for climate and soil variability in the study area.
+Additionally, we quantified the effect of AGC on SOC stocks: in line with @Ma2020, we expected a positive relationship of AGC with SOC in the topsoil layer (0-10 cm), but no relationship with SOC below 10 cm.
 
@@
 
-where $Corg_{a,b}$ is the average SOC content between depths a and b (see rationale for equation \@ref(eq:intersoc) in Supplemental Information).
+where $Corg_{a,b}$ is the average SOC content between depths a and b (see rationale for equation \@ref(eq:soc) in Supplemental Information).
 
@@
 
-To test the relative effect of AFS type, climate and soil on carbon stocks, we modelled both AGC and SOC stocks as:
+We developed a Bayesian modelling framework to quantify the relative effect of AFS type, climate and soil on carbon stocks. We modelled both AGC and SOC stocks as:
 
 $$Y_{i,j,k} = \mu +  \sum_{l} \left( \theta_l \cdot X_{l,i} \right) + \alpha_{j} + \beta_{k} + \epsilon_{i}$$
-with $Y_{i,j,k}$ the stocks (either AGC or SOC) of plot $i$ in site $k$ and AFS type $j$.
-
-$\mu$ is the average stocks in our study area.
+with $Y_{i,j,k}$ the stocks (either AGC or SOC) of plot $i$ of AFS type $j$, in site $k$ .
+
+$\mu$ is the average stocks in the study area.
 
 $\theta_l$ is the effect of covariate $l$ on variable $Y$, and $X_{l,i}$ is the value of covariate $l$ in plot $i$.
-We included the following covariates: elevation [from NASA SRTM; @SRTM2024] to represent the climate gradient in our study area, and clay fraction (as measured in the Cirad laboratories; see “Field inventories” section) to represent the soil gradient in our study area.
+We included the following covariates: mean annual temperature [extracted from CHELSA V2.1; @Karger2017] to represent the climate gradient in the study area, and clay content (as measured in the Cirad laboratories; see “Field inventories” section) to represent the soil gradient in the study area.
 The rationale for the choice of these variables is given in the Supplemental Information.
-In addition, we included AGC stock as a covariate in the SOC stock model, as we wanted to quantify its effect on SOC stocks (hypotheses H2a and H2b).
+In addition, we included AGC stock as a covariate in the SOC stock model, as we wanted to quantify its effect on SOC stocks.
 Covariates were all centred and scaled.
@@
 
+We chose the standard half-normal distribution as the prior for all standard deviation parameters, and the standard normal distribution as the prior for all $\theta$ parameters.
+
 ### Model calibration
@@
 Calibration was carried out using an adapted form of the Hamiltonian Monte Carlo using Stan’s programming language [@Carpenter2017].
-We used the standard half-normal distribution as the prior for all standard deviation parameters, and the standard normal distribution for all $\theta$ parameters.
 To propagate uncertainties on the model outputs, we calibrated the model for each of the 1000 iterations and sampled one set of parameter values in the posterior distributions for each iteration.
@@
 AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-AGC stocks varied substantially between AFS, but not between sites (Figure \@ref(fig:effect-afs-site)).
+AGC stocks varied substantially between AFS, but not between sites (Figure \@ref(fig:effect-afs-site)a,c).
 The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
@@
 SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-SOC stocks varied substantially between sites, but not between AFS (Figure \@ref(fig:effect-afs-site)).
+SOC stocks varied substantially between sites, but not between AFS (Figure \@ref(fig:effect-afs-site)b,d).
 The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$) than those in Osso-Luga (`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$; Figure \@ref(fig:barplot)).
@@
 
-## Effect of mean annual temperature and clay fraction on AGC and SOC stocks
-
-Zero was included in the 80% credibility intervals for the effect of mean annual temperature and clay fraction on AGC, meaning that no strong effect was detected in our data (Figure \@ref(fig:effect-agc-clay)).
-Similarly, the clay fraction did not show a strong effect on SOC stocks, as zero was included in its 80% credible intervals (Figure \@ref(fig:effect-agc-clay)).
-
-Mean annual temperature, on the contrary, had a stronger negative effect on SOC stocks at both 0-10 cm and 0-30 cm depth (Figure \@ref(fig:effect-agc-clay)).
+## Effect of mean annual temperature and clay content on AGC and SOC stocks
+
+Zero was included in the 80% credibility intervals for the effect of mean annual temperature and clay content on AGC, meaning that no strong effect was detected in our data (Figure \@ref(fig:effect-agc-clay)a).
+Similarly, the clay content did not show a strong effect on SOC stocks, as zero was included in its 80% credible intervals (Figure \@ref(fig:effect-agc-clay)b,c).
+
+Mean annual temperature, on the contrary, had a stronger negative effect on SOC stocks at both 0-10 cm and 0-30 cm depth (Figure \@ref(fig:effect-agc-clay)b,c).
 Values of $\theta_{MAT}$ were negative in `r sum(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value<0)/10`% of cases for SOC at 0-30 cm depth, and in `r sum(subset(data_par, stock=="soc10" & param == "theta" & level == "mat")$value<0)/10`% of cases for SOC at 0-10 cm depth.
@@
 
-The relationship between AGC and SOC stocks was weak in our dataset (Figure  \@ref(fig:plot-reg)), as the 80% credibility interval of the associated parameter ($\theta_{AGC}$) included zero (Figure \@ref(fig:effect-agc-clay)).
+The relationship between AGC and SOC stocks was weak in our dataset (Figure  \@ref(fig:plot-reg)), as the 80% credibility interval of the associated parameter ($\theta_{AGC}$) included zero (Figure \@ref(fig:effect-agc-clay)b,c; Figure \@ref(fig:plot-reg)).
 
@@
 In a meta-analysis by @Shi2018, it was estimated that AFS stored around 50 Mg C ha$^{-1}$ at a 0-30 cm depth.
-Such large SOC stocks could be due to the high exchangeable Ca^2+^ content at the sites included in this study (Table \@ref(tab:site-soil)), as Ca^2+^ is frequently attributed to enhanced organic matter stabilization [@Wiesmeier2019].
+The higher SOC stocks in this study could be due to the high exchangeable Ca^2+^ content at the sites included in this study (Table \@ref(tab:site-soil)), as Ca^2+^ is frequently attributed to enhanced soil organic matter stabilization [@Wiesmeier2019].
 In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]` Mg C ha$^{-1}$) were similar to the average of 74 Mg C ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
@@
 
-The AFS studied were found to differ significantly in terms of AGC stocks, in agreement with our hypothesis H1a.
+The AFS studied were found to differ significantly in terms of AGC stocks.
 This result reflects the density and size of trees in the plot, which in turn depend on the importance of trees in the production system.
@@
 This more limited role for trees may explain why their total biomass is lower in these systems.
-Indeed, in CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun, while in SP systems, the growth of grazing grasses is favored by reducing the number of trees.
+Indeed, in CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun.
+In SP systems, the growth of grazing grasses is favored by reducing the number of trees; however, the leaves and fruits of some trees (e.g.
+*Ziziphus mauritiana*, *Tamarindus indica*, *Schleichera oleiosa*) can provide additional fodder, especially during the dry season.
 
@@
 
-We found no significant difference in SOC between the AFS (H1b rejected).
+We found no significant difference in SOC between the AFS.
 This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017; @Wiesmeier2019], and with previous studies that show significant differences in SOC stocks between AFS [e.g., @Ramesh2015].
 However, it is worth noting that some single-site studies [e.g., @Betemariyam2020; @GamaRodrigues2010] and at least two meta-analyses failed to find any sizable effect of AFS on SOC stocks.
-@Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by the type of AFS, and @Shi2018 found no significant differences in SOC stocks between four different AFS in 354 previous studies.
+Indeed, @Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by the type of AFS, and @Shi2018 found no significant differences in SOC stocks between four different AFS in 354 previous studies.
 We propose three main explanations to reconcile these different findings.
@@
 Increasing the number of replicates for each AFS could have increased the power of the tests, and potentially resulted in more significant differences between AFS.
-Third, other pedoclimatic and historical factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g., within-site variability in climate, soil texture (Figures S\@ref(fig:clay-corg), S\@ref(fig:altitude-corg)) or previous land use may have significantly influenced SOC stocks [@Wiesmeier2019; @Ma2020].
+Third, other factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g., within-site variability in land-use history.
 Our experimental design aimed to limit as much as possible the influence of external co-factors at each site (e.g., variability in farmers' management practices, micro-topography, etc.), but much of the variability cannot be controlled in farmers' plots.
 
-We found a significant difference in SOC between the two sites, with higher SOC stocks at Gariuai than at Osso-Luga.
-A priori, three factors that differentiate the sites could explain these differences in SOC stocks: climate (through elevation; Figure \@ref(fig:map-sites); Figure S\@ref(fig:altitude-corg)), soil texture (Table S\@ref(fig:site-soil); Figure S\@ref(fig:clay-corg)) and history.
-First, we found that SOC stocks were positively correlated with elevation, both between and within sites (Figure S\@ref(fig:altitude-corg)).
-This result is consistent with those of @Raich2006, who found higher SOC at higher elevations in tropical evergreen forests in various countries (e.g., Congo, Thailand).
-As changes in elevation are associated with variations in climate (precipitation, temperature), our result is also consistent with previous studies that found higher SOC in areas with higher precipitation [@Schuur2001] and lower temperature [@Raich2006; @MarinSpiotta2013].
-Second, soil texture is a key driving process for SOC storage in the tropics [@Don2010; @Wiesmeier2019; @Matus2021].
-In particular, many studies have found a strong positive correlation between SOC stocks and the fine mineral fractions of soils [i.e., clay or clay + fine silt content; @Zinn2005].
-In our study, we found no significant correlation between SOC stocks and clay + fine silt content when site and depth were included in the model (Figure S\@ref(fig:clay-corg)).
+We found that SOC stocks were negatively correlated with mean annual temperature (Figure \@ref(fig:effect-agc-clay)b,c), reflecting the climatic and elevational gradient in the study sites (Figure S\@ref(fig:choice-clim-vars)).
+This result is consistent with that of @Raich2006, who found higher SOC stocks at higher elevations in tropical evergreen forests in different countries (e.g., Congo, Thailand), and with that of @MarinSpiotta2013, who found mean annual temperature to be the main driver of SOC stocks in successional and planted forests.
+
+Surprisingly, we found no significant correlation between SOC stocks and clay content (Figure \@ref(fig:effect-agc-clay)b,c).
+Soil texture is a key driving process for SOC storage in the tropics [@Don2010; @Wiesmeier2019; @Matus2021].
+In particular, many studies have found a strong positive correlation between SOC stocks and the fine mineral fractions of soils [i.e., clay or clay content; @Zinn2005].
 Furthermore, the Gariuai soils, which had the highest SOC stocks, were also the least clayey.
 This contradicts the typical (positive) relationship between the fine mineral fraction and SOC stocks.
-It is therefore unlikely that the difference in SOC stocks between the two sites can be attributed to soil texture.
-Thirdly, history could be a significant driving factor.
-Gariuai was continuously inhabited for over 50 years, whereas Osso-Luga was abandoned for almost three decades.
+It is therefore unlikely that the variability in SOC stocks in this study can be attributed to soil texture.
+
+After controlling for agricultural practices, and for the main pedoclimatic gradients in the study area, we still found a substantial difference in SOC between the two sites, with higher SOC stocks at Gariuai than at Osso-Luga (Figure \@ref(fig:effect-afs-site)d).
+Land-use history could be a significant driving factor: Gariuai was continuously inhabited for over 50 years, whereas Osso-Luga was abandoned for almost three decades.
 The management and care of the AFS in Gariuai, such as pruning, organic inputs related to the presence of livestock, as well as optimization of the different strata in terms of C input, may have led to an increase in soil organic carbon compared to abandoned systems [@Lorenz2014].
@@
 
-We found no significant relationship between AGC and SOC stocks in either the topsoil layer 0-10 cm depth (H2a refuted) or at 0-30 cm depth (H2b confirmed).
+We found no significant relationship between AGC and SOC stocks in either the topsoil layer 0-10 cm depth or at 0-30 cm depth.
 Overall, AGC stocks were a poor predictor of SOC stocks.
@@
 On the one hand, the addition of easily decomposable litter could stimulate microbial decomposition and turnover of old stored SOC [i.e., priming effect; @Sayer2019].
-On the other hand, the capacity of the soil to store SOC through organo-mineral associations would be limited by its fine element content (clay + fine silt) (i.e., saturation concept).
+On the other hand, the capacity of the soil to store SOC through organo-mineral associations would be limited by its fine element content (clay + fine silt; i.e., saturation concept).
 
@@
   geom_text(data = data_letter, x = -Inf, y = Inf, hjust = -1, vjust = 1, col = 1, aes(label = lab)) +
+  scale_x_continuous(breaks = seq(20, 100, 10)) +
   facet_grid(
     labels_effect(effect) ~ labels_stock(stock), 
-    scales = "free", switch = "both", labeller = label_parsed) +
+    scales = "free", space = "free", switch = "both", labeller = label_parsed) +
   labs(x=NULL, y = NULL) +
@@
 
-```{r fig.id = "effect-agc-clay", fig.width = 7, fig.height = 2, fig.cap = "Effect of covariates (MAT: mean annual temperature, Clay: clay fraction, AGC: aboveground carbon) on carbon stocks (SOC: soil organic carbon, at 0-10 cm or 0-30 cm depth).
+```{r fig.id = "effect-agc-clay", fig.width = 7, fig.height = 2, fig.cap = "Effect of covariates (MAT: mean annual temperature, Clay: clay content, AGC: aboveground carbon) on carbon stocks (SOC: soil organic carbon, at 0-10 cm or 0-30 cm depth).
 Covariates are all centred and scaled; positive values indicate that C stocks tend to be higher for higher values of the covariate.
@@
 labels_depth <-
-  function(x) ifelse(grepl(10, x), "b. SOC (0-10 cm)", "a. SOC (0-30 cm)")
+  function(x) ifelse(grepl(10, x), "b. Depth: 0-10 cm", "a. Depth: 0-30 cm")
 
@@
 
-pca_soil <- data_soil_melt[, c("pH\n", "CEC\n",  "CLAY", "silt", "sand")] |> 
+pca_soil <- data_soil_melt[, c("pH\n", "CLAY", "silt", "sand")] |> 
   scale() |> princomp()
@@
 We selected the following set of variables to represent the effect of climate: mean annual temperature (BIO1), mean annual precipitation (BIO12), annual precipitation variability (BIO15) over the period 1981-2010.
-These bioclimatic variables were extracted from CHELSA V2.1 [@Krager2017].
+These bioclimatic variables were extracted from CHELSA V2.1 [@Karger2017].
 These variables were all highly correlated in our dataset (Figure S\@ref(fig:choice-clim-vars)a).
@@
 
-We selected the following set of variables to represent the effect of soil properties on C stocks: clay fraction (<2 $\mu$m), silt fraction (2-50 $\mu$m), sand fraction (50-2000 $\mu$m), pH and cation exchange capacity xx.
-These edaphic variables were measured on our soil samples analysed in the Cirad laboratories (see section "Field inventories").
-After scaling and centering these variables, we performed a principal component analysis: the first component explained `r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance in our data set (Figure S\@ref(fig:choice-soil-vars)b).
-We therefore decided to retain only one variable to represent this edaphic gradient: clay fraction, which was one of the most highly correlated with the first component of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been shown to influence C stocks in other contexts [@Wiesmeier2019].
-
-\## Tables
+We selected the following set of variables to represent the effect of soil properties on C stocks: clay content (<2 $\mu$m), silt content (2-50 $\mu$m), sand content (50-2000 $\mu$m), and pH.
+These variables were measured on our soil samples analysed in the Cirad laboratories (see section "Field inventories").
+After scaling and centering these variables, we performed a principal component analysis: the first component explained `r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance in our dataset (Figure S\@ref(fig:choice-soil-vars)b).
+We therefore decided to retain only one variable to represent this edaphic gradient: clay content, which was the most highly correlated with the first component of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been shown to be a good predictor of SOC stocks in other contexts [@Wiesmeier2019].
+
+## Supplemental tables
 
@@
 
-\## Figures
+## Supplemental figures
 
@@
 
-<!-- ```{r fig.id="altitude-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap = "Effect of plot elevation (x-axis; in m) on SOC stocks (y-axis; in\ Mg\ C\ ha^-1^).
-Points are colored by site; the point corresponds to the mean prediction at that site, while the error bars correspond to the 95% confidence interval.
-We performed a linear mixed effects regression with log-transformed SOC stocks as the dependent variable, log-transformed elevation as the independent variable, and site as random effect.
-Results are presented as regression lines and the p-value of the relationship between SOC stocks and elevation.", fig.cap.pre="Figure S"} -->
-<!-- dataC_plot <- read.csv("dataC_plot.csv") |> setDT() -->
-<!-- dataCelev <- merge(dataC_plot[variable=="SOC", c("plot", "site", "mean", "upr", "lwr")],  -->
-<!--                    data_plot[, c("N_PLOT", "long", "lat")], by.x = "plot", by.y = "N_PLOT") -->
-
-<!-- # extract elevation -->
-<!-- dataCelev[, elev := terra::extract( -->
-<!--   terra::rast("data/srtm_62_14.tif"),  -->
-<!--   cbind(long, lat) -->
-<!-- )] -->
-
-<!-- regElev = lmer(log(mean) ~ log(elev) + (1|site), data = dataCelev) |>  -->
-<!--   summary() -->
-<!-- pvalueE = signif(regElev$coefficients[2, 5], 2) -->
-<!-- slp = regElev$coefficients[2, 1] -->
-
-<!-- dfElev = dataCelev[, .( -->
-<!--   int = mean(log(mean)) - slp * mean(log(elev)),  -->
-<!--   elev = min(elev), elevMax = max(elev), -->
-<!--   slp = slp), .(site)]  -->
-<!-- dfElev[, mean := exp(slp*log(elev) + int)] -->
-<!-- dfElev[, predMax := exp(slp*log(elevMax) + int)] -->
-
-<!-- ggplot(dataCelev, aes(x = elev, y = mean, col = site)) + -->
-<!--   geom_pointrange(aes(ymin=lwr, ymax = upr)) +  -->
-<!--   geom_segment(data = dfElev, aes(xend = elevMax, yend = predMax)) + -->
-<!--   theme_classic() + -->
-<!--   annotate(geom = "text", label = paste("p-value =", pvalueE), x = 500, y = 32) + -->
-<!--   labs(x = "Elevation (m)", y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), col = "Site") + -->
-<!--   scale_x_log10() + scale_y_log10() -->
-<!-- ``` -->
-
-<!-- </br> -->
-
-<!-- ```{r fig.id = "clay-corg", fig.height = 4, fig.width = 6, fig.lp="supp-fig", fig.cap="Distribution of SOC content (y-axis; in %) values as a function of fine silt and clay content (x-axis; in %), soil depth (point color; in cm), and site (point shape).
-We performed a linear mixed effects regression with log-transformed SOC content as the dependent variable, log-transformed fine soil texture as the independent variable, and soil depth and site as random effects.
-Results are presented as regression lines and the p-value of the relationship between SOC content and fine soil texture.", fig.cap.pre="Figure S"} -->
-<!-- data_soil[, site := substr(ID_VILLAGE, 1, 1)] -->
-<!-- data_soil[site == "S", site := "Osso-Luga"] -->
-<!-- data_soil[site == "G", site := "Gariuai"] -->
-
-<!-- regText = lmer( -->
-<!--   log(Corg) ~ log(as.numeric(CLAY) + as.numeric(`FINE SILT`)) +  -->
-<!--     (1|site)+  -->
-<!--     (1|SAMPLE_DEPTH),  -->
-<!--   data = data_soil[SAMPLE_DEPTH != "10-20"]) |>  -->
-<!--   summary() -->
-<!-- pvalueT = signif(regText$coefficients[2, 5], 3) -->
-<!-- slp = regText$coefficients[2, 1] -->
-
-<!-- dfText = data_soil[SAMPLE_DEPTH != "10-20" & !is.na(CLAY), .( -->
-<!--   int = mean(log(Corg)) - slp * mean(log(as.numeric(CLAY) + as.numeric(`FINE SILT`))),  -->
-<!--   xmin = min(as.numeric(CLAY) + as.numeric(`FINE SILT`)),  -->
-<!--   xmax = max(as.numeric(CLAY) + as.numeric(`FINE SILT`)), -->
-<!--   slp = slp), .(site, SAMPLE_DEPTH)]  -->
-<!-- dfText[, predMin := exp(slp*log(xmin) + int)] -->
-<!-- dfText[, predMax := exp(slp*log(xmax) + int)] -->
-
-<!-- ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, shape = SAMPLE_DEPTH, col = site)) + -->
-<!--   geom_point() + -->
-<!--   geom_segment(data = dfText, aes(x = xmin, y = predMin, xend = xmax, yend = predMax, lty = SAMPLE_DEPTH)) + -->
-<!--   scale_x_log10() + scale_y_log10() + -->
-<!--   scale_linetype_manual(values=c("twodash", "dotted")) + -->
-<!--   labs(shape = "Soil depth (cm)", lty = "Soil depth (cm)",  -->
-<!--        x = "Clay and fine silt content (%)",  -->
-<!--        y = "SOC content (%)",  -->
-<!--        col = "Site") + -->
-<!--   annotate(geom = "text", label = paste("p-value =", pvalueT), x = 45, y = 1) + -->
-<!--   theme_classic() -->
-<!-- ``` -->
-
 </br>
 
-```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesised effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
+```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesized effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
 # save_png <- function(plot, path){
@@
 
-```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with four soil variables: clay, silt and sand fraction, and soil pH.
+```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with four soil variables: clay, silt and sand content, and soil pH.
 Values for the soil variables were measured from in situ soil samples for each of the 30 study plots.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}

### 2025-04-14 manuscript.Rmd
manuscript after R1

@@
 abstract: |
-  Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and/or other woody perennials into agricultural land thereby  diversifying landscapes.
+  Agroforestry has the potential to make agriculture more resilient while improving carbon sequestration by incorporating trees and/or other woody perennials into agricultural land, thereby diversifying landscapes.
 Traditional agricultural systems in tropical areas often include trees, but their carbon sequestration potential is not always well described, hindering their inclusion in climate change mitigation strategies.
@@
 Our results show that these traditional AFS can store large amounts of carbon, with the average being 155\ Mg\ C\ ha$^{-1}$.
-The AFS with the highest carbon stocks (FG) stored an average of 210\ Mg\ C\ ha$^{-1}$,  close to the values of old-growth tropical forests.
-Other AFS stored on average: 107\ Mg\ C\ ha$^{-1}$ in CF, 158\ Mg\ C\ ha$^{-1}$ in SP, 132\ Mg\ C\ ha$^{-1}$ in YA, and 167\ Mg\ C\ ha$^{-1}$ in HG.
-Biomass carbon stocks varied significantly between AFS (which differed in tree cover), while soil carbon stocks were less variable between AFS but more dependent on site.
+The AFS with the highest total carbon stocks (FG) stored an average of 213\ Mg\ C\ ha$^{-1}$.
+The average stocks stored by the other AFS were 108\ Mg\ C\ ha$^{-1}$ in CF, 158\ Mg\ C\ ha$^{-1}$ in SP, 134\ Mg\ C\ ha$^{-1}$ in YA, and 171\ Mg\ C\ ha$^{-1}$ in HG.
+Biomass carbon stocks varied substantially between AFS (which differed in tree cover), while soil carbon stocks were less variable between AFS but more site-dependent.
 We found no relationship between the amount of carbon stored in biomass and soil.
@@
 nocite: |
+  @Nair1993, @Nair2021
 csl: misc/springer-basic-author-date.csl
@@
 
-Camille Piponiot(\*) - Marguerite Cogné - Vincent Freycon - Régis Peltier
+Camille Piponiot(\*) - Marguerite Cogné - Vincent Freycon - Régis
+Peltier
 
@@
 
-(\*) email: [camille.piponiot-laroche\@cirad.fr](mailto:camille.piponiot-laroche@cirad.fr){.email}
+(\*) email:
+[camille.piponiot-laroche\@cirad.fr](mailto:camille.piponiot-laroche@cirad.fr){.email}
 
@@
 
-ABSys, CIRAD, INRAE, Institut Agro, Université de Montpellier, Montpellier, France
+ABSys, CIRAD, INRAE, Institut Agro, Université de Montpellier,
+Montpellier, France
 
@@
 
-Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022; @Gibbs2010; @Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO~2~e/yr over 2010-2014; @Pendrill2019].
+Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022; @Gibbs2010; @Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO~2~e yr$^{-1}$ over 2010-2014; @Pendrill2019].
 The demand for tropical agricultural land is expected to continue to grow as the world’s population increases [@UN2022].
@@
 
-Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Albrecht2003; @Verchot2007; @Lasco2014; @Cardinael2021].
-Agroforestry is a collective name for land use systems and technologies where woody perennials (trees, shrubs, bamboos, creepers, etc.) are deliberately used on the same land management units as agricultural crops and/or animals, in some form of spatial arrangement or temporal sequence [@Lundgren1982].
-Therefore, agroforestry practices allow for the diversification of agricultural production, making agricultural systems more resilient to climate change [@Duffy2021; @Terasaki2023].
-
-Agroforestry systems (AFS) have the potential to store large amounts of carbon in the vegetation biomass and soil, and this carbon storage capacity depends on the type of the AFS in addition to environmental conditions [@Albrecht2003; @Feliciano2018].
-Since trees can store large amounts of carbon, the carbon stored in vegetation aboveground biomass of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
+Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation[@Albrecht2003; @Verchot2007; @Lasco2014; @Cardinael2021].
+Agroforestry is the collective name for land-use systems and technologies where woody perennials (trees, shrubs, bamboos, creepers, etc.) are deliberately used on the same land management units as agricultural crops and/or animals in some form of spatial arrangement or temporal sequence [@Lundgren1982].
+Agroforestry practices thus allow for the diversification of agricultural production, making agricultural systems more resilient to climate change [@Duffy2021; @Terasaki2023].
+Agroforestry practices have long been used in traditional agricultural systems for both environmental and economic benefits [@Viswanath2017].
+
+Agroforestry systems (AFS) have the potential to store large amounts of carbon in vegetation biomass and soil.
+This carbon storage capacity depends on the type of the AFS as well as environmental conditions [@Albrecht2003; @Feliciano2018].
+Since trees can store large amounts of carbon, the carbon stored in the vegetation aboveground biomass of AFS is largely dependent on the number and size of trees in the system [@Ma2020].
 However, the effect of AFS on soil organic carbon (SOC) is more variable.
-Although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e, planting trees with intercropping during the establishment phase) may decrease SOC stocks compared to non-AFS [@Feliciano2018], while AFS such as alley cropping and home gardens have been shown to increase SOC stocks [@Oelbermann2004;@Shi2018], and other AFS such as silvopastures have mixed results [@Feliciano2018;@Shi2018].
-Other factors such as climate, soil characteristics and previous land use type can also influence SOC in AFS, which may explain the variability of observations reported in the literature [@Nair2009;@Feliciano2018;@Ma2020;@Martin2020].
-Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS was only evidenced to a depth of 10 cm [@Ma2020].
+Although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e.
+planting trees with intercropping during the establishment phase) may decrease SOC stocks, possibly due to the initial soil disturbance caused by tree planting [@Feliciano2018].
+Conversely, AFS such as alley cropping and home gardens have been shown to increase SOC stocks [@Oelbermann2004; @Shi2018], while other AFS such as silvopastures have shown mixed results [@Feliciano2018; @Shi2018].
+Other factors such as climate, soil characteristics and previous land-use type can also influence SOC in AFS, which may explain the variability of the observations reported in the literature [@Nair2009; @Feliciano2018; @Ma2020; @Martin2020].
+Furthermore, the relationship between the amount of carbon stored in aboveground vegetation and soil in AFS has only been evidenced to a depth of 10 cm [@Ma2020].
 
@@
 However, population growth, increasing pressure on natural resources, and an increasingly variable climate with an unreliable rainfall regime are threatening food security [@WB2009; @Molyneux2012].
-In this context, productive, resilient, and sustainable agriculture is critical for the country and the livelihood of its people.
-To address this, AFS have been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022; @Cogne2024].
+In this context, productive, resilient, and sustainable agriculture is critical for the country and the livelihoods of its people.
+AFS have been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022; @Cogne2024].
 Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022; @Cogne2024] and can be key elements for biodiversity conservation [@Perfecto2008; @Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
@@
 
-In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying carbon storage in both vegetation biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
-To do this, we quantified carbon stored in biomass and soil in 30 agroforestry plots across 2 sites in the Baucau district of Timor-Leste.
-We developed a modeling framework to quantify the effect of agricultural practices (here, different types of traditional AFS) on vegetation aboveground carbon (AGC) and SOC stocks, while controlling for climate and soil variability in the study area.
-Additionally, we quantified the effect of AGC on SOC stocks: in line with @Ma2020, we expected a positive relationship of AGC with SOC in the topsoil layer (0-10 cm), but no relationship with SOC below 10 cm.
+In this study, we aim to better understand traditional AFS in Timor-Leste by quantifying carbon storage in both the vegetation biomass and soil of five traditional AFS: cropping systems with fallow, silvopastures, young agroforests, home gardens, and forest gardens.
+To do this, we quantified carbon stored in biomass and soil in 30 agroforestry plots on two sites in the Baucau municipality of Timor-Leste.
+We developed a modelling framework to quantify the effect of agricultural practices (here, different types of traditional AFS) on vegetation aboveground carbon (AGC) and SOC stocks, while controlling for climate and soil variability in the study area.
+In addition, we quantified the effect of AGC on SOC stocks.
+In line with @Ma2020, we expected a positive relationship of AGC with SOC in the topsoil layer (0-10 cm), but no relationship with SOC at 0-30 cm depth.
 
@@
 site_summary = site_env[, .(
-  coord = paste(round(mean(lat),2), round(mean(long),2), sep = ", "), 
+  coord = paste0(abs(round(mean(lat),2)), "°S, ", round(mean(long),2), "°E"), 
   prec = round(mean(prec)), 
@@
 
-The research was carried out in the center of the Baucau municipality, located on the north-eastern coast of Timor-Leste.
+The research was carried out in the centre of the Baucau municipality, located on the north-eastern coast of Timor-Leste.
 Two communities, or *suco* (a group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases.
-The first, Gariuai *suco* (Lat-Lon coordinates: `r site_summary[site=="Gariuai", coord]`), is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging between `r site_summary[site=="Gariuai", elev]` m above sea level (masl).
-The second, the village of Osso-Luga in Samalari *suco* (Lat-Lon coordinates: `r site_summary[site=="Osso-Luga", coord]`), is located on the Matebian foothills based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso-Luga", prec]` mm and an altitude ranging between `r site_summary[site=="Osso-Luga", elev]` masl (Figure \@ref(fig:map-sites)).
+The first, Gariuai *suco* (`r site_summary[site=="Gariuai", coord]`), is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging between `r site_summary[site=="Gariuai", elev]` m above sea level (masl).
+The second, the village of Osso-Luga in Samalari *suco* (`r site_summary[site=="Osso-Luga", coord]`), is located on the Matebian foothills based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso-Luga", prec]` mm and an altitude ranging between `r site_summary[site=="Osso-Luga", elev]` masl (Figure 1).
 The soils of the Baucau municipality are classified using the WRB soil classification [@Mantel2023] as Lithosols and Calcic Luvisols [@FAO2012].
@@
 The AFS in Osso-Luga lay fallow during the period of abandonment.
-The land use of these 2 villages is typical of the Baucau district.
-Land use in these two villages is typical of the Baucau district.
-The 5 AFS described in this article [see detailed description in @Cogne2024] are found in the valleys and high plateaux, while non-AFS such as rainfed rice and horticultural plots are more likely to be found in the lowlands.
-Other non-agricultural land uses are also present, such as limestone and limestone quarrying.
-Some natural forest parks are also preserved around mountain peaks, such as Matebian or Mundo Perdido.
-Villages are concentrated near paved roads and on hilltops, more rarely near rivers because of the risk of flooding.
-
+The land use of these two villages is typical of the Baucau municipality.
+The five AFS described in this article [see detailed description in @Cogne2024] are found in the valleys and high plateaus, while non-AFS such as rainfed rice and horticultural plots are more likely to be found in the lowlands.
+Other non-agricultural land uses are also present, such as limestone quarrying.
+Some natural forest parks are also preserved around mountain peaks, such as Matebian and Mundo Perdido.
+Due to the risk of flooding, villages are rarely located near rivers, and are concentrated instead near paved roads and on hilltops.
 ## Typology of agroforestry systems
@@
 We used the typology of AFS described in @Cogne2024, which was developed in the same study area.
-This typology was based on the one proposed by Nair *et al.* (1993, 2021) and adapted to the local context with semi-structured interviews previously conducted in another study at the same sites [@Cogne2024].
-The five AFS identified were as follows: Crop systems including a wooded Fallow phase (CF), Silvopastures (SP), Young Agroforests (YA), Home Gardens (HG), and Forest Gardens (FG).
+This typology was based on the one proposed by Nair et al.
+(1993, 2021) and adapted to the local context with semi-structured interviews previously conducted in another study at the same sites [@Cogne2024].
+The five AFS identified were as follows: Cropping systems with Fallow (CF), Silvopastures (SP), Young Agroforests (YA), Home Gardens (HG), and Forest Gardens (FG).
 The AFS are presented in more detail in the Supplemental Information (Table S\@ref(tab:afs-typology)) and in @Cogne2024.
 
-Briefly, the CF are plantations of a few non-perennial crops (e.g., maize [*Zea mays*], sweet potatoes [*Ipomoea batatas*], groundnuts [*Arachis hypogaea*], and squash gourd [*Cucurbita moschata*]) which are left fallow after one to three years of cropping.
-SP are vast areas (generally 200 to 500 ha) where cattle (*Bos taurus*, *Bos domesticus*, *Bubalus bubalis*) are grazed under the shade of palms, like coconut (*Cocos nucifera*), and trees, like tamarind (*Tamarindus indica*) or eucalyptus (*Eucalyptus alba*) and jujube (*Ziziphus sp.*).
+Briefly, CF are plantations of a few non-perennial crops (e.g., maize [*Zea mays*], sweet potatoes [*Ipomoea batatas*], groundnuts [*Arachis hypogaea*], and squash gourd [*Cucurbita moschata*]) which are left fallow after one to three years of cropping.
+SP are vast areas (generally 200 to 500 ha) where cattle (*Bos taurus*, *Bos domesticus*, *Bubalus bubalis*) are grazed under the shade of scattered palms and trees such as coconut (*Cocos nucifera*), tamarind (*Tamarindus indica*), eucalyptus (*Eucalyptus alba*) and jujube (*Ziziphus jujuba*).
 YA are agricultural plots where farmers plant rows of trees, like teak (*Tectona grandis*) and mahogany (*Swietenia mahagoni*), spaced approximately 5 x 10 m apart, which are generally left to grow to become HG or FG (see Figure S\@ref(fig:seq-diagram)).
-HG are plots of land located close to houses where crops and livestock (poultry [*Gallus gallus domesticus*], pigs [*Sus domesticus*], cattle, etc.) are kept in the shade of numerous trees, like Breadfruit (*Artocarpus altilis*), candlenut (*Aleurites moluccanus*) and palms, like areca palm (*Areca catechu*), for multiple uses (wood, fruit, fiber, etc.).
-FG are generally former HG (over 50 years old) which have been abandoned, for example as a result of the war, and are now used mainly for gathering fruits (e.g.
-mango [*Mangifera indica*]) and non-timber forest products (e.g.
-candlenut) with limited human influence.
+HG are plots of land located close to houses where crops and livestock (poultry [*Gallus gallus domesticus*], pigs [*Sus domesticus*], cattle [*Bos taurus*, *Bos domesticus*, *Bubalus bubalis*], etc.) are kept in the shade of numerous trees and palms, like breadfruit (*Artocarpus altilis*), candlenut (*Aleurites moluccanus*) and areca palm (*Areca catechu*), all of which are put to multiple uses (wood, fruit, fibre, etc.).
+FG are generally former HG (over 50 years old) which have been abandoned, for example as a result of the war, and are now used mainly for gathering fruits (e.g., mango [*Mangifera indica*]) and non-timber forest products (e.g., candlenut) with limited human influence.
 
@@
 Sampling plots were defined as the entire agricultural plot (area between 0.06 and 1.25 ha) for all AFS except SP plots, which were generally much larger than 1 ha.
-In SP plots, a 1 ha sampling plot was delineated within the agricultural plot.
+In SP plots, a 1 ha sampling plot was delineated at the centre of the agricultural plot.
 Tree and soil measurements (described below) were carried out between June and August 2021.
@@
 Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
-Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Figure S\@ref(fig:design)).
+Five subplots of 10x10 m (0.01 ha) were established at each corner and in the centre of the sampling plot (Figure S\@ref(fig:design)).
 Within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
@@
 
-Taxonomic identification was carried out using Tetum (a national official language) vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014] as well as the Australian Northern Territory Government's Native Plants Herbarium [@NT2013].
+Taxonomic identification was carried out using Tetum (a national official language) vernacular names, which were then linked to scientific names using the iNaturalist and Pl\@ntNet applications, Geoffrey Hull’s monograph [@Hull2006], the “Useful tropical plants” database [@Fern2014], and the Australian Northern Territory Government’s Native Plants Herbarium [@NT2013].
 The correspondence between Tetum names and scientific names is provided in the Supplemental Information (Table S\@ref(tab:tetum-names)).
 
-In each of the 30 sample plots, a composite of three samples was taken from three of the five 10 x 10 m subplots (including the plot center and two opposite corners; Figure S\@ref(fig:design)), covering the intra-plot variability and at a minimum distance of 1 m from any tree.
-Soils were sampled with a soil auger (Model Bor Tanah - Dormer Standard Soil Auger, Adiguna Karya Persada, Bogor, Indonesia) at 0-10 cm and 20-30 cm depth.
-As the length of the head auger (15 cm) was longer than our target soil depths (10 cm), we could not accurately sample the 10-20 cm soil layer.
-Soils samples were then air-dried and sieved at 2 mm.
-They underwent standard physical and chemical analyses at Cirad laboratories in Montpellier, France (Table S\@ref(tab:site-soil)).
-Total carbon content was determined by dry combustion, following the ISO 10694:1995 standard method, and measured with a thermal conductive detector (EA 1112 Elemental Analyser, Thermo Fischer Scientific, Netherlands).
-Carbonate content was determined with a calcimeter, following the ISO 10693 standard method.
-Soil organic carbon content was calculated as the difference of the total carbon and the carbonate contents.
-Soil texture (proportion of clay, silt and sand), and pH were measured as xxx.
-
-In addition, at one location in the center of the plot, three 250 cm$^3$ soil samples were taken with a soil sampling cylinder (5 cm in height, 7.98 cm inner diameter, and 8.4 cm outer diameter; Eijkelkamp Soil & Water, Giesbeek, Netherlands)  at depths of 2-7 cm, 12-17 cm and 22-27 cm.
-These depths were chosen to take account of the height of the cylinder (i.e.
-5 cm), the destruction of the soil below the cylinder during its extraction, and to be associated with depths 0-10 cm, 10-20 cm and 20-30 cm, the depths at which the organic carbon content was calculated or estimated.
-Each sample was dried at 105°C in a classic oven until a stable dry mass was reached and then sieved to 2 mm.
+In each of the 30 sample plots, a composite of three samples was taken from three of the five 10x10 m subplots (including the plot centre and the two opposite corners along the plot's longest diagonal; Figure S\@ref(fig:design)), covering the intra-plot variability and at a minimum distance of 1 m from any tree.
+Soils were sampled with a soil auger (Model Bor Tanah - Dormer Standard Soil Auger, Adiguna Karya Persada, Bogor, Indonesia) at depths of 0-10 cm and 20-30 cm.
+The sample from the 10-20 cm soil layer was omitted due to the difficulty of sampling with the 15-cm soil auger.
+Samples underwent standard physical and chemical analyses at CIRAD laboratories in Montpellier, France (Table S\@ref(tab:site-soil)).
+The soil organic carbon content of the samples collected with the soil auger was calculated as the difference between the total carbon and the inorganic carbon content of the soil.
+Total carbon content was determined using the Dumas dry combustion method, following the ISO 10694:1995 standard method, and measured with a thermal conductive detector (EA 1112 Elemental Analyser, Thermo Fischer Scientific, Netherlands).
+Soil inorganic content was determined using a calcimeter, and following the ISO 10693 standard method.
+Particle size distribution (proportion of clay, silt and sand) was determined using the pipette method and three classes: clay (\< 2 μm), silt (2–50 μm) and sand (50 μm - 2 mm).
+Soil pH was measured with a soil:deionized water ratio of 1:2.5.
+
+In addition, at the centre of the plot, three 250 cm^3 soil samples were taken with a soil sampling cylinder (5 cm in height, 7.98 cm inner diameter, and 8.4 cm outer diameter; Eijkelkamp Soil & Water, Giesbeek, Netherlands) at depths of 2-7 cm, 12-17 cm and 22-27 cm.
+These depths were chosen to take into account the height of the cylinder (i.e., 5 cm), and to be associated with depths 0-10 cm, 10-20 cm and 20-30 cm, the depths at which the organic carbon content was calculated or estimated.
+Each sample was dried at 105°C in a classic oven and weighed repeatedly until a stable dry mass was reached, and then sieved to 2 mm.
 The fine fraction (\< 2 mm) was weighed to be used in the SOC stocks calculation (equation \@ref(eq:soc)).
@@
                iter == "mean", 
-               HEIGHT*.9* 100,  ## height in cm
-               rnorm(length(HEIGHT), HEIGHT, HEIGHT*cvH)*
-                 runif(length(HEIGHT), .8, 1)* 100
+               HEIGHT,  ## height in m
+               rnorm(length(HEIGHT), HEIGHT, HEIGHT*cvH)
              )]
@@
                           rnorm(length(meanWD), meanWD,sdWD))]
-
-data_sim_agc[GENUS %in% palm_genera, 
-             AGC := WD * HEIGHT * (DIAM_130_CM/2)^2 * pi * 
-               0.89 * # volumetric shrinkage
-               1e-6 * # convert from g to tons
-               0.4713 # conversion factor from biomass to carbon
-]
+# palm biomass with Goodman et al 2013 equation that includes diameter, height
+# and dry mass fraction; we used an average dmf of 0.37 which 
+data_sim_agc[
+  GENUS %in% palm_genera, 
+  AGC := ((0.463 * DIAM_130_CM^2 * HEIGHT)^0.25 * 0.55512) ^ 4	* 0.4713 *1e-3]
 ```
@@
 Other taxa were assigned a plot-level average wood density.
-The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density.
-The volume of the palm trunks was estimated as a cylinder, using the height and diameter measured *in situ*, with a volumetric shrinkage of 11% [@WD2007].
-The height of the palms was measured above the leaves and was therefore multiplied by a conversion factor of 0.9 to obtain the trunk height.
-Biomass was then converted into carbon with a factor of 0.4713 [@Thomas2012].
+Palm biomass was estimated using the family level allometric equation of @Goodman2013 with DBH, stem height and dry mass fraction as predictors.
+As we didn't have species-specific values for dry mass fraction, we used the average value of 0.463 from @Goodman2013.
+Biomass was converted into carbon with a factor of 0.4713 [@Thomas2012].
 
@@
 This ratio was then multiplied by the estimated AGC stock to obtain tree-by-tree BGC stock.
+Plot-level AGC and BGC stocks were then obtained by summing tree-level stocks and dividing by the plot area.
 
@@
 # use CWD values from Chave et al., 2014
-
-data_sim_agc[, RS := exp(- 1.2312 - 0.0215 * DIAM_130_CM + 0.0002 * DIAM_130_CM^2 - 0.0007 * cwd)]
+# download from figshare https://figshare.com/articles/dataset/Root_Shoot_global_data_individual_trees/5144164?file=8754550
+# rs <- readxl::read_excel("C:/Users/piponiot-laroche/Downloads/S1Database.xlsx",
+#                          sheet = "DATA")
+# rs$BGB_kg <- as.numeric(rs$BGB_kg)
+# rs$wd <- as.numeric(rs$wd)
+# rs$dbh2 <- rs$DBH_cm^2
+# lm(log(BGB_kg/AGB_kg) ~ DBH_cm + dbh2 + wd,
+#    data = subset(rs, !is.na(BGB_kg) & BGB_kg != 0))
+
+data_sim_agc[, RS := exp(- 1.2312 - 0.0215 * DIAM_130_CM + 
+                           0.0002 * DIAM_130_CM^2 - 0.0007 * cwd)]
 data_sim_agc[, BGC := RS * AGC]
 ```
-
-Plot-level AGC and BGC stocks were then obtained by summing tree-level stocks and dividing by the plot area.
 
@@
 \begin{equation}
-SOC_p = \frac{\sum_i Corg_i \cdot (mfine_i \cdot 10^{-6}) \cdot (Splot \cdot \Delta d)}{Vcyl} (\#eq:soc)
+SOC_p = \sum_i Corg_{i,p} \cdot BDfine_{i,p}\cdot (1- \frac{Vcoarse_{i,p}}{Vcyl})\cdot (Surf \cdot \Delta d) \cdot 10^{-6}
 \end{equation}
 
-Where $Corg_i$ is the SOC content at depth i (either 0-10, 10-20 or 20-30 cm) as measured in the laboratory from the soil sampled by the auger and estimated at depth 10-20 cm (see details below); $mfine_i$ is the mass of the fine fraction of the soil in the sampling cylinder at depth i (in g), converted into Mg with a factor of 10$^-6$; $Surf\ =\ 10^8$ is the surface conversion factor from cm$^2$ to ha; $\Delta d$ is the difference in sample depths (= 10 cm) and $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$).
-This calculation corresponds to the method M4 recommended by @Poeplau2017, which correctly considers the coarse fraction of the soil.
-We considered that $mfine_i$ at depths of 2-7 cm, 12-17 cm and 22-27 cm was associated with $Corg_i$ at depths of 0-10, 10-20 cm and 20-30 cm, respectively.
-At depth $i$ = 10-20 cm, where the carbon content was not measured (see “Field inventories” section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a standard negative exponential model of SOC content decrease with soil depth [@Mishra2009;@Murphy2019]:
+Where $ Corg_{i,p}$ is the SOC content at depth $i$ (either 0-10, 10-20 or 20-30 cm) as measured in the laboratory from the soil sampled by the auger (depths 0-10 cm and 20-30 cm) and estimated at depth 10-20 cm (see details below).
+$BDfine_{i,p}$ is the bulk density of the fine fraction of the soil sampled with the cylinder at depth $i$ (in g cm$^{-3}$):
+
+$$BDfine_{i,p} = \frac{mfine_{i,p}}{Vcyl – Vcoarse_{i,p}}$$
+$mfine_{i,p}$ is the mass of the fine fraction (\< 2 mm) of the sampled soil; Vcyl is the volume of the sampling cylinder (250 cm$^3$); and $Vcoarse_{i,p}$ is the volume of the coarse fraction in the sampling cylinder.
+$Surf = 10^8$ is the surface conversion factor from cm$^2$ to ha; $\Delta d$ is the difference in sample depths (= 10 cm); 10^{-6}$ is the conversion factor from g to Mg.
+
+This equation could be reformulated [@Poeplau2017] as:
 
 \begin{equation}
-Corg_{10,20}=Corg_{0,10}\cdot\sqrt{\frac{Corg_{20,30}}{Corg_{0,10}}}
-(\#eq:soc)
+SOC_p = \sum_i Corg_{i,p} \cdot \frac{mfine_{i,p}}{Vcyl} \cdot (Surf \cdot \Delta d) \cdot 10^{-6} (\#eq:soc)
 \end{equation}
 
-where $Corg_{a,b}$ is the average SOC content between depths a and b (see rationale for equation \@ref(eq:soc) in Supplemental Information).
+We considered that $mfine_{i,p}$ at depths of 2-7 cm, 12-17 cm and 22-27 cm was associated with $Corg_{i,p}$ at depths of 0-10, 10-20 cm and 20-30 cm, respectively.
+At depth $i$ = 10-20 cm, where the carbon content was not measured (see “Field inventories” section), we estimated SOC content as a function of SOC content at 0-10 cm and 20-30 cm, assuming a standard negative exponential model of SOC content decrease with soil depth [@Mishra2009; @Murphy2019]:
+
+\begin{equation}
+Corg_{10-20,p}=Corg_{0-10,p}\cdot\sqrt{\frac{Corg_{20-30,p}}{Corg_{0-10,p}}}
+(\#eq:soc_depth)
+\end{equation}
+
+where $Corg_{a-b,p}$ is the average SOC content in plot $p$ between depths $a$ and $b$
+(see rationale for equation \@ref(eq:soc_depth) in Supplemental Information).
 
@@
 
-All statistical analyses were carried out using the R software [@R2023].
+All statistical analyses were carried out using R software [@R2023].
 
@@
 
-The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo function from the R package BIOMASS, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013] and large and small errors for 5 and 95% of trees, respectively [@Chave2004; @Rejou-Mechain2017].
-To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (as recommended by Cirad laboratories in Montpellier, France), and 10% for the mass of the fine fraction [@Page1999].
+The uncertainty on plot-level AGC was calculated with the AGBmonteCarlo function from the R package BIOMASS, using a (rather conservative) coefficient of variation of 15% on height measurements [@Hunter2013] and large and small errors for 5 and 95% of trees, respectively [@Chave2004;
+@Rejou-Mechain2017].
+To propagate uncertainty on SOC estimates, we used a coefficient of variation of 12% for organic carbon (as recommended by CIRAD laboratories in Montpellier, France), and 10% for the mass of the fine fraction [@Page1999].
 
@@
 For single plot estimates of carbon stocks, medians and 95% confidence intervals were estimated as quantiles of the results of the 1000 iterations.
-For multiplot estimates of carbon stocks, values within each grouping factor were additionally bootstrapped before calculating quantiles (median and 95% confidence intervals).
+For multiplot estimates of carbon stocks, values within each grouping factor were also bootstrapped before calculating quantiles (median and 95% confidence intervals).
 
@@
 
-$$Y_{i,j,k} = \mu +  \sum_{l} \left( \theta_l \cdot X_{l,i} \right) + \alpha_{j} + \beta_{k} + \epsilon_{i}$$
-with $Y_{i,j,k}$ the stocks (either AGC or SOC) of plot $i$ of AFS type $j$, in site $k$ .
+$$Y_{p,j,k} = \mu +  \sum_{l} \left( \theta_l \cdot X_{l,p} \right) + \alpha_{j} + \beta_{k} + \epsilon_{p}$$
+
+with $Y_{i,j,k}$ the stocks (either AGC or SOC) of plot $p$ of AFS type
+$j$, in site $k$ .
 
@@
 
-$\theta_l$ is the effect of covariate $l$ on variable $Y$, and $X_{l,i}$ is the value of covariate $l$ in plot $i$.
-We included the following covariates: mean annual temperature [extracted from CHELSA V2.1; @Karger2017] to represent the climate gradient in the study area, and clay content (as measured in the Cirad laboratories; see “Field inventories” section) to represent the soil gradient in the study area.
+$\theta_l$ is the effect of covariate $l$ on variable $Y$, and $X_{l,p}$
+is the value of covariate $l$ in plot $p$.
+We included the following covariates: mean annual temperature [extracted from CHELSA V2.1; @Karger2017] to represent the climate gradient in the study area, and clay content (as measured in the CIRAD laboratories; see “Field inventories” section) to represent the soil gradient in the study area.
 The rationale for the choice of these variables is given in the Supplemental Information.
@@
 
-$\alpha_{j}$ and $\beta_{k}$ are the random AFS and site effects, respectively distributed as:
+$\alpha_{j}$ and $\beta_{k}$ are the random AFS and site effects,
+respectively distributed as:
 
@@
 $$\beta_{k} \sim \mathcal{N}(0,\sigma_{\beta}^2)$$
-And $\epsilon_{i}$ the random error, distributed as:
-
-$$\epsilon_{i} \sim \mathcal{N}(0,\sigma^2)$$
-$\sigma$, $\sigma_{\alpha}$, and $\sigma_{\beta}$ are the standard deviation of the random error, the AFS effect and the site effect respectively.
-
-We chose the standard half-normal distribution as the prior for all standard deviation parameters, and the standard normal distribution as the prior for all $\theta$ parameters.
+
+And $\epsilon_{p}$
+the random error, distributed as:
+
+$$\epsilon_{p} \sim \mathcal{N}(0,\sigma^2)$$ $\sigma$,
+$\sigma_{\alpha}$, and $\sigma_{\beta}$ are the standard deviation of
+the random error, the AFS effect and the site effect respectively.
+
+We chose the standard half-normal distribution as the prior for all
+standard deviation parameters, and the standard normal distribution as
+the prior for all $\theta$ parameters.
 
@@
 
-model {
+transformed parameters {
   vector[N] pred_a;
@@
   }
+}
+
+model {
 
@@
               file = paste0("outputs/data_par_", n, ".csv"), row.names = FALSE)
+
   }
@@
 
-data_par <- read.csv("outputs/data_par.csv") |> setDT()
+data_par <- read.csv("outputs/data_par.csv") |> setDT() |> 
+  subset(!grepl("pred", param))
 
@@
 
-AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and `r dataC_plot[variable=="AGC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-AGC stocks varied substantially between AFS, but not between sites (Figure \@ref(fig:effect-afs-site)a,c).
-The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-
-SOC stocks varied between `r dataC_plot[variable=="SOC", round(min(mean))]` and `r dataC_plot[variable=="SOC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="SOC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-SOC stocks varied substantially between sites, but not between AFS (Figure \@ref(fig:effect-afs-site)b,d).
-The SOC stocks in Gariuai were on average higher (`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$) than those in Osso-Luga (`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]` Mg C ha$^{-1}$; Figure \@ref(fig:barplot)).
-
-Total carbon stocks ranged between `r dataC_plot[variable=="TOT", round(min(mean))]` and `r dataC_plot[variable=="TOT", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="TOT", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`% (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
-The AFS with the lowest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest total carbon stocks were the `r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of `r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+AGC stocks ranged between `r dataC_plot[variable=="AGC", round(min(mean))]` and
+`r dataC_plot[variable=="AGC", round(max(mean))]` Mg C ha$^{-1}$, with an average of `r dataC_plot[variable=="AGC", round(mean(mean))]` Mg C ha$^{-1}$ and a coefficient of variation of `r dataC_plot[variable=="AGC", round(sd(mean)/mean(mean)*100)]`% (Table
+\@ref(tab:summary); Figure \@ref(fig:barplot)).
+AGC stocks varied substantially between AFS, but not between sites: there was no overlap between the 95% credibility intervals of the lowest and highest AGC stocks of different AFS, but strong overlap between sites (Figure
+\@ref(fig:effect-afs-site)a,c).
+The AFS with the lowest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.min(mean)]]`, with an average of `r dataC_afs[variable=="AGC", round(mean[which.min(mean)])]` Mg C ha$^{-1}$; the AFS with the highest AGC stocks were the `r dataC_afs[variable=="AGC", afs[which.max(mean)]]` with an average of
+`r dataC_afs[variable=="AGC", round(mean[which.max(mean)])]` Mg C ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
+
+SOC stocks (0-30 cm) varied between
+`r dataC_plot[variable=="SOC", round(min(mean))]` and
+`r dataC_plot[variable=="SOC", round(max(mean))]` Mg C ha$^{-1}$, with
+an average of `r dataC_plot[variable=="SOC", round(mean(mean))]` Mg C
+ha$^{-1}$ and a coefficient of variation of
+`r dataC_plot[variable=="SOC", round(sd(mean)/mean(mean)*100)]`% (Table
+\@ref(tab:summary); Figure \@ref(fig:barplot)).
+SOC stocks varied substantially between sites, but not between AFS: there was no overlap between the 95% credibility intervals of the two sites, but strong overlap between AFS (Figure
+\@ref(fig:effect-afs-site)b,d). The SOC stocks in Gariuai were on
+average higher
+(`r dataC_site[site=="Gariuai" & variable=="SOC", round(mean)]` Mg C
+ha$^{-1}$) than those in Osso-Luga
+(`r dataC_site[site=="Osso-Luga" & variable=="SOC", round(mean)]` Mg C
+ha$^{-1}$; Figure \@ref(fig:barplot)).
+
+Total carbon stocks (AGC+BGC+SOC) ranged between
+`r dataC_plot[variable=="TOT", round(min(mean))]` and
+`r dataC_plot[variable=="TOT", round(max(mean))]` Mg C ha$^{-1}$, with
+an average of `r dataC_plot[variable=="TOT", round(mean(mean))]` Mg C
+ha$^{-1}$ and a coefficient of variation of
+`r dataC_plot[variable=="TOT", round(sd(mean)/mean(mean)*100)]`% (Table
+\@ref(tab:summary); Figure \@ref(fig:barplot)). The AFS with the lowest
+total carbon stocks were the
+`r dataC_afs[variable=="TOT", afs[which.min(mean)]]` with an average of
+`r dataC_afs[variable=="TOT", round(mean[which.min(mean)])]` Mg C
+ha$^{-1}$; the AFS with the highest total carbon stocks were the
+`r dataC_afs[variable=="TOT", afs[which.max(mean)]]` with an average of
+`r dataC_afs[variable=="TOT", round(mean[which.max(mean)])]` Mg C
+ha$^{-1}$ (Table \@ref(tab:summary); Figure \@ref(fig:barplot)).
 
@@
 
-Zero was included in the 80% credibility intervals for the effect of mean annual temperature and clay content on AGC, meaning that no strong effect was detected in our data (Figure \@ref(fig:effect-agc-clay)a).
-Similarly, the clay content did not show a strong effect on SOC stocks, as zero was included in its 80% credible intervals (Figure \@ref(fig:effect-agc-clay)b,c).
-
-Mean annual temperature, on the contrary, had a stronger negative effect on SOC stocks at both 0-10 cm and 0-30 cm depth (Figure \@ref(fig:effect-agc-clay)b,c).
-Values of $\theta_{MAT}$ were negative in `r sum(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value<0)/10`% of cases for SOC at 0-30 cm depth, and in `r sum(subset(data_par, stock=="soc10" & param == "theta" & level == "mat")$value<0)/10`% of cases for SOC at 0-10 cm depth.
-They averaged `r round(mean(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value), 2)` at 0-30 cm depth, meaning that SOC stocks were estimated to decrease by `r abs(round(mean(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value), 2))` Mg C ha$^{-1}$ for an increase in mean annual temperature of one standard deviation, i.e.
-`r round(sd(data_plot$temp), 2)` °C.
+Zero was included in the 80% credibility intervals for the effect of
+mean annual temperature and clay content on AGC, meaning that no strong
+effect was detected in our data (Figure \@ref(fig:effect-agc-clay)a).
+Similarly, the clay content did not show a strong effect on SOC stocks,
+as zero was included in its 80% credible intervals (Figure
+\@ref(fig:effect-agc-clay)b,c).
+
+Mean annual temperature, on the contrary, had a stronger negative effect
+on SOC stocks at both 0-10 cm and 0-30 cm depth (Figure
+\@ref(fig:effect-agc-clay)b,c). Values of the effect of mean annual temperature $\theta_{MAT}$ were negative in
+`r sum(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value<0)/10`%
+of cases for SOC at 0-30 cm depth, and in
+`r sum(subset(data_par, stock=="soc10" & param == "theta" & level == "mat")$value<0)/10`%
+of cases for SOC at 0-10 cm depth. They averaged
+`r round(mean(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value), 2)`
+at 0-30 cm depth, meaning that SOC stocks were estimated to decrease by
+`r abs(round(mean(subset(data_par, stock=="soc" & param == "theta" & level == "mat")$value), 2))`
+Mg C ha$^{-1}$ for an increase in mean annual temperature of one
+standard deviation, i.e. `r round(sd(data_plot$temp), 2)` °C.
 
@@
 The relationship between AGC and SOC stocks was weak in our dataset (Figure \@ref(fig:plot-reg)), as the 80% credibility interval of the associated parameter ($\theta_{AGC}$) included zero (Figure \@ref(fig:effect-agc-clay)b,c; Figure \@ref(fig:plot-reg)).
+
+## Goodness of fit
+
+```{r gof}
+data_pred <- read.csv("outputs/data_par.csv") |> setDT() |> 
+  subset(grepl("pred", param))
+data_pred[, variable := ifelse(grepl("_a", param), "AGC", "SOC")]
+data_pred[, variable := ifelse(grepl("_s10", param), "SOC10", variable)]
+data_pred[, iter := paste0("V", iter)]
+data_pred[, pred := value]
+data_pred <- merge(
+  data_pred, 
+  data.frame(level_num = paste0("V", 1:30), 
+             N_PLOT = data_plot$N_PLOT)
+)
+
+df_op <- merge(data_pred[, c("N_PLOT", "variable", "iter", "pred")], 
+              data_sim[, c("N_PLOT", "variable", "iter", "value")])
+gof <- df_op[, .(
+  rmse = sqrt((sum(pred - value)^2)/length(pred)), 
+  r2b = var(pred)/(var(pred) + var(value-pred)), 
+  rho = cor(pred, value)
+), .(variable, iter)] |> 
+  melt(id.vars = c("iter", "variable"), variable.name = "meas")
+gof <- gof[, .(med = mean(value)), .(meas, variable)] |> 
+  dcast(variable ~ meas)
+```
+
+The root mean square error (median over all 1000 iterations) was `r gof[variable=="AGC", signif(rmse, 2)]` Mg C ha$^{-1}$ for AGC stands, `r gof[variable=="SOC", signif(rmse, 2)]` Mg C ha$^{-1}$ for SOC stands at 0-30 cm, and `r gof[variable=="SOC10", signif(rmse, 2)]` Mg C ha$^{-1}$ for SOC stands at 0-10 cm.
+The median Bayesian coefficient of determination [as defined by @Gelman2017] was `r gof[variable=="AGC", signif(r2b, 2)]` for AGC stocks, `r gof[variable=="SOC", signif(r2b, 2)]` for SOC stocks at 0-30 cm, and `r gof[variable=="SOC10", signif(r2b, 2)]` for SOC stocks at 0-10 cm.
 
@@
 In a meta-analysis by @Shi2018, it was estimated that AFS stored around 50 Mg C ha$^{-1}$ at a 0-30 cm depth.
-The higher SOC stocks in this study could be due to the high exchangeable Ca^2+^ content at the sites included in this study (Table \@ref(tab:site-soil)), as Ca^2+^ is frequently attributed to enhanced soil organic matter stabilization [@Wiesmeier2019].
-In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]` Mg C ha$^{-1}$) were similar to the average of 74 Mg C ha$^{-1}$ estimated by @Ma2020, although some AFS had particularly high biomass values.
-The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`, had total carbon stocks (`r dataC_afs[variable=="TOT", round(max(mean))]` Mg C ha$^{-1}$) close to those of old-growth tropical forests [270 Mg C ha$^{-1}$, @Lal2005].
+The higher SOC stocks in this study could be due to the relatively high exchangeable Ca^2+^ content at the sites included in this study compared with the majority of tropical acid soils, and the presence of carbonates (Table S\@ref(tab:site-soil)).
+Indeed, Ca^2+^ and carbonates are frequently attributed to enhanced soil organic matter stabilization [@Rowley2018; @Wiesmeier2019].
+In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`
+Mg C ha$^{-1}$) were similar to the average of 74 Mg C ha$^{-1}$
+estimated by @Ma2020, although some AFS had particularly high biomass values. The AFS with the highest carbon stocks, `r dataC_afs[variable=="TOT", afs[which.max(mean)]]`, had total carbon
+stocks (`r dataC_afs[variable=="TOT", round(max(mean))]` Mg C ha$^{-1}$)
+close to those of old-growth tropical forests [270 Mg C ha$^{-1}$,
+@Lal2005].
 The high carbon stocks in FG can be attributed to the length of time they had been established, which was generally longer than 50 years, with little human intervention (Table S\@ref(tab:afs-typology)).
@@
 Although agroforestry has been identified as one of the possible levers to increase carbon storage in the Timorese agricultural landscape [@UNFCCC2020; @Paudel2022], to our knowledge no study has quantified carbon stocks in these systems.
-Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste's climate change mitigation commitments [@UNFCCC2020] and the potential carbon gains from a wider adoption of these techniques.
+Our study could therefore help to better quantify the national contribution of agroforestry to Timor-Leste’s climate change mitigation commitments [@UNFCCC2020] and the potential carbon gains from a wider adoption of these techniques.
 Further studies should help to increase the representativeness of the AFS sampled by including a greater diversity of AFS and environmental conditions.
@@
 
-The AFS studied were found to differ significantly in terms of AGC stocks.
+The AFS studied were found to differ substantially in terms of AGC stocks, which explains their differences in total carbon stocks.
 This result reflects the density and size of trees in the plot, which in turn depend on the importance of trees in the production system.
@@
 This more limited role for trees may explain why their total biomass is lower in these systems.
-Indeed, in CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the planted annual crops to receive full sun.
-In SP systems, the growth of grazing grasses is favored by reducing the number of trees; however, the leaves and fruits of some trees (e.g.
-*Ziziphus mauritiana*, *Tamarindus indica*, *Schleichera oleiosa*) can provide additional fodder, especially during the dry season.
+Indeed, in CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the annual crops planted to receive full sun.
+In SP systems, the growth of grazing grasses is favoured by reducing the number of trees; however, the leaves and fruit of some trees (e.g.
+*Ziziphus mauritiana*, *Tamarindus
+indica*, *Schleichera oleiosa*) can provide additional fodder, especially during the dry season.
 
@@
 The AFS described in this study reflect different stages of transition in land use, which follow the growth of their trees (Figure S\@ref(fig:seq-diagram)).
-For example, a farmer might plant trees in a CF field, which will then become a YA that can still produce annual crops for the first 2 to 4 years, before the shade from the trees becomes too great.
-The trees are then used as property markers and, as they mature, produce wood and fruits; the YA then becomes an HG or FG.
+For example, a farmer might plant trees in a CF field, which will then become a YA that can still produce annual crops for the first two to four years, before the shade from the trees becomes too great.
+The trees are then used as property markers and, as they mature, produce wood and fruit; the YA then becomes an HG or FG.
 Similarly, our results show that AGC stocks increase from CF to YA, HG and FG (Table \@ref(tab:summary)).
@@
 The same applies to FG, which are former HG that have been more or less abandoned, and YA, where the farmer has chosen the site with the intention of turning it into a HG.
-We therefore expect HG, FG and YA to be located in areas that are more favorable for tree growth [@Wagner2012; @Madejon2016; @Treuer2018].
+We therefore expect HG, FG and YA to be located in areas that are more favourable for tree growth [@Wagner2012; @Madejon2016; @Treuer2018].
 Conversely, CF and SP are mainly located on limestone plateaus where water tables are deeper and soils are shallower, which could reduce tree growth and biomass [@Imada2008].
@@
 
-We found no significant difference in SOC between the AFS.
-This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017; @Wiesmeier2019], and with previous studies that show significant differences in SOC stocks between AFS [e.g., @Ramesh2015].
-However, it is worth noting that some single-site studies [e.g., @Betemariyam2020; @GamaRodrigues2010] and at least two meta-analyses failed to find any sizable effect of AFS on SOC stocks.
+We found no substantial difference in SOC between the AFS.
+This result is inconsistent with our expectation that agroforestry practices affect SOC stocks [@Minasny2017; @Wiesmeier2019] and with previous studies that show significant differences in SOC stocks between AFS [e.g.,
+@Ramesh2015].
+However, it is worth noting that some single-site studies [@Betemariyam2020; @GamaRodrigues2010] and at least two meta-analyses failed to find any sizable effect of AFS on SOC stocks.
 Indeed, @Feliciano2018 highlighted that only 3.2% of the variance in soil carbon sequestration was explained by the type of AFS, and @Shi2018 found no significant differences in SOC stocks between four different AFS in 354 previous studies.
@@
 First, in a global meta-analysis, @Ma2020 estimated that it takes only five years for SOC stocks in tropical AFS to reach an equilibrium value.
-In our study, at least three of the five AFS (namely YA, HG and FG) represent different stages of an AFS transition depending on the age of the trees (Figure S\@ref(fig:seq-diagram)).
+In our study, at least three of the five AFS (namely YA, HG and FG) represent different stages of an AFS transition depending on the age of the trees (Figure S2).
 Most of these AFS are more than five years old and may have already reached their equilibrium SOC stocks, which could explain why there was little difference between them.
@@
 Third, other factors can influence SOC stocks, sometimes to a greater extent than the type of AFS, e.g., within-site variability in land-use history.
-Our experimental design aimed to limit as much as possible the influence of external co-factors at each site (e.g., variability in farmers' management practices, micro-topography, etc.), but much of the variability cannot be controlled in farmers' plots.
+Our experimental design aimed to limit as much as possible the influence of external co-factors at each site (e.g., variability in farmers’ management practices, micro-topography, etc.), but much of the variability cannot be controlled in farmers’ plots.
 
 We found that SOC stocks were negatively correlated with mean annual temperature (Figure \@ref(fig:effect-agc-clay)b,c), reflecting the climatic and elevational gradient in the study sites (Figure S\@ref(fig:choice-clim-vars)).
-This result is consistent with that of @Raich2006, who found higher SOC stocks at higher elevations in tropical evergreen forests in different countries (e.g., Congo, Thailand), and with that of @MarinSpiotta2013, who found mean annual temperature to be the main driver of SOC stocks in successional and planted forests.
+This result is consistent with that of @Raich2006, who found higher SOC stocks at higher elevations (i.e., lower mean annual temperatures) in tropical evergreen forests in different countries (e.g., Congo, Thailand), and with that of @MarinSpiotta2013, who found mean annual temperature to be the main driver of SOC stocks in successional and planted forests.
 
@@
 After controlling for agricultural practices, and for the main pedoclimatic gradients in the study area, we still found a substantial difference in SOC between the two sites, with higher SOC stocks at Gariuai than at Osso-Luga (Figure \@ref(fig:effect-afs-site)d).
-Land-use history could be a significant driving factor: Gariuai was continuously inhabited for over 50 years, whereas Osso-Luga was abandoned for almost three decades.
+Land-use history could be a significant driving factor.
+Gariuai was continuously inhabited for over 50 years, whereas Osso-Luga was abandoned for almost three decades.
 The management and care of the AFS in Gariuai, such as pruning, organic inputs related to the presence of livestock, as well as optimization of the different strata in terms of C input, may have led to an increase in soil organic carbon compared to abandoned systems [@Lorenz2014].
@@
 
-We found no significant relationship between AGC and SOC stocks in either the topsoil layer 0-10 cm depth or at 0-30 cm depth.
+We found no relationship between AGC and SOC stocks in either the topsoil layer (0-10 cm depth) or at 0-30 cm depth.
 Overall, AGC stocks were a poor predictor of SOC stocks.
@@
 On the one hand, the addition of easily decomposable litter could stimulate microbial decomposition and turnover of old stored SOC [i.e., priming effect; @Sayer2019].
-On the other hand, the capacity of the soil to store SOC through organo-mineral associations would be limited by its fine element content (clay + fine silt; i.e., saturation concept).
+On the other hand, the capacity of the soil to store SOC through organo-mineral associations could be limited by its fine element content (clay + fine silt; i.e., saturation concept).
 
@@
 This is the case of the traditional AFS studied here in Timor-Leste.
-Their heritage and socio-economic values [as shown by @Cogne2024], in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally.
+Their heritage and socio-economic values (as shown by @Cogne2024), in addition to their high carbon stock value, confirm the importance of their conservation, both nationally and internationally.
 
@@
 However, this practice unfortunately results in a significant loss of biodiversity and carbon stored in the biomass, and can lead to an irreversible loss of fertility on the plot through soil degradation and erosion.
-This phenomenon has been documented in a similar context in Sumatra (Indonesia), where @Gouyon1993 highlighted the fact that old agroforests, which they called “jungle-rubber”, and which are quite similar in structure to “forest gardens” (except for the rubber tree component, which has been replaced by other species in Timor-Leste), were being cleared in favor of more intensive production systems.
+This phenomenon has been documented in a similar context in Sumatra (Indonesia), where @Gouyon1993 highlighted the fact that old agroforests, which they called “jungle-rubber”, and which are quite similar in structure to “forest gardens” (except for the rubber tree component, which has been replaced by other species in Timor-Leste), were being cleared in favour of more intensive production systems.
 This trend toward clearing jungle-rubber was subsequently confirmed by @Ekadinata2011, and its impact on carbon stock decline was quantified by @Villamor2014.
@@
 Strategies to avoid the clearing of traditional AFS could involve new ways of intensifying production while maintaining high carbon and biodiversity values, and gradually renewing the tree population in the plots.
-For example, it would be interesting to improve tree regeneration by clearing about one hundred square meters (corresponding to the felling of 1 to 10 trees, depending on their crown area).
+For example, it would be interesting to improve tree regeneration by clearing about one hundred square metres (corresponding to the felling of 1 to 10 trees, depending on their crown area).
 In these clearings, young trees of species considered interesting for their future production could be associated with first heliophile crops (e.g., cereals), and then shade-tolerant crops (e.g., taro).
@@
 
-Agroforestry systems in the Baucau municipality of Timor-Leste had variable but overall high carbon stocks.
-These systems have only recently been described in the scientific literature, and more research is needed to improve our knowledge of their diversity and functioning.
-It is therefore important that all stakeholders, including local authorities, universities, and NGOs, are aware of their existence, diversity, and environmental and social value, in order to propose possible innovations to maintain or improve their functioning.
+Agroforestry systems in the Baucau municipality of Timor-Leste had
+variable but overall high carbon stocks. These systems have only
+recently been described in the scientific literature, and more research
+is needed to improve our knowledge of their diversity and functioning.
+It is therefore important that all stakeholders, including local
+authorities, universities, and NGOs, are aware of their existence,
+diversity, and environmental and social value, in order to propose
+possible innovations to maintain or improve their functioning.
 
@@
 
-All authors contributed to the design of the sampling protocol.
-MC carried out the fieldwork.
-CP analyzed the data with the help of MC and VF.
-CP prepared the manuscript and all authors contributed to the drafting and revision of the manuscript.
+All authors contributed to the design of the sampling protocol. MC
+carried out the fieldwork. CP analyzed the data with the help of MC and
+VF. CP prepared the manuscript and all authors contributed to the
+drafting and revision of the manuscript.
 
@@
 
-We are grateful to Alain Rival (Cirad) for his facilitation of the field studies and critical revision of the paper, and to the trainees who helped with field data collection: Juvencio Dos Santos (vegetation) and Teofilio Gomes (soil).
-We would also like to thank the 31 farmers who gave their time to take part in the interviews and field measurements.
-A first English version of this paper was proofread by Grace Delobel.
-
-This project was co-funded by the European Union and the German Federal Ministry for Economic Cooperation and Development (BMZ) as part of the Partnership for Sustainable Agroforestry (PSAF) - Ai ba Futuru project, implemented by the Deutsche Gesellschaft für Internationale Zusammenarbeit (GIZ).
-Additional funding for this study was provided by the French Embassy in Jakarta and the French Institute of Timor-Leste.
+We are grateful to Alain Rival (Cirad) for his facilitation of the field
+studies and critical revision of the paper, and to the trainees who
+helped with field data collection: Juvencio Dos Santos (vegetation) and
+Teofilio Gomes (soil). We would also like to thank the 31 farmers who
+gave their time to take part in the interviews and field measurements. A
+first English version of this paper was proofread by Grace Delobel.
+
+This project was co-funded by the European Union and the German Federal
+Ministry for Economic Cooperation and Development (BMZ) as part of the
+Partnership for Sustainable Agroforestry (PSAF) - Ai ba Futuru project,
+implemented by the Deutsche Gesellschaft für Internationale
+Zusammenarbeit (GIZ). Additional funding for this study was provided by
+the French Embassy in Jakarta and the French Institute of Timor-Leste.
 
@@
 
-All data collected for this study is available in the following Dataverse repository \_\_\_\_\_\_.
-
-The R codes used to analyze the data and produce the manuscript (including figures and tables) are available on \_\_\_\_\_\_.
+All data collected for this study is available in the following
+Dataverse repository \_\_\_\_\_\_.
+
+The R codes used to analyze the data and produce the manuscript
+(including figures and tables) are available on \_\_\_\_\_\_.
 
@@
 
-```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (in parentheses) of total, aboveground carbon (AGC), belowground carbon (BGC), and soil organic carbon (SOC) stocks in each agroforestry system."}
+```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (in parentheses) of total, aboveground carbon (AGC), belowground carbon (BGC), and soil organic carbon (SOC; 0-30 cm) stocks in each type of agroforestry system (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden)."}
 # calculate median and 95% confidence interval at site level
-dataC_afs[, value := paste0(
+dataC_afs[, `:=`(value = paste0(
 round(mean), " (",
 round(lwr_mean), "-", 
-  round(upr_mean), ")")] |> 
+round(upr_mean), ")"), 
+afs = factor(afs, levels = c("CF", "SP", "YA", "HG", "FG")))] |> 
 subset(variable != "SOC10") |> 
 dcast(afs ~ variable, value.var = "value") |> 
+setorder() |> 
 flextable(col_keys = c("afs", "TOT", "AGC", "BGC", "SOC"), 
@@
 
+```{r tab.id = "abbrev", results = "show", tab.cap = "List of abbreviations used in the paper and their definitions."}
+data.frame(
+Abbreviation = c("AFS", "AGC", "BGC", "CBH", "CF", "FG", "HG", "SOC", "SP", "YA"), 
+Definition = c("Agroforestry system", "Aboveground carbon", "Belowground carbon",
+"Circumference at breast height", "Crop and fallow", "Forest garden",
+"Home garden", "Soil organic carbon", "Silvopastoral", 
+"Young agroforest")
+) |> 
+flextable(cwidth = c(1, 2)) 
+```
+
 <!-- ```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAs with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
@@
 Bold p-values are lower than 0.05, and the mean difference between groups is thus considered significantly different than zero."} -->
+
 <!-- anova_table$factor = factor(anova_table$factor) -->
+
 <!-- levels(anova_table$factor) = c("AFS", "site") -->
+
 <!-- flextable(anova_table, col_keys = c("variable", "factor", "pvalue"),  -->
+
 <!--           cwidth = c(1, 1, 0.8)) |>  -->
+
 <!--   bold(i = anova_table$pvalue < 0.05, j = 2:3) |>  -->
+
 <!--   set_header_labels(values = c("Dependent variable", "Independent variable", "p-value")) |>  -->
+
 <!--   merge_at(i = 1:2, j = 1) |> merge_at(i = 3:4, j = 1) |>  -->
+
 <!--   fix_border_issues() |> hline(i=2) -->
+
 <!-- ``` -->
@@
 
-```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Distribution of carbon stocks in each type of AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden), per site.
-AGC: Aboveground (biomass) carbon; BGC: belowground (biomass) carbon; SOC: soil organic carbon."}
+```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Distribution of carbon stocks in each type of AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden), per site (Gariuai, Osso-Luga).
+AGC: Aboveground (biomass) carbon; BGC: belowground (biomass) carbon; SOC: soil organic carbon (between 0-30 cm)."}
 ## reorder factors 
 dataC_afs_site[, afs := factor(afs, levels = c("CF", "SP", "YA", "HG", "FG"))]
-
-# add facet letter
-data_letter <- data.frame( 
-  site = c("Gariuai", "Osso-Luga"), 
-  lab = letters[1:2]
-)
 
@@
 geom_bar(position="stack", stat="identity", aes(fill = variable)) + 
-  geom_text(data = data_letter, x = -Inf, y = Inf, hjust = -1, vjust = 1, 
-            aes(label = lab)) +
 facet_wrap(~site, nrow = 1, strip.position = "bottom") +
@@
 
-```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3, fig.cap = "Effect of (a, b) AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden) and (c, d) site (G: Gariuai, OL: Osso-Luga) on (a, c) aboveground carbon (AGC, in Mg C ha$^{-1}$) and (b, d) soil organic carbon (SOC, in Mg C ha$^{-1}$).
+```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3, fig.cap = "Effect of (a, b) AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden) and (c, d) site (G: Gariuai, OL: Osso-Luga) on (a, c) aboveground carbon (AGC, in Mg C ha$^{-1}$) and (b, d) soil organic carbon (SOC, between 0-30 cm, in Mg C ha$^{-1}$).
 These values were obtained as the sum of the mean predicted stocks ($\\mu$) and the AFS or site random effects.
@@
 geom_point(aes(x = mean), size = 2) +
-  geom_text(data = data_letter, x = -Inf, y = Inf, hjust = -1, vjust = 1, col = 1, aes(label = lab)) +
+geom_text(data = data_letter, 
+x = -Inf, y = Inf, hjust = -1, vjust = 1, col = 1, 
+aes(label = lab)) +
 scale_x_continuous(breaks = seq(20, 100, 10)) +
@@
 
-During field data collection, because the auger length (15 cm) didn't match the target soil sampling depths (10 cm each), we could only accurately sample and analyse the 0-10 cm and 20-30 cm soil depths for SOC content.
-The SOC content of the intermediate 10-20 cm soil depth had to be estimated based on these first two measurements.
-
-To reduce the bias in this estimation, we used the standard assumption that SOC content $c(x)$ decreases as an exponential function of soil depth $x$:
-
-$$c(x) = C_0 \cdot exp(-\lambda\cdot x) $$ where $C_0$ is the carbon content at the surface (depth $x= 0$) and $\lambda$ is the rate at which the carbon content decreases with soil depth.
-Under this assumption, we can calculate the mean SOC content between depths $x$ and $x+\Delta$ as:
+During field data collection, because the auger length (15 cm) didn't
+match the target soil sampling depths (10 cm each), we could only
+accurately sample and analyse the 0-10 cm and 20-30 cm soil depths for
+SOC content. The SOC content of the intermediate 10-20 cm soil depth had
+to be estimated based on these first two measurements.
+
+To reduce the bias in this estimation, we used the standard assumption
+that SOC content $c(x)$ decreases as an exponential function of soil
+depth $x$:
+
+$$c(x) = C_0 \cdot exp(-\lambda\cdot x) $$ where $C_0$ is the carbon
+content at the surface (depth $x= 0$) and $\lambda$ is the rate at which
+the carbon content decreases with soil depth. Under this assumption, we
+can calculate the mean SOC content between depths $x$ and $x+\Delta$ as:
 
@@
 
-$$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{\lambda \cdot \Delta} exp(-\lambda\cdot x)\left[1- exp(-\lambda\cdot\Delta) \right]$$ The average SOC contents between 0-10 cm, 10-20 cm and 20-30 cm can respectively be written as:
+$$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{\lambda \cdot \Delta} exp(-\lambda\cdot x)\left[1- exp(-\lambda\cdot\Delta) \right]$$
+The average SOC contents between 0-10 cm, 10-20 cm and 20-30 cm can
+respectively be written as:
 
@@
 
-$$C_{20,30} = \frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)$$ We can thus write:
-
-$$\frac{C_{20,30}}{C_{0,10}} = \frac{\frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)}{\frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)} = exp(-20\lambda)$$ $$ \Rightarrow \lambda = \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20}$$ We can also write:
-
-$$ \frac{C_{10,20}}{C_{0,10}} = exp(-10\lambda) $$ $$\Rightarrow C_{10,20} = C_{0,10} \cdot exp(-10 \cdot \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20})$$
+$$C_{20,30} = \frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)$$
+We can thus write:
+
+$$\frac{C_{20,30}}{C_{0,10}} = \frac{\frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)}{\frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)} = exp(-20\lambda)$$
+$$ \Rightarrow \lambda = \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20}$$ We
+can also write:
+
+$$ \frac{C_{10,20}}{C_{0,10}} = exp(-10\lambda) $$
+$$\Rightarrow C_{10,20} = C_{0,10} \cdot exp(-10 \cdot \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20})$$
 
 $$ \Rightarrow C_{10,20} = C_{0,10} \cdot \left( \frac{C_{20,30}}{C_{0,10}}\right)^{\frac{1}{2}}$$
-### Choice of climate and soil covariates
-
-As described in the "Carbon stocks model" section of the Materials and Methods, we hypothesized that C stocks (both AGC and SOC) are affected by climate and soil characteristics, together with land use (in this case, AFS type) and other spatial differences (in our case, a site effect).
+\### Choice of climate and soil covariates
+
+As described in the "Carbon stocks model" section of the Materials and
+Methods, we hypothesized that C stocks (both AGC and SOC) are affected
+by climate and soil characteristics, together with land use (in this
+case, AFS type) and other spatial differences (in our case, a site
+effect).
 
@@
 
-We selected the following set of variables to represent the effect of climate: mean annual temperature (BIO1), mean annual precipitation (BIO12), annual precipitation variability (BIO15) over the period 1981-2010.
-These bioclimatic variables were extracted from CHELSA V2.1 [@Karger2017].
-These variables were all highly correlated in our dataset (Figure S\@ref(fig:choice-clim-vars)a).
-We performed a principal component analysis after scaling and centering these three variables: the first component explained `r round(pca_clim$sdev[1]^2/sum(pca_clim$sdev^2)*100)`% of the variance in our dataset (Figure S\@ref(fig:choice-clim-vars)b).
-We therefore decided to keep only one variable to represent this climatic gradient: mean annual temperature, which was most correlated with the first component of the PCA (Figure S\@ref(fig:choice-clim-vars)b), and has been shown to influence C stocks in other contexts [@Wiesmeier2019;@Muller-Landau2021].
-
-We selected the following set of variables to represent the effect of soil properties on C stocks: clay content (<2 $\mu$m), silt content (2-50 $\mu$m), sand content (50-2000 $\mu$m), and pH.
-These variables were measured on our soil samples analysed in the Cirad laboratories (see section "Field inventories").
-After scaling and centering these variables, we performed a principal component analysis: the first component explained `r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance in our dataset (Figure S\@ref(fig:choice-soil-vars)b).
-We therefore decided to retain only one variable to represent this edaphic gradient: clay content, which was the most highly correlated with the first component of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been shown to be a good predictor of SOC stocks in other contexts [@Wiesmeier2019].
+We selected the following set of variables to represent the effect of
+climate: mean annual temperature (BIO1), mean annual precipitation
+(BIO12), annual precipitation variability (BIO15) over the period
+1981-2010. These bioclimatic variables were extracted from CHELSA V2.1
+[@Karger2017]. These variables were all highly correlated in our dataset
+(Figure S\@ref(fig:choice-clim-vars)a). We performed a principal
+component analysis after scaling and centering these three variables:
+the first component explained
+`r round(pca_clim$sdev[1]^2/sum(pca_clim$sdev^2)*100)`% of the variance
+in our dataset (Figure S\@ref(fig:choice-clim-vars)b). We therefore
+decided to keep only one variable to represent this climatic gradient:
+mean annual temperature, which was most correlated with the first
+component of the PCA (Figure S\@ref(fig:choice-clim-vars)b), and has
+been shown to influence C stocks in other contexts [@Wiesmeier2019;
+@Muller-Landau2021].
+
+We selected the following set of variables to represent the effect of
+soil properties on C stocks: clay content (\<2 $\mu$m), silt content
+(2-50 $\mu$m), sand content (50-2000 $\mu$m), and pH. These variables
+were measured on our soil samples analysed in the Cirad laboratories
+(see section "Field inventories"). After scaling and centering these
+variables, we performed a principal component analysis: the first
+component explained
+`r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance
+in our dataset (Figure S\@ref(fig:choice-soil-vars)b). We therefore
+decided to retain only one variable to represent this edaphic gradient:
+clay content, which was the most highly correlated with the first
+component of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been
+shown to be a good predictor of SOC stocks in other contexts
+[@Wiesmeier2019].
 
@@
 # choose variables to show
-vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n")
+vars = c("Corg\n", "N", "Corg_N", "TC", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n")
 
@@
 # convert sand, silt, clay and C content from % into g/kg
-data_soil_site_simu[ variable %in% c("Corg", "CLAY", "sand", "silt"), value := value*10]
+data_soil_site_simu[ variable %in% c("Corg", "CLAY", "sand", "silt", "TC"), value := value*10]
 
@@
 ## add unit row
-data_soil_site = data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "CEC", "Ca", "K", "Mg", "Na")]
-colnames(data_soil_site) = c(" ", "Clay", "Silt", "Sand", "C content", "N content", "C:N", "pH", "CEC", "Ca", "K", "Mg", "Na")
+data_soil_site = data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "TC", "pH", "CEC", "Ca", "K", "Mg", "Na")]
+colnames(data_soil_site) = c(" ", "Clay", "Silt", "Sand", "Organic C content", "Total N content", "C:N", "Carbonate content", "pH", "CEC", "Ca", "K", "Mg", "Na")
 ```
@@
 ```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (average over 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
-Units are: in g/kg for clay, silt, sand, C and N contents; in cmolc/kg for CEC, Ca, K, Mg and Na; and no units for C:N and pH.
+Units are: in g/kg for clay, silt, sand, carbonate, organic C and total N contents; in cmolc/kg for CEC, Ca, K, Mg and Na; and no units for C:N and pH.
 The confidence intervals were calculated by propagating a 12% error on the lab-provided values and obtaining site-level estimates through bootstrapping plot values (1000 simulations).
-Total organic carbon content (dry combustion method) ; total nitrogen content (dry combustion method) ; pH in water (soil:deionized water ratio of 1:2.5) ; CEC - cation exchangeable capacity - and exchangeable cations (Ca, K, Mg and Na) at soil pH [Cobaltihexamine method; Orsini and Remy, 1976]."}
+Organic carbon content (difference of the total carbon – dry combustion method – and the carbonates contents - calcimeter) ; total nitrogen content (dry combustion method) ; pH in water (soil:deionized water ratio of 1:2.5) ; CEC - cation exchangeable capacity - and exchangeable cations (Ca, K, Mg and Na) at soil pH [Cobaltihexamine method; Orsini and Remy, 1976]."}
 flextable(data_soil_site, cwidth = 0.5)
@@
 ```
+
 </br>
@@
 
-</br>
-
-```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesized effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
-# save_png <- function(plot, path){
-#   DiagrammeRsvg::export_svg(plot) %>%
-#     charToRaw() %>%
-#     rsvg::rsvg() %>%
-#     png::writePNG(path)
-# }
-# 
-# DiagrammeR::grViz("
-#       digraph {
-#       graph [layout = dot, rankdir = LR]
-# 
-#       node [shape = ellipse]
-#       AGC;
-#       SOC;
-#       node [shape = rectangle]
-#       climate;
-#       soil;
-#       AFS;
-# 
-#       # Connect nodes with edges and labels
-#       AFS -> AGC;
-#       climate -> AGC;
-#       AFS -> SOC;
-#       soil -> AGC;
-#       soil -> SOC;
-#       climate -> SOC;
-#       edge[style = 'dashed']
-#       AGC -> SOC;
-#        }
-# 
-#       ") %>%
-#   save_png("figures/dag.png")
-include_graphics("figures/dag.png") 
-```
+<!-- </br> -->
+
+<!-- ```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesized effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"} -->
+
+<!-- # save_png <- function(plot, path){ -->
+
+<!-- #   DiagrammeRsvg::export_svg(plot) %>% -->
+
+<!-- #     charToRaw() %>% -->
+
+<!-- #     rsvg::rsvg() %>% -->
+
+<!-- #     png::writePNG(path) -->
+
+<!-- # } -->
+
+<!-- #  -->
+
+<!-- # DiagrammeR::grViz(" -->
+
+<!-- #       digraph { -->
+
+<!-- #       graph [layout = dot, rankdir = LR] -->
+
+<!-- #  -->
+
+<!-- #       node [shape = ellipse] -->
+
+<!-- #       AGC; -->
+
+<!-- #       SOC; -->
+
+<!-- #       node [shape = rectangle] -->
+
+<!-- #       climate; -->
+
+<!-- #       soil; -->
+
+<!-- #       AFS; -->
+
+<!-- #  -->
+
+<!-- #       # Connect nodes with edges and labels -->
+
+<!-- #       AFS -> AGC; -->
+
+<!-- #       climate -> AGC; -->
+
+<!-- #       AFS -> SOC; -->
+
+<!-- #       soil -> AGC; -->
+
+<!-- #       soil -> SOC; -->
+
+<!-- #       climate -> SOC; -->
+
+<!-- #       edge[style = 'dashed'] -->
+
+<!-- #       AGC -> SOC; -->
+
+<!-- #        } -->
+
+<!-- #  -->
+
+<!-- #       ") %>% -->
+
+<!-- #   save_png("figures/dag.png") -->
+
+<!-- include_graphics("figures/dag.png")  -->
+
+<!-- ``` -->
 
@@
 <!-- ```{r dag-code} -->
+
 <!-- # dag { -->
+
 <!-- # bb="-2.799,-2.922,2.216,3.074" -->
+
 <!-- # afs [exposure,pos="0.265,-1.155"] -->
+
 <!-- # agc [exposure,pos="-0.591,0.840"] -->
+
 <!-- # climate [adjusted,pos="-1.049,-1.155"] -->
+
 <!-- # soc [outcome,pos="0.137,1.233"] -->
+
 <!-- # soil [adjusted,pos="-0.426,-1.182"] -->
+
 <!-- # afs -> agc -->
+
 <!-- # afs -> soc -->
+
 <!-- # agc -> soc -->
+
 <!-- # climate -> agc -->
+
 <!-- # climate -> soc -->
+
 <!-- # soil -> agc -->
+
 <!-- # soil -> soc -->
+
 <!-- # } -->
+
 <!-- ``` -->

### 2025-04-29 manuscript.Rmd
minor edits requested by AFS

@@
 In this study, we quantified carbon storage in both vegetation biomass and soil in five traditional agroforestry systems (AFS) in Timor-Leste, namely cropping systems with fallow (CF), silvopastures (SP), young agroforests (YA), home gardens (HG), and forest gardens (FG).
-Our results show that these traditional AFS can store large amounts of carbon, with the average being 155\ Mg\ C\ ha$^{-1}$.
+Our results show that these traditional AFS can store large amounts of carbon, with the average total carbon stocks (soil and biomass) being 155\ Mg\ C\ ha$^{-1}$.
 The AFS with the highest total carbon stocks (FG) stored an average of 213\ Mg\ C\ ha$^{-1}$.
@@
 csl: misc/springer-basic-author-date.csl
-keywords: "Traditional agroforestry, carbon stocks, soil, biomass"
 editor_options: 
@@
 
+**Keywords**: Vegetation biomass, soil organic carbon, farming systems, climate change mitigation, South East Asia.
+
 ```{r setup, include=FALSE}
-# Run with R version 4.3.2
-
-packages_needed = c(
-  "ggplot2",      # version 3.4.4
-  "ggrepel",      # version 0.9.4
-  "ggspatial",    # version 1.1.9
-  "ggpubr",       # version 0.6.0
-  "ggfortify",    # version 0.4.17
-  "maps",         # version 3.4.2
-  "rdryad",       # version 1.0.0
-  "readxl",       # version 1.4.3
-  "data.table",   # version 1.14.8
-  "knitr",        # version 1.45
-  "BIOMASS",      # version 2.1.11
-  "terra",        # version 1.7-55
-  "car",          # version 3.1-2
-  "lmerTest",     # version 3.1-3
-  "lme4",         # version 1.1-35.1
-  "Matrix",       # version 1.6
-  "performance",  # version 0.10.8
-  "flextable"     # version 0.9.5
-)
-
-packages_to_install = packages_needed[!( packages_needed %in% rownames(installed.packages()))]
-
-if (length(packages_to_install) > 0)
-  install.packages(packages_to_install)
+# Run with R version 4.4.1
+# package management is done with renv
+packages_needed <- c(
+  "ggplot2", "ggrepel",  "ggspatial",  "ggpubr",  "ggfortify",  "maps", 
+  "rdryad", "readxl", "data.table", "knitr", "BIOMASS", "terra", "car", 
+  "flextable"
+)
 
@@
 
-opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide")
+opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = "hide")
 ```
@@
 
-Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation[@Albrecht2003; @Verchot2007; @Lasco2014; @Cardinael2021].
+Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Albrecht2003; @Verchot2007; @Lasco2014; @Cardinael2021].
 Agroforestry is the collective name for land-use systems and technologies where woody perennials (trees, shrubs, bamboos, creepers, etc.) are deliberately used on the same land management units as agricultural crops and/or animals in some form of spatial arrangement or temporal sequence [@Lundgren1982].
@@
 However, the effect of AFS on soil organic carbon (SOC) is more variable.
-Although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e.
-planting trees with intercropping during the establishment phase) may decrease SOC stocks, possibly due to the initial soil disturbance caused by tree planting [@Feliciano2018].
+Although AFS appear to increase SOC stocks overall in tropical regions, some AFS such as woodlots (i.e., planting trees with intercropping during the establishment phase) may decrease SOC stocks, possibly due to the initial soil disturbance caused by tree planting [@Feliciano2018].
 Conversely, AFS such as alley cropping and home gardens have been shown to increase SOC stocks [@Oelbermann2004; @Shi2018], while other AFS such as silvopastures have shown mixed results [@Feliciano2018; @Shi2018].
@@
 AFS have been presented as a potential solution to many of the challenges facing Timorese agriculture [@Paudel2022; @Cogne2024].
-Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022; @Cogne2024] and can be key elements for biodiversity conservation [@Perfecto2008; @Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
+Although trees already form a large part of traditional agricultural systems in Timor-Leste [@Paudel2022; @Cogne2024; @Gusmao2025] and can be key elements for biodiversity conservation [@Perfecto2008; @Torquebiau1992], they have received little attention and, to our knowledge, no study has quantified carbon storage in traditional Timorese AFS.
 A precise quantification of these carbon stocks will improve understanding of the ecological importance of AFS at both local and global scales, and help integrate them into national and international climate change mitigation initiatives.
@@
   ## crop to reduce memory usage
-  box = c(124.5, 127.5, -9.2, -8.3) # Timor-Leste
+  box <- c(124.5, 127.5, -9.2, -8.3) # Timor-Leste
   rst <- terra::rast(destfile)
@@
 data_plot[, `:=`(
-  elev = terra::extract(terra::rast("data/srtm_62_14.tif"), 
-                        cbind(long, lat))[[1]], 
-  temp = terra::extract(terra::rast("data/CHELSA_bio1_1981-2010_V.2.1.tif"), 
-                        cbind(long, lat))[[1]], 
-  prec = terra::extract(terra::rast("data/CHELSA_bio12_1981-2010_V.2.1.tif"), 
-                        cbind(long, lat))[[1]], 
-  varp = terra::extract(terra::rast("data/CHELSA_bio15_1981-2010_V.2.1.tif"), 
-                        cbind(long, lat))[[1]], 
-  cwd = terra::extract(terra::rast("data/CWD.tif"), 
-                       cbind(long, lat))[[1]]
+  elev = terra::extract(
+    terra::rast("data/srtm_62_14.tif"),
+    cbind(long, lat)
+  )[[1]],
+  temp = terra::extract(
+    terra::rast("data/CHELSA_bio1_1981-2010_V.2.1.tif"),
+    cbind(long, lat)
+  )[[1]],
+  prec = terra::extract(
+    terra::rast("data/CHELSA_bio12_1981-2010_V.2.1.tif"),
+    cbind(long, lat)
+  )[[1]],
+  varp = terra::extract(
+    terra::rast("data/CHELSA_bio15_1981-2010_V.2.1.tif"),
+    cbind(long, lat)
+  )[[1]],
+  cwd = terra::extract(
+    terra::rast("data/CWD.tif"),
+    cbind(long, lat)
+  )[[1]]
 )]
@@
 # extract elevation and rainfall for the 2 sites
-site_env = data_plot[, as.list(expand.grid(seq(min(long), max(long), 0.001), 
-                                           seq(min(lat), max(lat), 0.001))),
-                     .(site)]
-
-colnames(site_env) = c("site", "long", "lat")
+site_env <- data_plot[
+  , as.list(expand.grid(
+    seq(min(long), max(long), 0.001),
+    seq(min(lat), max(lat), 0.001)
+  )),
+  .(site)
+]
+
+colnames(site_env) <- c("site", "long", "lat")
 
@@
 
-site_summary = site_env[, .(
-  coord = paste0(abs(round(mean(lat),2)), "°S, ", round(mean(long),2), "°E"), 
+site_summary <- site_env[
+  , .(
+    coord = paste0(abs(round(mean(lat), 2)), "°S, ", round(mean(long), 2), "°E"),
     prec = round(mean(prec)),
-  elev = paste(round(quantile(elev, c(0.025, 0.975))/10)*10, collapse = " and ")),
-  .(site)]
+    elev = paste(round(quantile(elev, c(0.025, 0.975)) / 10) * 10, collapse = " and ")
+  ),
+  .(site)
+]
 ```
@@
 Due to the risk of flooding, villages are rarely located near rivers, and are concentrated instead near paved roads and on hilltops.
+
 ## Typology of agroforestry systems
@@
 
-Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured.
+Within each sampling plot, all trees with a circumference at breast height greater than or equal to 100 cm were identified taxonomically and their circumference at breast height and height were measured.
 Five subplots of 10x10 m (0.01 ha) were established at each corner and in the centre of the sampling plot (Figure S\@ref(fig:design)).
-Within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured.
+Within these subplots, all trees with a circumference at breast height between 30 and 100 cm were taxonomically identified and their circumference at breast height and height were measured.
 Height measurements were taken with a dendrometer.
@@
 # remove NA or <30 cm circumference values or < 100 cm outside subplots
-data_tree <- subset(data_tree, 
+data_tree <- subset(
+  data_tree,
   !is.na(DIAM_130_CM) &
-                      DIAM_130_CM >= 30/pi &
-                      (DIAM_130_CM >= 100/pi | grepl("Q", N_Q_ID)))
+    DIAM_130_CM >= 30 / pi &
+    (DIAM_130_CM >= 100 / pi | grepl("Q", N_Q_ID))
+)
 
@@
 # get carbon content
-GLOWCAD_files = dryad_download("10.5061/dryad.18931zcxk")
-GLOWCAD = read.csv(GLOWCAD_files[[1]][3]) |> 
+GLOWCAD_files <- dryad_download("10.5061/dryad.18931zcxk")
+GLOWCAD <- read.csv(GLOWCAD_files[[1]][3]) |>
   subset(grepl("tree", growth.form)) |>
@@
 data_ccontent <-
-
   # estimate mean AGC
@@
     H = data_tree$HEIGHT,
-    errH = data_tree$HEIGHT*cvH, 
+    errH = data_tree$HEIGHT * cvH,
     Dpropag = "chave2004",
@@
 ## height in cm, with a conversion factor between 0.8 and 1
-data_sim_agc[GENUS %in% palm_genera, 
+data_sim_agc[
+  GENUS %in% palm_genera,
   HEIGHT := ifelse(
@@
     HEIGHT, ## height in m
-               rnorm(length(HEIGHT), HEIGHT, HEIGHT*cvH)
-             )]
+    rnorm(length(HEIGHT), HEIGHT, HEIGHT * cvH)
+  )
+]
 
 ## add WD data
-data_wd_palms = unique(data_wd[data_wd$genus %in% palm_genera, c("genus", "meanWD", "sdWD")])
-data_sim_agc = merge(
+data_wd_palms <- unique(data_wd[data_wd$genus %in% palm_genera, c("genus", "meanWD", "sdWD")])
+data_sim_agc <- merge(
   data_sim_agc,
@@
 )
-data_sim_agc[GENUS %in% palm_genera, 
+data_sim_agc[
+  GENUS %in% palm_genera,
   WD := ifelse(iter == "mean",
     meanWD,
-                          rnorm(length(meanWD), meanWD,sdWD))]
+    rnorm(length(meanWD), meanWD, sdWD)
+  )
+]
 # palm biomass with Goodman et al 2013 equation that includes diameter, height
@@
   GENUS %in% palm_genera,
-  AGC := ((0.463 * DIAM_130_CM^2 * HEIGHT)^0.25 * 0.55512) ^ 4	* 0.4713 *1e-3]
-```
-
-We estimated tree-level AGC stocks from the tree CBH and height measurements using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017].
+  AGC := ((0.463 * DIAM_130_CM^2 * HEIGHT)^0.25 * 0.55512)^4 * 0.4713 * 1e-3
+]
+```
+
+We estimated tree-level AGC stocks from the tree circumference at breast height and height measurements using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017].
 Each tree was assigned a mean and standard deviation of wood density at the species or genus level when at least one value was available from the Global Wood Density Database [@Zanne2009].
@@
 
-data_sim_agc[, RS := exp(- 1.2312 - 0.0215 * DIAM_130_CM + 
+data_sim_agc[, RS := exp(-1.2312 - 0.0215 * DIAM_130_CM +
   0.0002 * DIAM_130_CM^2 - 0.0007 * cwd)]
@@
 ## bootstrap AGC and BGC values in small plots (except for iter = mean)
-data_sim = 
+data_sim <-
   merge(
@@
           mean(sample(c(value, rep(0, 5 - length(value))), replace = TRUE))
-        )), .(N_PLOT, variable, iter)], 
-      data_sim[grepl("Q", N_Q_ID) & iter == "mean", 
-               .(valueSmall = sum(value / 5)), .(N_PLOT, variable, iter)]
+        )
+      ), .(N_PLOT, variable, iter)],
+      data_sim[
+        grepl("Q", N_Q_ID) & iter == "mean",
+        .(valueSmall = sum(value / 5)), .(N_PLOT, variable, iter)
+      ]
     ),
@@
 # use uniform distribution between zero and limit of detectability
-data_sim_empty = expand.grid(N_PLOT = setdiff(data_plot$N_PLOT, data_sim$N_PLOT), 
+data_sim_empty <- expand.grid(
+  N_PLOT = setdiff(data_plot$N_PLOT, data_sim$N_PLOT),
   iter = unique(data_sim$iter),
-                             variable = c("AGC", "BGC"))|> 
-  setDT() |> merge(data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")]) 
-data_sim_empty$value = 0
+  variable = c("AGC", "BGC")
+) |>
+  setDT() |>
+  merge(data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])
+data_sim_empty$value <- 0
 
@@
 
-data_sim = rbind(data_sim[, colnames(data_sim_empty), with = FALSE], data_sim_empty)
+data_sim <- rbind(data_sim[, colnames(data_sim_empty), with = FALSE], data_sim_empty)
 ```
@@
 \begin{equation}
-SOC_p = \sum_i Corg_{i,p} \cdot BDfine_{i,p}\cdot (1- \frac{Vcoarse_{i,p}}{Vcyl})\cdot (Surf \cdot \Delta d) \cdot 10^{-6}
+SOC_p = \sum_i Corg_{i,p} \cdot BDfine_{i,p}\cdot (1- \frac{Vcoarse_{i,p}}{Vcyl})\cdot (Surf \cdot \Delta d) \cdot 10^{-6} (\#eq:socbd)
 \end{equation}
 
-Where $ Corg_{i,p}$ is the SOC content at depth $i$ (either 0-10, 10-20 or 20-30 cm) as measured in the laboratory from the soil sampled by the auger (depths 0-10 cm and 20-30 cm) and estimated at depth 10-20 cm (see details below).
+Where $Corg_{i,p}$ is the SOC content at depth $i$ (either 0-10, 10-20 or 20-30 cm) as measured in the laboratory from the soil sampled by the auger (depths 0-10 cm and 20-30 cm) and estimated at depth 10-20 cm (see details below).
 $BDfine_{i,p}$ is the bulk density of the fine fraction of the soil sampled with the cylinder at depth $i$ (in g cm$^{-3}$):
@@
 $$BDfine_{i,p} = \frac{mfine_{i,p}}{Vcyl – Vcoarse_{i,p}}$$
-$mfine_{i,p}$ is the mass of the fine fraction (\< 2 mm) of the sampled soil; Vcyl is the volume of the sampling cylinder (250 cm$^3$); and $Vcoarse_{i,p}$ is the volume of the coarse fraction in the sampling cylinder.
-$Surf = 10^8$ is the surface conversion factor from cm$^2$ to ha; $\Delta d$ is the difference in sample depths (= 10 cm); 10^{-6}$ is the conversion factor from g to Mg.
+$mfine_{i,p}$ is the mass of the fine fraction (\< 2 mm) of the sampled soil; $Vcyl$ is the volume of the sampling cylinder (250 cm$^3$); and $Vcoarse_{i,p}$ is the volume of the coarse fraction in the sampling cylinder.
+$Surf = 10^8$ is the surface conversion factor from cm$^2$ to ha; $\Delta d$ is the difference in sample depths (= 10 cm); 10$^{-6}$ is the conversion factor from g to Mg.
 
@@
 Corg_{10-20,p}=Corg_{0-10,p}\cdot\sqrt{\frac{Corg_{20-30,p}}{Corg_{0-10,p}}}
-(\#eq:soc_depth)
+(\#eq:socdepth)
 \end{equation}
@@
 where $Corg_{a-b,p}$ is the average SOC content in plot $p$ between depths $a$ and $b$
-(see rationale for equation \@ref(eq:soc_depth) in Supplemental Information).
+(see rationale for equation \@ref(eq:socdepth) in Supplemental Information).
 
@@
 
-data_soil <- merge(data_soil_lab, data_soil_field, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
+data_soil <- merge(data_soil_lab, data_soil_field, by = c("N_PLOT", "SAMPLE_DEPTH"), all = TRUE)
 setDT(data_soil)
@@
 # organic carbon content that decreases as a negative exponential of depth, we can deduce the carbon content at 10-20 cm with the following function: SOC[10-20] = SOC[0-10]*sqrt(SOC[20-30]/SOC[0-10])
-data_soil = merge(
+data_soil <- merge(
   data_soil,
-  data_soil[, .(
-    Corg1020 = Corg[SAMPLE_DEPTH=="0-10"]*
-      sqrt(Corg[SAMPLE_DEPTH=="20-30"]/Corg[SAMPLE_DEPTH=="0-10"])), 
-    .(N_PLOT)])
-
-data_soil[SAMPLE_DEPTH=="10-20", Corg := Corg1020]
+  data_soil[
+    , .(
+      Corg1020 = Corg[SAMPLE_DEPTH == "0-10"] *
+        sqrt(Corg[SAMPLE_DEPTH == "20-30"] / Corg[SAMPLE_DEPTH == "0-10"])
+    ),
+    .(N_PLOT)
+  ]
+)
+
+data_soil[SAMPLE_DEPTH == "10-20", Corg := Corg1020]
 ```
@@
 ## error propagation using a Monte Carlo method -------------
-data_sim_soil = data_soil[, .(mfine = c(Sample_DRY, rnorm(1000, Sample_DRY, Sample_DRY*0.1)), # 10% error on fine soil mass
-                              Corg = c(Corg, rnorm(1000, Corg, Corg*0.12)), # 12% error on C content measurement
-                              iter = c("mean", paste0("V", 1:1000))), 
-                          .(N_PLOT, SAMPLE_DEPTH, h1, h0)]
-
-data_sim_soil[, Cstock :=             ## in Mg/ha
+data_sim_soil <- data_soil[
+  , .(
+    mfine = c(Sample_DRY, rnorm(1000, Sample_DRY, Sample_DRY * 0.1)), # 10% error on fine soil mass
+    Corg = c(Corg, rnorm(1000, Corg, Corg * 0.12)), # 12% error on C content measurement
+    iter = c("mean", paste0("V", 1:1000))
+  ),
+  .(N_PLOT, SAMPLE_DEPTH, h1, h0)
+]
+
+data_sim_soil[
+  , Cstock := ## in Mg/ha
     (h1 - h0) * 1e8 * ## equivalent volume of a 1 ha soil sample, in cm3/ha
-                Corg/100 *            ## carbon organic content in fine soil (unitless)
+      Corg / 100 * ## carbon organic content in fine soil (unitless)
       (mfine * 1e-6) * ## mass of the fine fraction of the soil (Mg)
-                (1/250)        ## 1/volume of the cylinder, in cm-3
+      (1 / 250) ## 1/volume of the cylinder, in cm-3
 ]
@@
 # sum over depths + at 0-10 cm depth
-data_sim_soil = data_sim_soil[, .(
+data_sim_soil <- data_sim_soil[
+  , .(
     SOC = sum(Cstock),
-  SOC10 = Cstock[SAMPLE_DEPTH == "0-10"]),
+    SOC10 = Cstock[SAMPLE_DEPTH == "0-10"]
+  ),
   .(N_PLOT, iter)
@@
 ## melt data
-data_sim_soil = melt(data_sim_soil, id.vars = c("N_PLOT", "iter"))
+data_sim_soil <- melt(data_sim_soil, id.vars = c("N_PLOT", "iter"))
 ```
@@
 ## merge with biomass data
-data_sim = rbind(data_sim[, colnames(data_sim_soil), with = FALSE], data_sim_soil) |> 
+data_sim <- rbind(data_sim[, colnames(data_sim_soil), with = FALSE], data_sim_soil) |>
   merge(data_plot[, c("N_PLOT", "site", "ID_AF")], by = "N_PLOT")
@@
 # add total carbon stocks
-data_sim = rbind(data_sim, data_sim[, .(value = sum(value[variable != "SOC10"]), variable = "TOT"), .(N_PLOT, site, ID_AF, iter)])
+data_sim <- rbind(data_sim, data_sim[, .(value = sum(value[variable != "SOC10"]), variable = "TOT"), .(N_PLOT, site, ID_AF, iter)])
 ```
@@
   subset(!is.na(TC)) |>
-  melt(id.vars = c("N_PLOT", "SAMPLE_DEPTH"), 
-       measure.vars= c("CLAY", "FINE SILT", "pH\n", "Ca", "CEC\n"))  |> 
+  melt(
+    id.vars = c("N_PLOT", "SAMPLE_DEPTH"),
+    measure.vars = c("CLAY", "FINE SILT", "pH\n", "Ca", "CEC\n")
+  ) |>
   setDT()
@@
   for (n in setdiff(seq_len(100), pars_prev)) {
-    data_par <- lapply(paste0("V", seq((n-1)*10+1, n*10)), function(i) {
-
+    data_par <- lapply(paste0("V", seq((n - 1) * 10 + 1, n * 10)), function(i) {
       data_stan <- data_sim |>
-        subset(iter==i) |> 
+        subset(iter == i) |>
         dcast(N_PLOT + site + ID_AF + CLAY + temp ~ variable) |>
@@
             soc10 = SOC10
-          ))
-
-      fit <- stan(model_code = stan_model, data = data_stan, 
-                  iter = 60, warmup = 50, chains = 1)
+          )
+        )
+
+      fit <- stan(
+        model_code = stan_model, data = data_stan,
+        iter = 60, warmup = 50, chains = 1
+      )
       pars <- extract(fit) |>
@@
         rbindlist(idcol = "param")
-      return(pars[iter_stan==sample(iter_stan, 1), -"iter_stan"])
+      return(pars[iter_stan == sample(iter_stan, 1), -"iter_stan"])
     }) |> rbindlist(idcol = "iter")
@@
     write.csv(data_par,
-              file = paste0("outputs/data_par_", n, ".csv"), row.names = FALSE)
-
+      file = paste0("outputs/data_par_", n, ".csv"), row.names = FALSE
+    )
   }
@@
   data_par <- list.files("outputs", pattern = "data_par_", full.names = TRUE) |>
-    lapply(read.csv) |> rbindlist(idcol = "iter10") 
-  data_par[, `:=`(iter = (iter10-1)*10+iter, iter10 = NULL)]
+    lapply(read.csv) |>
+    rbindlist(idcol = "iter10")
+  data_par[, `:=`(iter = (iter10 - 1) * 10 + iter, iter10 = NULL)]
 
@@
 
-data_par <- read.csv("outputs/data_par.csv") |> setDT() |> 
+data_par <- read.csv("outputs/data_par.csv") |>
+  setDT() |>
   subset(!grepl("pred", param))
@@
 data_par[, effect := tstrsplit(param, "_")[[2]]]
-data_par[! effect %in% c("afs", "site"), effect := NA]
+data_par[!effect %in% c("afs", "site"), effect := NA]
 data_par[, stock := tstrsplit(param, "_")[[3]]]
@@
 data_par[, level_num := as.numeric(gsub("V", "", level_num))]
-data_par[effect == "afs" & !is.na(level_num), 
-         level := levels(as.factor(data_sim$ID_AF))[level_num]]
-data_par[effect == "site" & !is.na(level_num),
-         level := levels(as.factor(data_sim$site))[level_num]]
-data_par[param == "theta",
-         level := c("mat", "clay", "agc")[level_num]]
+data_par[
+  effect == "afs" & !is.na(level_num),
+  level := levels(as.factor(data_sim$ID_AF))[level_num]
+]
+data_par[
+  effect == "site" & !is.na(level_num),
+  level := levels(as.factor(data_sim$site))[level_num]
+]
+data_par[
+  param == "theta",
+  level := c("mat", "clay", "agc")[level_num]
+]
 
@@
 
-itermaxlp = data_par[param == "lp", iter[which.max(value)]]
+itermaxlp <- data_par[param == "lp", iter[which.max(value)]]
 pars_summary <- data_par[, .(
   mean = mean(value),
-  maxlp = value[iter==itermaxlp],
+  maxlp = value[iter == itermaxlp],
   low95 = quantile(value, 0.025),
@@
 # plot level statistics
-dataC_plot = data_sim[, .(
-  mean = value[iter=="mean"],
+dataC_plot <- data_sim[, .(
+  mean = value[iter == "mean"],
   lwr = quantile(value, 0.025),
-  upr= quantile(value, 0.975)
+  upr = quantile(value, 0.975)
 ), .(plot = N_PLOT, site, afs = ID_AF, variable)]
@@
 # AFS level statistics (with bootstrapping)
-dataC_afs = data_sim[, .(
+dataC_afs <- data_sim[, .(
   mean = mean(value),
+  sd = sd(value),
   meanBs = mean(sample(value, replace = TRUE))
 ), .(afs = ID_AF, variable, iter)][, .(
-  mean = mean[iter=="mean"],        
+  mean = mean[iter == "mean"],
   lwr_mean = quantile(meanBs, 0.025),
-  upr_mean = quantile(meanBs, 0.975)
+  upr_mean = quantile(meanBs, 0.975), 
+  se = sd[iter == "mean"]/sqrt(sum(iter == "mean"))
 ), .(afs, variable)]
@@
 # site level statistics (with bootstrapping)
-dataC_site = data_sim[, .(
+dataC_site <- data_sim[, .(
   mean = mean(value),
@@
 ), .(site, variable, iter)][, .(
-  mean = mean[iter=="mean"],    
+  mean = mean[iter == "mean"],
   lwr_mean = quantile(meanBs, 0.025),
@@
 # AFS + site level statistics (with bootstrapping)
-dataC_afs_site = data_sim[, .(
+dataC_afs_site <- data_sim[, .(
   mean = mean(value),
+  sd = sd(value),
   meanBs = mean(sample(value, replace = TRUE))
 ), .(afs = ID_AF, site, variable, iter)][, .(
-  mean = mean[iter=="mean"],        
+  mean = mean[iter == "mean"],
   lwr_mean = quantile(meanBs, 0.025),
-  upr_mean = quantile(meanBs, 0.975)
+  upr_mean = quantile(meanBs, 0.975), 
+  se = sd[iter == "mean"]/sqrt(sum(iter == "mean"))
 ), .(afs, site, variable)]
@@
 Similarly, the clay content did not show a strong effect on SOC stocks,
-as zero was included in its 80% credible intervals (Figure
+as zero was included in its 80% credibility intervals (Figure
 \@ref(fig:effect-agc-clay)b,c).
@@
 ```{r gof}
-data_pred <- read.csv("outputs/data_par.csv") |> setDT() |> 
+data_pred <- read.csv("outputs/data_par.csv") |>
+  setDT() |>
   subset(grepl("pred", param))
@@
   data_pred,
-  data.frame(level_num = paste0("V", 1:30), 
-             N_PLOT = data_plot$N_PLOT)
-)
-
-df_op <- merge(data_pred[, c("N_PLOT", "variable", "iter", "pred")], 
-              data_sim[, c("N_PLOT", "variable", "iter", "value")])
+  data.frame(
+    level_num = paste0("V", 1:30),
+    N_PLOT = data_plot$N_PLOT
+  )
+)
+
+df_op <- merge(
+  data_pred[, c("N_PLOT", "variable", "iter", "pred")],
+  data_sim[, c("N_PLOT", "variable", "iter", "value")]
+)
 gof <- df_op[, .(
-  rmse = sqrt((sum(pred - value)^2)/length(pred)), 
-  r2b = var(pred)/(var(pred) + var(value-pred)), 
+  rmse = sqrt((sum(pred - value)^2) / length(pred)),
+  r2b = var(pred) / (var(pred) + var(value - pred)),
   rho = cor(pred, value)
@@
 The higher SOC stocks in this study could be due to the relatively high exchangeable Ca^2+^ content at the sites included in this study compared with the majority of tropical acid soils, and the presence of carbonates (Table S\@ref(tab:site-soil)).
-Indeed, Ca^2+^ and carbonates are frequently attributed to enhanced soil organic matter stabilization [@Rowley2018; @Wiesmeier2019].
+Indeed, Ca^2+^ and carbonates are frequently attributed to enhanced soil organic matter stabilization [@Rowley2018; @Wiesmeier2019;@Yao2022].
 In contrast, the average biomass stocks (`r dataC_plot[variable%in%c("AGC", "BGC"), sum(mean), .(plot)][, round(mean(V1))]`
@@
 Indeed, in CF systems, the plot is periodically cleared and the remaining trees are pruned to allow the annual crops planted to receive full sun.
-In SP systems, the growth of grazing grasses is favoured by reducing the number of trees; however, the leaves and fruit of some trees (e.g.
-*Ziziphus mauritiana*, *Tamarindus
-indica*, *Schleichera oleiosa*) can provide additional fodder, especially during the dry season.
+In SP systems, the growth of grazing grasses is favoured by reducing the number of trees; however, the leaves and fruit of some trees (e.g., *Ziziphus mauritiana*, *Tamarindus indica*, *Schleichera oleiosa*) can provide additional fodder, especially during the dry season.
 
@@
 Clear-cutting old agroforests for full-sun crops can appear to be an attractive short-term strategy, providing a large cash inflow from the sale of wood products, followed by four or five years of agricultural production on fertile land.
-However, this practice unfortunately results in a significant loss of biodiversity and carbon stored in the biomass, and can lead to an irreversible loss of fertility on the plot through soil degradation and erosion.
+However, this practice unfortunately results in a significant loss of biodiversity and carbon stored in the biomass, and can lead to an irreversible loss of fertility on the plot through soil degradation and erosion [@Gusmao2025].
 This phenomenon has been documented in a similar context in Sumatra (Indonesia), where @Gouyon1993 highlighted the fact that old agroforests, which they called “jungle-rubber”, and which are quite similar in structure to “forest gardens” (except for the rubber tree component, which has been replaced by other species in Timor-Leste), were being cleared in favour of more intensive production systems.
@@
 All data collected for this study is available in the following
-Dataverse repository \_\_\_\_\_\_.
+Dataverse repository https://doi.org/10.18167/DVN1/QCXWIY.
 
@@
 
-```{r tab.id = "summary", results = "show", tab.cap = "Mean and associated 95% confidence intervals (in parentheses) of total, aboveground carbon (AGC), belowground carbon (BGC), and soil organic carbon (SOC; 0-30 cm) stocks in each type of agroforestry system (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden)."}
+```{r tab.id = "summary", results = "show", tab.cap = "Mean ± standard error of total, aboveground carbon (AGC), belowground carbon (BGC), and soil organic carbon (SOC; 0-30 cm) stocks in each type of agroforestry system (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden)."}
 # calculate median and 95% confidence interval at site level
-dataC_afs[, `:=`(value = paste0(
-round(mean), " (",
-round(lwr_mean), "-", 
-round(upr_mean), ")"), 
-afs = factor(afs, levels = c("CF", "SP", "YA", "HG", "FG")))] |> 
+dataC_afs[, `:=`(
+  value = paste(round(mean), "±", round(se)),
+  afs = factor(afs, levels = c("CF", "SP", "YA", "HG", "FG"))
+)] |>
   subset(variable != "SOC10") |>
@@
   setorder() |>
-flextable(col_keys = c("afs", "TOT", "AGC", "BGC", "SOC"), 
-cwidth = c(0.9, 1.2, 1.3, 1.3, 1.2)) |> 
+  flextable(
+    col_keys = c("afs", "TOT", "AGC", "BGC", "SOC"),
+    cwidth = c(0.9, 1.2, 1.3, 1.3, 1.2)
+  ) |>
   compose(i = 1, j = "afs", part = "header", value = as_paragraph("Agroforestry\nsystem")) |>
-compose(i = 1, j = "TOT", part = "header", value = as_paragraph( "Total C stocks\n(Mg C ha", as_sup("-1"), ")"))|> 
-compose(i = 1, j = "AGC", part = "header", value = as_paragraph( "AGC stocks\n(Mg C ha", as_sup("-1"), ")"))|> 
-compose(i = 1, j = "BGC", part = "header", value = as_paragraph( "BGC stocks\n(Mg C ha", as_sup("-1"), ")"))|> 
-compose(i = 1, j = "SOC", part = "header", value = as_paragraph( "SOC stocks\n(Mg C ha", as_sup("-1"), ")"))
+  compose(i = 1, j = "TOT", part = "header", value = as_paragraph("Total C stocks\n(Mg C ha", as_sup("-1"), ")")) |>
+  compose(i = 1, j = "AGC", part = "header", value = as_paragraph("AGC stocks\n(Mg C ha", as_sup("-1"), ")")) |>
+  compose(i = 1, j = "BGC", part = "header", value = as_paragraph("BGC stocks\n(Mg C ha", as_sup("-1"), ")")) |>
+  compose(i = 1, j = "SOC", part = "header", value = as_paragraph("SOC stocks\n(Mg C ha", as_sup("-1"), ")"))
 ```
@@
 
-```{r tab.id = "abbrev", results = "show", tab.cap = "List of abbreviations used in the paper and their definitions."}
+```{r tab.id = "abbrev", results = "show", tab.cap = "List of abbreviations used in the paper."}
 data.frame(
   Abbreviation = c("AFS", "AGC", "BGC", "CBH", "CF", "FG", "HG", "SOC", "SP", "YA"),
-Definition = c("Agroforestry system", "Aboveground carbon", "Belowground carbon",
+  Definition = c(
+    "Agroforestry system", "Aboveground carbon", "Belowground carbon",
     "Circumference at breast height", "Crop and fallow", "Forest garden",
     "Home garden", "Soil organic carbon", "Silvopastoral",
-"Young agroforest")
+    "Young agroforest"
+  )
 ) |>
@@
 ## site location: create box around plots
-data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(site)]
+data_box <- data_plot[, .(xc = mean(long), yc = mean(lat)), .(site)]
 data_box[site == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
 data_box[site == "Osso-Luga", w := 0.05]
-data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
+data_box[, `:=`(x1 = xc - w / 2, x2 = xc + w / 2, y1 = yc - w / 2, y2 = yc + w / 2)]
 
@@
 
-g1 = ggplot(timor_map) + 
-annotate(geom = "text", label = "a", x = -Inf, y = Inf, 
-hjust = -3, vjust = 1, size = 5) + 
-geom_polygon(aes(x = long, y = lat, group = group), fill="lightgray", colour = "grey") +
+g1 <- ggplot(timor_map) +
+  annotate(
+    geom = "text", label = "a", x = -Inf, y = Inf,
+    hjust = -3, vjust = 1, size = 5
+  ) +
+  geom_polygon(aes(x = long, y = lat, group = group), fill = "lightgray", colour = "grey") +
   geom_tile(data = data_box, aes(x = xc, y = yc, width = w, height = w), fill = NA, color = 1, linewidth = 0.5) +
@@
   labs(x = "", y = "") +
-theme(legend.position = "top", axis.title = element_blank(), 
+  theme(
+    legend.position = "top", axis.title = element_blank(),
     axis.ticks = element_blank(), axis.text = element_blank(),
-panel.background = element_blank())+
-coord_sf(crs = 'WGS84')
+    panel.background = element_blank()
+  ) +
+  coord_sf(crs = "WGS84")
 
 # get elevation information
-elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
+elev <- as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
 elevS <- subset(elev, x > data_box[site == "Osso-Luga"]$x1 &
@@
   y < data_box[site == "Osso-Luga"]$y2)
-elevS$site = "Osso-Luga"
+elevS$site <- "Osso-Luga"
 elevG <- subset(elev, x > data_box[site == "Gariuai"]$x1 &
@@
   y < data_box[site == "Gariuai"]$y2)
-elevG$site = "Gariuai"
-
-elevAll = rbind(elevS, elevG)
+elevG$site <- "Gariuai"
+
+elevAll <- rbind(elevS, elevG)
 
@@
 ## Map of sites and plots with elevation
-g2 = 
+g2 <-
   ggplot(data = elevAll) +
-geom_raster(aes(x = x, y = y, fill=srtm_62_14)) +
-geom_point(data = data_plot, 
-aes(x = long, y = lat, color = ID_AF), size = 2) + 
-geom_point(data = data_plot, 
-aes(x = long, y = lat), shape = 1, size = 2) + 
+  geom_raster(aes(x = x, y = y, fill = srtm_62_14)) +
+  geom_point(
+    data = data_plot,
+    aes(x = long, y = lat, color = ID_AF), size = 2
+  ) +
+  geom_point(
+    data = data_plot,
+    aes(x = long, y = lat), shape = 1, size = 2
+  ) +
   scale_fill_gradientn(colours = terrain.colors(10)) +
   scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
-labs(x = NULL, y = NULL, color = NULL, fill="Elevation (m)") +
-annotate(geom = "text", label = "b", x = -Inf, y = Inf, 
-hjust = -0.5, vjust = 1, size = 5) + 
-annotate(geom = "text", label = "c", x = Inf, y = -Inf, 
-hjust = 13.7, vjust = -10.8, size = 5) + 
-guides(color = FALSE, fill=guide_colorbar(title.position = "top")) +
-theme(legend.position = c(0.8, 0.85), legend.direction="horizontal", 
+  labs(x = NULL, y = NULL, color = NULL, fill = "Elevation (m)") +
+  annotate(
+    geom = "text", label = "b", x = -Inf, y = Inf,
+    hjust = -0.5, vjust = 1, size = 5
+  ) +
+  annotate(
+    geom = "text", label = "c", x = Inf, y = -Inf,
+    hjust = 13.7, vjust = -10.8, size = 5
+  ) +
+  guides(color = FALSE, fill = guide_colorbar(title.position = "top")) +
+  theme(
+    legend.position = c(0.8, 0.85), legend.direction = "horizontal",
     axis.title = element_blank(),
     plot.title = element_text(hjust = 0.5),
-panel.background = element_blank()) +
-scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
+    panel.background = element_blank()
+  ) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
   annotation_scale() +
-coord_sf(crs = 'WGS84')
+  coord_sf(crs = "WGS84")
 
 # get AFS legend
-g3 = ggplot(data_plot) +
+g3 <- ggplot(data_plot) +
   geom_point(aes(x = long, y = lat, color = ID_AF)) +
   labs(color = "AFS") +
-guides(color=guide_legend(ncol=2)) +
+  guides(color = guide_legend(ncol = 2)) +
   scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#E6AB02")) +
   theme_classic()
-l1 = get_legend(g3)
-
-ggarrange(ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))
+l1 <- get_legend(g3)
+
+ggarrange(ggarrange(g1, l1), g2, ncol = 1, heights = c(1, 2))
 ```
@@
 ```{r fig.id = "barplot", fig.height = 5, fig.width = 6, fig.cap = "Distribution of carbon stocks in each type of AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden), per site (Gariuai, Osso-Luga).
-AGC: Aboveground (biomass) carbon; BGC: belowground (biomass) carbon; SOC: soil organic carbon (between 0-30 cm)."}
+AGC: Aboveground (biomass) carbon; BGC: belowground (biomass) carbon; SOC: soil organic carbon (at 0-30 cm depth).
+Vertical segments represent the standard error of the total carbon stocks (AGB + BGC + SOC)."}
+
 ## reorder factors
@@
   ggplot(aes(x = afs, y = mean)) +
-geom_bar(position="stack", stat="identity", aes(fill = variable)) + 
+  geom_bar(position = "stack", stat = "identity", aes(fill = variable)) +
+  geom_errorbar(data = dataC_afs_site[variable == "TOT"], 
+                aes(ymin = mean, ymax = mean + se), width = 0) +
   facet_wrap(~site, nrow = 1, strip.position = "bottom") +
@@
   theme_classic() +
-labs(x = "", y = expression("Carbon stocks (Mg C"*" "*ha^{-1}*")"), fill = "Carbon\npool") +
+  labs(x = "", y = expression("Carbon stocks (Mg C" * " " * ha^{
+    -1
+  } * ")"), fill = "Carbon\npool") +
   scale_y_continuous(expand = c(0, 0)) +
@@
 
-```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3, fig.cap = "Effect of (a, b) AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden) and (c, d) site (G: Gariuai, OL: Osso-Luga) on (a, c) aboveground carbon (AGC, in Mg C ha$^{-1}$) and (b, d) soil organic carbon (SOC, between 0-30 cm, in Mg C ha$^{-1}$).
+```{r fig.id = "effect-afs-site", fig.width = 7, fig.height = 3, fig.cap = "Effect of (a, b) AFS (CF: crop and fallow, SP: silvopasture, YA: young agroforest, HG: home garden, FG: forest garden) and (c, d) site (G: Gariuai, OL: Osso-Luga) on (a, c) aboveground carbon (AGC, in Mg C ha$^{-1}$) and (b, d) soil organic carbon (SOC, at 0-30 cm depth, in Mg C ha$^{-1}$).
 These values were obtained as the sum of the mean predicted stocks ($\\mu$) and the AFS or site random effects.
@@
 labels_stock <- function(x) paste0(toupper(x), "~(Mg~C~ha^-1)")
-labels_effect <- function(x) ifelse(x=="afs", toupper(x), x)
+labels_effect <- function(x) ifelse(x == "afs", toupper(x), x)
 
@@
   geom_point(aes(x = mean), size = 2) +
-geom_text(data = data_letter, 
+  geom_text(
+    data = data_letter,
     x = -Inf, y = Inf, hjust = -1, vjust = 1, col = 1,
-aes(label = lab)) +
+    aes(label = lab)
+  ) +
   scale_x_continuous(breaks = seq(20, 100, 10)) +
@@
     labels_effect(effect) ~ labels_stock(stock),
-scales = "free", space = "free", switch = "both", labeller = label_parsed) +
-labs(x=NULL, y = NULL) +
+    scales = "free", space = "free", switch = "both", labeller = label_parsed
+  ) +
+  labs(x = NULL, y = NULL) +
   theme_bw() +
@@
 Black dots represent the mean of the 1000 iterations; red segments represent 80% confidence intervals and black segments represent 95% confidence intervals."}
-labels_stock = function(x)
-ifelse(x=="agc", "a. AGC", ifelse(grepl("10", x), "c. SOC (0-10 cm)", "b. SOC (0-30 cm)"))
+labels_stock <- function(x) {
+  ifelse(x == "agc", "a. AGC", ifelse(grepl("10", x), "c. SOC (0-10 cm)", "b. SOC (0-30 cm)"))
+}
 
@@
   facet_wrap(~stock, labeller = labeller(stock = labels_stock)) +
-labs(x="Effect of covariates on C stocks", y = NULL) +
-scale_y_discrete(labels= c("AGC", "Clay", "MAT")) +
+  labs(x = "Effect of covariates on C stocks", y = NULL) +
+  scale_y_discrete(labels = c("AGC", "Clay", "MAT")) +
   theme_bw() +
@@
 The dotted lines represent the mean predicted SOC stocks for a given AGC stock value, and the grey areas represent the 95% confidence interval of this prediction."}
-
-dfpred <- expand.grid(agc = seq(0, 170, length.out = 20), 
+dfpred <- expand.grid(
+  agc = seq(0, 170, length.out = 20),
   iter = unique(data_par$iter)
@@
   merge(
-data_par[grepl("soc", stock) & (param =="mu" | level == "agc")] |> 
+    data_par[grepl("soc", stock) & (param == "mu" | level == "agc")] |>
       dcast(iter + stock ~ param)
@@
   merge(
-data_sim[variable=="AGC", .(mua = mean(value), sda = sd(value)), 
-.(iter = as.numeric(as.factor(iter)))]
-) |> setDT()
-dfpred[, pred := (mu - mua*theta/sda) + (theta/sda) * agc]
+    data_sim[
+      variable == "AGC", .(mua = mean(value), sda = sd(value)),
+      .(iter = as.numeric(as.factor(iter)))
+    ]
+  ) |>
+  setDT()
+dfpred[, pred := (mu - mua * theta / sda) + (theta / sda) * agc]
 dfpred <- dfpred[, .(
@@
 ggplot(dataC_fig, aes(color = afs)) +
-geom_ribbon(data = dfpred, aes(x = agc, ymin = lower, ymax = upper), 
-col = NA, alpha = 0.1) +
+  geom_ribbon(
+    data = dfpred, aes(x = agc, ymin = lower, ymax = upper),
+    col = NA, alpha = 0.1
+  ) +
   geom_point(aes(x = mean_AGC, y = mean_SOC, shape = site), size = 2) +
@@
   geom_line(data = dfpred, aes(x = agc, y = mean), lty = 2, col = 1) +
-labs(color = "AFS", x = expression("AGC stocks (Mg C"*" "*ha^{-1}*")"), 
-y = expression("SOC stocks (Mg C"*" "*ha^{-1}*")"), shape = "Site") +
+  labs(
+    color = "AFS", x = expression("AGC stocks (Mg C" * " " * ha^{
+      -1
+    } * ")"),
+    y = expression("SOC stocks (Mg C" * " " * ha^{
+      -1
+    } * ")"), shape = "Site"
+  ) +
   expand_limits(x = 0, y = 0) +
@@
 $$ \Rightarrow C_{10,20} = C_{0,10} \cdot \left( \frac{C_{20,30}}{C_{0,10}}\right)^{\frac{1}{2}}$$
-\### Choice of climate and soil covariates
+### Choice of climate and soil covariates
 
@@
 pca_clim <- data_plot[, c("temp", "prec", "varp")] |>
-scale() |> princomp()
+  scale() |>
+  princomp()
 
@@
   data_soil |>
-melt(id.vars = c("N_PLOT", "SAMPLE_DEPTH"), 
+  melt(
+    id.vars = c("N_PLOT", "SAMPLE_DEPTH"),
     measure.vars = c(
       "pH\n", "CEC\n", "CLAY", "FINE SILT",
-"COARSE SILT", "FINE SAND", "COARSE SAND"))
+      "COARSE SILT", "FINE SAND", "COARSE SAND"
+    )
+  )
 
@@
   dcast(N_PLOT ~ variable)
-data_soil_melt[, silt := `COARSE SILT` +`FINE SILT`]
+data_soil_melt[, silt := `COARSE SILT` + `FINE SILT`]
 data_soil_melt[, sand := `FINE SAND` + `COARSE SAND`]
@@
 pca_soil <- data_soil_melt[, c("pH\n", "CLAY", "silt", "sand")] |>
-scale() |> princomp()
+  scale() |>
+  princomp()
 ```
@@
 # choose variables to show
-vars = c("Corg\n", "N", "Corg_N", "TC", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n")
-
-data_soil_plot = melt(data_soil_plot, 
+vars <- c("Corg\n", "N", "Corg_N", "TC", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n")
+
+data_soil_plot <- melt(data_soil_plot,
   id.vars = c("N_PLOT", "SAMPLE_DEPTH", "site"),
-                      measure.vars = vars)
+  measure.vars = vars
+)
 
@@
 # when value below detection limit: use half of detection value
-data_soil_plot[grepl("LD", value), value := as.numeric(strsplit(value, "=")[[1]][2])/2]
+data_soil_plot[grepl("LD", value), value := as.numeric(strsplit(value, "=")[[1]][2]) / 2]
 data_soil_plot[, value := as.numeric(value)]
@@
 # remove NA values
-data_soil_plot = subset(data_soil_plot, !is.na(value))
+data_soil_plot <- subset(data_soil_plot, !is.na(value))
 
@@
 # of variation (CV value from lab)
-data_soil_plot_simu = 
-  data_soil_plot[, .(
-    value = rnorm(1000, value, sd = 0.12*value), n = 1:1000), 
-    .(site, N_PLOT, SAMPLE_DEPTH, variable)]
+data_soil_plot_simu <-
+  data_soil_plot[
+    , .(
+      value = rnorm(1000, value, sd = 0.12 * value), n = 1:1000
+    ),
+    .(site, N_PLOT, SAMPLE_DEPTH, variable)
+  ]
 
 # bootstrap plot values to get mean value at the site level (for each simulation)
-data_soil_site_simu = 
-  data_soil_plot_simu[, .(
-    value = mean(sample(value, length(value), replace = TRUE))),
-    .(site, variable, n)]
+data_soil_site_simu <-
+  data_soil_plot_simu[
+    , .(
+      value = mean(sample(value, length(value), replace = TRUE))
+    ),
+    .(site, variable, n)
+  ]
 
 # convert sand, silt, clay and C content from % into g/kg
-data_soil_site_simu[ variable %in% c("Corg", "CLAY", "sand", "silt", "TC"), value := value*10]
+data_soil_site_simu[variable %in% c("Corg", "CLAY", "sand", "silt", "TC"), value := value * 10]
 
 # calculate mean and 95% confidence interval at site level
-data_soil_site  = data_soil_site_simu[, .(
+data_soil_site <- data_soil_site_simu[
+  , .(
     mean = signif(mean(value), 3),
@@
   ),
-.(site, variable)]
+  .(site, variable)
+]
 data_soil_site[, value := paste0(round(mean, 1), " (", round(lwr, 1), "-", round(upr, 1), ")")]
@@
 # change table shape to have one column per variable
-data_soil_site = dcast(data_soil_site, site ~ variable, value.var = "value")
+data_soil_site <- dcast(data_soil_site, site ~ variable, value.var = "value")
 
 ## add unit row
-data_soil_site = data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "TC", "pH", "CEC", "Ca", "K", "Mg", "Na")]
-colnames(data_soil_site) = c(" ", "Clay", "Silt", "Sand", "Organic C content", "Total N content", "C:N", "Carbonate content", "pH", "CEC", "Ca", "K", "Mg", "Na")
+data_soil_site <- data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "TC", "pH", "CEC", "Ca", "K", "Mg", "Na")]
+colnames(data_soil_site) <- c(" ", "Clay", "Silt", "Sand", "Organic C content", "Total N content", "C:N", "Carbonate content", "pH", "CEC", "Ca", "K", "Mg", "Na")
 ```
@@
 read_excel("tables/afs-typology.xlsx") |>
-  flextable(col_keys = c(
+  flextable(
+    col_keys = c(
       "Typology adapted from Nair (1993)",
@@
       "Time since installation",
-    "Previous land use"), 
-    cwidth = c(1.5, 1.5, 1, 1, 1.5)) 
+      "Previous land use"
+    ),
+    cwidth = c(1.5, 1.5, 1, 1, 1.5)
+  )
 ```
@@
 ```{r tetum-scientific-correspondance}
-sheets1 = grep("INV", excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
-sheets2 = grep("INV", excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)
+sheets1 <- grep("INV", excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
+sheets2 <- grep("INV", excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)
 
 vern1 <- lapply(sheets1, \(x){
-  df = read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
-  df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
-    unique() |>  subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
+  df <- read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
+  df <- df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
+    unique() |>
+    subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
 })
 vern2 <- lapply(sheets2, \(x){
-  df = read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
-  df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
-    unique() |>  subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
+  df <- read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
+  df <- df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
+    unique() |>
+    subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
 })
 
-vern = rbind(rbindlist(vern1), rbindlist(vern2)) |>  unique()
+vern <- rbind(rbindlist(vern1), rbindlist(vern2)) |> unique()
 vern <- subset(vern, !is.na(TETUM_NAME) & !is.na(SCIENTIFIC_NAME))
@@
 vern[species == "purpuroides", species := NA]
-corr = correctTaxo(genus = vern$genus, species = vern$species)
-vern = cbind(vern, corr)
+corr <- correctTaxo(genus = vern$genus, species = vern$species)
+vern <- cbind(vern, corr)
 
@@
 vern[, latin := gsub(" NA", " sp.", latin)]
-vern[,tetum := gsub("  ", " ", tolower(TETUM_NAME))]
+vern[, tetum := gsub("  ", " ", tolower(TETUM_NAME))]
 vern <- vern[, c("tetum", "latin", "genusCorrected", "speciesCorrected")]
-vern = unique(vern)
+vern <- unique(vern)
 
@@
 dupl_tetum <- vern[duplicated(tetum), unique(tetum)]
-vern = vern[!(tetum %in% dupl_tetum & is.na(speciesCorrected))]
+vern <- vern[!(tetum %in% dupl_tetum & is.na(speciesCorrected))]
 setorder(vern, latin)
 
-vern = vern[, .(tetum = paste(tetum, collapse = ", ")), .(latin)]
+vern <- vern[, .(tetum = paste(tetum, collapse = ", ")), .(latin)]
 ```
@@
   autofit() |>
-  italic(j=1)
+  italic(j = 1)
 ```
@@
   setnames(c("MAT", "MAP", "Pvar")) |>
-  cor() |> melt() |>
+  cor() |>
+  melt() |>
   ggplot(aes(x = Var1, y = Var2, label = round(value, 2), fill = value)) +
@@
   princomp() |>
-  autoplot(data = data_plot, colour = 'site', loadings.colour = 'black',
-           loadings.label = TRUE, loadings.label.col = 'black') +
+  autoplot(
+    data = data_plot, colour = "site", loadings.colour = "black",
+    loadings.label = TRUE, loadings.label.col = "black"
+  ) +
   theme_classic()
@@
   setnames(c("clay", "silt", "sand", "pH")) |>
-  cor() |> melt() |>
+  cor() |>
+  melt() |>
   ggplot(aes(x = Var1, y = Var2, label = round(value, 2), fill = value)) +

### 2025-04-29 script-agb.R
delete agb script (outdated)

@@ -1,249 +1,0 @@
-### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ###  ### 
-### Estimation of aboveground biomass in agroforestry systems in Timor Leste ###
-### Author: Camille Piponiot                                                 ###
-### Correspondance:           camille.piponiot-laroche@cirad.fr              ###
-### Date: 2022-06-21                                                         ###
-### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ###  ###
-
-
-## 1. Prepare R environment ####
-
-# R packages - installation 
-
-# required packages
-req_packages <- c("BIOMASS", "readxl", "data.table", "ggplot2")
-
-# install missing packages
-inst_packages <- req_packages[!( req_packages %in% rownames(installed.packages()))]
-
-if (length(inst_packages) > 0)
-  install.packages(inst_packages)
-
-# Set working directory to source file location 
-setwd(dirname(rstudioapi::getSourceEditorContext()$path))
-
-
-## 2. Load data from excel file ####
-
-# Open large tree data (> 100 cm circumference)
-data_large <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV >100CM")
-
-# Open large tree data (30-100 cm circumference)
-data_small <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV 30-100CM")
-
-# Open plot information
-data_plot <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")
-
-
-## 3. combine tables ####
-
-# create a "N_Q_ID" column in data_large, with corresponding quadrat number =
-# "TOT" (entire plot)
-data_large$N_Q_ID <- paste(data_large$N_PLOT, "TOT", sep = "_")
-
-# common columns to be kept
-col_keep <- c("N_PLOT", "N_Q_ID", "SCIENTIFIC_NAME", "DIAM_130_CM", "HEIGHT")
-
-# stack tree tables
-data_tot <- rbind(data_large[, col_keep], data_small[, col_keep])
-
-# merge with plot information
[…]

### 2025-04-29 manuscript.Rmd
remove si

@@
     reference_docx: misc/word-styles-ref-02.docx
-bibliography: misc/references.bib
+bibliography: references.bib
 nocite: |
@@
 # package management is done with renv
-packages_needed <- c(
-  "ggplot2", "ggrepel",  "ggspatial",  "ggpubr",  "ggfortify",  "maps", 
-  "rdryad", "readxl", "data.table", "knitr", "BIOMASS", "terra", "car", 
-  "flextable"
-)
-
-lapply(packages_needed, require, character.only = TRUE)
+packages_needed <- c("maps", "car")
+library(ggplot2)
+library(ggrepel)
+library(ggspatial)
+library(ggpubr)
+library(ggfortify)
+library(BIOMASS)
+library(data.table)
+library(flextable)
+library(terra)
+library(rdryad)
+library(knitr)
+library(readxl)
+library(dataverse)
 
@@
 data.frame(
-  Abbreviation = c("AFS", "AGC", "BGC", "CBH", "CF", "FG", "HG", "SOC", "SP", "YA"),
+  Abbreviation = c("AFS", "AGC", "BGC", "CF", "FG", "HG", "SOC", "SP", "YA"),
   Definition = c(
     "Agroforestry system", "Aboveground carbon", "Belowground carbon",
-    "Circumference at breast height", "Crop and fallow", "Forest garden",
+    "Crop and fallow", "Forest garden",
     "Home garden", "Soil organic carbon", "Silvopastoral",
@@
 :::
-
-# Supplemental Information
-
-## Supplemental methods
-
-### Calculation of 10-20 cm average carbon content
-
-During field data collection, because the auger length (15 cm) didn't
-match the target soil sampling depths (10 cm each), we could only
-accurately sample and analyse the 0-10 cm and 20-30 cm soil depths for
-SOC content. The SOC content of the intermediate 10-20 cm soil depth had
-to be estimated based on these first two measurements.
-
-To reduce the bias in this estimation, we used the standard assumption
-that SOC content $c(x)$ decreases as an exponential function of soil
-depth $x$:
-
-$$c(x) = C_0 \cdot exp(-\lambda\cdot x) $$ where $C_0$ is the carbon
-content at the surface (depth $x= 0$) and $\lambda$ is the rate at which
-the carbon content decreases with soil depth. Under this assumption, we
-can calculate the mean SOC content between depths $x$ and $x+\Delta$ as:
-
-$$\bar c(x, x+\Delta) = \frac{1}{\Delta}\int_{x}^{x+\Delta}  C_0 \cdot exp(-\lambda\cdot x) dx$$
-
-$$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{-\lambda \cdot \Delta} \left[ exp(-\lambda\cdot (x+\Delta)) - exp(-\lambda\cdot x) \right]$$
-
-$$ \Rightarrow \bar c(x, x+\Delta) = \frac{C_0}{\lambda \cdot \Delta} exp(-\lambda\cdot x)\left[1- exp(-\lambda\cdot\Delta) \right]$$
-The average SOC contents between 0-10 cm, 10-20 cm and 20-30 cm can
-respectively be written as:
-
-$$C_{0,10} = \frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)$$
-
-$$C_{10,20} = \frac{C_0}{10\lambda} exp(-10\lambda)\left(1- exp(-10\lambda) \right)$$
-
-$$C_{20,30} = \frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)$$
-We can thus write:
-
-$$\frac{C_{20,30}}{C_{0,10}} = \frac{\frac{C_0}{10\lambda} exp(-20\lambda)\left(1- exp(-10\lambda) \right)}{\frac{C_0}{10\lambda}\left(1- exp(-10\lambda) \right)} = exp(-20\lambda)$$
-$$ \Rightarrow \lambda = \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20}$$ We
-can also write:
-
-$$ \frac{C_{10,20}}{C_{0,10}} = exp(-10\lambda) $$
-$$\Rightarrow C_{10,20} = C_{0,10} \cdot exp(-10 \cdot \frac{log(\frac{C_{0,10}}{C_{20,30}})}{20})$$
-
-$$ \Rightarrow C_{10,20} = C_{0,10} \cdot \left( \frac{C_{20,30}}{C_{0,10}}\right)^{\frac{1}{2}}$$
-### Choice of climate and soil covariates
-
-As described in the "Carbon stocks model" section of the Materials and
-Methods, we hypothesized that C stocks (both AGC and SOC) are affected
-by climate and soil characteristics, together with land use (in this
-case, AFS type) and other spatial differences (in our case, a site
-effect).
-
-```{r pca-clim-soil}
-pca_clim <- data_plot[, c("temp", "prec", "varp")] |>
-  scale() |>
-  princomp()
-
-data_soil_melt <-
-  data_soil |>
-  melt(
-    id.vars = c("N_PLOT", "SAMPLE_DEPTH"),
-    measure.vars = c(
-      "pH\n", "CEC\n", "CLAY", "FINE SILT",
-      "COARSE SILT", "FINE SAND", "COARSE SAND"
-    )
-  )
-
-# make values numeric; when detectability limit is reached, used limit value
-data_soil_melt[grepl("LD", value), value := gsub(",", ".", tstrsplit(value, "LD=")[[2]])]
-
-data_soil_melt <-
-  data_soil_melt[, .(value = mean(as.numeric(value), na.rm = TRUE)), .(N_PLOT, variable)] |>
-  dcast(N_PLOT ~ variable)
-data_soil_melt[, silt := `COARSE SILT` + `FINE SILT`]
-data_soil_melt[, sand := `FINE SAND` + `COARSE SAND`]
-
-pca_soil <- data_soil_melt[, c("pH\n", "CLAY", "silt", "sand")] |>
-  scale() |>
-  princomp()
-```
-
-We selected the following set of variables to represent the effect of
-climate: mean annual temperature (BIO1), mean annual precipitation
-(BIO12), annual precipitation variability (BIO15) over the period
-1981-2010. These bioclimatic variables were extracted from CHELSA V2.1
-[@Karger2017]. These variables were all highly correlated in our dataset
-(Figure S\@ref(fig:choice-clim-vars)a). We performed a principal
-component analysis after scaling and centering these three variables:
-the first component explained
-`r round(pca_clim$sdev[1]^2/sum(pca_clim$sdev^2)*100)`% of the variance
-in our dataset (Figure S\@ref(fig:choice-clim-vars)b). We therefore
-decided to keep only one variable to represent this climatic gradient:
-mean annual temperature, which was most correlated with the first
-component of the PCA (Figure S\@ref(fig:choice-clim-vars)b), and has
-been shown to influence C stocks in other contexts [@Wiesmeier2019;
-@Muller-Landau2021].
-
-We selected the following set of variables to represent the effect of
-soil properties on C stocks: clay content (\<2 $\mu$m), silt content
-(2-50 $\mu$m), sand content (50-2000 $\mu$m), and pH. These variables
-were measured on our soil samples analysed in the Cirad laboratories
-(see section "Field inventories"). After scaling and centering these
-variables, we performed a principal component analysis: the first
-component explained
-`r round(pca_soil$sdev[1]^2/sum(pca_soil$sdev^2)*100)`% of the variance
-in our dataset (Figure S\@ref(fig:choice-soil-vars)b). We therefore
-decided to retain only one variable to represent this edaphic gradient:
-clay content, which was the most highly correlated with the first
-component of the PCA (Figure S\@ref(fig:choice-soil-vars)b) and has been
-shown to be a good predictor of SOC stocks in other contexts
-[@Wiesmeier2019].
-
-## Supplemental tables
-
-```{r site-soil-char}
-data_soil[, Corg := `Corg\n`]
-data_soil[, muCorg := mean(Corg, na.rm = TRUE), .(N_PLOT)]
-data_soil[is.na(Corg), Corg := muCorg]
-
-data_soil[, c("h0", "h1") := lapply(tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
-
-data_soil$silt <- as.numeric(data_soil$`FINE SILT`) + as.numeric(data_soil$`COARSE SILT`)
-data_soil$sand <- as.numeric(data_soil$`FINE SAND`) + as.numeric(data_soil$`COARSE SAND`)
-
-# add site info to data_soil
-data_soil_plot <- merge(data_soil, data_plot[, c("N_PLOT", "site")])
-
-# choose variables to show
-vars <- c("Corg\n", "N", "Corg_N", "TC", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n")
-
-data_soil_plot <- melt(data_soil_plot,
-  id.vars = c("N_PLOT", "SAMPLE_DEPTH", "site"),
-  measure.vars = vars
-)
-
-data_soil_plot[, variable := gsub("\n", "", variable)]
-
-# when value below detection limit: use half of detection value
-data_soil_plot[grepl("LD", value), value := as.numeric(strsplit(value, "=")[[1]][2]) / 2]
-data_soil_plot[, value := as.numeric(value)]
-
-# remove NA values
-data_soil_plot <- subset(data_soil_plot, !is.na(value))
-
-# simulation 1000 values per plot, soil depth and variable with a 12% coefficient
-# of variation (CV value from lab)
-data_soil_plot_simu <-
-  data_soil_plot[
-    , .(
-      value = rnorm(1000, value, sd = 0.12 * value), n = 1:1000
-    ),
-    .(site, N_PLOT, SAMPLE_DEPTH, variable)
-  ]
-
-# bootstrap plot values to get mean value at the site level (for each simulation)
-data_soil_site_simu <-
-  data_soil_plot_simu[
-    , .(
-      value = mean(sample(value, length(value), replace = TRUE))
-    ),
-    .(site, variable, n)
-  ]
-
-# convert sand, silt, clay and C content from % into g/kg
-data_soil_site_simu[variable %in% c("Corg", "CLAY", "sand", "silt", "TC"), value := value * 10]
-
-# calculate mean and 95% confidence interval at site level
-data_soil_site <- data_soil_site_simu[
-  , .(
-    mean = signif(mean(value), 3),
-    lwr = signif(quantile(value, 0.025), 3),
-    upr = signif(quantile(value, 0.975), 3)
-  ),
-  .(site, variable)
-]
-data_soil_site[, value := paste0(round(mean, 1), " (", round(lwr, 1), "-", round(upr, 1), ")")]
-data_soil_site[variable %in% c("K", "Mg", "Na"), value := paste0(round(mean, 2), " (", round(lwr, 2), "-", round(upr, 2), ")")]
-
-# change table shape to have one column per variable
-data_soil_site <- dcast(data_soil_site, site ~ variable, value.var = "value")
-
-## add unit row
-data_soil_site <- data_soil_site[, c("site", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "TC", "pH", "CEC", "Ca", "K", "Mg", "Na")]
-colnames(data_soil_site) <- c(" ", "Clay", "Silt", "Sand", "Organic C content", "Total N content", "C:N", "Carbonate content", "pH", "CEC", "Ca", "K", "Mg", "Na")
-```
-
-```{r tab.id = "site-soil", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Soil characteristics of the sites (average over 0-30 cm), presented as mean values with 95% confidence intervals (in parentheses).
-Units are: in g/kg for clay, silt, sand, carbonate, organic C and total N contents; in cmolc/kg for CEC, Ca, K, Mg and Na; and no units for C:N and pH.
-The confidence intervals were calculated by propagating a 12% error on the lab-provided values and obtaining site-level estimates through bootstrapping plot values (1000 simulations).
-Organic carbon content (difference of the total carbon – dry combustion method – and the carbonates contents - calcimeter) ; total nitrogen content (dry combustion method) ; pH in water (soil:deionized water ratio of 1:2.5) ; CEC - cation exchangeable capacity - and exchangeable cations (Ca, K, Mg and Na) at soil pH [Cobaltihexamine method; Orsini and Remy, 1976]."}
-flextable(data_soil_site, cwidth = 0.5)
-```
-
-</br>
-
-```{r tab.id = "afs-typology", tab.cap.pre="Table S", tab.cap = "Description of 5 agroforestry systems in Central Timor-Leste.", tab.autonum.start_at=1, tab.lp="supp-tab", results = "show"}
-read_excel("tables/afs-typology.xlsx") |>
-  flextable(
-    col_keys = c(
-      "Typology adapted from Nair (1993)",
-      "Usual location",
-      "Typical plot size (ha)",
-      "Time since installation",
-      "Previous land use"
-    ),
-    cwidth = c(1.5, 1.5, 1, 1, 1.5)
-  )
-```
-
-</br>
-
-```{r tetum-scientific-correspondance}
-sheets1 <- grep("INV", excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
-sheets2 <- grep("INV", excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)
-
-vern1 <- lapply(sheets1, \(x){
-  df <- read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
-  df <- df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
-    unique() |>
-    subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
-})
-vern2 <- lapply(sheets2, \(x){
-  df <- read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x)
-  df <- df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |>
-    unique() |>
-    subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
-})
-
-vern <- rbind(rbindlist(vern1), rbindlist(vern2)) |> unique()
-vern <- subset(vern, !is.na(TETUM_NAME) & !is.na(SCIENTIFIC_NAME))
-
-## correct typos in vernacular names
-vern[, TETUM_NAME := tolower(gsub(" oan| kain| restu| hun restu", "", TETUM_NAME))]
-vern[grep("kami", TETUM_NAME), TETUM_NAME := "ai-kamii"]
-vern[, TETUM_NAME := gsub("ai-haek", "ai-hanek", TETUM_NAME)]
-vern[, TETUM_NAME := gsub("ai-bobur", "ai-bubur", TETUM_NAME)]
-vern[, TETUM_NAME := gsub("duut", "du'ut", TETUM_NAME)]
-vern[, TETUM_NAME := gsub("aifarina", "ai-farina", TETUM_NAME)]
-vern[, TETUM_NAME := gsub("aitarak", "ai-tarak", TETUM_NAME)]
-vern[, TETUM_NAME := gsub("ainanas", "ai-nanas", TETUM_NAME)]
-vern[, TETUM_NAME := gsub("fore tali", "fore-tali", TETUM_NAME)]
-vern[, TETUM_NAME := gsub("kromolaina", "kromolaena", TETUM_NAME)]
-
-## correct scientific names
-vern[, c("genus", "species") := tstrsplit(SCIENTIFIC_NAME, " ")[1:2]]
-vern[species %in% c("sp", "sp."), species := NA]
-vern[species == "purpuroides", species := NA]
-corr <- correctTaxo(genus = vern$genus, species = vern$species)
-vern <- cbind(vern, corr)
-
-vern[, latin := paste(genusCorrected, speciesCorrected)]
-vern[, latin := gsub(" NA", " sp.", latin)]
-vern[, tetum := gsub("  ", " ", tolower(TETUM_NAME))]
-vern <- vern[, c("tetum", "latin", "genusCorrected", "speciesCorrected")]
-vern <- unique(vern)
-
-# check that all names are unique
-# if one tetum name has two levels of id (same genus), keep better identification
-dupl_tetum <- vern[duplicated(tetum), unique(tetum)]
-vern <- vern[!(tetum %in% dupl_tetum & is.na(speciesCorrected))]
-setorder(vern, latin)
-
-vern <- vern[, .(tetum = paste(tetum, collapse = ", ")), .(latin)]
-```
-
-```{r tab.id = "tetum-names", tab.lp="supp-tab", results = "show", tab.cap.pre="Table S", tab.cap = "Correspondence between Tetum names and scientific names used in this study."}
-flextable(vern) |>
-  set_header_labels(values = c("Scientific name", "Tetum name")) |>
-  autofit() |>
-  italic(j = 1)
-```
-
-## Supplemental figures
-
-```{r fig.id="design", fig.cap = "Plot sampling design.
-(a) Typical plot < 1 ha, in all AFS except SPs.
-(b) Typical 1-ha plot in SPs.
-Red areas are the five 10x10 m subplots, and blue dots are the three soil sampling locations.", fig.autonum.start_at=1, fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 5, fig.height = 3}
-include_graphics("figures/plot-design.jpg")
-```
-
-</br>
-
-```{r fig.id="seq-diagram", fig.cap = "Typical sequence of AFS in the sites included in this study.
-Transitions between two systems are represented by arrows: dashed arrows represent transitions that usually include tree removal, whereas solid arrows represent transitions where trees are planted and/or the AFS is left to regenerate naturally.
-These transitions may vary according to the farmer's strategy and available resources (Cogné and Lescuyer, 2024).", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S", fig.width = 8, fig.height = 2.2, out.width="50%"}
-# save_png <- function(plot, path){
-#   DiagrammeRsvg::export_svg(plot) %>%
-#     charToRaw() %>%
-#     rsvg::rsvg() %>%
-#     png::writePNG(path)
-# }
-#
-# DiagrammeR::grViz("
-#       digraph {
-#       graph [layout = dot, rankdir = LR]
-#
-#       node [shape = ellipse]
-#       CF[color = '#1B9E77'];
-#       SP[color = '#D95F02'];
-#       YA[color = '#7570B3'];
-#       HG[color = '#E7298A'];
-#       FG[color = '#E6AB02'];
-#
-#       NE[label = 'Forest or\nsavanna', shape = rectangle, fillcolor ='black'];
-#
-#       # Connect nodes with edges and labels
-#       CF -> YA -> HG -> FG;
-#       YA -> FG;
-#       edge[style = 'dashed']
-#       SP -> CF;
-#       FG -> CF
-#       FG -> YA
-#       NE -> SP
-#       NE -> CF
-#        }
-#
-#       ") %>%
-#   save_png("figures/seq_diagram.png")
-include_graphics("figures/seq_diagram.png")
-```
-
-<!-- </br> -->
-
-<!-- ```{r fig.id = "dag", fig.cap="Direct acyclic plot of the hypothesized effect of soil characteristics, climate and land use (here AFS) on AGC and SOC stocks.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"} -->
-
-<!-- # save_png <- function(plot, path){ -->
-
-<!-- #   DiagrammeRsvg::export_svg(plot) %>% -->
-
-<!-- #     charToRaw() %>% -->
-
-<!-- #     rsvg::rsvg() %>% -->
-
-<!-- #     png::writePNG(path) -->
-
-<!-- # } -->
-
-<!-- #  -->
-
-<!-- # DiagrammeR::grViz(" -->
-
-<!-- #       digraph { -->
-
-<!-- #       graph [layout = dot, rankdir = LR] -->
-
-<!-- #  -->
-
-<!-- #       node [shape = ellipse] -->
-
-<!-- #       AGC; -->
-
-<!-- #       SOC; -->
-
-<!-- #       node [shape = rectangle] -->
-
-<!-- #       climate; -->
-
-<!-- #       soil; -->
-
-<!-- #       AFS; -->
-
-<!-- #  -->
-
-<!-- #       # Connect nodes with edges and labels -->
-
-<!-- #       AFS -> AGC; -->
-
-<!-- #       climate -> AGC; -->
-
-<!-- #       AFS -> SOC; -->
-
-<!-- #       soil -> AGC; -->
-
-<!-- #       soil -> SOC; -->
-
-<!-- #       climate -> SOC; -->
-
-<!-- #       edge[style = 'dashed'] -->
-
-<!-- #       AGC -> SOC; -->
-
-<!-- #        } -->
-
-<!-- #  -->
-
-<!-- #       ") %>% -->
-
-<!-- #   save_png("figures/dag.png") -->
-
-<!-- include_graphics("figures/dag.png")  -->
-
-<!-- ``` -->
-
-</br>
-
-```{r fig.id="choice-clim-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with three climate variables: mean annual temperature (MAT), mean annual precipitation (MAP) and precipitation variability (PVar).
-Values for the climate variables were extracted from CHELSA V2.1 (Karger et al., 2017) for each of the 30 study plots.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
-g1 <- data_plot[, c("temp", "prec", "varp")] |>
-  setnames(c("MAT", "MAP", "Pvar")) |>
-  cor() |>
-  melt() |>
-  ggplot(aes(x = Var1, y = Var2, label = round(value, 2), fill = value)) +
-  geom_tile() +
-  geom_text() +
-  scale_fill_gradientn(colours = c("brown1", "#FFFFFF", "cornflowerblue")) +
-  theme_bw() +
-  labs(x = NULL, y = NULL, fill = "Pearson\nCorrelation\nCoefficient")
-
-g2 <- data_plot[, c("temp", "prec", "varp")] |>
-  setnames(c("MAT", "MAP", "Pvar")) |>
-  scale() |>
-  princomp() |>
-  autoplot(
-    data = data_plot, colour = "site", loadings.colour = "black",
-    loadings.label = TRUE, loadings.label.col = "black"
-  ) +
-  theme_classic()
-
-ggpubr::ggarrange(g1, g2, labels = "auto")
-```
-
-</br>
-
-```{r fig.id = "choice-soil-vars", fig.width = 10, fig.height=4, fig.cap="(a) Pearson coefficient of correlation matrix and (b) principal component analysis biplot, with four soil variables: clay, silt and sand content, and soil pH.
-Values for the soil variables were measured from in situ soil samples for each of the 30 study plots.", fig.lp="supp-fig", results = "show", fig.cap.pre="Figure S"}
-g1 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n")] |>
-  setnames(c("clay", "silt", "sand", "pH")) |>
-  cor() |>
-  melt() |>
-  ggplot(aes(x = Var1, y = Var2, label = round(value, 2), fill = value)) +
-  geom_tile() +
-  geom_text() +
-  scale_fill_gradientn(colours = c("brown1", "#FFFFFF", "cornflowerblue")) +
-  theme_bw() +
-  labs(x = NULL, y = NULL, fill = "Pearson\nCorrelation\nCoefficient")
-
-g2 <- data_soil_melt[, c("CLAY", "silt", "sand", "pH\n")] |>
-  setnames(c("clay", "silt", "sand", "pH")) |>
-  scale() |>
-  princomp() |>
-  autoplot(
-    data = data_plot, colour = "site", loadings.colour = "black",
-    loadings.label = TRUE, loadings.label.col = "black"
-  ) +
-  theme_classic()
-ggpubr::ggarrange(g1, g2, labels = "auto")
-```
-
-<!-- ```{r dag-code} -->
-
-<!-- # dag { -->
-
-<!-- # bb="-2.799,-2.922,2.216,3.074" -->
-
-<!-- # afs [exposure,pos="0.265,-1.155"] -->
-
-<!-- # agc [exposure,pos="-0.591,0.840"] -->
-
-<!-- # climate [adjusted,pos="-1.049,-1.155"] -->
-
-<!-- # soc [outcome,pos="0.137,1.233"] -->
-
-<!-- # soil [adjusted,pos="-0.426,-1.182"] -->
-
-<!-- # afs -> agc -->
-
-<!-- # afs -> soc -->
-
-<!-- # agc -> soc -->
-
-<!-- # climate -> agc -->
-
-<!-- # climate -> soc -->
-
-<!-- # soil -> agc -->
-
-<!-- # soil -> soc -->
-
-<!-- # } -->
-
-<!-- ``` -->

### 2025-04-29 manuscript.Rmd
remove unused code

@@
 # use CWD values from Chave et al., 2014
-# download from figshare https://figshare.com/articles/dataset/Root_Shoot_global_data_individual_trees/5144164?file=8754550
-# rs <- readxl::read_excel("C:/Users/piponiot-laroche/Downloads/S1Database.xlsx",
-#                          sheet = "DATA")
-# rs$BGB_kg <- as.numeric(rs$BGB_kg)
-# rs$wd <- as.numeric(rs$wd)
-# rs$dbh2 <- rs$DBH_cm^2
-# lm(log(BGB_kg/AGB_kg) ~ DBH_cm + dbh2 + wd,
-#    data = subset(rs, !is.na(BGB_kg) & BGB_kg != 0))
-
 data_sim_agc[, RS := exp(-1.2312 - 0.0215 * DIAM_130_CM +
@@
 
-# limit of detectability: tree of 30 cm CBH, mean WD and height (for this DBH)
-# LOD = computeAGB(D = 30/pi,
-#                  WD = unique(data_wd$meanWD[data_wd$genus=="NR"]),
-#                  H = data_tree[DIAM_130_CM == 30/pi, mean(HEIGHT)])
-# data_sim_empty[iter != "mean", AGC := runif(sum(iter != "mean"), 0, LOD)]
-# data_sim_empty[iter == "mean", AGC := LOD/2]
-# RS = mean(data_sim_agc$RS[data_sim_agc$DIAM_130_CM==30/pi])
-# data_sim_empty[, BGC := AGC * RS]
-
-# data_sim_empty = melt(data_sim_empty, measure.vars = c("AGC", "BGC"))
-
 data_sim <- rbind(data_sim[, colnames(data_sim_empty), with = FALSE], data_sim_empty)
@@
 ## add intermediate soil carbon in 10-20 cm depth under the assumption of an
-# organic carbon content that decreases as a negative exponential of depth, we can deduce the carbon content at 10-20 cm with the following function: SOC[10-20] = SOC[0-10]*sqrt(SOC[20-30]/SOC[0-10])
+# organic carbon content that decreases as a negative exponential of depth, 
+# we can deduce the carbon content at 10-20 cm with the following function: 
+# SOC[10-20] = SOC[0-10]*sqrt(SOC[20-30]/SOC[0-10])
 data_soil <- merge(
@@
   , .(
-    mfine = c(Sample_DRY, rnorm(1000, Sample_DRY, Sample_DRY * 0.1)), # 10% error on fine soil mass
-    Corg = c(Corg, rnorm(1000, Corg, Corg * 0.12)), # 12% error on C content measurement
+    # 10% error on fine soil mass
+    mfine = c(Sample_DRY, rnorm(1000, Sample_DRY, Sample_DRY * 0.1)), 
+    # 12% error on C content measurement
+    Corg = c(Corg, rnorm(1000, Corg, Corg * 0.12)), 
     iter = c("mean", paste0("V", 1:1000))
@@
 ```
-
-<!-- ```{r tab.id = "anova", results = "show", tab.cap = "Results of the two 2-way ANOVAs with log-transformed median AGC and SOC as dependent variables, and AFS and site as independent variables.
-The p-value gives the probability of the means of the categories of each independent variable being the same.
-Bold p-values are lower than 0.05, and the mean difference between groups is thus considered significantly different than zero."} -->
-
-<!-- anova_table$factor = factor(anova_table$factor) -->
-
-<!-- levels(anova_table$factor) = c("AFS", "site") -->
-
-<!-- flextable(anova_table, col_keys = c("variable", "factor", "pvalue"),  -->
-
-<!--           cwidth = c(1, 1, 0.8)) |>  -->
-
-<!--   bold(i = anova_table$pvalue < 0.05, j = 2:3) |>  -->
-
-<!--   set_header_labels(values = c("Dependent variable", "Independent variable", "p-value")) |>  -->
-
-<!--   merge_at(i = 1:2, j = 1) |> merge_at(i = 3:4, j = 1) |>  -->
-
-<!--   fix_border_issues() |> hline(i=2) -->
-
-<!-- ``` -->
 
